### 2022-08-23 strengthmating.Rmd
add rmarkdown for ms

---
title: "Grip strength and mating success in US adults"
author: "C. B. Smith and E. H. Hagen"
date: "8/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
source('analysis.R')
```

## Abstract

Previous research posits that the dramatic male bias in upper body strength is due to sexual selection, either from advantages conferred by greater strength in male-male contests or female preference for stronger mates. Others have proposed that strength is a cue of important mate qualities, such as health or immune function. The sexual selection hypothesis is supported by studies that have found that grip strength is associated with both mating success and reproductive success in men. Although there is some evidence that strength is associated with reproductive success in women, the relationship between female strength and mating outcomes is poorly understood.

Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey (NHANES), a large representative sample of the US (N = 3328), the current study found that grip strength was a significant positive predictor of several outcome measures of adult mating success (partnership status, lifetime number of sexual partners, and past year number of sexual partners), e.g., 3-5 more lifetime partners in those with strength 1 SD above vs. 1 SD below the mean, even controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds. However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive. Hence, grip strength is positively related to mating success for females as well as males, contrary to some versions of the sexual selection hypothesis. We discuss possible explanations for these findings.

#Introduction

#Methods

All measures were obtained from the NHANES 2011-2012 (G) and 2013-2014 (H) survey cycles (See SI for NHANES variable names).

## Outcome Variables: mating success

Mating success models included either umber of sexual partners, past year and lifetime, or partnered status. Participants self-reported the number of sexual partners from the past year and in their lifetime. We used a poisson distribution to model number of partners. All  models of number of sexual partners included partnered status as a predictor. Partnered status was a binary outcome variable assigned a value of 1 if participants reported being either married or living with a partner, and a 0 if participants reported being single, divorced, or separated.

## Outcome variables: health

We also modeled white blood cell count, a measure of innate immunity.

## Predictor Variables

Grip strength, an index of upper body strength and physical formidability, was measured 3 times on each hand using a dynamometer. We used combined grip strength, which was the sum of the highest reading of each hand. All regression models included sex (male/female), age, grip strength, and an age $\times$ strength interaction term, per the original paper, along with potential confounds described below [@hagen2016].

## Potential confounds

We specified four regression models in order to determine if the effect of grip strength on mating success was due to confounds with a range of anthropometric, socioeconomic, health, and hormone variables that have been associated with grip strength.

The anthropometric model included our main predictor variables plus standing height, weight, BMI≥30, and an interaction term BMI≥30:sex. The socioeconomic model included education level, whether the participant was living alone, and Poverty Income Ratio (PIR). PIR was calculated as the ratio of family income to the local poverty threshold for each participant, controlling for family size (range 0-5, 0 indicates no income; 1 is income equal to the poverty threshold; values greater than 5 set to 5 to protect anonymity). (Include ethnicity? need explicit justification)

The hormone model included only serum total testosterone because thyroid-stimulating hormone and free thyroxine (T4 free) were not collected in the 2013-2014 NHANES survey cycle.

The health model included white blood cell count, hemoglobin, and perceived abnormal weight. We use a single item, "Special Equipment" (PFQ054) (e.g., cane, walker), to control for physical disability. We also created two variables: "Chronic disease score" (CDS) and "Physical disease difficulty" (PDD) since there is a bidirectional relationship between depression and chronic illness [@katon2011]. The CDS was a score (0-6) of chronic diseases including diabetes, cancer, stroke, arthritis, heart disease, and respiratory disease (asthma, emphysema, or chronic bronchitis). A point was added for each disease a participant indicated they had been diagnosed with, regardless of any impairment due to the disease [@jokela2019]. Physical disease difficulty was calculated from a separate NHANES question, which asked participants to list up to five health conditions which specifically caused them to have difficulties with physical activities (0-5) (we excluded counts for "depression/anxiety/emotional problem" and for "other impairment/problem" because these conceptually overlapped with our depression outcome variable).

#Analysis

#Results
```{r summarytable}
designsG$d.design.adults %>%
  tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi, 
                             whitebloodcell, strength, testosterone, partnered, 
                             height, weight, special_equipment)) 
```

```{r anthropometric}

```

```{r }

```

### 2022-08-23 analysis.R

@@ -6,6 +6,7 @@ library(hagenutils)
 library(patchwork)
 library(foreign)
 library(effects)
+library(gtsummary)
 #library(sjstats)
 
 #remove 10 males who report taking testosterone supplements
@@ -549,3 +550,25 @@ plot(allEffects(m))
 #only asks about hormonal birth control pills and depo-provera :|
 
 
+designsG$d.design.adults %>%
+  tbl_svysummary(include = c(age, edu, maritalstatus, income, race, bmi,
+                             whitebloodcell, strength, testosterone, partnered,
+                             height, weight, special_equipment),
+                 statistic = list(all_continuous() ~ "{mean} ({sd})",  # stats and format for continuous columns
+                                  all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
+                 digits = all_continuous() ~ 2,          # rounding for continuous columns
+                 #type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
+                 type = list(where(is.numeric) ~ "continuous"),
+                 label  = list(                                   # display labels for column names
+                   age  = "Age (years)",
+                   edu = "Education (highest level completed)" ,
+                   maritalstatus = "Marital Status",
+                   income = "2021 Household Income ($)",
+                   bmi = "BMI (kg/m^2)"),
+                 missing_text = "Missing"             # how missing values should display
+  )
+
+designsG$d.design.adults %>%
+  tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi,
+                             whitebloodcell, strength, testosterone, partnered,
+                             height, weight, special_equipment))

### 2022-10-07 strengthmating.Rmd
added models

@@
 ```
+EHB Research Report
 

### 2022-10-07 analysis.R

@@ -76,14 +76,21 @@ m14 <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + height_cent
 summary(m14)
 plot(allEffects(m14))
 nobs(m14)
+nobs(mp)
+
+mp <- svyglm(partnered ~ sex + age, family = quasibinomial(), design=designsG$d.design)
+summary(mp) #no sex difference after controlling for age in adults 18-60
+
 
 # m14.1 <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + height_centered*sex + weight_centered * sex +
 #                   height_centered * weight_centered, family=quasibinomial(), design=designsG$d.design.adults)
 # summary(m14.1)
 
-m14.2 <- svyglm(partnered ~ strength_centered*sex + armlength + leglength + height_centered + age_centered*sex + weight_centered, family=quasibinomial(), design=designsG$d.design.adults)
+designsG$d.design.adults <- update(designsG$d.design.adults, leglenth_prop = leglength/height)
+m14.2 <- svyglm(partnered ~ strength_centered*sex + leglenth_prop*sex + age_centered*sex + weight_centered, family=quasibinomial(), design=designsG$d.design.adults)
 summary(m14.2)
-plot(allEffects(m14.2))
+plot(allEffects(m14.2)) #marginal negative effect of armlength on partnered, no theory
+Anova(m14.2, type = 3)
 
 # m15 <- svyglm(partnered ~ height_centered*sex + age_centered*sex, family=quasibinomial(), design=designsG$d.design.adults)
 # summary(m15)
@@ -92,6 +99,7 @@ plot(allEffects(m14.2))
 m13 <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + race, family=quasibinomial(), design=designsG$d.design.adults)
 summary(m13)
 nobs(m13)
+
 
 # visreg(m13, by = "race", xvar = "sex",  scale = "response", rug = FALSE, gg = TRUE) + coord_flip()
 #
@@ -173,11 +181,12 @@ plot(emmeans(m, specs = "strength_centered", by = "sex", type = "response"))
 # visreg(m)
 
 # lifetime partners (anthropometric model)
-manth1 <- svyglm(sex_partners ~ age_centered*sex + strength_centered*sex + height_centered*sex + weight_centered + bmi_centered*sex + partnered, family=quasipoisson(), design=designsG$d.design.adults)
[…]

### 2022-10-07 pre_reg.R

@@ -1,0 +1,179 @@
+library(survey)
+library(tidyverse)
+library(nhanesGH)
+library(visreg)
+library(hagenutils)
+library(patchwork)
+library(foreign)
+library(effects)
+library(gtsummary)
+
+# compute occupational expenditure per Lassek and Gaulin
+# we have MET from participant reported occ exertion, a better var than this, also NHANES recommneds using
+# diff weights for dietary data
+
+occ_exertion <- svyby(~avgcalories, ~census_code_best, design = designsG$d.design.adults, svymean)
+occ_exertion_dict <- occ_exertion$avgcalories
+names(occ_exertion_dict) <- occ_exertion$census_code_curr
+designsG$d.design.adults <- update(designsG$d.design.adults, occ_exertion = occ_exertion_dict[census_code_curr])
+
+
+x <- svyby(~avgcalories, ~race, design = designsG$d.design.adults, svymean, na.rm = T)
+
+
+# use both number of partners and num vaginal partners, missing 743 who report having vaginal sex but are missing number of vaginal sex partners
+# much fewer (~56) missing number of total sex partners
+# around 686 ppl skipped whole module?
+
+
+# conceptual rep of Lassek and Gaulin, using grip strength and including males and females
+
+# Model 1 (Table 2) Lifetime sexual partners ------------------------------
+
+m2_totalpartners_exact <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    sex * strength_centered +
+    partnered +
+    bmi_centered,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+summary(m2_totalpartners_exact)
+
+m2_totalpartners_mod <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered +
+    partnered +
+  #  bmi_centered*sex +
+  #  edu +
+    median_salary_current +
+    age_first_sex*sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+summary(m2_totalpartners_mod)
+plot(allEffects(m2_totalpartners_mod))
+
+ggplot(d_G, aes(age_first_sex, sex_partners)) + geom_count() + scale_y_log10() +facet_grid(I(cut(age, 5))~sex) +geom_smooth()
+
+m2 <- svyglm(
+  vaginal_sex_partners ~
+    age_centered * sex +
[…]

### 2023-03-30 ms_draft.Rmd
new analysis files and ms rmd

---
title: "Grip strength, mating success, and immune function in US adults: a registered report "
author: "C. B. Smith and E. H. Hagen"
date: 03/23/2023
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
library(weights)
library(jtools)
library(reshape2)
library(gtsummary)
library(corrplot)
source('draft_analysis.R')
```
EHB Research Report

## Abstract

Evidence suggests that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection. It is unknown if females experience a similar tradeoff. Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey, a large representative sample of the US, we will test whether grip strength, a proxy for upper body strength, is a positive predictor of self-reported outcome measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, and age at first intercourse) in both males and females, and negatively related to a measure of native immune function, white blood cell count, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds. These results replicate an earlier study of US males. However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive, evidence that a similar trade off in the costs and benefits of muscle exists for women.

## Methods

To assess the relationship between strength and reproductive and immune outcomes we used data from the Centers for Disease Control National Health and Nutrition Examination Survey (NHANES) from years 2011-2014. NHANES uses a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S. population. NHANES combines interview, examination, and laboratory data to assess health status and identify health risks for adults and children in the United States. Data collection occurs in new cycles every two years.

We will use data from years 2011-2014 since grip strength, our key predictor variable was collected in those years. We developed models based on previous research and through exploratory analyses of the 2011-2012 dataset to be tested in the 2013-2014 dataset. We include data from US adults between the ages of 18 and 65, the years when grip strength is the most stable. For each model, participants will be included if they have complete data for each predictor variable.

### Outcome variables: 
In industrialized populations like the U.S., widespread access to contraceptives unlinks reproductive success from sexual behavior. Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).Our main outcome was mating success, which we operationalized in three ways. 1) Total sexual partners was a count based on responses to the question: In your lifetime, with how many men/women have you had any kind of sex? 2) Total sexual partners in past year, from the question: In the past 12 months, with how many men/women have you had any kind of sex? Because of the way these questions are framed, (e.g. female participants were asked about their male partners and vice versa) these variables represent heterosexual partners, however it is not necessarily the case that all of the participants included in these models identified as heterosexual. 3) Age at first sexual intercourse reported from: How old were you when you had sex for the first time?. The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen. Only respondents who could self-report were asked these questions.

Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL). In the original paper, Lassek and Gaulin also use C-reactive protein, but CRP was not measured in the 2011-2014 data collection years and so is not included here.

### Predictor variables: 
Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume. We operationalized formidability as grip strength (kg), which was measured three times on each hand using a dynamometer. We used combined grip strength which was the sum of the highest reading of each hand. All models also included sex (male/female), age in years, partnered status (partnered included participants who reported either being married or living with a partner while unpartnered included participants who reported being single, widowed, divorced, or separated). The interaction between sex and strength was crucial for our models since we want to know whether the effect of strength on mating and immune outcomes is different for women than for men. Likewise, we also interacted age and partnered status with sex.

### Models:
We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.

### Baseline (anthropometric) model
We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could be related to either strength or mating outcomes.

### Additional control variables: 
The socioeconomic control model included education and race as categorical variables. The health model included a number of variables related to health, which could affect either strength or sexual behavior. White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection. Depression was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument. Each item represents a symptom of depression, and for each participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday). These ratings were summed to produce a depression score ranging from 0-27. Chronic illness was included using a 'chronic disease score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease. A point was added for each disease a participant indicated they had been diagnosed with, regardless of any impairment due to the disease. We then controlled for impairment resulting from chronic illness (Disease Impairment Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities. We also controlled for physical disability and using the item "special equipment" which referred to participants' report that they needed special equipment to walk. Finally the 'perceived abnormal weight' variable was coded as true or false depending on whether a participant reported that they perceived their weight to be abnormal.

The hormone control model included testosterone. Since the difference in mean testosterone between men and women is an order of magnitude apart, and testosterone was highly confounded with sex. We therefore centered testosterone at the mean and divided by two standard deviations within each sex. The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they did not.

Current salary Measures of wealth available in NHANES are collected at the household level, and as such do not accurately reflect wealth or poverty at the individual level. As a measure of wealth, we used occupation codes collected for each participant by NHANES and imputed the median salary calculated for that occupation category (source?).

###

In keeping with the outcome variables utilized in the original paper, we used number of partners for any kind of sex above. Since, however, NHANES also collects more specific data on sexual behavior we also tested Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex. In these questions participants were asked “In your lifetime, with how many men/women have you had vaginal sex?” and “In the past 12 months, with how many men/women have you had vaginal sex?”

Missing variables 
We do not have data on bone mineral density

We did not include a model of daily energy intake as reported by Lassek and Gaulin, because the gold-standard research utilizes fat-free mass as a predictor, which the original authors use. Grip strength is therefore unlikely to provide novel insight into explaining daily energy intake.

## Analysis 
Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models. Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models from the survey package to incorporate survey weights. For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data. Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.

## Results

### Exploratory Analysis

```{r summarytable, fig.width = 12, fig.height = 16}
d_G[d_G$age >= 18 &
      d_G$age <= 60, ]  %>% select(
        "sex_partners",
        "sex_partners_year",
        "age_first_sex",
        "sex",
        "strength",
        "age",
        "partnered",
        "bmi",
        "edu",
        "race",
        "whitebloodcell",
        "hemoglobin",
        "depression",
        "chronic_disease_score",
        "physical_disease_count",
        "special_equipment",
        "perceived_abnormal_weight",
        "testosterone",
        "vigorous_work",
        "vigorous_rec",
        "moderate_work",
        "moderate_rec"
      ) %>%
  mutate(race = factor(
    race,
    labels = c(
      "Mexican American",
      "Other Hispanic",
      "Non-Hispanic White",
      "Non-Hispanic Black",
      "Non-Hispanic Asian",
      "Other Race"
    )
  )) %>%
  mutate(edu = factor(
    edu,
    labels = c(
      "Less than 9th grade",
      "9-11th grade",
      "High school graduate/GED or equivalent",
      "Some college or AA degree",
      "College graduate or above"
    )
  )) %>%
  tbl_summary(
    by = sex,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous"
    ),
    label  = list(
      age_first_sex = "Age at first sex (years)",
      sex_partners = "Lifetime number of sexual partners",
      sex_partners_year = "Past year number of sexual partners",
      strength = "Combined Grip Strength (kg)",
      age = "Age (years)",
      bmi = "Body mass index (kg/m^2)",
      partnered = "Partnered",
      edu = "Education",
      race = "Race and Ethnicity",
      whitebloodcell = "White blood cell count (1000 cells/µL)",
      hemoglobin = "Hemoglobin (g/dL)",
      testosterone = "Testosterone (ng/dL)",
      perceived_abnormal_weight = "Perceived abnormal weight",
      special_equipment = "Special equipment needed to walk",
      chronic_disease_score = "Chronic Disease Score (0-6)",
      physical_disease_count = "Disease Impairment Score (0-5)",
      depression = "Depression Score (0-27)",
      vigorous_work = "Work involves vigorous activity",
      moderate_work = "Work involves moderate activity",
      vigorous_rec = "Recreation involves vigorous activity",
      moderate_rec = "Recreation involves moderate activity"
    ),
    missing_text = "Missing"
  ) %>% add_p() %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
  modify_caption("**Table 1. Participant Characteristics**") %>%
  bold_labels()

```

```{r}

designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous"
    ),
    label  = list(
      age_first_sex = "Age at first sex (years)",
      sex_partners = "Lifetime number of sexual partners",
      sex_partners_year = "Past year number of sexual partners",
      strength = "Combined Grip Strength (kg)",
      age = "Age (Years)",
      bmi = "Body mass index (kg/m^2)",
      height = "Height (cm)",
      weight = "Weight (kg)",
      partnered = "Partnered",
      edu = "Education",
      race = "Race and Ethnicity",
      whitebloodcell = "White blood cell count (1000 cells/µL)",
      hemoglobin = "Hemoglobin (g/dL)",
      testosterone = "Testosterone (ng/dL)",
      perceived_abnormal_weight = "Perceived abnormal weight",
      special_equipment = "Special equipment needed to walk",
      chronic_disease_score = "Chronic Disease Score (0-6)",
      physical_disease_count = "Disease Impairment Score (0-5)",
      depression = "Depression Score (0-27)",
      vigorous_work = "Work involves vigorous activity",
      moderate_work = "Work involves moderate activity",
      vigorous_rec = "Recreation involves vigorous activity",
      moderate_rec = "Recreation involves moderate activity"
    ),
    missing_text = "Missing",
    include = c(
      "sex_partners",
      "sex_partners_year",
      "age_first_sex",
      "sex",
      "strength",
      "age",
      "partnered",
      "bmi",
      "edu",
      "race",
      "whitebloodcell",
      "hemoglobin",
      "depression",
      "chronic_disease_score",
      "physical_disease_count",
      "special_equipment",
      "perceived_abnormal_weight",
      "testosterone",
      "vigorous_work",
      "vigorous_rec",
      "moderate_work",
      "moderate_rec"
    )
  ) %>%
  add_p() %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
  modify_caption("**Table 1. Participant Characteristics**") %>%
  bold_labels()

```

```{r exact, fig.width = 12, fig.height = 8}

forestplot(
  m_agefirst,
  m_lifetime,
  m_pastyear,
  intercept = F,
  facet = F,
  dodgewidth = .5,
  modelnames = mnames1,
  varnames = vnames
)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
  labs(title = "Exact Models: Regression coefficients from generalized linear models", color = "Model") +
  guides(colour = guide_legend(reverse = T))

#visualize coefs from models derived from lassek and gaulin. 
#add coef table below to match l and g

```

```{r table2_exact}

mm1 <- tbl_regression(m_lifetime, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm2 <- tbl_regression(m_pastyear, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm3 <- tbl_regression(m_agefirst, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI", 
                        edu = "Education",
                        total_work_MET = "Occupational exertion"
    #              label = list(
    # age_centered = "Age",
    # sex = "Sex",
    # strength_centered = "Grip strength",
    # partnered = " Currently partnered", 
    # bmi_centered = "BMI"
    # edu = "Education",
    # total_work_MET = "Occupational exertion"
  )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

  tbl_merge(
    tbls = list(mm1, mm2, mm3),
     tab_spanner = c("Lifetime number of sexual partners",
                     "Past year number of sexual partners",
                     "Age at first sexual intercourse")) %>% 
  as_gt() %>%
  gt::tab_header(title = "Table 2. Exact Models of Mating Success",
                 subtitle = "Standardized regression coefficents from generalized
                 linear models for three indices of mating success for men and women ages
                 18-59. Models derived from Lassek and Gaulin 2009")

```

```{r lifetime, fig.width = 12, fig.height = 42}

#include groups of potential confounds to explore lack of sex effect for all 4 mating outcomes

fig1 <- forestplot(
  manth1, msoc1, mheal1, mphys1, mhor1a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames, varnames = vnames)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Lifetime Number of Sexual Partners", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig2 <- forestplot(
  manth2,msoc2,mheal2, mphys2, mhor2a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames
)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(.8)) +
  labs(title = "Past Year Number of Sexual Partners", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig3 <- forestplot(
  manth3, msoc3, mheal3, mphys3, mhor3a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames)$plot +
  theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Age at First Sex", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig4 <- forestplot(
  manth4, msoc4, mheal4, mphys4, mhor4a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames)$plot +
  theme_minimal(20) + geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
  labs(title = "Currently Partnered", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig1234 <- ggarrange(fig1, fig2, fig3, fig4,
                     labels = "AUTO",
                     font.label = list(size = 20),
                     ncol = 1, nrow = 4,
                     common.legend = TRUE, legend="bottom")

 annotate_figure(fig1234,
                       top = text_grob("Figure 2. Regression coefficients from generalized linear models of mating success\n",
                                       size = 20, face = "bold",
                                       hjust = 0,
                                       x = 0.01),
                       bottom = text_grob("\nVariables labelled (S) have been standarized by 2 SD. Base level of race/ethnicity is Mexican American.",
                                          size = 18))

```

```{r table2_mod}
# 
# mm1_m <- tbl_regression(mm1_mod, 
#                         label = list(
#                           age_centered = "Age",
#                           sex = "Sex",
#                           strength_centered = "Grip strength",
#                           partnered = "Partnered",
#                           median_salary_current = "Median Salary",
#                           age_first_sex = "Age at first sex"
#                         ))  %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
# mm2_m <- tbl_regression(mm2_mod, 
#                         label = list(
#                           sex = "Sex",
#                           strength_centered = "Grip strength",
#                           partnered = "Partnered",
#                           bmi_centered = "BMI"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
#     
# mm3_m <- tbl_regression(mm3_mod, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           total_work_MET = "Occupational exertion",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
#   tbl_merge(
#     tbls = list(mm1_m, mm2_m, mm3_m),
#     tab_spanner = c("Lifetime number of sexual partners",
#                     "Past year number of sexual partners",
#                     "Age at first sexual intercourse")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 3. Modified Models of Mating Success",
#                  subtitle = "Standardized regression coefficents from generalized
#                  linear models for three indicies of mating success. Models derived
#                  from exploratory data")

```

```{r table2_vaginalsex}

# mm1_v <- tbl_regression(m1_vaginal_sex_partners, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
# mm2_v <- tbl_regression(m2_vaginal_pastyear, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
#   tbl_merge(
#     tbls = list(mm1_v, mm2_v),
#     tab_spanner = c("Lifetime number of vaginal sex partners",
#                     "Past year number of vaginal sex partners")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 4. Models of Mating Success (Vaginal Sex)",
#                  subtitle = "Standardized regression coefficents from generalized
#                  linear models for lifetime and past year vaginal sex partners")

```

```{r immunefig, fig.width = 12, fig.height = 8}
fig5 <- forestplot(
  mwbc,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  varnames = vnames)$plot +
  theme_minimal(25) +
  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
  labs(title = "White blood cell count") +
  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
  #      subtitle = "White Blood Cell Count",
  #      caption = "Variables with (S) have been standarized by 2 SD.") +
  theme(legend.position = "none")

 annotate_figure(fig5,
                           top = text_grob("Figure 3. Regression coefficients from generalized linear model of immune investment\n",
                                           size = 20, face = "bold",
                                           hjust = 0,
                                           x = 0.01),
                           bottom = text_grob("Variables labelled (S) have been standarized by 2 SD.",
                                              size = 18))

```

Regression coefficients for predictors of WBC count in men and women 18-59, exact model from lassek and gaulin

```{r immunemodel}

tbl_regression(
  mwbc,
  label = list(
    age_centered = "Age",
    bmi_centered = "BMI",
    sex = "Sex",
    strength_centered = "Grip strength"
  )
) %>%
  add_significance_stars(hide_se = TRUE,
                         pattern = "{estimate}{stars}<br>({std.error})") %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

# m_lymph <- tbl_regression(m_l, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# m_mono <- tbl_regression(m_m, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# m_neutro <- m_wbc <- tbl_regression(m_n, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
# m_eos <- tbl_regression(m_e, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# m_baso <- tbl_regression(m_b, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# tbl_merge(
#     tbls = list(m_wbc, m_lymph, m_mono, m_neutro, m_eos, m_baso),
#     tab_spanner = c("White blood cell count",
#                     "Lymphocyte count", 
#                     "Monocyte count", 
#                     "Neutrophil count", 
#                     "Eosinophil count", 
#                     "Basophil count")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 4. Models of Immune Function",
#                  subtitle = "Regression coefficents from generalized
#                  linear models of white blood cell count, and cell type counts")

```

```{r immunemodelalternate, fig.width = 12, fig.height = 8}

forestplot(
  mwbc_alt,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  varnames = vnames)$plot +
  theme_minimal(25) +
  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
  labs(title = "White blood cell count") +
  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
  #      subtitle = "White Blood Cell Count",
  #      caption = "Variables with (S) have been standarized by 2 SD.") +
  theme(legend.position = "none")

```

### 2023-03-30 draft_analysis.R

@@ -1,9 +1,115 @@
-# spsp poster forestplot
 library(survey)
 library(ggplot2)
 library(nhanesGH)
 library(hagenutils)
 library(ggpubr)
+
+
+# Baseline models ---------------------------------------------------------
+
+vnames <- c(
+  "strength_centered" = "Strength (S)",
+  "sexfemale" = "Sex (Female)",
+  "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "age_centered" = "Age (S)",
+  "partneredTRUE" = "Partnered",
+  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
+  "age_centered:sexfemale" = "Age x Sex (Female)",
+  "height_centered" = "Height (S)",
+  "weight_centered" = "Weight (S)",
+  "bmi_centered" = "BMI (S)",
+  "sexfemale:bmi_centered" = "Sex (Female) x BMI",
+  "sexfemale:height_centered" = "Sex (Female) x Height",
+  "sexfemale:weight_centered" = "Sex (Female) x Weight",
+  'perceived_abnormal_weightTRUE' = "Perceived Abnormal Weight",
+  "whitebloodcell_centered" = "White Blood Cell Count (S)",
+  "hemoglobin_centered" = "Hemoglobin (S)",
+  "special_equipmentTRUE" = "Need special equip to walk",
+  "chronic_disease_score" = "Chronic Disease Score",
+  "physical_disease_count" = "Physical Disease Count",
+  "depression" = "Depression Score",
+  "log(testosterone)" = "Testosterone (log)",
+  "sexfemale:log(testosterone)" = "Sex (Female) x Testosterone (log)",
+  "testosterone_sex_centered" = "Testosterone centered by sex",
+  "sexfemale:testosterone_sex_centered"= "sex (Female) x testosterone",
+  "vigorous_rec" = "Vigorous Recreation",
+  "moderate_rec" = "Moderate Recreation",
+  "vigorous_work" = "Vigorous Work",
+  "moderate_work" = "Moderate Work",
+  "edu" = "Education",
+  "raceOtherHispanic" = "Other Hispanic",
+  "raceNonHispanicBlack" = "Non-Hispanic Black",
+  "raceNonHispanicAsian" = "Non-Hispanic Asian",
+  "raceNonHispanicWhite" = "Non-Hispanic White",
+  "raceOtherRace" = "Other Race"
+)
+
+
+m_lifetime <- svyglm(
[…]

### 2023-04-04 draft3.23.23.Rmd
adding tables to ms

--- 2023-03-30 ms_draft.Rmd
@@
 
-## Abstract
+##Abstract
 
@@
 
-## Methods
-
-To assess the relationship between strength and reproductive and immune outcomes we used data from the Centers for Disease Control National Health and Nutrition Examination Survey (NHANES) from years 2011-2014.
-NHANES uses a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
-population.
-NHANES combines interview, examination, and laboratory data to assess health status and identify health risks for adults and children in the United States.
-Data collection occurs in new cycles every two years.
-
-We will use data from years 2011-2014 since grip strength, our key predictor variable was collected in those years.
-We developed models based on previous research and through exploratory analyses of the 2011-2012 dataset to be tested in the 2013-2014 dataset.
-We include data from US adults between the ages of 18 and 65, the years when grip strength is the most stable.
-For each model, participants will be included if they have complete data for each predictor variable.
-
-### Outcome variables: 
-In industrialized populations like the U.S., widespread access to contraceptives unlinks reproductive success from sexual behavior.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).Our main outcome was mating success, which we operationalized in three ways.
-1) Total sexual partners was a count based on responses to the question: In your lifetime, with how many men/women have you had any kind of sex?
-2) Total sexual partners in past year, from the question: In the past 12 months, with how many men/women have you had any kind of sex?
-Because of the way these questions are framed, (e.g.
-female participants were asked about their male partners and vice versa) these variables represent heterosexual partners, however it is not necessarily the case that all of the participants included in these models identified as heterosexual.
-3) Age at first sexual intercourse reported from: How old were you when you had sex for the first time?.
-The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen.
-Only respondents who could self-report were asked these questions.
-
-Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL).
-In the original paper, Lassek and Gaulin also use C-reactive protein, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
-
-### Predictor variables: 
-Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume.
-We operationalized formidability as grip strength (kg), which was measured three times on each hand using a dynamometer.
-We used combined grip strength which was the sum of the highest reading of each hand.
-All models also included sex (male/female), age in years, partnered status (partnered included participants who reported either being married or living with a partner while unpartnered included participants who reported being single, widowed, divorced, or separated).
-The interaction between sex and strength was crucial for our models since we want to know whether the effect of strength on mating and immune outcomes is different for women than for men.
-Likewise, we also interacted age and partnered status with sex.
-
-### Models:
-We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
-
-### Baseline (anthropometric) model
-We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could be related to either strength or mating outcomes.
-
-### Additional control variables: 
-The socioeconomic control model included education and race as categorical variables.
-The health model included a number of variables related to health, which could affect either strength or sexual behavior.
-White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument.
-Each item represents a symptom of depression, and for each participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
-These ratings were summed to produce a depression score ranging from 0-27.
-Chronic illness was included using a 'chronic disease score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease.
-A point was added for each disease a participant indicated they had been diagnosed with, regardless of any impairment due to the disease.
-We then controlled for impairment resulting from chronic illness (Disease Impairment Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities.
-We also controlled for physical disability and using the item "special equipment" which referred to participants' report that they needed special equipment to walk.
-Finally the 'perceived abnormal weight' variable was coded as true or false depending on whether a participant reported that they perceived their weight to be abnormal.
-
-The hormone control model included testosterone.
-Since the difference in mean testosterone between men and women is an order of magnitude apart, and testosterone was highly confounded with sex.
-We therefore centered testosterone at the mean and divided by two standard deviations within each sex.
-The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they did not.
-
-Current salary Measures of wealth available in NHANES are collected at the household level, and as such do not accurately reflect wealth or poverty at the individual level.
-As a measure of wealth, we used occupation codes collected for each participant by NHANES and imputed the median salary calculated for that occupation category (source?).
-
-###
-
-In keeping with the outcome variables utilized in the original paper, we used number of partners for any kind of sex above.
-Since, however, NHANES also collects more specific data on sexual behavior we also tested Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex.
-In these questions participants were asked “In your lifetime, with how many men/women have you had vaginal sex?” and “In the past 12 months, with how many men/women have you had vaginal sex?”
-
-Missing variables 
-We do not have data on bone mineral density
-
-We did not include a model of daily energy intake as reported by Lassek and Gaulin, because the gold-standard research utilizes fat-free mass as a predictor, which the original authors use.
-Grip strength is therefore unlikely to provide novel insight into explaining daily energy intake.
-
-## Analysis 
-Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models.
-Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models from the survey package to incorporate survey weights.
-For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
-Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.
-
-## Results
-
-### Exploratory Analysis
-
-```{r summarytable, fig.width = 12, fig.height = 16}
-d_G[d_G$age >= 18 &
-      d_G$age <= 60, ]  %>% select(
-        "sex_partners",
-        "sex_partners_year",
-        "age_first_sex",
-        "sex",
-        "strength",
-        "age",
-        "partnered",
-        "bmi",
-        "edu",
-        "race",
-        "whitebloodcell",
-        "hemoglobin",
-        "depression",
-        "chronic_disease_score",
-        "physical_disease_count",
-        "special_equipment",
-        "perceived_abnormal_weight",
-        "testosterone",
-        "vigorous_work",
-        "vigorous_rec",
-        "moderate_work",
-        "moderate_rec"
-      ) %>%
-  mutate(race = factor(
-    race,
-    labels = c(
-      "Mexican American",
-      "Other Hispanic",
-      "Non-Hispanic White",
-      "Non-Hispanic Black",
-      "Non-Hispanic Asian",
-      "Other Race"
-    )
-  )) %>%
-  mutate(edu = factor(
-    edu,
-    labels = c(
-      "Less than 9th grade",
-      "9-11th grade",
-      "High school graduate/GED or equivalent",
-      "Some college or AA degree",
-      "College graduate or above"
-    )
-  )) %>%
-  tbl_summary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} / {N} ({p}%)"
-    ),
+```{r summarytable}
+d_G[d_G$age>=18 & d_G$age<=60,]  %>% select("sex", "age", "race", "strength", "bmi", "partnered", "edu", "height", "weight", "whitebloodcell", "testosterone") %>%
+  tbl_summary(by = sex,
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} / {N} ({p}%)"),
               digits = all_continuous() ~ 2,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
+        #      type = list(where(is.numeric) ~ "continuous"),
+             # type = list(partnered ~ "categorical"),
+              type = all_dichotomous() ~ "categorical",
               label  = list(
-      age_first_sex = "Age at first sex (years)",
+                age = "Age (Years)",
+                race = "Race and Ethnicity",
+                strength = "Combined Grip Strength",
+                bmi = "BMI (kg/m^2)",
+                partnered = "Partnered Status",
+                edu = "Education", 
+                height = "Height (cm)", 
+                weight = "Weight (kg)", 
+                whitebloodcell = "WBCC (1000 cells/µL)", 
+                testosterone = "Testosterone (ng/dL)",
+                age_first_sex = "Age at first sex (Years)", 
                 sex_partners = "Lifetime number of sexual partners", 
-      sex_partners_year = "Past year number of sexual partners",
-      strength = "Combined Grip Strength (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
-    ),
+                sex_partners_year = "Past year number of sex partners"), 
               missing_text = "Missing"
@@
   modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>% 
-  modify_caption("**Table 1. Participant Characteristics**") %>%
-  bold_labels()
+    modify_caption("**Table 1. Participant Characteristics**")
+
+# designsG$d.design.adults %>%
+#   tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi, 
+#                              whitebloodcell, strength, testosterone, partnered, 
+#                              height, weight, special_equipment)) 
+```
+
+```{r cor, fig.width=12, fig.height=12 }
+# cordat <- d_G[, c("age", "sex1", "strength", "bmi")]
+# cordat <- cordat[cordat$age>=18 & cordat$age<60,]
+# cormat <- round(cor(cordat, use = "complete.obs"),2)
+
+cor_mat <- svycor(
+  ~ sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
+    physical_disease_count +
+    testosterone +
+    vigorous_rec +
+    moderate_rec +
+    vigorous_work +
+    moderate_work,
+  designsG$d.design.adults,
+  na.rm = T
+)
+
+cor_mat_cors <- cor_mat$cors
+corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
+
+# meltedcor_mat <- melt(cor_mat_cors)
+# 
+# ggplot(data = meltedcor_mat, aes(x=Var1, y=Var2, fill = value)) +
+#   geom_tile()
 
@@
 
-```{r}
+```{r exact, fig.width = 12, fig.height = 12}
 
-designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} / {N} ({p}%)"
-    ),
-    digits = all_continuous() ~ 2,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
-    label  = list(
-      age_first_sex = "Age at first sex (years)",
-      sex_partners = "Lifetime number of sexual partners",
-      sex_partners_year = "Past year number of sexual partners",
-      strength = "Combined Grip Strength (kg)",
-      age = "Age (Years)",
-      bmi = "Body mass index (kg/m^2)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
-    ),
-    missing_text = "Missing",
-    include = c(
-      "sex_partners",
-      "sex_partners_year",
-      "age_first_sex",
-      "sex",
-      "strength",
-      "age",
-      "partnered",
-      "bmi",
-      "edu",
-      "race",
-      "whitebloodcell",
-      "hemoglobin",
-      "depression",
-      "chronic_disease_score",
-      "physical_disease_count",
-      "special_equipment",
-      "perceived_abnormal_weight",
-      "testosterone",
-      "vigorous_work",
-      "vigorous_rec",
-      "moderate_work",
-      "moderate_rec"
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered  +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
 )
-  ) %>%
-  add_p() %>%
-  add_overall() %>%
-  add_n() %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
-  modify_caption("**Table 1. Participant Characteristics**") %>%
-  bold_labels()
 
-```
+m_pastyear <- svyglm(
+  sex_partners_year ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered * strength_centered +
+     bmi_centered * sex,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
 
-```{r exact, fig.width = 12, fig.height = 8}
+m_agefirst <-  svyglm(
+    age_first_sex ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered  +
+      bmi_centered * sex,
+    family = gaussian(),
+    design = designsG$d.design.adults
+  )
 
-forestplot(
-  m_agefirst,
-  m_lifetime,
-  m_pastyear,
-  intercept = F,
-  facet = F,
-  dodgewidth = .5,
-  modelnames = mnames1,
-  varnames = vnames
-)$plot + theme_minimal(20) +
+mnames1 <- c(
+  "Age at first intercourse",
+  "Lifetime Number of Sexual Partners",
+  "Past Year Number of Sexual Partners"
+)
+
+vnames1 <- c(
+  "strength_centered" = "Strength (S)",
+  "sexfemale" = "Sex (Female)",
+  "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "age_centered" = "Age (S)",
+  "partneredTRUE" = "Partnered",
+  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
+  "age_centered:sexfemale" = "Age x Sex (Female)",
+  "height_centered" = "Height (S)",
+  "weight_centered" = "Weight (S)",
+  "bmi_centered" = "BMI (S)",
+  "sexfemale:bmi_centered" = "Sex (Female) x BMI"
+)
+
+forestplot(m_agefirst, m_lifetime, m_pastyear, intercept = F, facet = F, dodgewidth = .5, modelnames = mnames1,
+  varnames = vnames1)$plot + theme_minimal(20) +
   geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
-  labs(title = "Exact Models: Regression coefficients from generalized linear models", color = "Model") +
-  guides(colour = guide_legend(reverse = T))
+labs(title = "Exact Models: Regression coefficients from generalized linear models", color = "Model") + guides(colour = guide_legend(reverse = T))
 
@@
 #add coef table below to match l and g 
+#decrease dodgewidth, reorder models to match points
 
@@
 
-```{r immunefig, fig.width = 12, fig.height = 8}
+```{r immunefig}
 fig5 <- forestplot(
@@
 ```
-
-```{r immunemodelalternate, fig.width = 12, fig.height = 8}
-
-forestplot(
-  mwbc_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  varnames = vnames)$plot +
-  theme_minimal(25) +
-  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
-  labs(title = "White blood cell count") +
-  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
-  #      subtitle = "White Blood Cell Count",
-  #      caption = "Variables with (S) have been standarized by 2 SD.") +
-  theme(legend.position = "none")
-
-```

### 2023-04-04 ms_draft.Rmd
adding tables to ms

@@
 Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.
+Analyses include adults ages 18-59 because this is period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
 
-## Results
+## 3. Results
 
-### Exploratory Analysis
+### 3.1 Exploratory Analysis
+
+#### 3.1.1 Descriptive statistics
+
+Table one shows weighted means and standard deviations of all variables for men and women ages 18-59.
+Mean age was 38 years for this sample.
+Combined grip strength, height, weight, and testosterone were significantly higher in males than females.
+while BMI was significantly higher in females than males (a reversal from Lassek and Gaulin).
+
+```{r table1b}
+
+# has ratio and cohen's d but not sds
+tabb <- d_G[d_G$age >= 18 &
+              d_G$age <= 59, ]  %>% select(
+                "sex",
+                "age",
+                "height",
+                "weight",
+                "bmi",
+                "strength",
+                "whitebloodcell",
+                "hemoglobin",
+                "testosterone",
+              ) %>%
+  tbl_summary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "cohens_d")
+
+#has sds and ratio but not cohen's d
+
+sd_gts <- function(data, variable, ...) {
+  sd(data[[variable]], na.rm = TRUE)
+}
+
+tabb2 <- d_G[d_G$age >= 18 &
+               d_G$age <= 59, ]  %>% select(
+                 "sex",
+                 "age",
+                 "height",
+                 "weight",
+                 "bmi",
+                 "strength",
+                 "whitebloodcell",
+                 "hemoglobin",
+                 "testosterone",
+               ) %>%
+  tbl_strata(
+  strata = sex,
+  ~ .x %>%
+    tbl_summary(
+ #     by = sex,
+      statistic = list(all_continuous() ~ "{mean}"),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
+        age = "Age (years)",
+        bmi = "Body mass index (kg/m^2)",
+        whitebloodcell = "White blood cell count (1000 cells/µL)",
+        hemoglobin = "Hemoglobin (g/dL)",
+        testosterone = "Testosterone (ng/dL)"
+      ),
+      missing = "no"
+    ) %>%
+   add_n() %>%
+    add_stat(fns = everything() ~ sd_gts)) %>%
+  # modify_fmt_fun(list(add_stat_1 ~ partial(style_number, digits = 2))) %>%
+  modify_header(list(stat_0_1 ~ "**Mean**",
+                     stat_0_2 ~ "**Mean**",
+                     add_stat_1_1 ~ "**SD**",
+                     add_stat_1_2 ~ "**SD**")) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_0_1)/as.numeric(stat_0_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**")
+
+# has all!
+tabb3 <- tabb2 %>% modify_table_body(
+  ~ .x %>%
+    dplyr::mutate(d = round(tabb$table_body$estimate, digits = 2))) %>%
+  modify_header(d = "**d**") %>%
+  modify_header(label ~ "**Variable**") %>%
+  modify_spanning_header(c("ratio", "d") ~ "**Sexual dimorphism**") %>%
+  modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-59**")
+
+tabb3
+
+```
+
+```{r tab1age1849}
+
+# has ratio and cohen's d but not sds
+tabba <- d_G[d_G$age >= 18 &
+              d_G$age <= 49, ]  %>% select(
+                "sex",
+                "age",
+                "height",
+                "weight",
+                "bmi",
+                "strength",
+                "whitebloodcell",
+                "hemoglobin",
+                "testosterone",
+              ) %>%
+  tbl_summary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "cohens_d")
+
+#has sds and ratio but not cohen's d
+
+sd_gts <- function(data, variable, ...) {
+  sd(data[[variable]], na.rm = TRUE)
+}
+
+tabb2a <- d_G[d_G$age >= 18 &
+               d_G$age <= 49, ]  %>% select(
+                 "sex",
+                 "age",
+                 "height",
+                 "weight",
+                 "bmi",
+                 "strength",
+                 "whitebloodcell",
+                 "hemoglobin",
+                 "testosterone",
+               ) %>%
+  tbl_strata(
+  strata = sex,
+  ~ .x %>%
+    tbl_summary(
+ #     by = sex,
+      statistic = list(all_continuous() ~ "{mean}"),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
+        age = "Age (years)",
+        bmi = "Body mass index (kg/m^2)",
+        whitebloodcell = "White blood cell count (1000 cells/µL)",
+        hemoglobin = "Hemoglobin (g/dL)",
+        testosterone = "Testosterone (ng/dL)"
+      ),
+      missing = "no"
+    ) %>%
+   add_n() %>%
+    add_stat(fns = everything() ~ sd_gts)) %>%
+  # modify_fmt_fun(list(add_stat_1 ~ partial(style_number, digits = 2))) %>%
+  modify_header(list(stat_0_1 ~ "**Mean**",
+                     stat_0_2 ~ "**Mean**",
+                     add_stat_1_1 ~ "**SD**",
+                     add_stat_1_2 ~ "**SD**")) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_0_1)/as.numeric(stat_0_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**")
+
+# has all!
+tabb3a <- tabb2a %>% modify_table_body(
+  ~ .x %>%
+    dplyr::mutate(d = round(tabba$table_body$estimate, digits = 2))) %>%
+  modify_header(d = "**d**") %>%
+  modify_header(label ~ "**Variable**") %>%
+  modify_spanning_header(c("ratio", "d") ~ "**Sexual dimorphism**") %>%
+  modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-49**")
+
+tabb3a
+```
+
+```{r table1survey}
+tabbsvy <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  add_n() %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "smd") %>%
+  modify_column_hide(columns = ci)
+
+#get sds
+
+  tabbsvysd <- designsG$d.design.adults %>%
+    tbl_svysummary(
+      by = sex,
+      include = c(age,
+                  height,
+                  weight,
+                  bmi,
+                  strength,
+                  whitebloodcell,
+                  hemoglobin,
+                  testosterone),
+      statistic = list(
+        all_continuous() ~ "{sd}"
+      ),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
+        age = "Age (years)",
+        bmi = "Body mass index (kg/m^2)",
+        whitebloodcell = "White blood cell count (1000 cells/µL)",
+        hemoglobin = "Hemoglobin (g/dL)",
+        testosterone = "Testosterone (ng/dL)"
+      ),
+      missing = "no"
+    )
+
+  tabb3svy <- tabbsvy %>% modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(sd1 = tabbsvysd$table_body$stat_1)
+  ) %>% modify_table_body(
+    ~ .x %>%
+    dplyr::mutate(sd2 = tabbsvysd$table_body$stat_2)
+    ) %>%
+    modify_header(list(sd1 = "**SD**",
+                       sd2 = "**SD**")) %>%
+    modify_table_body(~.x %>% dplyr::relocate(sd1, .after = stat_1)) %>%
+   modify_table_body(~.x %>% dplyr::relocate(sd2, .after = stat_2)) %>% 
+    modify_spanning_header(c("stat_1", "sd1") ~ "**Male**") %>% 
+    modify_spanning_header(c("stat_2", "sd2") ~ "**Female**") %>% 
+    modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-59 using population weights**")
+
+  tabb3svy
+```
 
@@
 d_G[d_G$age >= 18 &
-      d_G$age <= 60, ]  %>% select(
+      d_G$age <= 59, ]  %>% select(
         "sex_partners",
@@
   modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
-  modify_caption("**Table 1. Participant Characteristics**") %>%
+  modify_caption("**Table 1. Participant Characteristics for all variables**") %>%
   bold_labels() 
+# %>% 
+#  add_difference()
 
@@
   modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
-  modify_caption("**Table 1. Participant Characteristics**") %>%
+  modify_caption("**Table 1. Participant Characteristics for all variables using population weights**") %>%
   bold_labels()
@@
 
-```{r exact, fig.width = 12, fig.height = 8}
+```{r exact, fig.width = 12, fig.height = 8, warning = F}
 
@@
 ```
+
+We did not include an interaction between CGS and partnered status for models of age at first sex since, for the majority of participants their age at first sex was in the past and unlikely to be predicted by current grip strength or current partnered status.
 

### 2023-04-04 3.23.R

@@ -1,0 +1,160 @@
+library(qgraph)
+
+design_new <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
+
+cor_mat <- svycor(
+  ~ sex2 +
+    sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
+    physical_disease_count +
+    testosterone +
+    vigorous_rec +
+    moderate_rec +
+    vigorous_work +
+    moderate_work,
+  design_new,
+  na.rm = T
+)
+
+cor_mat_cors <- cor_mat$cors
+corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
+
+p <- EBICglasso(cor_mat_cors, 2800)
+qgraph(p, layout = "spring")
+
+
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered  +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+m_pastyear <- svyglm(
+  sex_partners_year ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered * strength_centered +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+m_agefirst <-  svyglm(
+  age_first_sex ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered  +
+    bmi_centered * sex,
+  family = gaussian(),
+  design = designsG$d.design.adults
+)
+
+
+mnames1 <- c(
+  "Age at first intercourse",
+  "Lifetime Number of Sexual Partners",
+  "Past Year Number of Sexual Partners"
+)
+
+vnames1 <- c(
+  "strength_centered" = "Strength (S)",
+  "sexfemale" = "Sex (Female)",
+  "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "age_centered" = "Age (S)",
+  "partneredTRUE" = "Partnered",
+  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
+  "age_centered:sexfemale" = "Age x Sex (Female)",
[…]

### 2023-04-04 draft_analysis.R

@@ -47,8 +47,8 @@ vnames <- c(
 
 m_lifetime <- svyglm(
   sex_partners ~
-    age_centered * sex +
     strength_centered * sex +
+    age_centered * sex +
     partnered * strength_centered +
     bmi_centered * sex,
   family = quasipoisson(),

### 2023-04-04 meetings_exploratory.R

@@ -1,0 +1,228 @@
+
+#03.31.23
+
+cordat <- d_G[, c("age", "strength", "bmi")]
+cordat <- cordat[cordat$age>=18 & cordat$age<60,]
+cormat <- round(cor(cordat, use = "complete.obs"),2)
+
+
+cordat <- d_Gadults[, c("age", "strength", "bmi", "testosterone")]
+cordat <- cordat[cordat$sex == "male",]
+cormat <- round(cor(cordat, use = "complete.obs"),2)
+
+
+ggplot(d_G) +
+     aes(x = "", y = testosterone) + facet_wrap(~sex) +
+      geom_boxplot(fill = "#0c4c8a") +
+      theme_minimal()
+
+d_Gadults <- d_G[d_G$age >= 18 & d_G$age <= 60,]
+tapply(d_Gadults$testosterone, d_Gadults$sex, summary)
+
+
+# 03.23.23 correlation matrix ------------------------------------------------------
+library(jtools)
+library(qgraph)
+library(corrplot)
+
+# cordat <- d_G[, c("age", "sex1", "strength", "bmi")]
+# cordat <- cordat[cordat$age>=18 & cordat$age<60,]
+# cormat <- round(cor(cordat, use = "complete.obs"),2)
+
+
+cor_mat <- svycor(
+  ~ sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
+    physical_disease_count +
+    testosterone +
+    vigorous_rec +
+    moderate_rec +
+    vigorous_work +
+    moderate_work,
+  designsG$d.design.adults,
+  na.rm = T
+)
+
+cor_mat_cors <- cor_mat$cors
+corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
+
+# meltedcor_mat <- melt(cor_mat_cors)
+#
+# ggplot(data = meltedcor_mat, aes(x=Var1, y=Var2, fill = value)) +
+#   geom_tile()
+
+
+
+design_new <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
+
+cor_mat <- svycor(
+  ~ sex2 +
+    sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
[…]

### 2023-04-04 svytable.R

@@ -1,0 +1,80 @@
+#figure out how to add ns for male and female
+
+tabbsvy <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  add_n() %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "smd") %>%
+  modify_column_hide(columns = ci)
+
+tabbsvysd <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{sd}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  )
+
+tabb3svy <- tabbsvy %>% modify_table_body(
+  ~ .x %>%
+    dplyr::mutate(sd1 = tabbsvysd$table_body$stat_1)
[…]

### 2023-04-04 table1.R

@@ -1,0 +1,552 @@
+
+# Table 1. sex differences usings sample data -----------------------------
+
+
+# has ratio and cohen's d but not sds
+tabb <- d_G[d_G$age >= 18 &
+              d_G$age <= 60, ]  %>% select(
+                "sex",
+                "age",
+                "height",
+                "weight",
+                "bmi",
+                "strength",
+                "whitebloodcell",
+                "hemoglobin",
+                "testosterone",
+              ) %>%
+  tbl_summary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "cohens_d")
+
+
+#has sds and ratio but not cohen's d
+
+sd_gts <- function(data, variable, ...) {
+  sd(data[[variable]], na.rm = TRUE)
+}
+
+
+tabb2 <- d_G[d_G$age >= 18 &
+               d_G$age <= 60, ]  %>% select(
+                 "sex",
+                 "age",
+                 "height",
+                 "weight",
+                 "bmi",
+                 "strength",
+                 "whitebloodcell",
+                 "hemoglobin",
+                 "testosterone",
+               ) %>%
+  tbl_strata(
+  strata = sex,
+  ~ .x %>%
+    tbl_summary(
+ #     by = sex,
+      statistic = list(all_continuous() ~ "{mean}"),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
[…]

### 2023-04-11 ms_draft.Rmd
added intro to draft, edits to table 1

@@
 
+1.1	Sexual dimorphism in modern humans:
+
+Whether modern human sexual size dimorphism is inherited or derived remains debated (Plavcan, 2012).
+Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs (Plavcan, 2001).
+Sexual size dimorphism in humans begins in the first trimester in utero where male embryos measure larger than female (Bukowski et al., 2007).
+Newborn boys are more likely to be larger for gestational age than female newborns, while female newborns exhibit higher average skinfold difference and higher relative fat mass.
+However, no sex typical differences in stature are observed during early childhood and the prepubertal period.
+In adults, there is a difference of between 12 and 25% in body mass (Lassek & Gaulin, 2022).
+Humans are slightly more dimorphic in body mass than gibbons, but slightly less than chimpanzees (Smith & Jungers, 1997).
+A modest 7-8% height dimorphism in humans, a reduction from ancestral species (Gray & Wolf, 1980; Gaulin & Boster, 1985).
+Other traits that are dimorphic in humans include digit ratio, voice pitch, facial features, and canine length.
+
+While humans are only moderately dimorphic in terms of overall body mass, this is not the case of allocation of fat and muscle, which are highly dimorphic (Puts, 2010).
+This is partly due to women’s copious storage of fat, unique among primates.
+Human males have between 30 and 42% more fat free mass (Lassek & Gaulin, 2022).
+On average, men have 61% more overall muscle mass and 78% more muscle mass in the upper arms.
+This concentrated muscle dimorphism in the arms and back translates to 90% greater upper body strength in men than women (Lassek & Gaulin, 2009).
+This difference in muscle and strength is not accounted for by size difference alone (Lassek & Gaulin, 2022).
+
+1.2	Sexual dimorphism hypotheses
+
+A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success via sexual selection, either through attracting mates or winning access through contests or both(.
+Two sexual selection hypotheses dominate the literature regarding explanations for human sexual dimorphism, both asserting that men with more masculine traits had higher biological fitness over evolutionary history: intra-sexual selection if more physically formidably men were able to outcompete other men for access to mates and inter-sexual selection if females preferentially mated with men who display masculine traits, either because they are a costly signal of genetic quality, in conjunction with intra-sexual selection if women preferentially mate with more formidable men.
+
+1.2.1	Intrasexual selection: male-male competition
+
+The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past (Puts, 2010), and argues that many masculine traits, including aggression, large body size, and strength, are best explained by an evolutionary history of contest competition for mate access (Puts, 2016, Hill et al., 2017).
+They argue that intrasexual selection due to an evolutionary history of male-male contests is the most parsimonious explanation for the constellation of sexually dimorphic traits in humans, while acknowledging that some degree of sexual size dimorphism could have been inherited from an ancestral species.
+
+Masculine traits including, increased musculature, could arise through male-male competition if formidable men were more successful at attaining resources and mates, either directly through physical contests or indirectly.
+It also could be the case that female choice still plays a role if women preferentially mate with more formidable men.
+While it could be the case that female choice still plays a role in selection for masculine traits, proponents of this hypothesis point to evidence that mating success due to formidability is better explained by male-rated dominance and not their attractiveness to women.
+
+Sexual dimorphism in formidability is taken as evidence of sexual selection, as larger size or strength, typically in males, indicates competition for access to mates.
+In extant primate species, sexual dimorphism in body weight and canine size are strongly associated with degree of intrasexual competition, and the same relationship is used to infer the sexual behavior of extinct australopithecines (Plavcan & van Shaik, 1997).
+Furthermore, contests override other mechanisms of sexual selection, because when males can exclude other males from mating opportunities mate choice and sperm competition become irrelevant (Puts, 2010).
+
+1.2.2 The immunocompetence handicap hypothesis
+
+In contrast to the intra-sexual selection hypothesis, others argue that sexual dimorphism in humans is largely the result of female choice, not for more formidable or competitive men, but for immunocompetent men.
+The immunocompetence handicap hypothesis relies on evidence that testosterone is immunosuppressive, and therefore only men with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine traits.
+Therefore, masculine traits like strength, could be costly signals of good genetic quality, and masculine men should produce higher quality offspring (Folstad & Karter, 1992).
+Under this hypothesis, masculine traits arise through female choice if females preferentially choose to mate with males displaying these traits indicating heritable phenotypic quality, therefore passing those traits to their offspring.
+A few major problems arise for this hypothesis.
+First, evidence that testosterone is actually immunosuppressive is mixed (Nowak et al., 2018).
+Secondly, it is unclear to what extent women are attracted to testosterone-dependent traits, for instance facial masculinity.
+
+Strength and reproductive success
+
+In a meta-analysis, Lidborg et al (2022) found that while voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations.
+The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported on men (Gallup & Fink, ).
+Of the four studies which have tested the association between grip strength and sexual behavior, only two included both women and men.
+In all four, hand grip strength was positively related to sexual behavior including number of sex partners for men, but there was either no relationship or a negative relationship for women.
+
+Previous research has shown that greater muscle mass in males predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (C-reactive protein and white blood cell count) (Lassek & Gaulin, 2009), consistent with a tradeoff in mating success and immunity.
+However, these relationships were not tested in women, where the sexual selection hypothesis predicts no association.
+We plan to expand these analyses to include women using a different proxy for muscle mass, grip strength.
+
+1.3	Experimental aims and hypotheses
+
+Based on results from exploratory analyses, we predict:
+
+1.	Strength will be positively related to the number of self-reported heterosexual sex partners, both total and past year, and negatively related to the age of first intercourse for both men and women.
+2.	Strength will be positively related to currently partnered status for both men and women. 
+3.	Strength will be negatively related to white blood cell count, a measure of native immunity for both men and women.
+
+We will also control for a range of potential confounds including anthropometric, socioeconomic, hormone, health, and physical activity.
+
 ## Methods
@@
 
-### Baseline (anthropometric) model
+#### Baseline (anthropometric) model
 We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could be related to either strength or mating outcomes.
 
-### Additional control variables: 
+#### Additional control variables: 
 The socioeconomic control model included education and race as categorical variables.
@@
 
-## Analysis 
+#### Analysis 
 Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models.
-Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models from the survey package to incorporate survey weights.
+Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples (Smith, 2018), we use generalized linear models from the survey package to incorporate survey weights.
 For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
@@
 
-## 3. Results
-
-### 3.1 Exploratory Analysis
-
-#### 3.1.1 Descriptive statistics
+### 3. Results: Exploratory Analysis
+
+#### 3.1 Descriptive statistics
 
@@
 Mean age was 38 years for this sample.
-Combined grip strength, height, weight, and testosterone were significantly higher in males than females.
-while BMI was significantly higher in females than males (a reversal from Lassek and Gaulin).
+Sexual dimorphism is given as a ration of male mean:female mean.
+Larger male size dimorphism is reflected in 8% for height and 16% for weight, females, however had a 1% higher BMI than males on average.
+There is a much larger difference in the dimorphism in combined grip strength (57%).
+
+Hemoglobin is higher in males than females (14%), driven separately by an inhibitory effect of estrogen and stimulatory effect of androgens, possibly the result of selection on diverting resources that would otherwise be used to increase hemoglobin levels to the male mean to offspring instead (Murphy, 2014).
+White blood cell count, however, is 3% higher in women.
 
@@
   modify_spanning_header(c("ratio", "d") ~ "**Sexual dimorphism**") %>%
+  modify_footnote(update = everything() ~ NA) %>%
   modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-59**")
@@
 
-```
-
-```{r tab1age1849}
-
-# has ratio and cohen's d but not sds
-tabba <- d_G[d_G$age >= 18 &
-              d_G$age <= 49, ]  %>% select(
-                "sex",
-                "age",
-                "height",
-                "weight",
-                "bmi",
-                "strength",
-                "whitebloodcell",
-                "hemoglobin",
-                "testosterone",
-              ) %>%
-  tbl_summary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**") %>%
-  add_difference(everything() ~ "cohens_d")
-
-#has sds and ratio but not cohen's d
-
-sd_gts <- function(data, variable, ...) {
-  sd(data[[variable]], na.rm = TRUE)
-}
-
-tabb2a <- d_G[d_G$age >= 18 &
-               d_G$age <= 49, ]  %>% select(
-                 "sex",
-                 "age",
-                 "height",
-                 "weight",
-                 "bmi",
-                 "strength",
-                 "whitebloodcell",
-                 "hemoglobin",
-                 "testosterone",
-               ) %>%
-  tbl_strata(
-  strata = sex,
-  ~ .x %>%
-    tbl_summary(
- #     by = sex,
-      statistic = list(all_continuous() ~ "{mean}"),
-      digits = all_continuous() ~ 2,
-      label  = list(
-        strength = "Combined Grip Strength (kg)",
-        height = "Height (cm)",
-        weight = "Weight (kg)",
-        age = "Age (years)",
-        bmi = "Body mass index (kg/m^2)",
-        whitebloodcell = "White blood cell count (1000 cells/µL)",
-        hemoglobin = "Hemoglobin (g/dL)",
-        testosterone = "Testosterone (ng/dL)"
-      ),
-      missing = "no"
-    ) %>%
-   add_n() %>%
-    add_stat(fns = everything() ~ sd_gts)) %>%
-  # modify_fmt_fun(list(add_stat_1 ~ partial(style_number, digits = 2))) %>%
-  modify_header(list(stat_0_1 ~ "**Mean**",
-                     stat_0_2 ~ "**Mean**",
-                     add_stat_1_1 ~ "**SD**",
-                     add_stat_1_2 ~ "**SD**")) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_0_1)/as.numeric(stat_0_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**")
-
-# has all!
-tabb3a <- tabb2a %>% modify_table_body(
-  ~ .x %>%
-    dplyr::mutate(d = round(tabba$table_body$estimate, digits = 2))) %>%
-  modify_header(d = "**d**") %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c("ratio", "d") ~ "**Sexual dimorphism**") %>%
-  modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-49**")
-
-tabb3a
 ```
@@
   ) %>%
-  add_n() %>%
+  add_n(statistic = "{N_nonmiss_unweighted}", 
+        col_label = "**Sample N**") %>% 
   modify_table_body(
@@
     missing_text = "Missing"
-  ) %>% add_p() %>%
-  add_overall() %>%
+  ) %>% 
+ # add_p() %>% 
+ # add_overall() %>%
   add_n() %>%
@@
   ) %>%
-  add_p() %>%
-  add_overall() %>%
+ # add_p() %>%
+ # add_overall() %>%
   add_n() %>%
@@
 ```
+
+Generalized linear regression was used to assess the effect of key predictors on outcome measures of mating success.
+In the baseline (anthropometric control) model, strength was associated with higher numbers of lifetime and past year sex partners and an earlier age at first sex, without an interaction with sex (Table 2).
 
@@
   varnames = vnames
-)$plot + theme_minimal(20) +
+)$plot + theme_minimal(15) +
   geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
-  labs(title = "Exact Models: Regression coefficients from generalized linear models", color = "Model") +
+  labs(title = "Baseline Models: Regression coefficients from generalized linear models", color = "Model") +
   guides(colour = guide_legend(reverse = T))
@@
 
-We did not include an interaction between CGS and partnered status for models of age at first sex since, for the majority of participants their age at first sex was in the past and unlikely to be predicted by current grip strength or current partnered status.
+We did not include an interaction between CGS and partnered status for models of age at first sex since, for the majority of participants their age at first sex was not in the recent past and unlikely to be predicted by current grip strength or current partnered status.
 
@@
 
-mm1 <- tbl_regression(m_lifetime, 
+mm1 <- tbl_regression(
+  m_lifetime,
   label = list(
@@
     bmi_centered = "BMI"
-                      )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
+  )
+) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
@@
 
-mm2 <- tbl_regression(m_pastyear, 
+mm2 <- tbl_regression(
+  m_pastyear,
   label = list(
@@
 
-mm3 <- tbl_regression(m_agefirst, 
+mm3 <- tbl_regression(
+  m_agefirst,
   label = list(
@@
     partnered = "Currently partnered",
-                        bmi_centered = "BMI", 
-                        edu = "Education",
-                        total_work_MET = "Occupational exertion"
-    #              label = list(
-    # age_centered = "Age",
-    # sex = "Sex",
-    # strength_centered = "Grip strength",
-    # partnered = " Currently partnered", 
-    # bmi_centered = "BMI"
-    # edu = "Education",
-    # total_work_MET = "Occupational exertion"
-  )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
+    bmi_centered = "BMI"
+  )
+) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
@@
   tbls = list(mm1, mm2, mm3),
-     tab_spanner = c("Lifetime number of sexual partners",
+  tab_spanner = c(
+    "Lifetime number of sexual partners",
     "Past year number of sexual partners",
-                     "Age at first sexual intercourse")) %>% 
+    "Age at first sexual intercourse"
+  )
+) %>%
   as_gt() %>%
-  gt::tab_header(title = "Table 2. Exact Models of Mating Success",
+  gt::tab_header(
+    title = "Table 2. Baseline Models of Mating Success",
     subtitle = "Standardized regression coefficents from generalized
                  linear models for three indices of mating success for men and women ages
-                 18-59. Models derived from Lassek and Gaulin 2009")
-
-```
+                 18-59. Models derived from Lassek and Gaulin 2009"
+  )
+
+```
+
+We then used generalized linear regression to evaluate the impact of a host of potential confounds on these results.
+For lifetime number of sex partners, strength remained a significant positive predictor when controlling for Socioeconomic, physical activity, health, and hormone variables, with no interaction with sex.
+This was also the case for past year number of sex partners, except in the case of the hormonal control, testosterone.
+Strength predicted an earlier age at first sex for all models except the socioeconomic.
+There was no significant interaction between sex and strength in any of these models.
+
+We also modeled current partnership status.
+While strength was a significant predictor of currently reporting being partnered, a sex and strength interaction emerged in the physical activity, and hormonal control models such that the strength effect was reduced for women.
 
@@
 fig1 <- forestplot(
-  manth1, msoc1, mheal1, mphys1, mhor1a,
+   msoc1, mheal1, mphys1, mhor1,
   intercept = F,
@@
 fig2 <- forestplot(
-  manth2,msoc2,mheal2, mphys2, mhor2a,
+ msoc2,mheal2, mphys2, mhor2,
   intercept = F,
@@
 fig3 <- forestplot(
-  manth3, msoc3, mheal3, mphys3, mhor3a,
+  msoc3, mheal3, mphys3, mhor3,
   intercept = F,
@@
 fig4 <- forestplot(
-  manth4, msoc4, mheal4, mphys4, mhor4a,
+  msoc4, mheal4, mphys4, mhor4,
   intercept = F,
@@
                                        x = 0.01),
-                       bottom = text_grob("\nVariables labelled (S) have been standarized by 2 SD. Base level of race/ethnicity is Mexican American.",
+                       bottom = text_grob("\nVariables labelled (S) have been centered and standarized by 2 SD. Base level of race/ethnicity is Mexican American.",
                                           size = 18))
 
+```
+
+```{r}
+l1 <- tbl_regression(msoc1)
+
+l2 <- tbl_regression(mheal1)
+
+l3 <- tbl_regression(mhor1)
+
+l4 <- tbl_regression(mphys1) 
+  tbl_merge(
+    tbls = list(l1, l2, l3, l4),
+    tab_spanner = c("Socioeconomic",
+                    "Health",
+                    "Hormone",
+                    "Physical Activity")) %>%
+   as_gt() %>%
+  gt::tab_header(title = "Lifetime number of partners control models",
+                 subtitle = "Standardized regression coefficents from generalized
+                 linear models")
+
+```
+
+```{r}
+# l1 <- tbl_regression(msoc1)  %>%
+#   add_significance_stars(
+#     hide_se = TRUE,
+#     pattern = "{estimate}{stars}<br>({std.error})"
+#   ) %>%
+#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
+#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+# 
+# l2 <- tbl_regression(mheal1
+#                         ) %>%
+#   add_significance_stars(
+#     hide_se = TRUE,
+#     pattern = "{estimate}{stars}<br>({std.error})"
+#   ) %>%
+#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
+#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+# 
+# 
+# 
+# l3 <- tbl_regression(mhor1) %>%
+#   add_significance_stars(
+#     hide_se = TRUE,
+#     pattern = "{estimate}{stars}<br>({std.error})"
+#   ) %>%
+#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
+#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+# 
+# l4 <- tbl_regression(mphys1) %>%
+#   add_significance_stars(
+#     hide_se = TRUE,
+#     pattern = "{estimate}{stars}<br>({std.error})"
+#   ) %>%
+#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
+#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+# 
+#   tbl_merge(
+#     tbls = list(l1, l2, l3, l4),
+#     tab_spanner = c("Socioeconomic",
+#                     "Health",
+#                     "Hormone", 
+#                     "Physical Activity")) %>%
+#    as_gt() %>%
+#   gt::tab_header(title = "Lifetime number of partners control models",
+#                  subtitle = "Standardized regression coefficents from generalized
+#                  linear models")
 ```
@@
 ```
+
+### Discussion
+
+Under what conditions would we still expect sexual dimorphism in strength even when it is positively associated with reproductive success for both males and females?
+
+Association between strength and fitness for women

### 2023-04-11 draft_analysis.R

@@ -14,13 +14,13 @@ vnames <- c(
   "age_centered" = "Age (S)",
   "partneredTRUE" = "Partnered",
   "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
-  "age_centered:sexfemale" = "Age x Sex (Female)",
+  "age_centered:sexfemale" = "Age (S) x Sex (Female)",
   "height_centered" = "Height (S)",
   "weight_centered" = "Weight (S)",
   "bmi_centered" = "BMI (S)",
-  "sexfemale:bmi_centered" = "Sex (Female) x BMI",
-  "sexfemale:height_centered" = "Sex (Female) x Height",
-  "sexfemale:weight_centered" = "Sex (Female) x Weight",
+  "sexfemale:bmi_centered" = "Sex (Female) x BMI (S)",
+  "sexfemale:height_centered" = "Sex (Female) x Height (S)",
+  "sexfemale:weight_centered" = "Sex (Female) x Weight (S)",
   'perceived_abnormal_weightTRUE' = "Perceived Abnormal Weight",
   "whitebloodcell_centered" = "White Blood Cell Count (S)",
   "hemoglobin_centered" = "Hemoglobin (S)",
@@ -30,8 +30,8 @@ vnames <- c(
   "depression" = "Depression Score",
   "log(testosterone)" = "Testosterone (log)",
   "sexfemale:log(testosterone)" = "Sex (Female) x Testosterone (log)",
-  "testosterone_sex_centered" = "Testosterone centered by sex",
-  "sexfemale:testosterone_sex_centered"= "sex (Female) x testosterone",
+  "testosterone_sex_centered" = "Testosterone (S by Sex)",
+  "sexfemale:testosterone_sex_centered"= "Sex (Female) x Testosterone (S by Sex)",
   "vigorous_rec" = "Vigorous Recreation",
   "moderate_rec" = "Moderate Recreation",
   "vigorous_work" = "Vigorous Work",
@@ -114,17 +114,17 @@ mnames1 <- c(
 # Lifetime partner models -------------------------------------------------
 
 
-# lifetime partners (anthropometric model)
-manth1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered * strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year
-      bmi_centered * sex,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
[…]

### 2023-04-11 meetings_exploratory.R

@@ -1,4 +1,35 @@
 
+m_lifetime <- svyglm(
+  sex_partners ~
+    strength_centered * sex +
+    age_centered * sex +
+    partnered +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+sum <- summary(m_lifetime, df.resid = Inf)
+
+
+
+
+#04.07.2023
+library(hagenutils)
+fem <- svysmooth(whitebloodcell~age, designsG$d.design.adult.female)
+
+mal <- svysmooth(whitebloodcell~age, designsG$d.design.adult.male)
+
+d <- svysmooth2df(female = fem, male = mal)
+
+ggplot(d, aes(age, whitebloodcell, colour = Smooth)) + geom_line()
+
+
+library(mgcv)
+
+m <- gam(whitebloodcell ~ s(strength, by = sex) + s(age, by = sex), data = d_G)
+
+
 #03.31.23
 
 cordat <- d_G[, c("age", "strength", "bmi")]

### 2023-07-11 ms_draft.Rmd
delete old

--- 2023-03-30 ms_draft.Rmd
@@
 
-## Abstract
-
-Evidence suggests that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-It is unknown if females experience a similar tradeoff.
-Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey, a large representative sample of the US, we will test whether grip strength, a proxy for upper body strength, is a positive predictor of self-reported outcome measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, and age at first intercourse) in both males and females, and negatively related to a measure of native immune function, white blood cell count, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds.
-These results replicate an earlier study of US males.
-However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive, evidence that a similar trade off in the costs and benefits of muscle exists for women.
-
 ## Methods
 
-To assess the relationship between strength and reproductive and immune outcomes we used data from the Centers for Disease Control National Health and Nutrition Examination Survey (NHANES) from years 2011-2014.
-NHANES uses a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
+To assess the relationship between strength and mating and immune outcomes we used data from the Centers for Disease Control (CDC) National Health and Nutrition Examination Survey (NHANES) from years 2011-2014.
+NHANES utilizes a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
 population.
@@
 
-We will use data from years 2011-2014 since grip strength, our key predictor variable was collected in those years.
-We developed models based on previous research and through exploratory analyses of the 2011-2012 dataset to be tested in the 2013-2014 dataset.
-We include data from US adults between the ages of 18 and 65, the years when grip strength is the most stable.
+We will use data from years 2011-2014 since grip strength, our key predictor variable, was only collected in those years.
+We developed models based on previous research and through exploratory analyses of the 2011-2012 dataset to be tested in the 2013-2014 dataset which was held out pending in principle acceptance of this registered report.
+We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable .
 For each model, participants will be included if they have complete data for each predictor variable.
@@
 In industrialized populations like the U.S., widespread access to contraceptives unlinks reproductive success from sexual behavior.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).Our main outcome was mating success, which we operationalized in three ways.
-1) Total sexual partners was a count based on responses to the question: In your lifetime, with how many men/women have you had any kind of sex?
-2) Total sexual partners in past year, from the question: In the past 12 months, with how many men/women have you had any kind of sex?
-Because of the way these questions are framed, (e.g.
-female participants were asked about their male partners and vice versa) these variables represent heterosexual partners, however it is not necessarily the case that all of the participants included in these models identified as heterosexual.
-3) Age at first sexual intercourse reported from: How old were you when you had sex for the first time?.
-The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen.
+Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).
+Our main outcome variable was mating success, which we operationalized in three ways: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
+female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
+3) Age at first sexual intercourse reported from: "How old were you when you had sex for the first time?" The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen.
 Only respondents who could self-report were asked these questions.
@@
 Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL).
-In the original paper, Lassek and Gaulin also use C-reactive protein, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
+In the original paper, Lassek and Gaulin also modelled C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
 
@@
 Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume.
-We operationalized formidability as grip strength (kg), which was measured three times on each hand using a dynamometer.
-We used combined grip strength which was the sum of the highest reading of each hand.
-All models also included sex (male/female), age in years, partnered status (partnered included participants who reported either being married or living with a partner while unpartnered included participants who reported being single, widowed, divorced, or separated).
-The interaction between sex and strength was crucial for our models since we want to know whether the effect of strength on mating and immune outcomes is different for women than for men.
+We operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
+All models also included sex (male/female), age in years, partnered status (partnered included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated).
+The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
 Likewise, we also interacted age and partnered status with sex.
+We interacted age with sex since menopause occurs within the age range of our participants.
+For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
 
 ### Models:
-We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
 
-### Baseline (anthropometric) model
-We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could be related to either strength or mating outcomes.
+#### Baseline (anthropometric) model
+We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and current partnered status.
+This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
+We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.
 
-### Additional control variables: 
+#### Additional control variables: 
+to additionally test if the association of strength with mating success shown in Lassek and Gaulin (2009) was due to some unmeasured variable, we fit regression models with a much wider range of potential confounds.
+We specified four additional models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with anthropometric, socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
+
 The socioeconomic control model included education and race as categorical variables.
+
 The health model included a number of variables related to health, which could affect either strength or sexual behavior.
@@
 Depression was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument.
-Each item represents a symptom of depression, and for each participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
+Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
 These ratings were summed to produce a depression score ranging from 0-27.
@@
 A point was added for each disease a participant indicated they had been diagnosed with, regardless of any impairment due to the disease.
-We then controlled for impairment resulting from chronic illness (Disease Impairment Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities.
+We then controlled for impairment resulting from chronic illness (Disability Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities.
 We also controlled for physical disability and using the item "special equipment" which referred to participants' report that they needed special equipment to walk.
@@
 
-The hormone control model included testosterone.
-Since the difference in mean testosterone between men and women is an order of magnitude apart, and testosterone was highly confounded with sex.
-We therefore centered testosterone at the mean and divided by two standard deviations within each sex.
-The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they did not.
+The hormone control model included serum total testosterone.
+Since male mean testosterone is an order of magnitude higher than female,  testosterone was highly confounded with sex.
+We therefore computed a testosterone value as follows: we centered testosterone at the mean and divided by two standard deviations within each sex.
 
-Current salary Measures of wealth available in NHANES are collected at the household level, and as such do not accurately reflect wealth or poverty at the individual level.
-As a measure of wealth, we used occupation codes collected for each participant by NHANES and imputed the median salary calculated for that occupation category (source?).
+The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
 
-###
-
-In keeping with the outcome variables utilized in the original paper, we used number of partners for any kind of sex above.
-Since, however, NHANES also collects more specific data on sexual behavior we also tested Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex.
+Our outcome variables were identical to those used by Lassk and Gaulin (2009) ie lifetime, past year number of partners and age at first sex.
+However, the NHANES questions include any kind of sex (e.g.
+oral, vaginal, anal).
+Since, however, NHANES also collects more specific data on sexual behavior we also fit Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex.
 In these questions participants were asked “In your lifetime, with how many men/women have you had vaginal sex?” and “In the past 12 months, with how many men/women have you had vaginal sex?”
-
-Missing variables 
-We do not have data on bone mineral density
 
@@
 
-## Analysis 
+#### Analysis 
 Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models.
-Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models from the survey package to incorporate survey weights.
+Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples (Smith, 2018), we use generalized linear models from the survey package to incorporate survey weights.
 For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
 Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.
+Analyses include adults ages 18-59 because this is period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
 
-## Results
+### 3. Results: Exploratory Analysis
 
-### Exploratory Analysis
+#### 3.1 Descriptive statistics
 
-```{r summarytable, fig.width = 12, fig.height = 16}
-d_G[d_G$age >= 18 &
-      d_G$age <= 60, ]  %>% select(
-        "sex_partners",
-        "sex_partners_year",
-        "age_first_sex",
-        "sex",
-        "strength",
-        "age",
-        "partnered",
-        "bmi",
-        "edu",
-        "race",
-        "whitebloodcell",
-        "hemoglobin",
-        "depression",
-        "chronic_disease_score",
-        "physical_disease_count",
-        "special_equipment",
-        "perceived_abnormal_weight",
-        "testosterone",
-        "vigorous_work",
-        "vigorous_rec",
-        "moderate_work",
-        "moderate_rec"
+Table 1 shows weighted means and standard deviations of all variables for men and women ages 18-59.
+Mean age was 38 years for this sample.
+Sexual dimorphism is given as a ratio of male mean:female mean.
+Larger male size dimorphism is reflected in 8% for height and 16% for weight, females, however had a 1% higher BMI than males on average.
+There is a much larger difference in the dimorphism in combined grip strength (57%).
+Hemoglobin is higher in males than females (14%) while white blood cell count is 3% higher in women.
+
+```{r sexualdimorphism}
+
+getn <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{N_nonmiss_unweighted}"
+    ),
+    digits = all_continuous() ~ 0,
+    missing = "no"
+  )
+
+getratio <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    missing = "no"
   ) %>%
-  mutate(race = factor(
-    race,
-    labels = c(
-      "Mexican American",
-      "Other Hispanic",
-      "Non-Hispanic White",
-      "Non-Hispanic Black",
-      "Non-Hispanic Asian",
-      "Other Race"
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
   )
-  )) %>%
-  mutate(edu = factor(
-    edu,
-    labels = c(
-      "Less than 9th grade",
-      "9-11th grade",
-      "High school graduate/GED or equivalent",
-      "Some college or AA degree",
-      "College graduate or above"
-    )
-  )) %>%
-  tbl_summary(
+
+tab1 <- designsG$d.design.adults %>%
+  tbl_svysummary(
     by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} / {N} ({p}%)"
+    include = c(
+      age,
+      height,
+      weight,
+      bmi,
+      strength,
+      whitebloodcell,
+      hemoglobin,
+      testosterone
     ),
+    statistic = list(all_continuous() ~ "{mean} ({sd})"),
     digits = all_continuous() ~ 2,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
     label  = list(
-      age_first_sex = "Age at first sex (years)",
-      sex_partners = "Lifetime number of sexual partners",
-      sex_partners_year = "Past year number of sexual partners",
       strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
       age = "Age (years)",
       bmi = "Body mass index (kg/m^2)",
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
       whitebloodcell = "White blood cell count (1000 cells/µL)",
       hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
+      testosterone = "Testosterone (ng/dL)"
     ),
-    missing_text = "Missing"
-  ) %>% add_p() %>%
-  add_overall() %>%
-  add_n() %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
-  modify_caption("**Table 1. Participant Characteristics**") %>%
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
+      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
+      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
+      dplyr::relocate(n_male, .before = stat_1) %>%
+      dplyr::relocate(n_female, .before = stat_2)
+  )  %>%
+  add_difference(test = everything() ~ "smd",
+                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
+  modify_column_hide(columns = ci) %>%
+  modify_header(
+    label = "**Variable**",
+    ratio = "**Ratio**",
+    n_male = "**N**",
+    n_female = "**N**",
+    stat_1 = "**Mean (SD)**",
+    stat_2 = "**Mean (SD)**",
+    estimate = "*d*"
+  ) %>%
+  modify_spanning_header(list(
+    c("stat_1", "n_male") ~ "**Male**",
+    c("stat_2", "n_female") ~ "**Female**",
+    c("ratio", "estimate") ~ "**Sexual Dimorphism**"
+  )) %>%
+  modify_footnote(update = everything() ~ NA) %>%
+  modify_caption("**Table 1. Sex differences for participants ages 18-59 using population weights**") %>%
   bold_labels()
+
+tab1
 
@@
 
-```{r}
+Generalized linear regression was used to assess the effect of key predictors on outcome measures of mating success.
+In the baseline (anthropometric control) model, strength was significantly associated with higher numbers of lifetime and past year sex partners and an earlier age at first sex, without an interaction with sex (Table 2).
 
-designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} / {N} ({p}%)"
-    ),
-    digits = all_continuous() ~ 2,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
-    label  = list(
-      age_first_sex = "Age at first sex (years)",
-      sex_partners = "Lifetime number of sexual partners",
-      sex_partners_year = "Past year number of sexual partners",
-      strength = "Combined Grip Strength (kg)",
-      age = "Age (Years)",
-      bmi = "Body mass index (kg/m^2)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
-    ),
-    missing_text = "Missing",
-    include = c(
-      "sex_partners",
-      "sex_partners_year",
-      "age_first_sex",
-      "sex",
-      "strength",
-      "age",
-      "partnered",
-      "bmi",
-      "edu",
-      "race",
-      "whitebloodcell",
-      "hemoglobin",
-      "depression",
-      "chronic_disease_score",
-      "physical_disease_count",
-      "special_equipment",
-      "perceived_abnormal_weight",
-      "testosterone",
-      "vigorous_work",
-      "vigorous_rec",
-      "moderate_work",
-      "moderate_rec"
-    )
-  ) %>%
-  add_p() %>%
-  add_overall() %>%
-  add_n() %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
-  modify_caption("**Table 1. Participant Characteristics**") %>%
-  bold_labels()
-
-```
-
-```{r exact, fig.width = 12, fig.height = 8}
+```{r exact, fig.width = 12, fig.height = 8, warning = F}
 
@@
 
-```{r table2_exact}
+We then used generalized linear regression to evaluate the impact of a host of potential confounds on these results.
+For lifetime number of sex partners, strength remained a significant positive predictor when controlling for Socioeconomic, physical activity, health, and hormone variables, with no interaction with sex.
+This was also the case for past year number of sex partners, except in the case of the hormonal control, testosterone.
+Strength predicted an earlier age at first sex for all models except the socioeconomic.
+There was no significant interaction between sex and strength in any of these models.
 
-mm1 <- tbl_regression(m_lifetime, 
-                      label = list(
-                        age_centered = "Age",
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Currently partnered",
-                        bmi_centered = "BMI"
-                      )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm2 <- tbl_regression(m_pastyear, 
-                      label = list(
-                        age_centered = "Age",
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Currently partnered",
-                        bmi_centered = "BMI"
-                      )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm3 <- tbl_regression(m_agefirst, 
-                      label = list(
-                        age_centered = "Age",
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Currently partnered",
-                        bmi_centered = "BMI", 
-                        edu = "Education",
-                        total_work_MET = "Occupational exertion"
-    #              label = list(
-    # age_centered = "Age",
-    # sex = "Sex",
-    # strength_centered = "Grip strength",
-    # partnered = " Currently partnered", 
-    # bmi_centered = "BMI"
-    # edu = "Education",
-    # total_work_MET = "Occupational exertion"
-  )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-  tbl_merge(
-    tbls = list(mm1, mm2, mm3),
-     tab_spanner = c("Lifetime number of sexual partners",
-                     "Past year number of sexual partners",
-                     "Age at first sexual intercourse")) %>% 
-  as_gt() %>%
-  gt::tab_header(title = "Table 2. Exact Models of Mating Success",
-                 subtitle = "Standardized regression coefficents from generalized
-                 linear models for three indices of mating success for men and women ages
-                 18-59. Models derived from Lassek and Gaulin 2009")
-
-```
+We also modeled current partnership status.
+While strength was a significant predictor of currently reporting being partnered, a sex and strength interaction emerged in the physical activity, and hormonal control models such that the strength effect was reduced for women.
 
@@
 ```
+
+```{r}
+
+designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      all_categorical() ~ "{n} / {N} ({p}%)"
+    ),
+    digits = all_continuous() ~ 2,
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous"
+    ),
+    label  = list(
+      age_first_sex = "Age at first sex (years)",
+      sex_partners = "Lifetime number of sexual partners",
+      sex_partners_year = "Past year number of sexual partners",
+      strength = "Combined Grip Strength (kg)",
+      age = "Age (Years)",
+      bmi = "Body mass index (kg/m^2)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      partnered = "Partnered",
+      edu = "Education",
+      race = "Race and Ethnicity",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)",
+      perceived_abnormal_weight = "Perceived abnormal weight",
+      special_equipment = "Special equipment needed to walk",
+      chronic_disease_score = "Chronic Disease Score (0-6)",
+      physical_disease_count = "Disease Impairment Score (0-5)",
+      depression = "Depression Score (0-27)",
+      vigorous_work = "Work involves vigorous activity",
+      moderate_work = "Work involves moderate activity",
+      vigorous_rec = "Recreation involves vigorous activity",
+      moderate_rec = "Recreation involves moderate activity"
+    ),
+    missing_text = "Missing",
+    include = c(
+      "sex_partners",
+      "sex_partners_year",
+      "age_first_sex",
+      "sex",
+      "strength",
+      "age",
+      "partnered",
+      "bmi",
+      "edu",
+      "race",
+      "whitebloodcell",
+      "hemoglobin",
+      "depression",
+      "chronic_disease_score",
+      "physical_disease_count",
+      "special_equipment",
+      "perceived_abnormal_weight",
+      "testosterone",
+      "vigorous_work",
+      "vigorous_rec",
+      "moderate_work",
+      "moderate_rec"
+    )
+  ) %>%
+ # add_p() %>%
+ # add_overall() %>%
+  add_n() %>%
+  modify_header(label ~ "**Variable**") %>%
+  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>%
+  modify_caption("**Table 1. Participant Characteristics for all variables using population weights**") %>%
+  bold_labels()
+
+```
+
+### Discussion
+
+Under what conditions would we still expect sexual dimorphism in strength even when it is positively associated with reproductive success for both males and females?
+
+Association between strength and fitness for women

### 2023-07-11 3.23.R

@@ -1,160 +1,0 @@
-library(qgraph)
-
-design_new <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
-
-cor_mat <- svycor(
-  ~ sex2 +
-    sex_partners +
-    sex_partners_year +
-    age_first_sex +
-    age +
-    strength +
-    bmi +
-    edu +
-    hemoglobin +
-    whitebloodcell +
-    depression +
-    chronic_disease_score +
-    physical_disease_count +
-    testosterone +
-    vigorous_rec +
-    moderate_rec +
-    vigorous_work +
-    moderate_work,
-  design_new,
-  na.rm = T
-)
-
-cor_mat_cors <- cor_mat$cors
-corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
-
-p <- EBICglasso(cor_mat_cors, 2800)
-qgraph(p, layout = "spring")
-
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered  +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-m_pastyear <- svyglm(
-  sex_partners_year ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered * strength_centered +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-m_agefirst <-  svyglm(
-  age_first_sex ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered  +
-    bmi_centered * sex,
-  family = gaussian(),
-  design = designsG$d.design.adults
-)
-
-
-mnames1 <- c(
-  "Age at first intercourse",
-  "Lifetime Number of Sexual Partners",
-  "Past Year Number of Sexual Partners"
-)
-
-vnames1 <- c(
-  "strength_centered" = "Strength (S)",
-  "sexfemale" = "Sex (Female)",
-  "sexfemale:strength_centered" = "Sex (Female) x Strength",
-  "age_centered" = "Age (S)",
-  "partneredTRUE" = "Partnered",
-  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
-  "age_centered:sexfemale" = "Age x Sex (Female)",
[…]

### 2023-07-11 descriptivestats.R

@@ -1,35 +1,0 @@
-designsG$d.design.adults %>%
-  tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi,
-                                       whitebloodcell, strength, testosterone, partnered,
-                                       height, weight, special_equipment))
-
-
-
-d_G[d_G$age>=18 & d_G$age<=60,]  %>% select("sex", "age", "race", "strength", "bmi", "partnered", "edu", "height", "weight", "whitebloodcell", "testosterone") %>%
-  tbl_summary(by = sex,
-              statistic = list(all_continuous() ~ "{mean} ({sd})",
-                               all_categorical() ~ "{n} / {N} ({p}%)"),
-              digits = all_continuous() ~ 2,
-        #      type = list(where(is.numeric) ~ "continuous"),
-             # type = list(partnered ~ "categorical"),
-              type = all_dichotomous() ~ "categorical",
-              label  = list(
-                age = "Age (Years)",
-                race = "Race and Ethnicity",
-                strength = "Combined Grip Strength",
-                bmi = "BMI (kg/m^2)",
-                partnered = "Partnered Status",
-                edu = "Education",
-                height = "Height (cm)",
-                weight = "Weight (kg)",
-                whitebloodcell = "WBCC (1000 cells/µL)",
-                testosterone = "Testosterone (ng/dL)"),
-              missing_text = "Missing"
-  ) %>% add_p() %>%
-  add_overall() %>%
-  add_n() %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**")
-
-
-

### 2023-07-11 table1.R

@@ -1,552 +1,0 @@
-
-# Table 1. sex differences usings sample data -----------------------------
-
-
-# has ratio and cohen's d but not sds
-tabb <- d_G[d_G$age >= 18 &
-              d_G$age <= 60, ]  %>% select(
-                "sex",
-                "age",
-                "height",
-                "weight",
-                "bmi",
-                "strength",
-                "whitebloodcell",
-                "hemoglobin",
-                "testosterone",
-              ) %>%
-  tbl_summary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**") %>%
-  add_difference(everything() ~ "cohens_d")
-
-
-#has sds and ratio but not cohen's d
-
-sd_gts <- function(data, variable, ...) {
-  sd(data[[variable]], na.rm = TRUE)
-}
-
-
-tabb2 <- d_G[d_G$age >= 18 &
-               d_G$age <= 60, ]  %>% select(
-                 "sex",
-                 "age",
-                 "height",
-                 "weight",
-                 "bmi",
-                 "strength",
-                 "whitebloodcell",
-                 "hemoglobin",
-                 "testosterone",
-               ) %>%
-  tbl_strata(
-  strata = sex,
-  ~ .x %>%
-    tbl_summary(
- #     by = sex,
-      statistic = list(all_continuous() ~ "{mean}"),
-      digits = all_continuous() ~ 2,
-      label  = list(
-        strength = "Combined Grip Strength (kg)",
-        height = "Height (cm)",
-        weight = "Weight (kg)",
[…]

### 2023-07-11 dailyenergy.R
updated manuscripts

@@ -1,0 +1,173 @@
+library(fitdistrplus)
+
+# Daily Energy Intake (Table 3) -------------------------------------------
+
+# bone mineral density not collected in G, collected in H in altered sample
+
+# compute occupational expenditure per Lassek and Gaulin
+# we have MET from participant reported occ exertion, a better var than this, also NHANES recommends using
+# diff weights for dietary data
+#
+# occ_exertion <- svyby(~avgcalories, ~census_code_best, design = designsG$d.design.adults, svymean)
+# occ_exertion_dict <- occ_exertion$avgcalories
+# names(occ_exertion_dict) <- occ_exertion$census_code_curr
+# designsG$d.design.adults <- update(designsG$d.design.adults, occ_exertion = occ_exertion_dict[census_code_curr])
+#
+#
+# x <- svyby(~avgcalories, ~race, design = designsG$d.design.adults, svymean, na.rm = T)
+
+# not including in rep because we do not have ffm
+
+
+x <- c(na.omit(d_G$avgcalories))
+x <- c(na.omit(d_G$whitebloodcell))
+
+descdist(x, discrete = T)
+
+
+adults = d_G$age>=18 & d_G$age<=60
+
+d.design.dietary1 <-
+  svydesign(
+    id = ~SDMVPSU ,
+    strata = ~SDMVSTRA ,
+    nest = TRUE ,
+    weights = ~WTMEC2YR, #~WTINT2YR ,
+    data = d_G
+  )
+
+d.design.dietary.adults1 <-
+  subset(
+    d.design.dietary1,
+    adults
+  )
+
+# m <- svyglm(
+#   log(avgcalories) ~
+#     age +
+#     sex * tot_MET +
+#     strength +
+#     weight +
+#     height,
+#   family = gaussian(),
+#   design = d.design.dietary.adults
+# )
+# summary(m)
+# plot(allEffects(m))
+#
+
+
+m <- svyglm(
+  log(avgcalories) ~
+    age +
+    total_work_MET  +
+    total_rec_MET +
+    wob_MET +
+    strength +
+    height +
+    weight,
+  family = gaussian(),
+  design = d.design.dietary.adults1
+)
+summary(m, df.resid = Inf)
+plot(allEffects(m))
+
+m_energy1 <- svyglm( #use this
+  avgcalories ~
+    age +
+    total_work_MET  +
+    total_rec_MET +
+    wob_MET +
+    strength +
+    height +
+    weight,
+  family = gaussian(),
[…]

### 2023-07-11 differencesall.R

@@ -1,0 +1,177 @@
+getn <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age_first_sex,
+                sex_partners,
+                sex_partners_year,
+                strength,
+                age,
+                bmi,
+                height,
+                weight,
+                partnered,
+                edu,
+                race,
+                whitebloodcell,
+                hemoglobin,
+                testosterone,
+                perceived_abnormal_weight,
+                special_equipment,
+                chronic_disease_score,
+                physical_disease_count,
+                depression,
+                vigorous_work,
+                moderate_work,
+                vigorous_rec,
+                moderate_rec,
+                avgcalories),
+    statistic = list(
+      all_continuous() ~ "{N_nonmiss_unweighted}",
+      all_categorical() ~ "{N_nonmiss_unweighted}"
+    ),
+    digits = all_continuous() ~ 0,
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous"
+    ),
+    missing = "no"
+  )
+
+getratio <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age_first_sex,
+                sex_partners,
+                sex_partners_year,
+                strength,
+                age,
+                bmi,
+                height,
+                weight,
+                partnered,
+                edu,
+                race,
+                whitebloodcell,
+                hemoglobin,
+                testosterone,
+                perceived_abnormal_weight,
+                special_equipment,
+                chronic_disease_score,
+                physical_disease_count,
+                depression,
+                vigorous_work,
+                moderate_work,
+                vigorous_rec,
+                moderate_rec,
+                avgcalories),
+    statistic = list(
+      all_continuous() ~ "{mean}",
+      all_categorical() ~ "{p}"
+    ),
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous",
+      avgcalories ~ "continuous"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
[…]

### 2023-07-11 draft_analysis.R

@@ -47,9 +47,9 @@ vnames <- c(
 
 m_lifetime <- svyglm(
   sex_partners ~
-    strength_centered * sex +
     age_centered * sex +
-    partnered * strength_centered +
+    strength_centered * sex +
+    partnered +
     bmi_centered * sex,
   family = quasipoisson(),
   design = designsG$d.design.adults
@@ -59,7 +59,7 @@ m_pastyear <- svyglm(
   sex_partners_year ~
     age_centered * sex +
     strength_centered * sex +
-    partnered * strength_centered +
+    partnered * strength_centered + #keeping partnered x strength interaction only for this model
     bmi_centered * sex,
   family = quasipoisson(),
   design = designsG$d.design.adults
@@ -133,7 +133,7 @@ msoc1 <-
     sex_partners ~
       age_centered * sex +
       strength_centered * sex +
-      partnered * strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year
+      partnered + #* strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year; took out 4.6
       edu +
       race,
     family = quasipoisson(),
@@ -147,7 +147,7 @@ mheal1 <-
     sex_partners ~
       age_centered * sex +
       strength_centered * sex +
-      partnered * strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year
+      partnered +
       perceived_abnormal_weight +
       whitebloodcell_centered +
       hemoglobin_centered +
@@ -158,19 +158,19 @@ mheal1 <-
     family = quasipoisson(),
     design = designsG$d.design.adults
   )
-summary(mheal1)
+summary(mheal1, df.resid = Inf)
 
 mhor1 <-
   svyglm(
     sex_partners ~
       age_centered * sex +
       strength_centered * sex +
-      partnered * strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year
+      partnered +
       testosterone_sex_centered * sex,
     family = quasipoisson(),
     design = designsG$d.design.adults
   )
-summary(mhor1, df = Inf)
+summary(mhor1, df.resid = Inf)
 
 #phys activity
 mphys1 <-
@@ -178,7 +178,7 @@ mphys1 <-
     sex_partners ~
       age_centered * sex +
       strength_centered * sex +
[…]

### 2023-07-11 meetings_exploratory.R

@@ -1,6 +1,37 @@
 
 m_lifetime <- svyglm(
   sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+summary(m_lifetime, df.resid = Inf)
+
+
+
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered +
+    weight_centered * sex +
+    height_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+summary(m_lifetime, df.resid = Inf)
+
+
+
+#04.11.2023
+library(jtools)
+
+m_lifetime <- svyglm(
+  sex_partners ~
     strength_centered * sex +
     age_centered * sex +
     partnered +
@@ -9,9 +40,10 @@ m_lifetime <- svyglm(
   design = designsG$d.design.adults
 )
 
-sum <- summary(m_lifetime, df.resid = Inf)
+summary(m_lifetime, df.resid = Inf)
+sum <- summary(m_lifetime, df.resid = Inf)$coefficients
 
-
+summ(m_lifetime, df.resid = Inf)
 
 
 #04.07.2023
@@ -225,6 +257,170 @@ mwbc1 <- svyglm(whitebloodcell ~
                  testosterone_centered * sex,
                family= quasipoisson(),
                design=designsG$d.design.adults)
+summary(mwbc1)
+
+
+vnames <- c(
+  "strength_centered" = "Strength (S)",
+  "sexfemale" = "Sex (Female)",
+  "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "age_centered" = "Age (S)",
+  "partneredTRUE" = "Partnered",
+  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
+  "age_centered:sexfemale" = "Age x Sex (Female)",
+  "height_centered" = "Height (S)",
+  "weight_centered" = "Weight (S)",
+  "bmi_centered" = "BMI (S)",
+  "sexfemale:bmi_centered" = "Sex (Female) x BMI",
+  "testosterone_centered" = "Testosterone",
+  "sexfemale:testosterone_centered" = "Testosterone x Sex Female"
+)
+
+forestplot(
+  mwbc1,
+  intercept = F,
+  facet = F,
+  dodgewidth = .8,
[…]

### 2023-07-12 ms_draft07.06.Rmd
edits to models, added discussion to draft

---
title: "Grip strength, mating success, and immune function in US adults: a registered report "
author: "C. B. Smith and E. H. Hagen"
date: "`r Sys.Date()`"
output:
 bookdown::html_document2:
   number_sections: false
 bookdown::pdf_document2:
   number_sections: false
   latex_engine: xelatex
 bookdown::word_document2:
   number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
library(weights)
library(jtools)
library(reshape2)
library(gtsummary)
library(corrplot)
source('draft_analysis.R')
```
EHB Registered Report

### 1.1 Sexual dimorphism in modern humans:

The mechanisms that have shaped human sexual dimorphism and whether they have primarily operated on male or female traits remain debated. Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs (Plavcan, 2001). Unlike most animals, mammalian sexual size dimorphism tends to be male biased (Andersson, 1994). In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos (Bukowski et al., 2007). Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass (Kirchengast, 2010). At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass (Lassek & Gaulin, 2022), which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees (Plavcan, 2012; Smith & Jungers, 1997). Men also have larger and stronger bones (Wells, 2007). Other traits that are sexually dimorphic in humans include digit ratio, voice pitch, facial features, body and facial hair growth, and canine length.

While humans are only moderately dimorphic in terms of overall body mass, this is not the case for fat and muscle allocation, which are highly dimorphic (Puts, 2010) due to women’s copious storage of fat and male investment in muscle mass. Men have between 30% and 42% more fat-free mass (Lassek & Gaulin, 2022), 61% more overall muscle mass, and 78% more muscle mass in the upper arms. This concentrated muscle dimorphism in the arms and back translates to greater upper body strength in men than women. Studies of sex differences in strength show that men are stronger on average than women on all tests of muscle strength, but especially for tests of upper-body muscle strength, where female upper-body strength is 50-60% of male upper-body strength while lower-body strength is 60-70% of male values, and trunk strength is 60% of male values (Nuzzo, 2023). Muscle mass alone does not explain the sex difference in strength, as strength assessments are often greater in males than females even when pair-matched on muscle thickness (Kataoka et al., 2023), and likewise men have greater strength-to-body mass ratios than women (Nuzzo, 2023).

There is no evidence for a sex difference in the ability of the nervous system to drive the muscle (voluntary action) but instead differences in strength lie in muscle characteristics including mass, size, and fiber type. Men have both more absolute muscle mass and greater mass proportional to body size, and these muscles have higher volume and cross-sectional area size. These differences are greatest in the upper body. Finally, a greater proportion of male to female muscle is occupied by type II muscle fibers, which create greater force than type I (Nuzzo, 2023).

### 1.2  Sexual dimorphism hypotheses

Darwin defined sexual selection as selection that acts, not to fit individuals to their environment, but instead as to confer advantage over others in competition for mates (Darwin, 1871). Modern definitions distinguish pre- and post-copulatory competition (Andersson and Simmons, 2006). A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success, either through attracting mates or winning access to mates and fertilizations through contests, or both (Andersson, 1994). Two hypotheses for the evolution of human sexual size dimorphism and body composition dimorphism dominate the literature: 1) intrasexual selection through male contests and 2) intersexual selection through female mate choice. Both assert that men with more masculine traits had higher biological fitness over evolutionary history: intrasexual selection if more formidable men were able to physically outcompete other men for access to mates, and intersexual selection if females preferentially mated with men who displayed masculine traits. The male contest and the female choice hypotheses concur that sexual size and strength dimorphism are proximately (that is, developmentally) caused by sex differences in androgen hormones in the uterine environment and during differentiation at puberty. They differ, however, in the ultimate (functional) explanations for these differences, as discussed below.

#### 1.2.1 	Male contest competition

The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past (Puts, 2010), and that instead male fitness was determined by physical contests with other males for access to mates. Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success (Puts, 2016, Hill et al., 2017; Plavcan, 2012). Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression (Archer, 2009). Particular emphasis is placed on sexual dimorphism in the upper-body due to the important role upper-body strength plays in both armed and unarmed fighting ability (Sell, et al., 2012) and where sex differences in strength and muscularity are greatest as discussed above.

In extant nonhuman primate species, sexual dimorphism in body weight and canine size are strongly associated with the degree of intrasexual competition, a relationship which, if used to infer the behavior of extinct australopithecines, suggests a high degree of intrasexual competition in ancestral species (Plavcan & van Schaik, 1997). Evidence of sexual size dimorphism in Homo is less certain, but in general shows reduced size dimorphism compared to early species (Plavcan, 2012).

#### 1.2.2 	Female choice for masculine traits

In contrast to the intrasexual selection hypothesis, others argue that sexual dimorphism in humans is largely the result of female choice, not for more formidable or competitive mates, but for high quality males. The immunocompetence handicap hypothesis relies on evidence that testosterone is immunosuppressive, and therefore only males with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine secondary sexual traits (Folstad & Karter, 1992). Therefore, androgen-dependent traits could be costly signals of good genetic quality, in which case females could increase their long-term fitness by preferentially mating with males displaying these traits thus conferring heritable immunity to their offspring and securing investment from a healthy partner (Folstad & Karter, 1992). Similar arguments have been proposed which argue that masculine traits could be more general cues of quality and condition other than immunocompetence, or ability to bear other costs (Kokko et al., 2003).

In humans, the immunocompetence hypothesis has primarily been leveraged to explain male facial masculinity and voice pitch but it and more general cost models have also been argued to explain apparent female preference for muscular male body types (Boothroyd et al., 2013).

A few problems arise for this hypothesis. First, evidence that testosterone is actually immunosuppressive is mixed (Nowak et al., 2018). Likewise, masculine trait expression is not consistently related to retrospective or prospective health (Boothroyd et al., 2013). Secondly, it is unclear to what extent women are attracted to testosterone-dependent traits, for instance facial masculinity. Although testosterone is associated with mating success, it does not seem to be through the mechanism of female choice for testosterone-derived face or body features since testosterone does not predict female rated masculinity or attractiveness (Peters et al., 2008).

Females might also choose males who could provide more resources, which in ancestral hunter-gatherer populations might have been physically stronger men. It could also have been the case that mate choice was heavily influenced by parents, in which case parental preferences would have been important (see Hill et al. 2017 for a brief review of alternative hypotheses for sexual strength dimorphism).

More generally, multiple mechanisms of sexual selection could have played a role in the evolution of human sexual dimorphism. Nevertheless, there is evidence that mating success due to formidability is mediated by male-rated dominance and not by female-rated attractiveness (Kordsmeyer et al., 2018; Hill et al., 2013).

### 1.3 Strength and reproductive success

In a meta-analysis, Lidborg et al (2022) found that while voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations. The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported for men (Gallup & Fink, 2018). Of the four studies which have tested the association between grip strength and sexual behavior, only two included both women and men (Gallup et al., 2007; Varella et al., 2014) and the other two only men (Shoup and Gallup, 2008; Sneade and Furnham, 2016). In all four, hand grip strength was positively related to mating success (including number of sex partners) for men, but there was either no relationship or a negative relationship for women. These four studies were in convenience samples.

There have been very few tests of the hypotheses that some female traits, such as breast size and waist-hip ratio (WHR), have evolved by sexual selection. In a meta-analysis, Lidborg and Boothroyd (2023) found weak evidence that more feminine digit ratios predicted higher fertility, insufficient evidence for voice pitch, and none regarding facial femininity, breast size or waist-hip ratio (WHR). At present there is mixed evidence that strength is related to reproductive success in women (Lidborg & Boothroyd, 2023). One study among the Himba showed that women with higher strength/muscularity had more living children and grandchildren (Atkinson et al., 2012), while a similar study among the Hadza showed no relationship between strength and reproductive outcomes (Smith et al., 2017). The dearth of research on female traits and reproductive outcomes leaves us unable to draw any firm conclusions about their selection pressures over time.

Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (operationalized as C-reactive protein and white blood cell count) (Lassek & Gaulin, 2009), consistent with a tradeoff in mating success, immunity and energy costs. However, these relationships were not tested in women, for whom both selection hypotheses predict no association.

### 1.4 Experimental aims and hypotheses

The aim of this study is to replicate the costs and benefits of formidability reported by Lassek and Gaulin (2009) using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format. It also aims to investigate if these costs and benefits are also experienced by women.

In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success, with a significant interaction with sex such that strength will be a stronger predictor of male than female mating success. We also predict that there will be a negative association between male formidabillity and immune function, but little or no association for females. Finally, we predict that formidability will be positively associated with dietary energy intake for both males and females.

## 2. Pilot study

In order to refine our hypotheses and statistical models for our confirmatory study, which will use data we have not yet observed, we first conducted a pilot study whose results we report here.

### 2.1 Methods

To assess the relationship between formidability and mating benefits, and immune and energy costs, we used data from the Centers for Disease Control (CDC) National Health and Nutrition Examination Survey (NHANES). NHANES utilizes a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S. population. NHANES combines interview, examination, and laboratory data to assess health status and identify health risks for adults and children in the United States. Data collection occurs in new cycles every two years.

The pilot study uses the 2011-2012 dataset, whereas the 2013-2014 dataset has been held out for confirmatory analysis pending in principle acceptance of this registered report. Grip strength, our key predictor variable, was only collected in these years. We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable (Hogrel, 2015). For each model, participants will be included if they have complete data for each predictor and outcome variable.

#### 2.1.2 Key predictor variables: 
Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume. Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer. All regression models also included sex (male/female), age in years, and partnered status (not included in models where partnered status was the outcome variable). The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men. Likewise, we also interacted age and partnered status with sex. We interacted age with sex since menopause occurs within the age range of our participants. For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength. Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable (Gelman, CITE).

#### 2.1.1 Outcome variables: 
In industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior. Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993). Our main outcome was mating success, which we operationalized in three ways following Lassek and Gaulin (2009): 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g. female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual). 3) Age at first sexual intercourse reported from: "How old were you when you had sex for the first time?" The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen. Only respondents who could self-report were asked these questions.

Unlike Lassek and Gaulin (2009), we also operationalized mating success in a fourth outcome measure as partnered status, as a proxy for longterm mating strategy since number of partners emphasizes success through short term mating strategies. Partnered status included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated.

Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL). In the original paper, Lassek and Gaulin (2009) also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.

Finally, to assess costs of greater strength, we investigated the relationship between strength and dietary energy intake. Participants reported all food and beverages they consumed in the 24 hour period prior to their interview. This interview was repeated for a different 24 hour period 3-10 days later to obtain dietary recalls for two separate days. These data were then used to estimate energy intake in the form of kilocalories (kcal) calculated by matching reported foods to the USDA's Food and Nutrient Database for Dietary Studies. For the outcome measure in this model we used the average number of calories per day calculated for each participant across their two recall days.

### Models:
We attempted to replicate the regression models in Lassek and Gaulin as closely as possible. There were some unavoidable differences, however. First, our study aimed to investigate the role of sex; thus we include sex as a main effect as well as in interactions with key predictor variables. Second, some variables included in Lassek and Gaulin were not available in our data. Third, we included additional control variables as noted below. Finally, we treated our pilot study as an exploratory study in which we fit numerous models not reported here. The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in Lassek and Gaulin (2009) based on the theories we have described above and given our constraints.

#### Control models: 
Our modelling strategy takes inspiration from Lassek and Gaulin (2009). However, Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples (Smith, 2018). Instead, we will fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses. Control variables included a much wider range of theoretically motivated potential confounds than used by Lassek and Gaulin (2009), organized into themes. We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated with strength and sexual behavior.

#### Anthropometric control model
We first derived a simple model based on that reported in Table 2 in Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status. This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes. We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.

#### Socioeconomic control model
The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength (Johnson & Wilson, 2019) and sexual behavior (Fenton et al., 2005). Likewise, there is also variation in sexual behavior by education (Chandra et al., 2011). physical fitness (CITE).

#### Health control model
An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability (Soysal et al., 2021). Furthermore, health status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women (Lindau et al., 2009). Another study of men aged 45-59 found that lower sexual activity was associated with increased mortality (Smith et al., 1997). The health model, therefore, included a number of variables related to health. White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection. Depression, which is negatively related to strength (Smith et al., 2022), was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument. Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday). These ratings were summed to produce a depression score ranging from 0-27. Chronic illness was included using Chronic Disease Score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease. A point was added for each disease a participant reported being diagnosed with, regardless of any impairment due to the disease. We then controlled for impairment resulting from chronic illness (Disease Impairment Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities. We also controlled for physical disability and using the item "special equipment" which referred to participants' report that they needed special equipment to walk. Finally the 'perceived abnormal weight' variable was coded as true or false depending on whether a participant reported that they perceived their weight to be abnormal.

#### Hormone control model
There is evidence that testosterone is associated with higher numbers of partners, but mixed evidence that it is related to sexual desire (van Anders, 2013). Likewise, evidence of a relationship between circulating testosterone and handgrip strength is also mixed (Nowak-Kornicka et al., 2020). The hormone control model included serum total testosterone (ng/dL). Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex. We therefore computed a sex-specific testosterone value as follows: for each sex we centered testosterone at the sex-specific mean and divided by two times the sex-specific standard deviation. This differed from other centered and standardized variables which we centered and standardized across all male and female values.

#### Physical activity control model
The physical activity control model included four dichotomous variables of vigorous and moderate work and recreational activity coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.

#### Costs of strength
Following Lassek and Gaulin (2009) we also developed models of native immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day). For each outcome we developed a model as similar as possible to those reported in Lassek and Gaulin (2009), except that we always include sex and its interaction with other predictors where we have theoretical reasons to expect sex differences in the effect of the predictor on the outcome. We then also specified an alternative model for each outcome with additional control variables that could confound strength and energy or immune investment. These models were based in part on exploratory analyses of multiple models not reported here.

#### Native Immune Function

Following the significant predictors reported in Lassek and Gaulin (2009, Table 4), we first specified a model with age, sex, strength, and BMI. We also included an age:sex interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women. Finally, we included a strength:sex interaction to test the hypothesis that the immune costs of strength differ by sex.

Based on exploratory analysis we developed a second model of WBCC that included a few additional controls. First, we controlled for sex-specific testosterone and an interaction of sex and sex-specific testosterone since testosterone possibly has immunosuppressive effects, which may differ by sex (CITE). We also controlled for three variables related to energy availability that could impact immune investment: food insecurity, total MET, and average calories per day.

#### Dietary Energy Intake
Based on Lassek and Gaulin (2009, Table 3) we developed a model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET. We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.

We then developed a second model that also controlled for WBCC and food insecurity, since either could influence energy intake and strength.

### Analysis 
All analyses were completed in `r R.version.string`, using the survey package (Lumley, CITE) in order to incorporate the survey sampling weights and to preserve the representative structure of the sample. For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data. Models of age at first sexual intercourse and white blood cell count utilized a normal distribution. Analyses include adults ages 18-59 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.

### 2.2 Pilot Study Results

#### 2.2.1 Descriptive statistics

Table \@ref(tab:sexualdimorphism) shows weighted means and standard deviations of all variables for men and women. Mean age was 38 years for this sample. Sexual dimorphism is given as a ratio of male mean:female mean. Larger male size dimorphism is reflected in an 8% higher value for height and a 16% higher value for weight; females, however, had a 1% higher BMI than males on average. There is a much larger degree of dimorphism in combined grip strength (57%). Hemoglobin is 14% higher in males than females while white blood cell count is 3% higher in women.

```{r sexualdimorphism}

getn <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age,
                height,
                weight,
                bmi,
                strength,
                whitebloodcell,
                hemoglobin,
                testosterone),
    statistic = list(
      all_continuous() ~ "{N_nonmiss_unweighted}"
    ),
    digits = all_continuous() ~ 0,
    missing = "no"
  )

getratio <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age,
                height,
                weight,
                bmi,
                strength,
                whitebloodcell,
                hemoglobin,
                testosterone),
    statistic = list(
      all_continuous() ~ "{mean}"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
  )

tab1 <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(
      age,
      height,
      weight,
      bmi,
      strength,
      whitebloodcell,
      hemoglobin,
      testosterone
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    digits = all_continuous() ~ 2,
    label  = list(
      strength = "Combined Grip Strength (kg)",
      height = "Height (cm)",
      weight = "Weight (kg)",
      age = "Age (years)",
      bmi = "Body mass index (kg/m^2^)",
      whitebloodcell = "White blood cell count (1000 cells/µL)",
      hemoglobin = "Hemoglobin (g/dL)",
      testosterone = "Testosterone (ng/dL)"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
      dplyr::relocate(n_male, .before = stat_1) %>%
      dplyr::relocate(n_female, .before = stat_2)
  )  %>%
  add_difference(test = everything() ~ "smd",
                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
  modify_column_hide(columns = ci) %>%
  modify_header(
    label = "**Variable**",
    ratio = "**Ratio**",
    n_male = "**N**",
    n_female = "**N**",
    stat_1 = "**Mean (SD)**",
    stat_2 = "**Mean (SD)**",
    estimate = "*d*"
  ) %>%
  modify_spanning_header(list(
    c("stat_1", "n_male") ~ "**Male**",
    c("stat_2", "n_female") ~ "**Female**",
    c("ratio", "estimate") ~ "**Sexual Dimorphism**"
  )) %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_caption("**Sex differences for participants ages 18-59 using population weights**") %>%
  bold_labels()

tab1

```

#### 2.2.2 Anthropometric models of mating success

In the anthropometric control model (Figure \@ref(fig:baseline)), strength was significantly associated with higher numbers of lifetime and past year sex partners and an earlier age at first sex, and the interaction with sex was not significant.

```{r baseline, fig.width = 12, fig.height = 8, fig.cap = "Anthropometric Models: Coefficent plot from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standarized by 2 SD."}

forestplot(
  m_agefirst,
  m_lifetime,
  m_pastyear,
  m_partnered,
  intercept = F,
  facet = F,
  dodgewidth = .5,
  modelnames = mnames1,
  varnames = vnames
) + theme_minimal(15) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
#  labs(title = "Figure 1. Baseline Models: Regression coefficients from generalized linear models", color = "Model") +
  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
# 
# 
# annotate_figure(fig1,
#                            top = text_grob("Figure 1. Baseline Models: Regression coefficients from generalized linear models\n",
#                                            size = 20, face = "bold",
#                                            hjust = 0,
#                                            x = 0.01),
#                            bottom = text_grob("Variables labelled (S) have been centered at the mean and standarized by 2 SD",
#                                               size = 12))

#visualize coefs from models derived from lassek and gaulin. 
#add coef table below to match l and g

```

#### 2.2.3 Models of mating success with socioeconomic, health, hormone, and physical activity controls
We then used generalized linear regression to evaluate the impact of a host of other potential confounds on these results. For lifetime number of sex partners, strength remained a significant positive predictor when controlling for socioeconomic, physical activity, health, and hormone variables, with no interaction with sex. This was also the case for past year number of sex partners, except in the case of the hormone control model. Strength predicted an earlier age at first sex for all models except the socioeconomic. There was no significant interaction between sex and strength in any model.

We also modeled partnered status. While strength was a significant predictor of currently being partnered in each model, a significant sex:strength interaction emerged in the physical activity and hormonal control models such that the strength effect was reduced for women. Coefficients are shown in Figure \@ref(fig:controls).

```{r controls, fig.width = 12, fig.height = 42, fig.cap = "Control Models: Coefficent plots from generalized linear models. Variables labelled (S) have been centered at the mean and standarized by 2 SD."}

#include groups of potential confounds to explore lack of sex effect for all 4 mating outcomes

fig1 <- forestplot(
   msoc1, mheal1, mphys1, mhor1,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames, varnames = vnames) + theme_minimal() +
  theme(axis.text.y = element_text(size = 12)) +
 # geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Lifetime Number of Sexual Partners") 
 # coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig2 <- forestplot(
 msoc2,mheal2, mphys2, mhor2,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames
) + theme_minimal() +
#  geom_pointrange(size = 1.2, position = position_dodge(.8)) +
  labs(title = "Past Year Number of Sexual Partners")  +
  theme(axis.text.y = element_blank())
#  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig3 <- forestplot(
  msoc3, mheal3, mphys3, mhor3,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames) +
  theme_minimal() +
#  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Age at First Sex")  +
  theme(axis.text.y = element_blank())
 # coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig4 <- forestplot(
  msoc4, mheal4, mphys4, mhor4,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames) +
  theme_minimal() + 
  #geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
  labs(title = "Currently Partnered")  +
  theme(axis.text.y = element_blank())
#  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

(fig1 + fig2 + fig3 + fig4) + plot_layout(guides = "collect", ncol = 4) & theme(legend.position = "bottom")

# ggarrange(fig1, fig2, fig3, fig4,
#                      labels = "AUTO",
#                      font.label = list(size = 20),
#                      ncol = 2, nrow = 2,
#                      common.legend = TRUE, legend="bottom")

 # annotate_figure(fig1234,
 #                       top = text_grob("Figure 2. Regression coefficients from generalized linear models of mating success\n",
 #                                       size = 20, face = "bold",
 #                                       hjust = 0,
 #                                       x = 0.01),
 #                       bottom = text_grob("\nVariables labelled (S) have been centered at the mean and standarized by 2 SD Base level of race/ethnicity is Mexican American.",
 #                                          size = 12))

```

#### 2.2.4 Immune costs of strength

Previous research found that higher musculature was associated with decreased investment in native immunity. We developed two models of immune investment in white blood cell count (\@ref(fig:immune)). In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI. In the expanded control model, this pattern remained even with additional controls (see Table ).

```{r immune, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of immune investment. Variables labelled (S) have been centered at the mean and standarized by 2 SD."}
forestplot(
  mwbc,
  mwbc_alt,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  varnames = vnames) +
  theme_minimal(25) +
  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
  labs(title = "White blood cell count") 
  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
  #      subtitle = "White Blood Cell Count",
  #      caption = "Variables with (S) have been standarized by 2 SD.") +
  # guides(colour = guide_legend(reverse = T))

```

#### 2.2.4 Model of energy intake

Finally, we tested whether strength was related to increased energy intake. We found that strength was a significant, positive predictor of energy intake controlling for sex. In the expanded control model we found the same pattern while controlling for additional variables.

### 2.3 Pilot Results Discussion

In this pilot study using grip strength as a proxy, we replicated findings from Lassek and Gaulin (2009) that muscularity is associated with increased mating success, as well as with immune and energy costs, even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables. The only exception was that strength was not a positive predictor of past year partners in the hormone control model; the coefficient, however, remained similar in magnitude to the significant strength coefficients in other models of past year partners. In addition, strength was a positive predictor of partnered status, an additional index of mating success.

We found that the negative effect of strength on age at first sex may be due to a confound with ethnicity. Specifically, Asian Americans have markedly lower grip strength compared to Mexican Americans and later age at first sex. For this reason, we do not predict an effect of strength on age at first sex controlling for ethnicity.

We extended these findings to include women in order to test the hypothesis that sexual dimorphism in strength is due to sexual selection for formidability. At the outset of our study we predicted that strength would be a strong predictor of mating success for men and a significantly weaker predictor for women. Contrary to predictions, in our pilot study we found no significant interaction between strength and sex for lifetime number of partners or past year number of partners. This calls into question the sexual selection hypothesis for sexual dimorphism in upper body strength, because our results confirm that strength has immune and energy costs that would be offset by the benefits of multiple mates for men but less plausibly for women. We did, however, find evidence that strength was a significantly stronger predictor of partnered status for men than for women.  In humans biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships. We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability clearly applies to long-term partnerships (proxied by our partnered variable), where strength was a stronger predictor for men than women. Partner numbers are arguably better proxies for short term mating success.

The lack of a sex difference in the significant positive effect of strength on short term mating success is puzzling. Because men have greater strength than women their partner numbers are higher. Still, women with higher grip strength report more partners than women with lower grip strength even after controlling for a host of potential confounds. It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men. It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently their (stronger) mates would also necessarily have more mates. It may be that stronger women require less male investment and so instead benefit from greater partner numbers through e.g. genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for high quality long term mate, or through avoiding costly long term partnerships.  It might also be the case that there are some sex-specific confounds that we failed to control for. Finally, greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.

## 3. Confirmatory Analysis [not yet conducted]

We indicate our predictions for the results of the confirmatory tests in Table X. Specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength. We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models. If the effect of strength on the outcome variable is not significant in any single model, especially if the size of the coefficient is markedly reduced, then we will consider that evidence against the sexual selection hypothesis.

### 3.1 Method

Our methods and models will remain identical to those reported in the pilot study, using unobserved data from the 2013-2014 NHANES collection cycle.

### 2023-07-12 dailyenergy.R

@@ -142,8 +142,9 @@ d.design.dietary.adults <-
   )
 
 m1 <- svyglm(
-  log(avgcalories) ~
+  avgcalories ~
     age +
+    sex +
     total_work_MET  +
     total_rec_MET +
     wob_MET +
@@ -151,23 +152,74 @@ m1 <- svyglm(
     height +
     weight,
   family = gaussian(),
-  design = d.design.dietary.adults
+  design = designsG$d.design.dietary.adults
 )
 summary(m1, df.resid = Inf)
 plot(allEffects(m))
 
 m_energy <- svyglm( #use this
   avgcalories ~
+    age_centered +
+    total_work_MET_centered  +
+    total_rec_MET_centered +
+    wob_MET_centered +
+    strength_centered +
+    sex +
+    bmi_centered +
+    whitebloodcell +
+    foodinsecurity_adult +
+    depression,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_energy, df.resid = Inf)
+plot(m)
+
+m_energy <- svyglm( #use this
+  avgcalories ~
+    age_centered +
+    total_work_MET_centered  +
+    total_rec_MET_centered +
+    wob_MET_centered +
+    strength_centered +
+    sex +
+    bmi_centered +
+    whitebloodcell +
+    foodinsecurity_adult,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_energy, df.resid = Inf)
+
+m_energy <- svyglm( #use this
+  avgcalories ~
+    age_centered +
+    tot_MET +
+    strength_centered +
+    sex +
+    bmi_centered +
+    whitebloodcell +
+    foodinsecurity_adult,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_energy, df.resid = Inf)
+
+m_energy <- glm( #use this
+  avgcalories ~
     age +
     total_work_MET  +
     total_rec_MET +
     wob_MET +
-    strength +
+    log(strength) +
+    sex +
     height +
     weight,
   family = gaussian(),
-  design = d.design.dietary.adults
+  d_G[d_G$age>=18 & d_G$age<=60,]
 )
-summary(m_energy, df.resid = Inf)
-plot(m)
+
+mean_workMET <- svymean(~total_work_MET, designsG$d.design.dietary.adults, na.rm=T)[[1]]
[…]

### 2023-07-12 differencesall.R

@@ -154,9 +154,8 @@ designsG$d.design.adults %>%
       dplyr::relocate(n_male, .before = stat_1) %>%
       dplyr::relocate(n_female, .before = stat_2)
   )  %>%
-  add_difference(test = everything() ~ "smd", ,
-                 estimate_fun = purrr::partial(style_ratio, digits = 2)),
-) %>%
+  add_difference(test = everything() ~ "smd",
+                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
   modify_column_hide(columns = ci) %>%
   modify_header(
     label = "**Variable**",

### 2023-07-12 draft_analysis.R

@@ -5,45 +5,7 @@ library(hagenutils)
 library(ggpubr)
 
 
-# Baseline models ---------------------------------------------------------
-
-vnames <- c(
-  "strength_centered" = "Strength (S)",
-  "sexfemale" = "Sex (Female)",
-  "sexfemale:strength_centered" = "Sex (Female) x Strength",
-  "age_centered" = "Age (S)",
-  "partneredTRUE" = "Partnered",
-  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
-  "age_centered:sexfemale" = "Age (S) x Sex (Female)",
-  "height_centered" = "Height (S)",
-  "weight_centered" = "Weight (S)",
-  "bmi_centered" = "BMI (S)",
-  "sexfemale:bmi_centered" = "Sex (Female) x BMI (S)",
-  "sexfemale:height_centered" = "Sex (Female) x Height (S)",
-  "sexfemale:weight_centered" = "Sex (Female) x Weight (S)",
-  'perceived_abnormal_weightTRUE' = "Perceived Abnormal Weight",
-  "whitebloodcell_centered" = "White Blood Cell Count (S)",
-  "hemoglobin_centered" = "Hemoglobin (S)",
-  "special_equipmentTRUE" = "Need special equip to walk",
-  "chronic_disease_score" = "Chronic Disease Score",
-  "physical_disease_count" = "Physical Disease Count",
-  "depression" = "Depression Score",
-  "log(testosterone)" = "Testosterone (log)",
-  "sexfemale:log(testosterone)" = "Sex (Female) x Testosterone (log)",
-  "testosterone_sex_centered" = "Testosterone (S by Sex)",
-  "sexfemale:testosterone_sex_centered"= "Sex (Female) x Testosterone (S by Sex)",
-  "vigorous_rec" = "Vigorous Recreation",
-  "moderate_rec" = "Moderate Recreation",
-  "vigorous_work" = "Vigorous Work",
-  "moderate_work" = "Moderate Work",
-  "edu" = "Education",
-  "raceOtherHispanic" = "Other Hispanic",
-  "raceNonHispanicBlack" = "Non-Hispanic Black",
-  "raceNonHispanicAsian" = "Non-Hispanic Asian",
-  "raceNonHispanicWhite" = "Non-Hispanic White",
-  "raceOtherRace" = "Other Race"
-)
-
+# Anthropometric models ---------------------------------------------------------
 
 m_lifetime <- svyglm(
   sex_partners ~
[…]

### 2023-07-14 analysis.R
commented out some code

@@ -7,6 +7,8 @@ library(patchwork)
 library(foreign)
 library(effects)
 library(gtsummary)
+library(car)
+library(emmeans)
 #library(sjstats)
 
 #remove 10 males who report taking testosterone supplements
@@ -76,10 +78,10 @@ m14 <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + height_cent
 summary(m14)
 plot(allEffects(m14))
 nobs(m14)
-nobs(mp)
 
 mp <- svyglm(partnered ~ sex + age, family = quasibinomial(), design=designsG$d.design)
 summary(mp) #no sex difference after controlling for age in adults 18-60
+nobs(mp)
 
 
 # m14.1 <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + height_centered*sex + weight_centered * sex +
@@ -223,7 +225,7 @@ Anova(manth2.1, type = 3)
 plot(svysmooth(strength~age, designsG$d.design.adult.female)) #inverted U relationship for both males and females
 #bimodalstrengthXsex partner svysmooth relationship, both males and females
 
-m <- svyglm(sex_partners ~ poly(strength_centered, 3)*sex + poly(age_centered, 2)*sex, family = quasipoisson(), design=designsG$d.design.adults)
+# m <- svyglm(sex_partners ~ poly(strength_centered, 3)*sex + poly(age_centered, 2)*sex, family = quasipoisson(), design=designsG$d.design.adults)
 
 
 #age not linearly related to sex partners, nor is strength
@@ -364,7 +366,7 @@ d <-
 
 x <- d$value
 names(x) <- d$name
-ggdotchart(x[x< -.2])
+# ggdotchart(x[x< -.2])
 
 d_female <-
   d_G %>%
@@ -397,8 +399,8 @@ mal_pos <- ggdotchart(x[x > .1])
 (fem_pos + mal_pos) /
   (fem_neg + mal_neg)
 
-paq <- read.xport("../nhanesGH/data-raw/NHANES data/PAQ_G.XPT")
-names(paq)
+# paq <- read.xport("../nhanesGH/data-raw/NHANES data/PAQ_G.XPT")
+# names(paq)
 # m <- prcomp(na.omit(paq[-c(1,21)]), scale. = TRUE)
 
 
@@ -462,18 +464,18 @@ mpy <- svyglm(sex_partners_year ~ age_centered*sex  + bmi_centered + edu + stren
 summary(mpy)
 plot(allEffects(mpy))
 nobs(mpy)
[…]

### 2023-07-14 spsp_poster.R

@@ -443,9 +443,9 @@ fig1 <- forestplot(
   intercept = F,
   facet = F,
   dodgewidth = .8,
-  modelnames = mnames, varnames = vnames)$plot + theme_minimal(20) +
+  modelnames = mnames, varnames = vnames) + theme_minimal(20) +
   geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
-labs(title = "Lifetime Number of Sexual Partners", color = "Controls") +
+  labs(title = "Lifetime Number of Sexual Partners", color = "Controls") +
   coord_cartesian(clip = "off")
 # + xlim(-2.5, 3.55)
 
@@ -456,7 +456,7 @@ fig2 <- forestplot(
   dodgewidth = .8,
   modelnames = mnames,
   varnames = vnames
-)$plot + theme_minimal(20) +
+) + theme_minimal(20) +
   geom_pointrange(size = 1.2, position = position_dodge(.8)) +
   labs(title = "Past Year Number of Sexual Partners", color = "Controls") +
   coord_cartesian(clip = "off")
@@ -468,7 +468,7 @@ fig3 <- forestplot(
   facet = F,
   dodgewidth = .8,
   modelnames = mnames,
-  varnames = vnames)$plot +
+  varnames = vnames) +
   theme_minimal(20) +
   geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
   labs(title = "Age at First Sex", color = "Controls") +
@@ -481,7 +481,7 @@ fig4 <- forestplot(
   facet = F,
   dodgewidth = .8,
   modelnames = mnames,
-  varnames = vnames)$plot +
+  varnames = vnames) +
   theme_minimal(20) + geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
   labs(title = "Currently Partnered", color = "Controls") +
   coord_cartesian(clip = "off")
@@ -510,7 +510,7 @@ fig5 <- forestplot(
   intercept = F,
   facet = F,
   dodgewidth = .8,
-  varnames = vnames)$plot +
+  varnames = vnames) +
   theme_minimal(25) +
   geom_pointrange(size = 2, position = position_dodge(width = .8)) +
   labs(title = "White blood cell count") +

### 2023-07-14 ms_draft07.06.Rmd

@@
 ---
-title: "Grip strength, mating success, and immune function in US adults: a registered report "
+title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
 author: "C. B. Smith and E. H. Hagen"
@@
    number_sections: false
+csl: apa.csl
+bibliography: regrep.bib
 ---
@@
 library(corrplot)
+library(patchwork)
 source('draft_analysis.R')
@@
 EHB Registered Report
+
+## Abstract
+
+Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
+It is unknown if females experience a similar tradeoff.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
+We will also test if there is a positive relationship between grip strength and dietary energy intake, and a negative relationship between grip strength and native immune function.
+This study replicates an earlier study of US males (Lassek & Gaulin, 2009) and additionally tests these relationships in females.
 
@@
 
-In humans, the immunocompetence hypothesis has primarily been leveraged to explain male facial masculinity and voice pitch but it and more general cost models have also been argued to explain apparent female preference for muscular male body types (Boothroyd et al., 2013).
+In humans, the immunocompetence hypothesis has primarily been leveraged to explain male facial masculinity and voice pitch but it and more general cost models have also been argued to explain apparent female preference for muscular male body types (Frederick & Haselton, 2007).
+Cues of upper body strength indeed account for most of the variation in female-rated male body attractiveness (Sell et al., 2017).
 
@@
 For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
-Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable (Gelman, CITE).
+Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable (Gelman, 2008).
 
@@
 The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength (Johnson & Wilson, 2019) and sexual behavior (Fenton et al., 2005).
-Likewise, there is also variation in sexual behavior by education (Chandra et al., 2011).
-physical fitness (CITE).
+Likewise, there is also variation in sexual behavior by education (Chandra et al., 2013), likewise there is variation in physical activity by education (He & Baker, 2005).
 
@@
 An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability (Soysal et al., 2021).
-Furthermore, health status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women (Lindau et al., 2009).
+Furthermore, health status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women (Lindau & Gavrilova, 2010).
 Another study of men aged 45-59 found that lower sexual activity was associated with increased mortality (Smith et al., 1997).
@@
 White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression, which is negatively related to strength (Smith et al., 2022), was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument.
+Depression, which is negatively related to strength (Smith et al., 2022), was measured using the Patient Health Questionnaire-9 (PHQ-9; Kroenke et al., 2001), a validated nine-item screening instrument.
 Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
@@
 #### Physical activity control model
-The physical activity control model included four dichotomous variables of vigorous and moderate work and recreational activity coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
+Physical activity is related to sexual behavior and function in adults (Morris et al., 2022).
+Therefore, the physical activity control model included four dichotomous variables of vigorous and moderate work and recreational activity coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
 
@@
 
-Following the significant predictors reported in Lassek and Gaulin (2009, Table 4), we first specified a model with age, sex, strength, and BMI.
+Following the significant predictors reported in Lassek and Gaulin (2009, Table 4), we first specified a replication model with age, sex, strength, and BMI.
 We also included an age:sex interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
@@
 
-Based on exploratory analysis we developed a second model of WBCC that included a few additional controls.
-First, we controlled for sex-specific testosterone and an interaction of sex and sex-specific testosterone since testosterone possibly has immunosuppressive effects, which may differ by sex (CITE).
-We also controlled for three variables related to energy availability that could impact immune investment: food insecurity, total MET, and average calories per day.
+Based on exploratory analysis we developed an expanded control model of WBCC that included a few additional controls.
+First, we controlled for sex-specific testosterone and an interaction of sex and sex-specific testosterone since testosterone possibly has immunosuppressive effects, which may differ by sex.
+We also controlled for three variables related to energy availability that could impact immune investment.
+First, food security status for adults in the participant's household was assessed using 10 items from the U.S.
+Food Security Survey Module.
+The resulting value can range from 1-4, where 1 represents full food security and 4 represents very low food security.
+Second, the total metabolic equivalent (total MET) was calculated based on participants' responses to questions about the minutes they spent walking or bicycling, engaging in vigorous and moderate work, and vigorous and moderate recreation per day, using MET scores provided by NHANES.
+Finally we included dietary energy intake (average calories per day as discussed above).
 
 #### Dietary Energy Intake
-Based on Lassek and Gaulin (2009, Table 3) we developed a model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
+Based on Lassek and Gaulin (2009, Table 3) we developed a replication model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
 We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.
 
-We then developed a second model that also controlled for WBCC and food insecurity, since either could influence energy intake and strength.
+We then developed an expanded control model that also controlled for WBCC and food insecurity in addition to the variables in the model above, since either could influence energy intake and strength.
 
 ### Analysis 
-All analyses were completed in `r R.version.string`, using the survey package (Lumley, CITE) in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
+All analyses were completed in `r R.version.string`, using the survey package (version 4.1-1; Lumley, 2021) in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
 For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
 Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.
-Analyses include adults ages 18-59 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
+Analyses include adults ages 18-60 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
 
@@
 
-```{r sexualdimorphism}
-
-getn <- designsG$d.design.adults %>%
+```{r sexual dimorphism}
+getncat <- designsG$d.design.adults %>%
   tbl_svysummary(
     by = sex,
-    include = c(age,
+    include = c(
+                partnered,
+                edu,
+                race,
+                perceived_abnormal_weight,
+                special_equipment,
+                vigorous_work,
+                moderate_work,
+                vigorous_rec,
+                moderate_rec),
+    statistic = list(
+      all_categorical() ~ "{N_nonmiss_unweighted}"
+    ) , missing = "no" )
+
+cat <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(
+      partnered,
+      edu,
+      race,
+      perceived_abnormal_weight,
+      special_equipment,
+      vigorous_work,
+      moderate_work,
+      vigorous_rec,
+      moderate_rec
+    ),
+    label  = list(
+      partnered = "Partnered",
+      edu = "Education",
+      race = "Race and Ethnicity",
+      perceived_abnormal_weight = "Perceived abnormal weight",
+      special_equipment = "Special equipment needed to walk",
+      vigorous_work = "Work involves vigorous activity",
+      moderate_work = "Work involves moderate activity",
+      vigorous_rec = "Recreation involves vigorous activity",
+      moderate_rec = "Recreation involves moderate activity"
+    ),
+    statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(n_male = getncat$table_body$stat_1, digits = 2) %>%
+      dplyr::mutate(n_female = getncat$table_body$stat_2, digits = 2) %>%
+      dplyr::relocate(n_male, .before = stat_1) %>%
+      dplyr::relocate(n_female, .before = stat_2)
+  ) %>% 
+  modify_header(
+    label = "**Variable**",
+    n_male = "**N**",
+    n_female = "**N**",
+    stat_1 = "**Mean (SD) or n (%)**",
+    stat_2 = "**Mean (SD) or n (%)**",
+  ) %>%
+  modify_spanning_header(list(
+    c("stat_1", "n_male") ~ "**Male**",
+    c("stat_2", "n_female") ~ "**Female**"
+  )) %>% 
+  bold_labels()
+
+#####
+
+getncon <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age_first_sex,
+                sex_partners,
+                sex_partners_year,
+                strength,
+                age,
+                bmi,
                 height,
                 weight,
-                bmi,
-                strength,
                 whitebloodcell,
                 hemoglobin,
-                testosterone),
+                testosterone, 
+                chronic_disease_score,
+                physical_disease_count,
+                depression,
+                avgcalories, 
+                foodinsecurity_adult, 
+                tot_MET),
     statistic = list(
-      all_continuous() ~ "{N_nonmiss_unweighted}"
+      all_continuous() ~ "{N_nonmiss_unweighted}",
+      all_categorical() ~ "{N_nonmiss_unweighted}"
     ),
     digits = all_continuous() ~ 0,
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous", 
+      foodinsecurity_adult ~ "continuous"
+    ),
     missing = "no"
@@
     by = sex,
-    include = c(age,
+    include = c(age_first_sex,
+                sex_partners,
+                sex_partners_year,
+                strength,
+                age,
+                bmi,
                 height,
                 weight,
-                bmi,
-                strength,
                 whitebloodcell,
                 hemoglobin,
-                testosterone),
+                testosterone, 
+                chronic_disease_score,
+                physical_disease_count,
+                depression,
+                avgcalories, 
+                foodinsecurity_adult, 
+                tot_MET),
     statistic = list(
-      all_continuous() ~ "{mean}"
+      all_continuous() ~ "{mean}",
+      all_categorical() ~ "{p}"
+    ),
+    digits = list(
+      all_continuous() ~ c(2)
+                  ), 
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous",
+      foodinsecurity_adult ~ "continuous"
     ),
@@
 
-tab1 <- designsG$d.design.adults %>%
+con <- designsG$d.design.adults %>%
   tbl_svysummary(
@@
     include = c(
+      age_first_sex,
+      sex_partners,
+      sex_partners_year,
+      strength,
       age,
+      bmi,
       height,
       weight,
-      bmi,
-      strength,
       whitebloodcell,
       hemoglobin,
-      testosterone
+      testosterone, 
+      chronic_disease_score,
+      physical_disease_count,
+      depression,
+      avgcalories, 
+      foodinsecurity_adult, 
+      tot_MET
     ),
-    statistic = list(all_continuous() ~ "{mean} ({sd})"),
-    digits = all_continuous() ~ 2,
     label  = list(
+      age_first_sex = "Age at first sex (years)",
+      sex_partners = "Lifetime number of sexual partners",
+      sex_partners_year = "Past year number of sexual partners",
       strength = "Combined Grip Strength (kg)",
+      age = "Age (Years)",
+      bmi = "Body mass index (kg/m^2)",
       height = "Height (cm)",
       weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2^)",
       whitebloodcell = "White blood cell count (1000 cells/µL)",
       hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
+      testosterone = "Testosterone (ng/dL)",
+      chronic_disease_score = "Chronic Disease Score (0-6)",
+      physical_disease_count = "Disease Impairment Score (0-5)",
+      depression = "Depression Score (0-27)",
+      avgcalories = "Dietary energy intake (kcals)", 
+      foodinsecurity_adult = "Food Insecurity Rating (1-4)", 
+      tot_MET = "Total MET"
+    ),
+    statistic = list(all_continuous() ~ "{mean} ({sd})",
+                     all_categorical() ~ "{n_unweighted} ({p}%)"),
+    digits = all_continuous() ~ 2,
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous", 
+      foodinsecurity_adult ~ "continuous"
     ),
@@
       dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
-      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
-      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
+      dplyr::mutate(n_male = getncon$table_body$stat_1, digits = 2) %>%
+      dplyr::mutate(n_female = getncon$table_body$stat_2, digits = 2) %>%
       dplyr::relocate(n_male, .before = stat_1) %>%
@@
   )) %>%
-  modify_footnote(update = everything() ~ NA) %>%
-  modify_caption("**Sex differences for participants ages 18-59 using population weights**") %>%
+  modify_footnote(all_stat_cols() ~ "Weighted means and standard deviations shown for continuous variables, unweighted n (%) for categorical variables") %>%
+  modify_caption("**Sex differences for participants ages 18-60 using population weights**") %>%
   bold_labels()
 
-tab1
+tbl_stack(list(con, cat))
+```
+
+```{r sexualdimorphism}
+
+# getn <- designsG$d.design.adults %>%
+#   tbl_svysummary(
+#     by = sex,
+#     include = c(age,
+#                 height,
+#                 weight,
+#                 bmi,
+#                 strength,
+#                 whitebloodcell,
+#                 hemoglobin,
+#                 testosterone),
+#     statistic = list(
+#       all_continuous() ~ "{N_nonmiss_unweighted}"
+#     ),
+#     digits = all_continuous() ~ 0,
+#     missing = "no"
+#   )
+# 
+# getratio <- designsG$d.design.adults %>%
+#   tbl_svysummary(
+#     by = sex,
+#     include = c(age,
+#                 height,
+#                 weight,
+#                 bmi,
+#                 strength,
+#                 whitebloodcell,
+#                 hemoglobin,
+#                 testosterone),
+#     statistic = list(
+#       all_continuous() ~ "{mean}"
+#     ),
+#     missing = "no"
+#   ) %>%
+#   modify_table_body(
+#     ~ .x %>%
+#       dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+#   )
+# 
+# 
+# 
+# tab1 <- designsG$d.design.adults %>%
+#   tbl_svysummary(
+#     by = sex,
+#     include = c(
+#       age,
+#       height,
+#       weight,
+#       bmi,
+#       strength,
+#       whitebloodcell,
+#       hemoglobin,
+#       testosterone
+#     ),
+#     statistic = list(all_continuous() ~ "{mean} ({sd})"),
+#     digits = all_continuous() ~ 2,
+#     label  = list(
+#       strength = "Combined Grip Strength (kg)",
+#       height = "Height (cm)",
+#       weight = "Weight (kg)",
+#       age = "Age (years)",
+#       bmi = "Body mass index (kg/m^2^)",
+#       whitebloodcell = "White blood cell count (1000 cells/µL)",
+#       hemoglobin = "Hemoglobin (g/dL)",
+#       testosterone = "Testosterone (ng/dL)"
+#     ),
+#     missing = "no"
+#   ) %>%
+#   modify_table_body(
+#     ~ .x %>%
+#       dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
+#       dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
+#       dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
+#       dplyr::relocate(n_male, .before = stat_1) %>%
+#       dplyr::relocate(n_female, .before = stat_2)
+#   )  %>%
+#   add_difference(test = everything() ~ "smd",
+#                  estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
+#   modify_column_hide(columns = ci) %>%
+#   modify_header(
+#     label = "**Variable**",
+#     ratio = "**Ratio**",
+#     n_male = "**N**",
+#     n_female = "**N**",
+#     stat_1 = "**Mean (SD)**",
+#     stat_2 = "**Mean (SD)**",
+#     estimate = "*d*"
+#   ) %>%
+#   modify_spanning_header(list(
+#     c("stat_1", "n_male") ~ "**Male**",
+#     c("stat_2", "n_female") ~ "**Female**",
+#     c("ratio", "estimate") ~ "**Sexual Dimorphism**"
+#   )) %>%
+#   modify_footnote(update = everything() ~ NA) %>%
+#   modify_caption("**Sex differences for participants ages 18-59 using population weights**") %>%
+#   bold_labels()
+# 
+# tab1
 
@@
 
-```{r baseline, fig.width = 12, fig.height = 8, fig.cap = "Anthropometric Models: Coefficent plot from generalized linear models (Pilot).
-Variables labelled (S) have been centered at the mean and standarized by 2 SD."}
+```{r baseline, fig.width = 12, fig.height = 8, fig.cap = "Anthropometric models of mating success (Pilot): Coefficient plot from generalized linear models.
+Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
 
-forestplot(
+Figure1 <- forestplot(
   m_agefirst,
@@
   modelnames = mnames1,
-  varnames = vnames
-) + theme_minimal(15) +
-  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
+  varnames = vnames, 
+  size = 1.2, 
+  linewidth = 1
+) + theme_minimal(17) +
+#  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
 #  labs(title = "Figure 1. Baseline Models: Regression coefficients from generalized linear models", color = "Model") +
@@
 
+ggsave("Figures/Figure1.pdf", Figure1, width = 14, height = 7)
+
+Figure1
 ```
@@
 
-```{r controls, fig.width = 12, fig.height = 42, fig.cap = "Control Models: Coefficent plots from generalized linear models.
-Variables labelled (S) have been centered at the mean and standarized by 2 SD."}
+```{r controls, fig.width = 12, fig.height = 42, fig.cap = "Additional control models (Pilot): Coefficient plots for predictors of mating success from generalized linear models.
+Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
 
@@
   dodgewidth = .8,
-  modelnames = mnames, varnames = vnames) + theme_minimal() +
-  theme(axis.text.y = element_text(size = 12)) +
+  modelnames = mnames, 
+  varnames = vnames, 
+  size = 0.7, 
+  linewidth = 0.7) + theme_minimal() +
+  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11)) +
  # geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
@@
   modelnames = mnames,
-  varnames = vnames
+  varnames = vnames, 
+ size = 0.7, 
+  linewidth = 0.7
 ) + theme_minimal() +
@@
   labs(title = "Past Year Number of Sexual Partners")  +
-  theme(axis.text.y = element_blank())
+  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
 #  coord_cartesian(clip = "off")
@@
   modelnames = mnames,
-  varnames = vnames) +
+  varnames = vnames, 
+  size = 0.7, 
+  linewidth = 0.7) +
   theme_minimal() +
@@
   labs(title = "Age at First Sex")  +
-  theme(axis.text.y = element_blank())
+  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
  # coord_cartesian(clip = "off")
@@
   modelnames = mnames,
-  varnames = vnames) +
+  varnames = vnames, 
+  size = 0.7, 
+  linewidth = 0.7) +
   theme_minimal() + 
@@
   labs(title = "Currently Partnered")  +
-  theme(axis.text.y = element_blank())
+  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
 #  coord_cartesian(clip = "off")
@@
 
-(fig1 + fig2 + fig3 + fig4) + plot_layout(guides = "collect", ncol = 4) & theme(legend.position = "bottom")
+Figure2 <- (fig1 + fig2 + fig3 + fig4) + plot_layout(guides = "collect", ncol = 4) & theme(legend.position = "bottom")
 
@@
 
+ggsave("Figures/Figure2.pdf", Figure2, width = 14, height = 7)
+
+Figure2
+
 ```
@@
 Previous research found that higher musculature was associated with decreased investment in native immunity.
-We developed two models of immune investment in white blood cell count (\@ref(fig:immune)).
+We developed two models of immune investment in white blood cell count (Figure \@ref(fig:immune)).
 In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
@@
 
-```{r immune, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of immune investment. Variables labelled (S) have been centered at the mean and standarized by 2 SD."}
-forestplot(
+```{r immune, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
+Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
+Figure3 <- forestplot(
   mwbc,
@@
   facet = F,
-  dodgewidth = .8,
-  varnames = vnames) +
-  theme_minimal(25) +
-  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
-  labs(title = "White blood cell count") 
+  dodgewidth = .4,
+  varnames = vnames, 
+  modelnames = mmnames) +
+  theme_minimal(17) +
+ # geom_pointrange(size = 2, position = position_dodge(width = .8)) +
+ # labs(title = "White blood cell count") +
+  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
   # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
@@
 
+ggsave("Figures/Figure3.pdf", Figure3, width = 14, height = 7)
+
+Figure3
+
 ```
@@
 
-Finally, we tested whether strength was related to increased energy intake.
+Finally, we tested whether strength was related to increased energy intake (Figure \@ref(fig:dietary)).
 We found that strength was a significant, positive predictor of energy intake controlling for sex.
 In the expanded control model we found the same pattern while controlling for additional variables.
+
+```{r dietary, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of predictors of dietary energy intake (kcals) from generalized linear models (Pilot).
+Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
+Figure4 <- forestplot(
+  m_energy,
+  m_energy_alt,
+  intercept = F,
+  facet = F,
+  dodgewidth = .4,
+  varnames = vnames, 
+  modelnames = mmnames) +
+  theme_minimal(17) +
+  #geom_pointrange(size = 2, position = position_dodge(width = .8)) +
+  #labs(title = "Dietary energy intake") + 
+  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
+  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
+  #      subtitle = "White Blood Cell Count",
+  #      caption = "Variables with (S) have been standarized by 2 SD.") +
+  # guides(colour = guide_legend(reverse = T))
+
+ggsave("Figures/Figure4.pdf", Figure4, width = 14, height = 7)
+
+Figure4 
+```
 
@@
 
-We indicate our predictions for the results of the confirmatory tests in Table X.
-Specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength.
+We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength, in Table X.
 We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models.
@@
 Our methods and models will remain identical to those reported in the pilot study, using unobserved data from the 2013-2014 NHANES collection cycle.
+
+#### References

### 2023-07-15 paper.Rmd
Renamed ms and script file to indicate submitted versions

--- 2023-07-14 ms_draft07.06.Rmd
@@
    number_sections: false
-csl: apa.csl
-bibliography: regrep.bib
+# csl: apa.csl
+# bibliography: regrep.bib
 ---
@@
 library(patchwork)
+
+# Needed to save figure pdf's
+if (!file.exists('Figures')){
+  dir.create('Figures')
+}
+
 source('draft_analysis.R')
@@
   add_difference(test = everything() ~ "smd",
-                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
+                 estimate_fun = list(~purrr::partial(style_ratio, digits = 2))) %>%
   modify_column_hide(columns = ci) %>%
@@
 
-## 3. Confirmatory Analysis [not yet conducted]
+## 3. Confirmatory Analysis [data not yet observed]
 

### 2023-10-01 paper.Rmd
Fixed plotting issue

@@
 title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
-author: "C. B. Smith and E. H. Hagen"
+author: "Caroline B. Smith and Edward H. Hagen"
 date: "`r Sys.Date()`"
@@
 
-source('draft_analysis.R')
+source('analysis.R')
 ```
@@
   size = 1.2, 
-  linewidth = 1
+  linewidth = 1,
+  ddf = Inf
 ) + theme_minimal(17) +
@@
 
-fig1 <- forestplot(
-   msoc1, mheal1, mphys1, mhor1,
+# "Partnered x Strength (S)"
+
+forestplot_custom <- function(m1, m2, m3, m4, nm, title, y_axis=F){
+  d <- forestplot(
+    m1, m2, m3, m4,
     intercept = F,
-  facet = F,
-  dodgewidth = .8,
     modelnames = mnames, 
     varnames = vnames,
-  size = 0.7, 
-  linewidth = 0.7) + theme_minimal() +
-  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11)) +
- # geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
-  labs(title = "Lifetime Number of Sexual Partners") 
- # coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
+    ddf = Inf,
+    draw = F
+  )
 
-fig2 <- forestplot(
+  if (!is.null(nm)){
+    for (n in nm){
+      d <- bind_rows(
+        d,
+        tibble(
+          Model = c('Health', 'Hormone', 'Physical Activity', 'Socioeconomic'),
+          term = factor(rep(n, 4)),
+          estimate = c(NA, NA, NA, NA),
+          std.error = c(NA, NA, NA, NA),
+          statistic = c(NA, NA, NA, NA),
+          p.value = c(NA, NA, NA, NA),
+          conf.low = c(NA, NA, NA, NA),
+          conf.high = c(NA, NA, NA, NA)
+        )
+      )
+    }
+
+  }
+  p <- 
+    d %>% 
+    ggplot(aes(estimate, term, xmin=conf.low, xmax=conf.high, colour=Model, shape=Model)) + 
+    geom_pointrange(size = 0.7, linewidth = 0.7, position = position_dodge(width = 0.8)) +
+    geom_vline(xintercept=0, linetype='longdash') +
+    labs(title = title, y='', x='Estimate (95% CI)', ) +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 11))
+
+    if (!y_axis){
+      p <- p + theme(axis.text.y = element_blank())
+    }
+  return(p)
+}
+
+fig1 <- forestplot_custom(
   msoc2,mheal2, mphys2, mhor2,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
- size = 0.7, 
-  linewidth = 0.7
-) + theme_minimal() +
-#  geom_pointrange(size = 1.2, position = position_dodge(.8)) +
-  labs(title = "Past Year Number of Sexual Partners")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
-#  coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
+  NULL, 
+  "Past Year Number of Sexual Partners",
+  y_axis = T
+)
 
-fig3 <- forestplot(
+fig2 <- forestplot_custom(
+  msoc1, mheal1, mphys1, mhor1, 
+  'Partnered x Strength (S)', 
+  "Lifetime Number of Sexual Partners"
+)
+
+fig3 <- forestplot_custom(
   msoc3, mheal3, mphys3, mhor3,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
-  size = 0.7, 
-  linewidth = 0.7) +
-  theme_minimal() +
-#  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
-  labs(title = "Age at First Sex")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
- # coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
+  "Partnered x Strength (S)", 
+  "Age at First Sex"
+)
 
-fig4 <- forestplot(
+fig4 <- forestplot_custom(
   msoc4, mheal4, mphys4, mhor4,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
-  size = 0.7, 
-  linewidth = 0.7) +
-  theme_minimal() + 
-  #geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
-  labs(title = "Currently Partnered")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
-#  coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
+  c("Partnered", "Partnered x Strength (S)"), 
+  "Currently Partnered"
+)
 
 Figure2 <- (fig1 + fig2 + fig3 + fig4) + plot_layout(guides = "collect", ncol = 4) & theme(legend.position = "bottom")
-
-# ggarrange(fig1, fig2, fig3, fig4,
-#                      labels = "AUTO",
-#                      font.label = list(size = 20),
-#                      ncol = 2, nrow = 2,
-#                      common.legend = TRUE, legend="bottom")
-
- # annotate_figure(fig1234,
- #                       top = text_grob("Figure 2. Regression coefficients from generalized linear models of mating success\n",
- #                                       size = 20, face = "bold",
- #                                       hjust = 0,
- #                                       x = 0.01),
- #                       bottom = text_grob("\nVariables labelled (S) have been centered at the mean and standarized by 2 SD Base level of race/ethnicity is Mexican American.",
- #                                          size = 12))
 
@@
   varnames = vnames, 
-  modelnames = mmnames) +
+  modelnames = mmnames,
+  ddf = Inf) +
   theme_minimal(17) +

### 2023-10-01 analysis.R

@@ -1,7 +1,7 @@
 library(survey)
 library(ggplot2)
 library(nhanesGH)
-library(hagenutils)
+library(hagenutils) # has a version of tidy modified to handle df.resid arg of summary.svyglm
 library(ggpubr)
 
 
@@ -460,7 +460,7 @@ summary(m_energy_alt, df.resid = Inf)
 
 mnames <- c(
  # "Anthropometric", baseline model has bmi
-  "Socieoeconomic",
+  "Socioeconomic",
   "Health",
   # "Hormone", old model removed since log(testosterone) is confounded with sex
   "Physical Activity",

### 2023-11-09 old ms drafts/ms_draft05.12.Rmd
organize files

--- 2023-04-11 ms_draft.Rmd
@@
 
-## Abstract
-
-Evidence suggests that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-It is unknown if females experience a similar tradeoff.
-Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey, a large representative sample of the US, we will test whether grip strength, a proxy for upper body strength, is a positive predictor of self-reported outcome measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, and age at first intercourse) in both males and females, and negatively related to a measure of native immune function, white blood cell count, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds.
-These results replicate an earlier study of US males.
-However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive, evidence that a similar trade off in the costs and benefits of muscle exists for women.
-
-1.1	Sexual dimorphism in modern humans:
-
-Whether modern human sexual size dimorphism is inherited or derived remains debated (Plavcan, 2012).
-Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs (Plavcan, 2001).
-Sexual size dimorphism in humans begins in the first trimester in utero where male embryos measure larger than female (Bukowski et al., 2007).
-Newborn boys are more likely to be larger for gestational age than female newborns, while female newborns exhibit higher average skinfold difference and higher relative fat mass.
-However, no sex typical differences in stature are observed during early childhood and the prepubertal period.
-In adults, there is a difference of between 12 and 25% in body mass (Lassek & Gaulin, 2022).
-Humans are slightly more dimorphic in body mass than gibbons, but slightly less than chimpanzees (Smith & Jungers, 1997).
-A modest 7-8% height dimorphism in humans, a reduction from ancestral species (Gray & Wolf, 1980; Gaulin & Boster, 1985).
-Other traits that are dimorphic in humans include digit ratio, voice pitch, facial features, and canine length.
-
-While humans are only moderately dimorphic in terms of overall body mass, this is not the case of allocation of fat and muscle, which are highly dimorphic (Puts, 2010).
-This is partly due to women’s copious storage of fat, unique among primates.
-Human males have between 30 and 42% more fat free mass (Lassek & Gaulin, 2022).
-On average, men have 61% more overall muscle mass and 78% more muscle mass in the upper arms.
-This concentrated muscle dimorphism in the arms and back translates to 90% greater upper body strength in men than women (Lassek & Gaulin, 2009).
-This difference in muscle and strength is not accounted for by size difference alone (Lassek & Gaulin, 2022).
-
-1.2	Sexual dimorphism hypotheses
-
-A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success via sexual selection, either through attracting mates or winning access through contests or both(.
-Two sexual selection hypotheses dominate the literature regarding explanations for human sexual dimorphism, both asserting that men with more masculine traits had higher biological fitness over evolutionary history: intra-sexual selection if more physically formidably men were able to outcompete other men for access to mates and inter-sexual selection if females preferentially mated with men who display masculine traits, either because they are a costly signal of genetic quality, in conjunction with intra-sexual selection if women preferentially mate with more formidable men.
-
-1.2.1	Intrasexual selection: male-male competition
-
-The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past (Puts, 2010), and argues that many masculine traits, including aggression, large body size, and strength, are best explained by an evolutionary history of contest competition for mate access (Puts, 2016, Hill et al., 2017).
-They argue that intrasexual selection due to an evolutionary history of male-male contests is the most parsimonious explanation for the constellation of sexually dimorphic traits in humans, while acknowledging that some degree of sexual size dimorphism could have been inherited from an ancestral species.
-
-Masculine traits including, increased musculature, could arise through male-male competition if formidable men were more successful at attaining resources and mates, either directly through physical contests or indirectly.
-It also could be the case that female choice still plays a role if women preferentially mate with more formidable men.
-While it could be the case that female choice still plays a role in selection for masculine traits, proponents of this hypothesis point to evidence that mating success due to formidability is better explained by male-rated dominance and not their attractiveness to women.
-
-Sexual dimorphism in formidability is taken as evidence of sexual selection, as larger size or strength, typically in males, indicates competition for access to mates.
-In extant primate species, sexual dimorphism in body weight and canine size are strongly associated with degree of intrasexual competition, and the same relationship is used to infer the sexual behavior of extinct australopithecines (Plavcan & van Shaik, 1997).
-Furthermore, contests override other mechanisms of sexual selection, because when males can exclude other males from mating opportunities mate choice and sperm competition become irrelevant (Puts, 2010).
-
-1.2.2 The immunocompetence handicap hypothesis
-
-In contrast to the intra-sexual selection hypothesis, others argue that sexual dimorphism in humans is largely the result of female choice, not for more formidable or competitive men, but for immunocompetent men.
-The immunocompetence handicap hypothesis relies on evidence that testosterone is immunosuppressive, and therefore only men with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine traits.
-Therefore, masculine traits like strength, could be costly signals of good genetic quality, and masculine men should produce higher quality offspring (Folstad & Karter, 1992).
-Under this hypothesis, masculine traits arise through female choice if females preferentially choose to mate with males displaying these traits indicating heritable phenotypic quality, therefore passing those traits to their offspring.
-A few major problems arise for this hypothesis.
-First, evidence that testosterone is actually immunosuppressive is mixed (Nowak et al., 2018).
-Secondly, it is unclear to what extent women are attracted to testosterone-dependent traits, for instance facial masculinity.
-
-Strength and reproductive success
-
-In a meta-analysis, Lidborg et al (2022) found that while voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations.
-The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported on men (Gallup & Fink, ).
-Of the four studies which have tested the association between grip strength and sexual behavior, only two included both women and men.
-In all four, hand grip strength was positively related to sexual behavior including number of sex partners for men, but there was either no relationship or a negative relationship for women.
-
-Previous research has shown that greater muscle mass in males predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (C-reactive protein and white blood cell count) (Lassek & Gaulin, 2009), consistent with a tradeoff in mating success and immunity.
-However, these relationships were not tested in women, where the sexual selection hypothesis predicts no association.
-We plan to expand these analyses to include women using a different proxy for muscle mass, grip strength.
-
-1.3	Experimental aims and hypotheses
-
-Based on results from exploratory analyses, we predict:
-
-1.	Strength will be positively related to the number of self-reported heterosexual sex partners, both total and past year, and negatively related to the age of first intercourse for both men and women.
-2.	Strength will be positively related to currently partnered status for both men and women. 
-3.	Strength will be negatively related to white blood cell count, a measure of native immunity for both men and women.
-
-We will also control for a range of potential confounds including anthropometric, socioeconomic, hormone, health, and physical activity.
-
 ## Methods
@@
 To assess the relationship between strength and reproductive and immune outcomes we used data from the Centers for Disease Control National Health and Nutrition Examination Survey (NHANES) from years 2011-2014.
-NHANES uses a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
+NHANES utilizes a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
 population.
@@
 
-We will use data from years 2011-2014 since grip strength, our key predictor variable was collected in those years.
+We will use data from years 2011-2014 since grip strength, our key predictor variable, was collected in those years.
 We developed models based on previous research and through exploratory analyses of the 2011-2012 dataset to be tested in the 2013-2014 dataset.
-We include data from US adults between the ages of 18 and 65, the years when grip strength is the most stable.
+We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable.
 For each model, participants will be included if they have complete data for each predictor variable.
@@
 In industrialized populations like the U.S., widespread access to contraceptives unlinks reproductive success from sexual behavior.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).Our main outcome was mating success, which we operationalized in three ways.
+Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).Our main outcome variable was mating success, which we operationalized in three ways.
 1) Total sexual partners was a count based on responses to the question: In your lifetime, with how many men/women have you had any kind of sex?
@@
 Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume.
-We operationalized formidability as grip strength (kg), which was measured three times on each hand using a dynamometer.
-We used combined grip strength which was the sum of the highest reading of each hand.
+We operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
 All models also included sex (male/female), age in years, partnered status (partnered included participants who reported either being married or living with a partner while unpartnered included participants who reported being single, widowed, divorced, or separated).
@@
 ### Models:
-We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
+We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with anthropometric, socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
 
 #### Baseline (anthropometric) model
-We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could be related to either strength or mating outcomes.
+We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by grip strength, age, sex, This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
 
@@
 
-The hormone control model included testosterone.
-Since the difference in mean testosterone between men and women is an order of magnitude apart, and testosterone was highly confounded with sex.
+The hormone control model included serum total testosterone.
+Since the difference in mean testosterone between men and women is an order of magnitude apart testosterone was highly confounded with sex.
 We therefore centered testosterone at the mean and divided by two standard deviations within each sex.
-The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they did not.
+The physical activity control model included four dichotomous variables coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
 
@@
 Mean age was 38 years for this sample.
-Sexual dimorphism is given as a ration of male mean:female mean.
+Sexual dimorphism is given as a ratio of male mean:female mean.
 Larger male size dimorphism is reflected in 8% for height and 16% for weight, females, however had a 1% higher BMI than males on average.
@@
   modify_footnote(update = everything() ~ NA) %>%
-  modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-59**")
+  modify_caption("**Table 1. Sexual dimorphism in continuous variables for participants ages 18-59**")
 
@@
 
+```{r}
+summary(m_lifetime, df.resid = Inf)
+```
+
+```{r}
+summ(m_lifetime, df.resid = Inf)
+```
+
 ```{r table2_exact}

### 2023-11-09 old ms drafts/ms_draft_04.06.Rmd
organize files

--- 2023-04-11 ms_draft.Rmd
@@
 Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models.
-Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples (Smith, 2018), we use generalized linear models from the survey package to incorporate survey weights.
+Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models from the survey package to incorporate survey weights.
 For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
@@
 
+```{r tab1age1849}
+
+# has ratio and cohen's d but not sds
+tabba <- d_G[d_G$age >= 18 &
+              d_G$age <= 49, ]  %>% select(
+                "sex",
+                "age",
+                "height",
+                "weight",
+                "bmi",
+                "strength",
+                "whitebloodcell",
+                "hemoglobin",
+                "testosterone",
+              ) %>%
+  tbl_summary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "cohens_d")
+
+#has sds and ratio but not cohen's d
+
+sd_gts <- function(data, variable, ...) {
+  sd(data[[variable]], na.rm = TRUE)
+}
+
+tabb2a <- d_G[d_G$age >= 18 &
+               d_G$age <= 49, ]  %>% select(
+                 "sex",
+                 "age",
+                 "height",
+                 "weight",
+                 "bmi",
+                 "strength",
+                 "whitebloodcell",
+                 "hemoglobin",
+                 "testosterone",
+               ) %>%
+  tbl_strata(
+  strata = sex,
+  ~ .x %>%
+    tbl_summary(
+ #     by = sex,
+      statistic = list(all_continuous() ~ "{mean}"),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
+        age = "Age (years)",
+        bmi = "Body mass index (kg/m^2)",
+        whitebloodcell = "White blood cell count (1000 cells/µL)",
+        hemoglobin = "Hemoglobin (g/dL)",
+        testosterone = "Testosterone (ng/dL)"
+      ),
+      missing = "no"
+    ) %>%
+   add_n() %>%
+    add_stat(fns = everything() ~ sd_gts)) %>%
+  # modify_fmt_fun(list(add_stat_1 ~ partial(style_number, digits = 2))) %>%
+  modify_header(list(stat_0_1 ~ "**Mean**",
+                     stat_0_2 ~ "**Mean**",
+                     add_stat_1_1 ~ "**SD**",
+                     add_stat_1_2 ~ "**SD**")) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_0_1)/as.numeric(stat_0_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**")
+
+# has all!
+tabb3a <- tabb2a %>% modify_table_body(
+  ~ .x %>%
+    dplyr::mutate(d = round(tabba$table_body$estimate, digits = 2))) %>%
+  modify_header(d = "**d**") %>%
+  modify_header(label ~ "**Variable**") %>%
+  modify_spanning_header(c("ratio", "d") ~ "**Sexual dimorphism**") %>%
+  modify_footnote(update = everything() ~ NA) %>%
+  modify_caption("**Table 1. Sex differences based on Lassek and Gaulin Table 1 for ages 18-49**")
+
+tabb3a
+```
+
 ```{r table1survey}
@@
   ) %>%
-  add_n(statistic = "{N_nonmiss_unweighted}", 
-        col_label = "**Sample N**") %>% 
+  add_n() %>%
   modify_table_body(

### 2023-11-09 old table scripts/descriptivestats.R

@@ -1,0 +1,35 @@
+designsG$d.design.adults %>%
+  tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi,
+                                       whitebloodcell, strength, testosterone, partnered,
+                                       height, weight, special_equipment))
+
+
+
+d_G[d_G$age>=18 & d_G$age<=60,]  %>% select("sex", "age", "race", "strength", "bmi", "partnered", "edu", "height", "weight", "whitebloodcell", "testosterone") %>%
+  tbl_summary(by = sex,
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} / {N} ({p}%)"),
+              digits = all_continuous() ~ 2,
+        #      type = list(where(is.numeric) ~ "continuous"),
+             # type = list(partnered ~ "categorical"),
+              type = all_dichotomous() ~ "categorical",
+              label  = list(
+                age = "Age (Years)",
+                race = "Race and Ethnicity",
+                strength = "Combined Grip Strength",
+                bmi = "BMI (kg/m^2)",
+                partnered = "Partnered Status",
+                edu = "Education",
+                height = "Height (cm)",
+                weight = "Weight (kg)",
+                whitebloodcell = "WBCC (1000 cells/µL)",
+                testosterone = "Testosterone (ng/dL)"),
+              missing_text = "Missing"
+  ) %>% add_p() %>%
+  add_overall() %>%
+  add_n() %>%
+  modify_header(label ~ "**Variable**") %>%
+  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**")
+
+
+

### 2023-11-09 old table scripts/tab1long.R

@@ -1,0 +1,77 @@
+
+
+designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(
+      age,
+      height,
+      weight,
+      bmi,
+      strength,
+      whitebloodcell,
+      hemoglobin,
+      testosterone,
+      depression,
+      chronic_disease_score,
+      physical_disease_count,
+      special_equipment,
+      perceived_abnormal_weight,
+      testosterone,
+      vigorous_work,
+      vigorous_rec,
+      moderate_work,
+      moderate_rec
+    ),
+    statistic = list(all_continuous() ~ "{mean} ({sd})"),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)",
+      depressiion = "Depression Score",
+      perceived_abnormal_weight = "Perceived abnormal weight",
+      special_equipment = "Special equipment needed to walk",
+      chronic_disease_score = "Chronic Disease Score (0-6)",
+      physical_disease_count = "Disease Impairment Score (0-5)",
+      depression = "Depression Score (0-27)",
+      vigorous_work = "Work involves vigorous activity",
+      moderate_work = "Work involves moderate activity",
+      vigorous_rec = "Recreation involves vigorous activity",
+      moderate_rec = "Recreation involves moderate activity"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
+      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
+      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
+      dplyr::relocate(n_male, .before = stat_1) %>%
+      dplyr::relocate(n_female, .before = stat_2)
+  )  %>%
+  add_difference(test = everything() ~ "smd",
+                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
[…]

### 2023-11-09 old table scripts/tab1simple.R

@@ -1,0 +1,99 @@
+getn <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{N_nonmiss_unweighted}"
+    ),
+    digits = all_continuous() ~ 0,
+    missing = "no"
+  )
+
+getratio <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  )
+
+
+
+tab1 <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(
+      age,
+      height,
+      weight,
+      bmi,
+      strength,
+      whitebloodcell,
+      hemoglobin,
+      testosterone
+    ),
+    statistic = list(all_continuous() ~ "{mean} ({sd})"),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
+      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
+      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
+      dplyr::relocate(n_male, .before = stat_1) %>%
[…]

### 2023-11-09 old table scripts/tabl1survey.R

@@ -1,0 +1,136 @@
+# has ratio and cohen's d but not sds
+
+tabbsvy <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age,
+                height,
+                weight,
+                bmi,
+                strength,
+                whitebloodcell,
+                hemoglobin,
+                testosterone),
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  add_n() %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "smd") %>%
+  modify_column_hide(columns = ci)
+
+
+
+#get sds
+
+  tabbsvysd <- designsG$d.design.adults %>%
+    tbl_svysummary(
+      by = sex,
+      include = c(age,
+                  height,
+                  weight,
+                  bmi,
+                  strength,
+                  whitebloodcell,
+                  hemoglobin,
+                  testosterone),
+      statistic = list(
+        all_continuous() ~ "{sd}"
+      ),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
+        age = "Age (years)",
+        bmi = "Body mass index (kg/m^2)",
+        whitebloodcell = "White blood cell count (1000 cells/µL)",
+        hemoglobin = "Hemoglobin (g/dL)",
+        testosterone = "Testosterone (ng/dL)"
+      ),
+      missing = "no"
+    )
+
+  tabb3svy <- tabbsvy %>% modify_table_body(
+    ~ .x %>%
[…]

### 2023-11-09 old table scripts/table1.R

@@ -1,0 +1,552 @@
+
+# Table 1. sex differences using sample data -----------------------------
+
+
+# has ratio and cohen's d but not sds
+tabb <- d_G[d_G$age >= 18 &
+              d_G$age <= 60, ]  %>% select(
+                "sex",
+                "age",
+                "height",
+                "weight",
+                "bmi",
+                "strength",
+                "whitebloodcell",
+                "hemoglobin",
+                "testosterone",
+              ) %>%
+  tbl_summary(
+    by = sex,
+    statistic = list(
+      all_continuous() ~ "{mean}"
+    ),
+    digits = all_continuous() ~ 2,
+    label  = list(
+      strength = "Combined Grip Strength (kg)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      age = "Age (years)",
+      bmi = "Body mass index (kg/m^2)",
+      whitebloodcell = "White blood cell count (1000 cells/µL)",
+      hemoglobin = "Hemoglobin (g/dL)",
+      testosterone = "Testosterone (ng/dL)"
+    ),
+    missing = "no"
+  ) %>%
+  modify_table_body(
+    ~ .x %>%
+      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
+  ) %>%
+  modify_header(ratio = "**Ratio**") %>%
+  add_difference(everything() ~ "cohens_d")
+
+
+#has sds and ratio but not cohen's d
+
+sd_gts <- function(data, variable, ...) {
+  sd(data[[variable]], na.rm = TRUE)
+}
+
+
+tabb2 <- d_G[d_G$age >= 18 &
+               d_G$age <= 60, ]  %>% select(
+                 "sex",
+                 "age",
+                 "height",
+                 "weight",
+                 "bmi",
+                 "strength",
+                 "whitebloodcell",
+                 "hemoglobin",
+                 "testosterone",
+               ) %>%
+  tbl_strata(
+  strata = sex,
+  ~ .x %>%
+    tbl_summary(
+ #     by = sex,
+      statistic = list(all_continuous() ~ "{mean}"),
+      digits = all_continuous() ~ 2,
+      label  = list(
+        strength = "Combined Grip Strength (kg)",
+        height = "Height (cm)",
+        weight = "Weight (kg)",
[…]

### 2023-11-09 meetings_exploratory.R
add corr matrices

@@ -1,4 +1,5 @@
 
+
 m_lifetime <- svyglm(
   sex_partners ~
     age_centered * sex +
@@ -93,7 +94,7 @@ library(corrplot)
 # cormat <- round(cor(cordat, use = "complete.obs"),2)
 
 
-cor_mat <- svycor(
+cor_mat_f <- svycor(
   ~ sex_partners +
     sex_partners_year +
     age_first_sex +
@@ -111,12 +112,43 @@ cor_mat <- svycor(
     moderate_rec +
     vigorous_work +
     moderate_work,
-  designsG$d.design.adults,
+  designsG$d.design.adult.female,
   na.rm = T
 )
 
-cor_mat_cors <- cor_mat$cors
-corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
+cor_mat_cors_f <- cor_mat_f$cors
+corr_f <- corrplot(cor_mat_cors_f, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
+         title = "Correlation matrix of coninuous study variables, Females 18-60")
+
+cor_mat_m <- svycor(
+  ~ sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
+    physical_disease_count +
+    testosterone +
+    vigorous_rec +
+    moderate_rec +
+    vigorous_work +
+    moderate_work,
+  designsG$d.design.adult.male,
+  na.rm = T
+)
+
+cor_mat_cors_m <- cor_mat_m$cors
+corrplot(cor_mat_cors_m, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)
+
+corr_m <- corrplot(cor_mat_cors_m, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
+         title = "Correlation matrix of coninuous study variables, Males 18-60")
+
+
 
 # meltedcor_mat <- melt(cor_mat_cors)
 #

### 2023-11-09 revisions.qmd
add new files

--- 2023-07-14 ms_draft07.06.Rmd
@@
 bibliography: regrep.bib
+crossref:
+  custom: 
+    - kind: float # crossreferenceable elements with captions are `float`
+      prefix: Supplementary Figure # used as the prefix for reference in output ("In Figure 1, ..")
+      name: Supplementary Figure # used as the prefix for captions ("Figure 1: ..")
+      ref-type: sfig # used as the prefix for the reference identifier ("In @fig-1, ...")
 ---
@@
 
+### Supplementary Information
+
+```{r, vaginalsex}
+
+```
+
 #### References

### 2023-11-09 test.qmd
add new files

---
title: "practice"
format: html
editor: visual
crossref:
  custom: 
    - kind: float # crossreferenceable elements with captions are `float`
      prefix: Supplementary Figure # used as the prefix for reference in output ("In Figure 1, ..")
      name: Supplementary Figure # used as the prefix for captions ("Figure 1: ..")
      ref-type: sfig # used as the prefix for the reference identifier ("In @fig-1, ...")
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed) see @fig-test.

```{r}
#| label: fig-test
library(ggplot2)
ggplot(mtcars, aes(hp, mpg)) + geom_point()
```

random text see @sfig-test2.

```{r}
#| label: sfig-test2
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) + geom_point()
```

### 2023-11-09 participantcharacteristics.R

@@ -1,0 +1,245 @@
+
+getn <- designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(age_first_sex,
+                sex_partners,
+                sex_partners_year,
+                strength,
+                age,
+                bmi,
+                height,
+                weight,
+                partnered,
+                edu,
+                race,
+                whitebloodcell,
+                hemoglobin,
+                testosterone,
+                perceived_abnormal_weight,
+                special_equipment,
+                chronic_disease_score,
+                physical_disease_count,
+                depression,
+                vigorous_work,
+                moderate_work,
+                vigorous_rec,
+                moderate_rec,
+                avgcalories),
+    statistic = list(
+      all_continuous() ~ "{N_nonmiss_unweighted}",
+      all_categorical() ~ "{N_nonmiss_unweighted}"
+    ),
+    digits = all_continuous() ~ 0,
+    type = list(
+      chronic_disease_score ~ "continuous",
+      physical_disease_count ~ "continuous"
+    ),
+    missing = "no"
+  )
+
+designsG$d.design.adults %>%
+  tbl_svysummary(
+    by = sex,
+    include = c(
+      age_first_sex,
+      sex_partners,
+      sex_partners_year,
+      strength,
+      age,
+      bmi,
+      height,
+      weight,
+      partnered,
+      edu,
+      race,
+      whitebloodcell,
+      hemoglobin,
+      testosterone,
+      perceived_abnormal_weight,
+      special_equipment,
+      chronic_disease_score,
+      physical_disease_count,
+      depression,
+      vigorous_work,
+      moderate_work,
+      vigorous_rec,
+      moderate_rec,
+      avgcalories
+    ),
+    label  = list(
+      age_first_sex = "Age at first sex (years)",
+      sex_partners = "Lifetime number of sexual partners",
+      sex_partners_year = "Past year number of sexual partners",
+      strength = "Combined Grip Strength (kg)",
+      age = "Age (Years)",
+      bmi = "Body mass index (kg/m^2)",
+      height = "Height (cm)",
+      weight = "Weight (kg)",
+      partnered = "Partnered",
+      edu = "Education",
+      race = "Race and Ethnicity",
[…]

### 2023-11-09 revisions-exploratory.R

@@ -1,0 +1,448 @@
+# new exploratory analyses based on reveiwers commments
+
+library(survey)
+library(tidyverse)
+library(nhanesGH)
+library(effects)
+
+# unweighted comparison ---------------------------------------------------
+# we cannot control for all variables accounted for through survey weights (eg response rate and income), and those we can control for (race/ethnicity)
+# are not accounted for in every model and would require completely restructuring our model design. Responded to reviewers that we will retain survey weights
+# but will report an unweighted analysis in the SI
+
+scale2 <- function(x) scale(x)[,1]/2
+
+m_lifetime_uw <- glm(
+  sex_partners ~
+    scale2(age) * sex +
+    scale2(strength) * sex +
+    partnered +
+    race +
+    scale2(bmi) * sex,
+  family = quasipoisson(),
+  data = d_G[d_G$age>=18 & d_G$age<=60,]
+)
+
+
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered +
+    race +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+summary(m_lifetime, df.resid = Inf)
+summary(m_lifetime_uw)
+
+# lifetime partners socieconomic model
+
+msoc1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered +
+      edu +
+      race,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+summary(msoc1, df.resid = Inf)
+
+
+msoc1_uw <-
+  glm(
+    sex_partners ~
+      scale2(age) * sex +
+      scale2(strength) * sex +
+      partnered +
+      edu +
+      race,
+    family = quasipoisson(),
+    data = d_G[d_G$age>=18 & d_G$age<=60,]
+  )
+summary(msoc1_uw)
+
+# correlation matrices ----------------------------------------------------
+# reported in SI
+# need sex specific testosterone?
+
+cor_mat_f <- svycor(
+  ~ sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
[…]

### 2023-11-09 tables_df_pvalues.R

@@ -1,0 +1,304 @@
+
+
+
+# anthropometric model ----------------------------------------------------
+
+
+mm1 <- tbl_regression(
+  m_lifetime,
+  label = vnames
+) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
+  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
+  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+
+
+mm2 <- tbl_regression(
+  m_pastyear,
+  label = vnames
+) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
+  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
+  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+
+mm3 <- tbl_regression(
+  m_agefirst,
+  label = vnames
+) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
+  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
+  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+
+mm4 <- tbl_regression(
+    m_partnered,
+    label = vnames
+  ) %>%
+  add_significance_stars(hide_se = TRUE,
+                         pattern = "{estimate}{stars}<br>({std.error})") %>%
+  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
+  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
+
+tbl_merge(
+  tbls = list(mm1, mm2, mm3, mm4),
+  tab_spanner = c(
+    "Lifetime number of sexual partners",
+    "Past year number of sexual partners",
+    "Age at first sexual intercourse",
+    "Currently partnered"
+  )
+) %>%
+  as_gt() %>%
+  gt::tab_header(
+    title = "Table 2. Baseline Models of Mating Success",
+    subtitle = "Regression coefficients from generalized
+                 linear models for three indices of mating success for men and women ages
+                 18-59. Models derived from Lassek and Gaulin 2009"
+  )
+
+summary(m_lifetime, df.resid = Inf)
[…]

### 2023-11-09 tmp.R

@@ -1,0 +1,10 @@
+forestplot(
+  mwbc,
+  mwbc_alt,
+  intercept = F,
+  facet = F,
+  dodgewidth = .8,
+  varnames = vnames)$plot +
+  theme_minimal(25) +
+  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
+  labs(title = "White blood cell count")

### 2023-11-09 unweighted.R

@@ -1,0 +1,56 @@
+scale2 <- function(x) scale(x)[,1]/2
+
+m_lifetime_uw <- glm(
+  sex_partners ~
+    scale2(age) * sex +
+    scale2(strength) * sex +
+    partnered +
+    race +
+    scale2(bmi) * sex,
+  family = quasipoisson(),
+  data = d_G[d_G$age>=18 & d_G$age<=60,]
+)
+
+
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered +
+    race +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+summary(m_lifetime, df.resid = Inf)
+summary(m_lifetime_uw)
+
+# lifetime partners socieconomic model
+
+msoc1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered +
+      edu +
+      race,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+summary(msoc1, df.resid = Inf)
+
+
+msoc1_uw <-
+  glm(
+    sex_partners ~
+      scale2(age) * sex +
+      scale2(strength) * sex +
+      partnered +
+      edu +
+      race,
+    family = quasipoisson(),
+    data = d_G[d_G$age>=18 & d_G$age<=60,]
+  )
+summary(msoc1_uw)
+

### 2023-11-15 compare.Rmd
comparing heterosexual models

---
title: "compare"
author: "C. B. Smith"
date: '2023-11-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
library(weights)
library(jtools)
library(reshape2)
library(gtsummary)
library(corrplot)
library(patchwork)
source('modelsregressiontables.R')
```

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control
```{r partners}
lifetime_compare
```

Models of lifetime sex partners with original sample, restricted sample
```{r past year}
pastyear_compare
```

Models of lifetime sex partners with original sample, restricted sample
```{r age first sex}
agefirst_compare
```

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control
```{r partnered}
partnered_compare
```

```{r married}

Married_compare2
```

### 2023-11-15 compare.qmd
comparing heterosexual models

---
title: "model comparisons"
format: html
editor: visual
---

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control

```{r partners}
lifetime_compare
```

Models of lifetime sex partners with original sample, restricted sample

```{r past year}
pastyear_compare
```

Models of lifetime sex partners with original sample, restricted sample

```{r age first sex}
agefirst_compare
```

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control

```{r partnered}
partnered_compare
```

```{r figs}

# Figures -----------------------------------------------------------------

mvnames <- c(
  "Anthropometric", #baseline model has bmi
  "Socieoeconomic",
  "Health",
  "Physical Activity",
  "Hormone" #sub new model with sex centered testosterone
)

fig1h <- forestplot(
  m_lifetime_h, msoc1_h, mheal1_h, mphys1_h, mhor1_h,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7) + theme_minimal() +
  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11)) +
  labs(title = "Lifetime Number of Sexual Partners (restricted to self-identified heterosexual)")

fig2h <- forestplot(
  m_pastyear_h, msoc2_h, mheal2_h, mphys2_h, mhor2_h,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Past Year Number of Sexual Partners (restricted to self-identified heterosexual)")  +
  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))

fig3h <- forestplot(
  m_agefirst_h, msoc3_h, mheal3_h, mphys3_h, mhor3_h,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Age at first sex (restricted to self-identified heterosexual)")  +
  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))

fig4h <- forestplot(
  m_partnered_h, msoc4_h, mheal4_h, mphys4_h, mhor4_h,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Currently partnered (restricted to self-identified heterosexual)")  +
  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))

#same models not restricted to heterosexual
fig1 <- forestplot(
  m_lifetime, msoc1, mheal1, mphys1, mhor1,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7) + theme_minimal() +
  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11)) +
  labs(title = "Lifetime Number of Sexual Partners")

fig2 <- forestplot(
  m_pastyear, msoc2, mheal2, mphys2, mhor2,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Past Year Number of Sexual Partners")  +
  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11))

fig3 <- forestplot(
  m_agefirst, msoc3, mheal3, mphys3, mhor3,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Age at first sex")  +
  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11))

fig4 <- forestplot(
  m_partnered, msoc4, mheal4, mphys4, mhor4,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mvnames,
  varnames = vnames,
  size = 0.7,
  linewidth = 0.7
) + theme_minimal() +
  labs(title = "Currently partnered")  +
  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11))

compare_any <- (fig1 + fig1h) + plot_layout(guides = "collect", ncol = 2) & theme(legend.position = "bottom")
compare_any

compare_pastyear <- (fig2 + fig2h) + plot_layout(guides = "collect", ncol = 2) & theme(legend.position = "bottom")
compare_pastyear

compare_agefirst <- (fig3 + fig3h) + plot_layout(guides = "collect", ncol = 2) & theme(legend.position = "bottom")
compare_agefirst

compare_partnered <- (fig4 + fig4h) + plot_layout(guides = "collect", ncol = 2) & theme(legend.position = "bottom")
compare_partnered

```

### 2023-11-15 draft_analysis.R

@@ -3,6 +3,7 @@ library(ggplot2)
 library(nhanesGH)
 library(hagenutils)
 library(ggpubr)
+library(tidyverse)
 
 
 # Anthropometric models ---------------------------------------------------------

### 2023-11-15 modelsregressiontables.R

@@ -1,0 +1,2648 @@
+library(survey)
+library(ggplot2)
+library(nhanesGH)
+library(hagenutils)
+library(ggpubr)
+library(tidyverse)
+library(gtsummary)
+
+# mating success models and tbl_regression objects using d.design.adults (NOT restricted to self-identified heterosexuals)
+
+
+# original models (adults with complete data 18-59) -------------------------------------------------
+
+# lifetime partners - baseline/anthropometric
+
+m_lifetime <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+# lifetime partners - socioeconomic
+msoc1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered + #* strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year; took out 4.6
+      edu +
+      race,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+
+
+# lifetime partners - health model
+mheal1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered +
+      perceived_abnormal_weight +
+      whitebloodcell_centered +
+      hemoglobin_centered +
+      special_equipment +
+      chronic_disease_score +
+      physical_disease_count +
+      depression,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+
+# lifetime partners - hormone
+mhor1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered +
+      testosterone_sex_centered * sex,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+
+
+# lifetime partners - phys activity
+mphys1 <-
+  svyglm(
+    sex_partners ~
+      age_centered * sex +
+      strength_centered * sex +
+      partnered +
+      vigorous_rec +
+      moderate_rec +
+      vigorous_work +
+      moderate_work,
+    family = quasipoisson(),
+    design = designsG$d.design.adults
+  )
+
+# past year partners - baseline/ anthropometric
+m_pastyear <- svyglm(
+  sex_partners_year ~
+    age_centered * sex +
[…]

### 2023-11-15 revisions-exploratory.R

@@ -446,3 +446,795 @@ mheal3_1 <-
     design = designsG$d.design.adults
   )
 summary(mheal3_1, df.resid = Inf)
+
+
+
+# partnered 2 -------------------------------------------------------------
+
+designsG$d.design.adults <- update(designsG$d.design.adults,
+                                   partnered3 = case_when(
+                                     maritalstatus == 2 |
+                                       maritalstatus == 3 |
+                                       maritalstatus == 4 |
+                                       maritalstatus == 5 ~ "Unpartnered",
+                                     maritalstatus == 6 ~ "Livingwithpartner",
+                                     maritalstatus == 1 ~ "Married"
+                                   )
+)
+
+
+m_lifetime_p2 <- svyglm(
+  sex_partners ~
+    age_centered * sex +
+    strength_centered * sex +
+    partnered3 * sex +
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+summary(m_lifetime, df.resid = Inf)
+summary(m_lifetime_p2, df.resid = Inf)
+
+
+m_lifetime_p3 <- svyglm(
+  sex_partners_year ~
+    age_centered * sex +
+    strength_centered * sex +
+    (maritalstatus == 1) * sex+
+    bmi_centered * sex,
+  family = quasipoisson(),
+  design = designsG$d.design.adults
+)
+
+
+
+m_partnered_1 <-
+  svyglm(
+    (maritalstatus == 1) ~
+      age_centered * sex +
+      strength_centered * sex +
+      bmi_centered * sex,
+    family = quasibinomial(),
+    design = designsG$d.design.adult.heterosexual
+  )
+
+
+#partnered status - hormone
+mhor4_1 <-
+  svyglm(
+    partnered ~
+      age_centered * sex +
+      strength_centered * sex +
+      testosterone_sex_centered * sex +
+      heterosexual,
+    family = quasibinomial(),
+    design = designsG$d.design.adults
+  )
+
+#partnered status - phys activity
+mphys4_1 <-
+  svyglm(
+    partnered ~
+      age_centered * sex +
+      strength_centered * sex +
[…]

### 2023-11-15 old/ms_draft.Rmd

--- 2023-07-11 ms_draft.Rmd
@@
 ```
-EHB Research Report
+EHB Registered Report
+
+1.4 Experimental aims and hypotheses
+
+In a large, nationally representative US sample, we predict that hand grip strength will be significantly positively associated with male mating success but not female mating success.
+We also predict that there will be a negative association between male grip strength and white blood cell count, but no association between female grip strength and white blood cell count.
+
+2. Exploratory analysis
+
+In order to inform our hypotheses and refine our models, we first conducted exploratory analyses.
 
@@
 
+Finally, we modeled dietary energy intake in the form of calories consumed.
+Participants reported all food and beverages they consumed in the 24 hours prior to their interview.
+This interview was repeated for a different 24 hour period 3-10 days later to obtain dietary recalls for two separate days.
+These data were then used to estimate energy intake in the form of kila-calories calculated by matching reported foods to the USDA's Food and Nutrient Database for Dietary Studies.
+We used the number of calories for each participant averaged across their two recall days.
+
 ### Predictor variables: 
@@
 #### Baseline (anthropometric) model
-We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and current partnered status.
+We first derived a simple baseline model from Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
 This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
@@
 #### Additional control variables: 
-to additionally test if the association of strength with mating success shown in Lassek and Gaulin (2009) was due to some unmeasured variable, we fit regression models with a much wider range of potential confounds.
-We specified four additional models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with anthropometric, socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
+In order to test if the association of strength with mating success shown in Lassek and Gaulin (2009) was due to some unmeasured variable, we fit regression models with a much wider range of potential confounds.
+We specified four additional models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated either with strength or sexual behavior.
 
-The socioeconomic control model included education and race as categorical variables.
+First, the socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength (Johnson & Wilson, 2019) and sexual behavior (Fenton et al., 2005).There is also variation in sexual behavior by education (Chandra et al., 2011).
 
@@
 White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument.
+Depression, which is negatively related to strength (Smith et al., 2022), was measured using the Patient Health Questionnaire-9 (PHQ-9), a validated nine-item screening instrument.
 Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
 These ratings were summed to produce a depression score ranging from 0-27.
-Chronic illness was included using a 'chronic disease score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease.
-A point was added for each disease a participant indicated they had been diagnosed with, regardless of any impairment due to the disease.
+Chronic illness was included using Chronic Disease Score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease.
+A point was added for each disease a participant reported being diagnosed with, regardless of any impairment due to the disease.
 We then controlled for impairment resulting from chronic illness (Disability Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities.
@@
 
-The hormone control model included serum total testosterone.
-Since male mean testosterone is an order of magnitude higher than female,  testosterone was highly confounded with sex.
-We therefore computed a testosterone value as follows: we centered testosterone at the mean and divided by two standard deviations within each sex.
+The hormone control model included serum total testosterone (ng/dL).
+Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex.
+We therefore computed a testosterone value as follows: we centered testosterone at the mean and divided by two standard deviations within each sex, differing from other centered and standardized variables which we centered and standardized across all male and female values.
 
@@
 
-Our outcome variables were identical to those used by Lassk and Gaulin (2009) ie lifetime, past year number of partners and age at first sex.
-However, the NHANES questions include any kind of sex (e.g.
-oral, vaginal, anal).
-Since, however, NHANES also collects more specific data on sexual behavior we also fit Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex.
-In these questions participants were asked “In your lifetime, with how many men/women have you had vaginal sex?” and “In the past 12 months, with how many men/women have you had vaginal sex?”
-
-We did not include a model of daily energy intake as reported by Lassek and Gaulin, because the gold-standard research utilizes fat-free mass as a predictor, which the original authors use.
-Grip strength is therefore unlikely to provide novel insight into explaining daily energy intake.
-
 #### Analysis 
+All analyses were completed in `r R.version.string`, using the survey package in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
 Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models.
@@
 
-### 3. Results: Exploratory Analysis
+### 3. Results
 
@@
 ```{r}
-l1 <- tbl_regression(msoc1)
-
-l2 <- tbl_regression(mheal1)
-
-l3 <- tbl_regression(mhor1)
-
-l4 <- tbl_regression(mphys1) 
-  tbl_merge(
-    tbls = list(l1, l2, l3, l4),
-    tab_spanner = c("Socioeconomic",
-                    "Health",
-                    "Hormone",
-                    "Physical Activity")) %>%
-   as_gt() %>%
-  gt::tab_header(title = "Lifetime number of partners control models",
-                 subtitle = "Standardized regression coefficents from generalized
-                 linear models")
+# l1 <- tbl_regression(msoc1)  
+# 
+# l2 <- tbl_regression(mheal1) 
+# 
+# l3 <- tbl_regression(mhor1) 
+# 
+# l4 <- tbl_regression(mphys1) 
+#   tbl_merge(
+#     tbls = list(l1, l2, l3, l4),
+#     tab_spanner = c("Socioeconomic",
+#                     "Health",
+#                     "Hormone",
+#                     "Physical Activity")) %>%
+#    as_gt() %>%
+#   gt::tab_header(title = "Lifetime number of partners control models",
+#                  subtitle = "Standardized regression coefficents from generalized
+#                  linear models")
 
@@
 
+## 4. Confirmatory Analysis
+
+### 4.1 Method
+
 ### Discussion

### 2023-11-15 old/strengthmating.Rmd

---
title: "Grip strength, mating success, and immune function in US adults: a registered report "
author: "C. B. Smith and E. H. Hagen"
date: "12/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
source('pre_reg.R')
```
EHB Research Report

##Abstract

Evidence suggests that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection. It is unknown if females experience a similar tradeoff. Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey, a large representative sample of the US, we will test whether grip strength, a proxy for upper body strength, is a positive predictor of self-reported outcome measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, and age at first intercourse) in both males and females, and negatively related to a measure of native immune function, white blood cell count, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds. These results replicate an earlier study of US males. However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive, evidence that a similar trade off in the costs and benefits of muscle exists for women.

#1.	Introduction
a.	Literature review

1.1	Sexual dimorphism in modern humans: height, weight, limb length, upper body strength, lower body strength. Relative to dimorphism in australopiths and early homo; Relative to: extant non-human primates

Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs (Plavcan, 2001). There is a modest 7-8% stature dimorphism in humans (Gray & Wolf, 1980; Gaulin & Boster, 1985) and a difference of about 15–20% in body mass (Mayhew & Salm, 1990). This is partly due to women’s copious storage of fat, unique among primates. While humans are only moderately dimorphic in terms of overall body mass, this is not the case of allocation of fat and muscle, which are highly dimorphic (Puts, 2010). Human males have 40% more fat free mass and 60% more total lean muscle mass. On average, men have 61% more overall muscle mass and 78% more muscle mass in the upper arms. This concentrated muscle dimorphism in the arms and back translates to 90% greater upper body strength in men than women (Lassek & Gaulin, 2009). Sexual size dimorphism in humans begins in utero where male embryos measure larger than female. Newborn boys are more likely to be larger for gestational age than female newborns, while female newborns exhibit higher average skinfold difference and higher relative fat mass. However, no sex typical differences in stature are observed during early childhood and the prepubertal period. Other traits that are dimorphic in humans include digit ratio, voice pitch, facial features, and canine length.

Whether modern human sexual size dimorphism is inherited or derived remains debated (Plavcan, 2012). Changes in body size and stature occurred over the course of human evolution, particularly the early homo species, characterized by taller stature, human-like limb proportions, and larger brains than the earlier australopithecines. Likely reflecting a shift from the plant-based diet of the australopiths to the scavenging strategy of early homo (Gallagher, 2013). Previous research posits that the dramatic male bias in upper body strength is due to sexual selection, either from advantages conferred by greater strength in male-male contests or female preference for stronger mates. Others have proposed that strength is a cue of important mate qualities, such as health or immune function. The sexual selection hypothesis is supported by studies that have found that grip strength is associated with both mating success and reproductive success in men. Although there is some evidence that strength is associated with reproductive success in women, the relationship between female strength and mating outcomes is poorly understood.

##1.2	Sexual dimorphism hypotheses

1.2.1 Sexual selection

Alternatively, masculine traits could arise through male-male competition if formidable men were more successful at attaining resources and mates, which could arise either directly through physical contests or indirectly. It also could be the case that female choice still plays a role if women preferentially mate with more formidable men.

Sexual dimorphism in formidability is taken as evidence of sexual selection, as larger size or strength, typically in males, indicates competition for access to mates.

Sexual size dimorphism also has implications for male reproductive success when it is skewed by the number of females a male can inseminate and male-male competition and mate guarding increase RS, as well as social behavior including degree of polygyny and competition in male dominance hierarchies.

Furthermore, contests override other mechanisms of sexual selection, because when males can exclude other males from mating opportunities mate choice and sperm competition become irrelevant (Puts, 2010).

In extant species, sexual dimorphism in body weight and canine size are strongly associated with degree of intrasexual competition, and the same relationship is used to infer about the sexual behavior of extinct australopithecines (Plavcan & van Shaik, 1997). 
Lidborg et al (2022) found that voice pitch, height and testosterone levels were associated with mating success in low fertility populations, but only muscularity was associated with actual reproductive success in high fertility populations.

Male anthropoid primates compete for access to mates and this competition results in reproductive skew, and it is presumed that formidability in body and canine size increase the chance of success in such competitions (Plavcan, 2012). In nonhuman primates, dimorphism in canine size is one example. Large canines are hypothesized to function as weapons, and although males and females may compete for different things (mates, resources) selection will still drive canine size as long as it relates to winning and losing such competitions. Because dimorphism is typically measured as a ratio of male to female size, it is equally important to consider female variation as male since that ratio will vary as a function of either. Female primates are sometimes involved in aggressive contests, which, following the logic applied to male contests, indicates that canines might be weapons that are useful in competition, as such factors that influence female canine size is as important as male. There is a high degree of variation in canine size of female primates.

Evidence that muscularity and strength are related to formidability:

Several studies have found that whereas females are more likely to experience anger, males are more likely to express anger (Campbell& Muncer, 2008; Barrett et al., 2013) and that there appears to be a cognitive mechanism for the assessment of others’ strength based on their bodies. Sell et al., (2009) found that perceived fighting ability was not due to variation in size, but variation in perceived upper body strength, such that men’s actual upper body strength predicted both their perceived strength and fighting ability while controlling for height, weight, and age. Although raters can estimate women’s strength from the same cues they do men’s, they are much less accurate.

1.2.1.1 Intrasexual competition in women

Because of the costs and constraints of anisogamy, internal gestation, and lactation women bear more of the costs of reproduction than men. This also means that women’s reproductive success is not constrained by number of mates, but typically by resource access. Therefore, sexual selection for contests is not expected to have shaped traits in women. Human males, however, invest extensively in offspring compared to other primates, forming a different basis for female mate competition. Furthermore, women may face a tradeoff in selecting mates based on their paternal investment or the health or formidability benefits of traits like muscularity since there is some evidence that more masculine men reduce investment in partners and offspring.

Limitations on energetic throughput (ie the amount of energy required to sustain more muscle is not possible during key life history events including gestation and lactation, but also baby carrying) evolution of homo, food sharing, investment in reproduction.

1.2.2 The immunocompetence handicap hypothesis

The immunocompetence handicap hypothesis states that secondary sexual features which are produced through testosterone, like strength, could be costly signals of good genetic quality and immunocompetence because only high-quality males can pay the immunosuppressive costs of testosterone (Folstad & Karter, 1992), although evidence that testosterone is actually immunosuppressive is mixed (Nowak et al., 2018). Under this hypothesis masculine traits arise through female choice if females preferentially choose to mate with males displaying these traits indicating heritable phenotypic quality, therefore passing those traits to their offspring. The relationship between testosterone levels in adolescence is only weakly related to testosterone levels in adulthood (van Bokhoven et al., 2006), and masculine traits are not consistently associated with testosterone (Peters et al., 2008).

1.3 Relationship between stature, strength, and reproductive/mating success; Measuring mating success v reproductive success; Link to strength and work effort, importance for child-rearing and reproductive success.

Dufour and Sauther (2002) argue that both human and nonhuman primates increase energy intake, rely on body tissue stores, reduce physical active and in some cases increase metabolic efficiency to meet the costs of lactation. Likewise, strength has been critical for childrearing in the ancestral past, linking strength and reproductive success in women.

Sexual dimorphism in stature and body composition is also influence by a number of social factors such as social status, disease burden, workload, and nutritional factors–many of which are associated with gender roles, such as division of labor and subsequent workload and food distribution. Sexual size dimorphism in humans is physiologically based in sex specific differences in biological factors including genes and hormones, the result of natural selection, but also in environmental factors that more closely track social status and gender roles, such as hormone secretion and nutrition (Kirchengast, 2014).

Previous research has shown that greater muscle mass in males predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (Lassek & Gaulin, 2009). We plan to expand these analyses to include women, and using a different proxy for muscle mass, grip strength.

What role do a history of mild polygyny, inter-male competition, but other contributing factors such as high levels of egalitarianism, phylogenetic inertia, and sex-biased parental investment play in modern levels of dimorphism?

##1.3	Experimental aims and hypotheses

1.	Is strength positively associated with reproductive success in women as well as men? 
2.	Why is strength positively associated with reproductive success in men and women?  
3.	Under what conditions would we still expect sexual dimorphism in strength even when it is positively associated with reproductive success for both males and females?

Based on pilot data, we predict that grip strength will be positively associated with mating success and negatively associated with white blood cell count, in women as well as men, although the coefficient size might be smaller.

#2	Method
3	Participants
NHANES uses a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S. population. NHANES combines interview, examination, and laboratory data to assess health status and identify health risks for adults and children in the United States. Data collection occurs in new cycles every two years. Participants are US adults between the ages of 18 and 65, the years when grip strength is the most stable. For each model, participants will be included if they have complete data for each predictor variable.

This study is a secondary registration of publicly available data collected for the Centers for Disease Control National Health and Nutrition Examination Survey (NHANES). We developed models through exploratory analyses of one dataset to be tested in a holdout dataset. 
Our key predictor variable for our models is combined grip strength, which was measured in NHANES cycles G (2011-2012) and H (2013-2014).  Our hypotheses were developed through exploratory analysis using 2011-2012 data series, we will use the 2013-2014 data for confirmatory analyses.

4	Procedures

Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models. Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models here. We defined our models using the models reported by Lassek and Gaulin using the predictors that were not eliminated by the step wise regression.

4.1 Measures

Predictor variable: 
Grip strength

Lassek and Gaulin operationalized formidability through fat-free mass and limb muscle volume. We operationalized formidability as grip strength, which was measured three times on each hand using a dynamometer. We used combined grip strength which was the sum of the highest reading of each hand. All models included sex (male/female) and age in years.

Control variables: 
We controlled for potential confounds where possible.

Body mass index (BMI), calculated as kg/m^2
Partnered status: Participants were coded as partnered if they reported being married or living with a partner, and unpartnered if they reported being widowed, divorced, separated, or never married. 
Education 
Current salary Measures of wealth available in NHANES are collected at the household level, and as such do not accurately reflect wealth or poverty at the individual level. As a measure of wealth, we used occupation codes collected for each participant by NHANES and imputed the median salary calculated for that occupation category by the 
MET

Outcome variables: 
Mating success
Total number of partners in lifetime (In your lifetime, with how many men/women have you had any kind of sex?)
Total number of sexual partners in past year (In the past 12 months, with how many women have you had any kind of sex?)
Age at first sexual intercourse (How old were you when you had sex for the first time?)

We used generalized linear models using the survey package to incorporate survey weights. For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data. Models of age at first sexual intercourse utilized a normal distribution.

In keeping with the outcome variables utilized in the original paper, we used number of partners for any kind of sex above. Since, however, NHANES also collects more specific data on sexual behavior we also tested Models 1 and 2 using lifetime and past year numbers of vaginal sex partners respectively, since heterosexual vaginal intercourse should be more directly related to reproductive success than other kinds of sex. In these questions participants were asked “In your lifetime, with how many men/women have you had vaginal sex?” and “In the past 12 months, with how many men/women have you had vaginal sex?”

Total number of vaginal sex partners lifetime 
Total number of vaginal sex partners past year

Because of the way these questions are framed, (e.g. female participants were asked about their male partners) these models predict heterosexual behavior, however it is not necessarily the case that all of the participants included in these models identified as heterosexual.

The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen. Only respondents who could self-report were asked these questions.

In industrialized populations like the U.S. widespread access to contraceptives unlinks reproductive success from sexual behavior. Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).

Native immunity
White blood cell count (1000 cells/ µL)

Missing variables 
We do not have data on bone mineral density

We did not include a model of daily energy intake as reported by Lassek and Gaulin, because the gold-standard research utilizes fat-free mass as a predictor, which the original authors use. Grip strength is therefore unlikely to provide novel insight into explaining daily energy intake.

Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models. Since stepwise regression has since been found to overfit data and therefore fail to replicate in future samples, we use generalized linear models here. We defined our models using the models reported by Lassek and Gaulin using the predictors that were not eliminated by the step wise regression.  
4.1	Pilot study

#4.1.1	Results

Descriptive statistics

Mating benefits model 1: Lifetime number of sexual partners
Mating benefits model 2: Past year number of sexual partners 
Mating benefits model 3: Age at first sexual intercourse
Immune costs model

Mating benefits model 1: Lifetime number of sexual partners
Using exploratory data, we added additional predictors to the model of lifetime sexual partners. We found that when age at first sexual intercourse was added as a predictor, strength was no longer a significant predictor. 
Mating benefits model 2: Past year number of sexual partners

Mating benefits model 3: Age at first sexual intercourse
Immune costs model

```{r summarytable}
d_G[d_G$age>=18 & d_G$age<=60,]  %>% select("sex", "age", "race", "strength", "bmi", "partnered", "edu", "height", "weight", "whitebloodcell", "testosterone") %>%
  tbl_summary(by = sex,
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} / {N} ({p}%)"),
              digits = all_continuous() ~ 2,
        #      type = list(where(is.numeric) ~ "continuous"),
             # type = list(partnered ~ "categorical"),
              type = all_dichotomous() ~ "categorical",
              label  = list(
                age = "Age (Years)",
                race = "Race and Ethnicity",
                strength = "Combined Grip Strength",
                bmi = "BMI (kg/m^2)",
                partnered = "Partnered Status",
                edu = "Education", 
                height = "Height (cm)", 
                weight = "Weight (kg)", 
                whitebloodcell = "WBCC (1000 cells/µL)", 
                testosterone = "Testosterone (ng/dL)",
                age_first_sex = "Age at first sex (Years)", 
                sex_partners = "Lifetime number of sexual partners", 
                sex_partners_year = "Past year number of sex partners"), 
              missing_text = "Missing"
  ) %>% add_p() %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>% 
    modify_caption("**Table 1. Participant Characteristics**")

# designsG$d.design.adults %>%
#   tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi, 
#                              whitebloodcell, strength, testosterone, partnered, 
#                              height, weight, special_equipment)) 
```

```{r table2_exact}

mm1 <- tbl_regression(mm1_exact, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm2 <- tbl_regression(mm2_exact, 
                      label = list(
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm3 <- tbl_regression(mm3_exact, 
                 label = list(
    age_centered = "Age",
    sex = "Sex",
    strength_centered = "Grip strength",
    partnered = " Currently partnered",
    edu = "Education",
    total_work_MET = "Occupational exertion"
  )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

  tbl_merge(
    tbls = list(mm1, mm2, mm3),
     tab_spanner = c("Lifetime number of sexual partners",
                     "Past year number of sexual partners",
                     "Age at first sexual intercourse")) %>% 
  as_gt() %>%
  gt::tab_header(title = "Table 2. Exact Models of Mating Success",
                 subtitle = "Standardized regression coefficents from generalized
                 linear models for three indices of mating success for men and women ages
                 18-59. Models derived from Lassek and Gaulin 2009")

```

```{r table2_mod}

mm1_m <- tbl_regression(mm1_mod, 
                        label = list(
                          age_centered = "Age",
                          sex = "Sex",
                          strength_centered = "Grip strength",
                          partnered = "Partnered",
                          median_salary_current = "Median Salary",
                          age_first_sex = "Age at first sex"
                        ))  %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm2_m <- tbl_regression(mm2_mod, 
                        label = list(
                          sex = "Sex",
                          strength_centered = "Grip strength",
                          partnered = "Partnered",
                          bmi_centered = "BMI"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm3_m <- tbl_regression(mm3_mod, 
                        label = list(
                          age_centered = "Age",
                          edu = "Education",
                          total_work_MET = "Occupational exertion",
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

  tbl_merge(
    tbls = list(mm1_m, mm2_m, mm3_m),
    tab_spanner = c("Lifetime number of sexual partners",
                    "Past year number of sexual partners",
                    "Age at first sexual intercourse")) %>% 
   as_gt() %>%
  gt::tab_header(title = "Table 3. Modified Models of Mating Success",
                 subtitle = "Standardized regression coefficents from generalized
                 linear models for three indicies of mating success. Models derived
                 from exploratory data")

```

```{r table2_vaginalsex}

mm1_v <- tbl_regression(m1_vaginal_sex_partners, 
                        label = list(
                          age_centered = "Age",
                          edu = "Education",
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm2_v <- tbl_regression(m2_vaginal_pastyear, 
                        label = list(
                          age_centered = "Age",
                          edu = "Education",
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

  tbl_merge(
    tbls = list(mm1_v, mm2_v),
    tab_spanner = c("Lifetime number of vaginal sex partners",
                    "Past year number of vaginal sex partners")) %>% 
   as_gt() %>%
  gt::tab_header(title = "Table 4. Models of Mating Success (Vaginal Sex)",
                 subtitle = "Standardized regression coefficents from generalized
                 linear models for lifetime and past year vaginal sex partners")

```

```{r immunemodel}

m_wbc <- tbl_regression(m_wbcc_exact, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

m_lymph <- tbl_regression(m_l, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
m_mono <- tbl_regression(m_m, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

m_neutro <- m_wbc <- tbl_regression(m_n, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

m_eos <- tbl_regression(m_e, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

m_baso <- tbl_regression(m_b, 
                        label = list(
                          bmi = "BMI", 
                          sex = "Sex",
                          strength_centered = "Grip strength"
                        )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

tbl_merge(
    tbls = list(m_wbc, m_lymph, m_mono, m_neutro, m_eos, m_baso),
    tab_spanner = c("White blood cell count",
                    "Lymphocyte count", 
                    "Monocyte count", 
                    "Neutrophil count", 
                    "Eosinophil count", 
                    "Basophil count")) %>% 
   as_gt() %>%
  gt::tab_header(title = "Table 4. Models of Immune Function",
                 subtitle = "Regression coefficents from generalized
                 linear models of white blood cell count, and cell type counts")

```

### 2024-03-23 revisions.qmd
Updated custom crossref YAML

--- 2023-07-14 ms_draft07.06.Rmd
@@
 bibliography: regrep.bib
+crossref:
+  custom:
+    - kind: float
+      reference-prefix: Figure S
+      key: sfig
+      space-before-numbering: false
 ---

### 2024-03-23 revisions.qmd
Update estimate_fun arg

@@
   add_difference(test = everything() ~ "smd",
-                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
+                 estimate_fun = list(all_continuous() ~ purrr::partial(style_ratio, digits = 2))) %>%
   modify_column_hide(columns = ci) %>%

### 2024-03-25 revisions.qmd
add dietary protein model, add depression to immune model, add limitations section and detail to discussion.

--- 2023-11-09 revisions.qmd
@@
 #### Hormone control model
-There is evidence that testosterone is associated with higher numbers of partners, but mixed evidence that it is related to sexual desire (van Anders, 2013).
-Likewise, evidence of a relationship between circulating testosterone and handgrip strength is also mixed (Nowak-Kornicka et al., 2020).
+There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women (Pollet et al., 2011; van Anders et al., 2007), but mixed evidence that it is related to sexual desire (van Anders, 2013).
+For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see van Anders, 2013.
+Likewise, although some studies have found a positive correlation between circulating testosterone and handgrip strength (Chiu et al., 2020; Gurlek et al., 2018) other studies have found no association (Gettler et al., 2010; Ribeiro et al., 2016).
+There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance (Guilherme et al., 2022).
 The hormone control model included serum total testosterone (ng/dL).
@@
                 avgcalories, 
+                avgprotein, 
                 foodinsecurity_adult, 
@@
                 avgcalories, 
+                avgprotein,
                 foodinsecurity_adult, 
@@
       avgcalories, 
+      avgprotein,
       foodinsecurity_adult, 
@@
       avgcalories = "Dietary energy intake (kcals)", 
+      avgprotein = "Dietary protein intake (grams)",
       foodinsecurity_adult = "Food Insecurity Rating (1-4)", 
@@
 
-#### 2.2.4 Model of energy intake
+#### 2.2.4 Models of energy and protein intake
 
-Finally, we tested whether strength was related to increased energy intake (Figure \@ref(fig:dietary)).
-We found that strength was a significant, positive predictor of energy intake controlling for sex.
+Finally, we tested whether strength was related to increased energy (Figure \@ref(fig:dietary)) or protein intake (Figure \@ref(fig:protein)).
+We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
 In the expanded control model we found the same pattern while controlling for additional variables.
@@
 Figure4 
+```
+
+```{r protein}
+Figure5 <- forestplot(
+  m_protein,
+  m_protein_alt,
+  intercept = F,
+  facet = F,
+  dodgewidth = .4,
+  varnames = vnames, 
+  modelnames = mmnames) +
+  theme_minimal(17) +
+  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
+
+ggsave("Figures/Figure5.pdf", Figure4, width = 14, height = 7)
+
+Figure5 
 ```
@@
 
+We now argue that these results emphasize the importance of biparental care in human reproduction (Gettler et al., 2020; Bribiescas et al., 2012).
+Across evolutionary time, fitness for both men and women was heavily dependent on obtaining successful long term partners, thereby reaping the efficiency benefits of a sexual division of labor (Quinlan, 2008; Quinlan & Quinlan, 2007).
+More specifically, humans require provisioning assistance for an extended period (Insert figure?) (Davison & Gurven, 2022; Davison & Gurven, 2021) and paternal care is an important predictor of child outcomes (Starkweather et al., 2021; Winking et al., 2011).
+Thus, our pilot study refines the sexual selection hypothesis for the relationship between strength and mating success to emphasize that mating success is best operationalized as long-term mating, rather than numbers of sexual partners.
+Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity and therefore the ability of stronger men to provision offspring through big-game hunting (Kaplan et al., 2000).
+
+### Limitations
+Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is greater for males than females.
+Our aim was not to test all possible byproduct hypotheses, nor is it possible to do so.
+Therefore, we are unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS).
+Likewise, NHANES does not collect data on sociosexuality.
+Likewise, as described in our measures, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior for both men and women.
+Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).
+
 ## 3. Confirmatory Analysis [not yet conducted]
@@
 
-```{r, vaginalsex}
+```{r vaginalsex}
 
@@
 
+```{r corrmatf}
+corr_f <- corrplot(cor_mat_cors_f, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
+                   title = "Correlation matrix of coninuous study variables, Females 18-60")
+
+```
+
+```{r corrmatm}
+
+corr_m <- corrplot(cor_mat_cors_m, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
+                   title = "Correlation matrix of coninuous study variables, Males 18-60")
+
+```
+
+```{r corrmatcom}
+
+corr_com <- corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75, 
+title = "Correlation matrix of coninuous study variables, Males and Females 18-60")
+```
+
 #### References

### 2024-03-25 draft_analysis.R

@@ -422,7 +422,8 @@ mwbc_alt <- svyglm(whitebloodcell ~
                  testosterone_sex_centered * sex +
                 foodinsecurity_adult  +
                   avgcalories_centered +
-                   tot_MET_centered,
+                   tot_MET_centered +
+                  depression,
                family= quasipoisson(),
                design=designsG$d.design.adults)
 
@@ -457,7 +458,37 @@ m_energy_alt <- svyglm( #use this
 summary(m_energy_alt, df.resid = Inf)
 
 
+# Protein -----------------------------------------------------------------
+
+m_protein <- svyglm( #use this
+  avgprotein ~
+    age_centered +
+    tot_MET_centered  +
+    strength_centered +
+    bmi_centered  +
+    sex,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_protein, df.resid = Inf)
+
+m_protein_alt <- svyglm( #use this
+  avgprotein ~
+    age_centered +
+    tot_MET_centered +
+    strength_centered +
+    sex +
+    bmi_centered +
+    whitebloodcell_centered +
+    foodinsecurity_adult,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_protein_alt, df.resid = Inf)
+
+
 # coef plots --------------------------------------------------------------
+
 
 mnames <- c(
  # "Anthropometric", baseline model has bmi
@@ -656,3 +687,89 @@ vnames <- c(
 #   theme(legend.position = "none")
 #
 
+# SI ----------------------------------------------------------------------
+library(jtools)
+library(qgraph)
+library(corrplot)
+
+# cordat <- d_G[, c("age", "sex1", "strength", "bmi")]
+# cordat <- cordat[cordat$age>=18 & cordat$age<60,]
+# cormat <- round(cor(cordat, use = "complete.obs"),2)
+
+
+cor_mat_f <- svycor(
+  ~ sex_partners +
+    sex_partners_year +
+    age_first_sex +
+    age +
+    strength +
+    bmi +
+    edu +
+    hemoglobin +
+    whitebloodcell +
+    depression +
+    chronic_disease_score +
+    physical_disease_count +
+    testosterone +
+    vigorous_rec +
+    moderate_rec +
+    vigorous_work +
+    moderate_work,
[…]

### 2024-03-25 revisions-exploratory.R

@@ -496,7 +496,17 @@ m_partnered_1 <-
       strength_centered * sex +
       bmi_centered * sex,
     family = quasibinomial(),
-    design = designsG$d.design.adult.heterosexual
+    design = designsG$d.design.adults
+  )
+
+m_partnered_2 <-
+  svyglm(
+    (maritalstatus == 6) ~
+      age_centered * sex +
+      strength_centered * sex +
+      bmi_centered * sex,
+    family = quasibinomial(),
+    design = designsG$d.design.adults
   )
 
 
@@ -1240,3 +1250,31 @@ m_pastyear_d <- svyglm(
   family = quasipoisson(),
   design = w2
 )
+
+# dietary protein ---------------------------------------------------------
+
+m_protein <- svyglm( #use this
+  avgprotein ~
+    age_centered +
+    tot_MET_centered  +
+    strength_centered +
+    bmi_centered  +
+    sex,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_protein, df.resid = Inf)
+
+m_protein_alt <- svyglm( #use this
+  avgprotein ~
+    age_centered +
+    tot_MET_centered +
+    strength_centered +
+    sex +
+    bmi_centered +
+    whitebloodcell_centered +
+    foodinsecurity_adult,
+  family = gaussian(),
+  design = designsG$d.design.dietary.adults
+)
+summary(m_protein_alt, df.resid = Inf)

### 2024-03-25 simulate.R
Added furrr code

@@ -3,6 +3,9 @@ library(tidyverse)
 library(nhanesGH)
 library(survey)
 library(boot)
+library(furrr)
+
+plan("multisession")
 
 d_adult <- d_G[d_G$age >= 18 & d_G$age < 60,]
 
@@ -87,7 +90,7 @@ year_params <- list(
   theta_year = 2.5
 )
 
-pvalues_year <- map_df(1:1000, ~getstats_year(year_params))
+pvalues_year <- future_map_dfr(1:1000, ~getstats_year(year_params))
 sum(pvalues_year$'strength_centered' < 0.05)/1000
 sum(pvalues_year$'sex:strength_centered' < 0.05)/1000
 
@@ -105,7 +108,7 @@ year_params <- list(
   theta_year = 2.5
 )
 
-pvalues_year <- map_df(1:1000, ~getstats_year(year_params))
+pvalues_year <- future_map_dfr(1:1000, ~getstats_year(year_params))
 sum(pvalues_year$'strength_centered' < 0.05)/1000
 sum(pvalues_year$'sex:strength_centered' < 0.05)/1000
 
@@ -148,7 +151,7 @@ life_params <- list(
   theta_life = 18
 )
 
-pvalues_life <- map_df(1:1000, ~getstats_life(life_params))
+pvalues_life <- future_map_dfr(1:1000, ~getstats_life(life_params))
 sum(pvalues_life$'strength_centered' < 0.05)/1000
 sum(pvalues_life$'sex:strength_centered' < 0.05)/1000
 
@@ -175,7 +178,7 @@ partnered_params <- list(
   b_sex_strength = -0.75, # We set this to half or full male value
   b_sex_age = -0.68
 )
-pvalues_partnered <- map_df(1:1000, ~getstats_partnered(partnered_params))
+pvalues_partnered <- future_map_dfr(1:1000, ~getstats_partnered(partnered_params))
 sum(pvalues_partnered$'strength_centered' < 0.05)/1000
 sum(pvalues_partnered$'sex:strength_centered' < 0.05)/1000
 

### 2024-03-25 revisions.qmd

@@
   custom:
-    - kind: float # crossreferenceable elements with captions are `float`
-      prefix: Supplementary Figure # used as the prefix for reference in output ("In Figure 1, ..")
-      name: Supplementary Figure # used as the prefix for captions ("Figure 1: ..")
-      ref-type: sfig # used as the prefix for the reference identifier ("In @fig-1, ...")
+    - kind: float
+      reference-prefix: Figure S
+      key: sfig
+      space-before-numbering: false
 ---

### 2024-03-25 paper.v2.qmd
Renamed revisions.qmd to paper.v2.qmd

--- 2024-03-25 revisions.qmd
@@
 title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
-author: "C. B. Smith and E. H. Hagen"
-date: "`r Sys.Date()`"
+author: "Caroline B. Smith and Edward H. Hagen"
+date: today
 output:

### 2024-03-25 paper.v2.qmd
Initial conversion to quarto elsevier format

--- 2024-03-25 revisions.qmd
@@
 title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
-author: "C. B. Smith and E. H. Hagen"
-date: "`r Sys.Date()`"
-output:
- bookdown::html_document2:
-   number_sections: false
- bookdown::pdf_document2:
-   number_sections: false
-   latex_engine: xelatex
- bookdown::word_document2:
-   number_sections: false
+author: 
+  - name: Caroline B. Smith
+    affiliation:
+     - id: wsu
+       department: Department of Anthropology
+       name: Washington State University
+  - name: Edward H. Hagen
+    affiliation:
+     - ref: wsu
+date: today
+format:
+  elsevier-html
 csl: apa.csl
@@
 library(patchwork)
-source('draft_analysis.R')
+source('analysis.v2.R')
 ```
-EHB Registered Report
 
-## Abstract
+# Abstract
 
@@
 We will also test if there is a positive relationship between grip strength and dietary energy intake, and a negative relationship between grip strength and native immune function.
-This study replicates an earlier study of US males (Lassek & Gaulin, 2009) and additionally tests these relationships in females.
+This study replicates an earlier study of US males [@lassek2009] and additionally tests these relationships in females.
 
-### 1.1 Sexual dimorphism in modern humans:
+## Sexual dimorphism in modern humans:
 
 The mechanisms that have shaped human sexual dimorphism and whether they have primarily operated on male or female traits remain debated.
-Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs (Plavcan, 2001).
-Unlike most animals, mammalian sexual size dimorphism tends to be male biased (Andersson, 1994).
-In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos (Bukowski et al., 2007).
-Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass (Kirchengast, 2010).
-At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass (Lassek & Gaulin, 2022), which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees (Plavcan, 2012; Smith & Jungers, 1997).
-Men also have larger and stronger bones (Wells, 2007).
+Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs [@plavcan2001].
+Unlike most animals, mammalian sexual size dimorphism tends to be male biased [@andersson1994].
+In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos [@Bukowski].
+Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass [@kirchengast2010].
+At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass [@lassek2022], which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees [@plavcan2012; @smith1997].
+Men also have larger and stronger bones [@wells2007].
 Other traits that are sexually dimorphic in humans include digit ratio, voice pitch, facial features, body and facial hair growth, and canine length.
@@
 While humans are only moderately dimorphic in terms of overall body mass, this is not the case for fat and muscle allocation, which are highly dimorphic (Puts, 2010) due to women’s copious storage of fat and male investment in muscle mass.
-Men have between 30% and 42% more fat-free mass (Lassek & Gaulin, 2022), 61% more overall muscle mass, and 78% more muscle mass in the upper arms.
+Men have between 30% and 42% more fat-free mass [@lassek2022], 61% more overall muscle mass, and 78% more muscle mass in the upper arms.
 This concentrated muscle dimorphism in the arms and back translates to greater upper body strength in men than women.
@@
 
-### 1.2  Sexual dimorphism hypotheses
+## Sexual dimorphism hypotheses
 
@@
 Modern definitions distinguish pre- and post-copulatory competition (Andersson and Simmons, 2006).
-A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success, either through attracting mates or winning access to mates and fertilizations through contests, or both (Andersson, 1994).
+A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success, either through attracting mates or winning access to mates and fertilizations through contests, or both [@andersson1994].
 Two hypotheses for the evolution of human sexual size dimorphism and body composition dimorphism dominate the literature: 1) intrasexual selection through male contests and 2) intersexual selection through female mate choice.
@@
 
-#### 1.2.1 	Male contest competition
+### Male contest competition
 
 The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past (Puts, 2010), and that instead male fitness was determined by physical contests with other males for access to mates.
-Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success (Puts, 2016, Hill et al., 2017; Plavcan, 2012).
+Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [Puts, 2016, Hill et al., 2017; @plavcan2012].
 Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression (Archer, 2009).
@@
 In extant nonhuman primate species, sexual dimorphism in body weight and canine size are strongly associated with the degree of intrasexual competition, a relationship which, if used to infer the behavior of extinct australopithecines, suggests a high degree of intrasexual competition in ancestral species (Plavcan & van Schaik, 1997).
-Evidence of sexual size dimorphism in Homo is less certain, but in general shows reduced size dimorphism compared to early species (Plavcan, 2012).
+Evidence of sexual size dimorphism in Homo is less certain, but in general shows reduced size dimorphism compared to early species [@plavcan2012].
 
-#### 1.2.2 	Female choice for masculine traits
+### Female choice for masculine traits
 
@@
 
-### 1.3 Strength and reproductive success
+## Strength and reproductive success
 
@@
 
-Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (operationalized as C-reactive protein and white blood cell count) (Lassek & Gaulin, 2009), consistent with a tradeoff in mating success, immunity and energy costs.
+Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (operationalized as C-reactive protein and white blood cell count) [@lassek2009], consistent with a tradeoff in mating success, immunity and energy costs.
 However, these relationships were not tested in women, for whom both selection hypotheses predict no association.
 
-### 1.4 Experimental aims and hypotheses
+## Experimental aims and hypotheses
 
-The aim of this study is to replicate the costs and benefits of formidability reported by Lassek and Gaulin (2009) using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
+The aim of this study is to replicate the costs and benefits of formidability reported by @@lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
 It also aims to investigate if these costs and benefits are also experienced by women.
@@
 
-## 2. Pilot study
+# Pilot study
 
@@
 
-### 2.1 Methods
+## Methods
 
@@
 
-#### 2.1.2 Key predictor variables: 
-Our main predictor variable is formidability, which Lassek and Gaulin operationalized as fat-free mass and limb muscle volume.
+### Key predictor variables
+
+Our main predictor variable is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
 Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
@@
 
-#### 2.1.1 Outcome variables: 
+### Outcome variables
+
 In industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior.
 Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).
-Our main outcome was mating success, which we operationalized in three ways following Lassek and Gaulin (2009): 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
+Our main outcome was mating success, which we operationalized in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
 female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
@@
 
-Unlike Lassek and Gaulin (2009), we also operationalized mating success in a fourth outcome measure as partnered status, as a proxy for longterm mating strategy since number of partners emphasizes success through short term mating strategies.
+Unlike @lassek2009, we also operationalized mating success in a fourth outcome measure as partnered status, as a proxy for longterm mating strategy since number of partners emphasizes success through short term mating strategies.
 Partnered status included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated.
@@
 Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL).
-In the original paper, Lassek and Gaulin (2009) also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
+In the original paper, @lassek2009 also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
 
@@
 
-### Models:
-We attempted to replicate the regression models in Lassek and Gaulin as closely as possible.
+## Models
+
+We attempted to replicate the regression models in @lassek2009 as closely as possible.
 There were some unavoidable differences, however.
 First, our study aimed to investigate the role of sex; thus we include sex as a main effect as well as in interactions with key predictor variables.
-Second, some variables included in Lassek and Gaulin were not available in our data.
+Second, some variables included in @lassek2009 were not available in our data.
 Third, we included additional control variables as noted below.
 Finally, we treated our pilot study as an exploratory study in which we fit numerous models not reported here.
-The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in Lassek and Gaulin (2009) based on the theories we have described above and given our constraints.
+The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we have described above and given our constraints.
 
-#### Control models: 
-Our modelling strategy takes inspiration from Lassek and Gaulin (2009).
-However, Lassek and Gaulin (2009) utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples (Smith, 2018).
+### Control models
+
+Our modelling strategy takes inspiration from @lassek2009.
+However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples (Smith, 2018).
 Instead, we will fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses.
-Control variables included a much wider range of theoretically motivated potential confounds than used by Lassek and Gaulin (2009), organized into themes.
+Control variables included a much wider range of theoretically motivated potential confounds than used by @lassek2009, organized into themes.
 We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated with strength and sexual behavior.
 
-#### Anthropometric control model
-We first derived a simple model based on that reported in Table 2 in Lassek and Gaulin (2009) for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
+### Anthropometric control model
+
+We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
 This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
@@
 
-#### Socioeconomic control model
+### Socioeconomic control model
+
 The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength (Johnson & Wilson, 2019) and sexual behavior (Fenton et al., 2005).
@@
 
-#### Health control model
+### Health control model
+
 An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability (Soysal et al., 2021).
@@
 
-#### Hormone control model
+### Hormone control model
+
 There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women (Pollet et al., 2011; van Anders et al., 2007), but mixed evidence that it is related to sexual desire (van Anders, 2013).
@@
 
-#### Physical activity control model
+### Physical activity control model
+
 Physical activity is related to sexual behavior and function in adults (Morris et al., 2022).
@@
 
-#### Costs of strength
-Following Lassek and Gaulin (2009) we also developed models of native immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day).
-For each outcome we developed a model as similar as possible to those reported in Lassek and Gaulin (2009), except that we always include sex and its interaction with other predictors where we have theoretical reasons to expect sex differences in the effect of the predictor on the outcome.
+### Costs of strength
+
+Following @lassek2009 we also developed models of native immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day).
+For each outcome we developed a model as similar as possible to those reported in @lassek2009, except that we always include sex and its interaction with other predictors where we have theoretical reasons to expect sex differences in the effect of the predictor on the outcome.
 We then also specified an alternative model for each outcome with additional control variables that could confound strength and energy or immune investment.
@@
 
-#### Native Immune Function
+### Native Immune Function
 
-Following the significant predictors reported in Lassek and Gaulin (2009, Table 4), we first specified a replication model with age, sex, strength, and BMI.
+Following the significant predictors reported in @lassek2009, Table 4, we first specified a replication model with age, sex, strength, and BMI.
 We also included an age:sex interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
@@
 
-#### Dietary Energy Intake
-Based on Lassek and Gaulin (2009, Table 3) we developed a replication model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
+### Dietary Energy Intake
+
+Based on @lassek2009, Table 3, we developed a replication model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
 We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.
@@
 
-### Analysis 
+## Analysis
+
 All analyses were completed in `r R.version.string`, using the survey package (version 4.1-1; Lumley, 2021) in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
@@
 
-### 2.2 Pilot Study Results
+## Pilot Study Results
 
-#### 2.2.1 Descriptive statistics
+### Descriptive statistics
 
@@
 
-#### 2.2.2 Anthropometric models of mating success
+### Anthropometric models of mating success
 
@@
 
-#### 2.2.3 Models of mating success with socioeconomic, health, hormone, and physical activity controls
+### Models of mating success with socioeconomic, health, hormone, and physical activity controls
+
 We then used generalized linear regression to evaluate the impact of a host of other potential confounds on these results.
@@
 
-#### 2.2.4 Immune costs of strength
+### Immune costs of strength
 
@@
 
-#### 2.2.4 Models of energy and protein intake
+### Models of energy and protein intake
 
@@
 
-### 2.3 Pilot Results Discussion
+## Pilot Results Discussion
 
-In this pilot study using grip strength as a proxy, we replicated findings from Lassek and Gaulin (2009) that muscularity is associated with increased mating success, as well as with immune and energy costs, even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
+In this pilot study using grip strength as a proxy, we replicated findings from @lassek2009 that muscularity is associated with increased mating success, as well as with immune and energy costs, even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
 The only exception was that strength was not a positive predictor of past year partners in the hormone control model; the coefficient, however, remained similar in magnitude to the significant strength coefficients in other models of past year partners.
@@
 
-### Limitations
+## Limitations
+
 Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is greater for males than females.
@@
 
-## 3. Confirmatory Analysis [not yet conducted]
+# Confirmatory Analysis [not yet conducted]
 
@@
 
-### 3.1 Method
+## Method
 
@@
 
-### Supplementary Information
+## Supplementary Information
 
@@
 
-#### References
+# References

### 2024-03-25 paper.v2.qmd
changes from switching to visual mode

@@
 
-The aim of this study is to replicate the costs and benefits of formidability reported by @@lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
+The aim of this study is to replicate the costs and benefits of formidability reported by \@@lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
 It also aims to investigate if these costs and benefits are also experienced by women.
@@
 We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
-This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
+This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m\^2, since body size could impact either strength or mating outcomes.
 We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.
@@
 
-# Confirmatory Analysis [not yet conducted]
+# Confirmatory Analysis \[not yet conducted\]
 

### 2024-03-30 paper.v2.qmd
Committing before replacing with text from google doc

--- 2024-03-25 paper.v2.qmd
@@
 format:
-  elsevier-html
+  html:
+    embed-resources: true
+execute:
+  echo: false
+  warning: false
 csl: apa.csl
@@
       space-before-numbering: false
+    - kind: float
+      reference-prefix: Table S
+      key: stbl
+      space-before-numbering: false
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 library(ggplot2)
@@
 library(gtsummary)
-library(corrplot)
+library(ggcorrplot)
 library(patchwork)
+library(modelsummary)
 source('analysis.v2.R')
@@
 We will also test if there is a positive relationship between grip strength and dietary energy intake, and a negative relationship between grip strength and native immune function.
-This study replicates an earlier study of US males [@lassek2009] and additionally tests these relationships in females.
+This study replicates an earlier study of US males [@lassek2009] with numerous additional control variables, and it additionally tests these relationships in females.
 
@@
 Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs [@plavcan2001].
-Unlike most animals, mammalian sexual size dimorphism tends to be male biased [@andersson1994].
-In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos [@Bukowski].
+Unlike most animals, mammalian sexual size dimorphism, when it exists, tends to be male biased [@andersson1994; @Tombak_2024].
+In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos [@bukowski2007].
 Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass [@kirchengast2010].
@@
 
-While humans are only moderately dimorphic in terms of overall body mass, this is not the case for fat and muscle allocation, which are highly dimorphic (Puts, 2010) due to women’s copious storage of fat and male investment in muscle mass.
+While humans are only moderately dimorphic in terms of overall body mass, this is not the case for fat and muscle allocation, which are highly dimorphic [@puts2010] due to women’s copious storage of fat and male investment in muscle mass.
 Men have between 30% and 42% more fat-free mass [@lassek2022], 61% more overall muscle mass, and 78% more muscle mass in the upper arms.
 This concentrated muscle dimorphism in the arms and back translates to greater upper body strength in men than women.
-Studies of sex differences in strength show that men are stronger on average than women on all tests of muscle strength, but especially for tests of upper-body muscle strength, where female upper-body strength is 50-60% of male upper-body strength while lower-body strength is 60-70% of male values, and trunk strength is 60% of male values (Nuzzo, 2023).
-Muscle mass alone does not explain the sex difference in strength, as strength assessments are often greater in males than females even when pair-matched on muscle thickness (Kataoka et al., 2023), and likewise men have greater strength-to-body mass ratios than women (Nuzzo, 2023).
+Studies of sex differences in strength show that men are stronger on average than women on all tests of muscle strength, but especially for tests of upper-body muscle strength, where female upper-body strength is 50-60% of male upper-body strength while lower-body strength is 60-70% of male values, and trunk strength is 60% of male values [@nuzzo2023].
+Muscle mass alone does not explain the sex difference in strength, as strength assessments are often greater in males than females even when pair-matched on muscle thickness [@kataoka2023], and likewise men have greater strength-to-body mass ratios than women [@nuzzo2023].
 
@@
 These differences are greatest in the upper body.
-Finally, a greater proportion of male to female muscle is occupied by type II muscle fibers, which create greater force than type I (Nuzzo, 2023).
+Finally, a greater proportion of male to female muscle is occupied by type II muscle fibers, which create greater force than type I [@nuzzo2023].
 
@@
 
-Darwin defined sexual selection as selection that acts, not to fit individuals to their environment, but instead as to confer advantage over others in competition for mates (Darwin, 1871).
-Modern definitions distinguish pre- and post-copulatory competition (Andersson and Simmons, 2006).
-A trait is likely to have been shaped by sexual selection when it is sexually dimorphic, develops at sexual maturity, and affects mating success, either through attracting mates or winning access to mates and fertilizations through contests, or both [@andersson1994].
 Two hypotheses for the evolution of human sexual size dimorphism and body composition dimorphism dominate the literature: 1) intrasexual selection through male contests and 2) intersexual selection through female mate choice.
@@
 
-The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past (Puts, 2010), and that instead male fitness was determined by physical contests with other males for access to mates.
-Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [Puts, 2016, Hill et al., 2017; @plavcan2012].
-Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression (Archer, 2009).
-Particular emphasis is placed on sexual dimorphism in the upper-body due to the important role upper-body strength plays in both armed and unarmed fighting ability (Sell, et al., 2012) and where sex differences in strength and muscularity are greatest as discussed above.
+The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past [@puts2010], and that instead male fitness was determined by physical contests with other males for access to mates.
+Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [@puts2016; @hill2017; @plavcan2012].
+Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression [@archer2009].
+Particular emphasis is placed on sexual dimorphism in the upper-body due to the important role upper-body strength plays in both armed and unarmed fighting ability [@sell2012] and where sex differences in strength and muscularity are greatest as discussed above.
 
-In extant nonhuman primate species, sexual dimorphism in body weight and canine size are strongly associated with the degree of intrasexual competition, a relationship which, if used to infer the behavior of extinct australopithecines, suggests a high degree of intrasexual competition in ancestral species (Plavcan & van Schaik, 1997).
-Evidence of sexual size dimorphism in Homo is less certain, but in general shows reduced size dimorphism compared to early species [@plavcan2012].
+In extant nonhuman primate species, sexual dimorphism in body weight and canine size are strongly associated with the degree of intrasexual competition, a relationship which, if used to infer the behavior of extinct australopithecines, suggests a high degree of intrasexual competition in ancestral species [@plavcan1997].
+Evidence of sexual size dimorphism in *Homo* is less certain, but in general shows reduced size dimorphism compared to earlier hominin species [@plavcan2012].
 
@@
 
-In contrast to the intrasexual selection hypothesis, others argue that sexual dimorphism in humans is largely the result of female choice, not for more formidable or competitive mates, but for high quality males.
-The immunocompetence handicap hypothesis relies on evidence that testosterone is immunosuppressive, and therefore only males with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine secondary sexual traits (Folstad & Karter, 1992).
-Therefore, androgen-dependent traits could be costly signals of good genetic quality, in which case females could increase their long-term fitness by preferentially mating with males displaying these traits thus conferring heritable immunity to their offspring and securing investment from a healthy partner (Folstad & Karter, 1992).
-Similar arguments have been proposed which argue that masculine traits could be more general cues of quality and condition other than immunocompetence, or ability to bear other costs (Kokko et al., 2003).
+Although there is evidence in Western young adult populations that male mating success is mediated by male-rated dominance and not by female-rated attractiveness [@kordsmeyer2018; @hill2013], other researchers nevertheless argue that sexual dimorphism in humans is largely the result of female choice for high quality males [which was possibly heavily influenced by parents, in which case parental preferences would also have been important, @apostolou2007].
+The immunocompetence handicap hypothesis, for example, relies on evidence that testosterone is immunosuppressive, and therefore only males with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine secondary sexual traits [@folstad1992].
+Androgen-dependent traits could thus be costly signals of good genetic quality, with females increasing their fitness by preferentially mating with males displaying these traits, thereby conferring heritable immunity to their offspring and securing investment from a healthy partner [@folstad1992].
+Evidence that testosterone is immunosuppressive is mixed, however [@nowak2018], and although testosterone is associated with mating success, it does not seem to be through the mechanism of female choice for testosterone-derived face or body features since testosterone does not predict female rated masculinity or attractiveness [@peters2008].
 
-In humans, the immunocompetence hypothesis has primarily been leveraged to explain male facial masculinity and voice pitch but it and more general cost models have also been argued to explain apparent female preference for muscular male body types (Frederick & Haselton, 2007).
-Cues of upper body strength indeed account for most of the variation in female-rated male body attractiveness (Sell et al., 2017).
+Similar arguments have therefore been proposed that masculine traits could be more general cues of quality and condition, or an ability to bear other costs [@kokko2003; @frederick2007].
+Cues of upper body strength indeed account for most of the variation in female-rated male body attractiveness [@sell2017].
+Yet masculine trait expression is not consistently related to retrospective or prospective health [@boothroyd2013].
 
-A few problems arise for this hypothesis.
-First, evidence that testosterone is actually immunosuppressive is mixed (Nowak et al., 2018).
-Likewise, masculine trait expression is not consistently related to retrospective or prospective health (Boothroyd et al., 2013).
-Secondly, it is unclear to what extent women are attracted to testosterone-dependent traits, for instance facial masculinity.
-Although testosterone is associated with mating success, it does not seem to be through the mechanism of female choice for testosterone-derived face or body features since testosterone does not predict female rated masculinity or attractiveness (Peters et al., 2008).
+Females might be choosing males who could provide more resources, or men who could better protect them from attacks by other men or by predators, which in ancestral hunter-gatherer populations would have been physically stronger men [@apicella2014; @willems2017b].
+There is substantial cross-cultural evidence of a female-biased preference for mates who can provide resources [@walter2020; @buss2019].
+Importantly, studies of contemporary hunter-gatherers and game-theoretic models find that a mutually beneficial sexual division of labor within long-term pair bonds is critical to provisioning offspring who remain dependent on caregivers for up to 20 years or more [@davison2021; @davison2022; @alger2020; @kaplan2000].
+See @fig-provisioning.
+Typically, forager women engage in lower risk, lower return foraging of plants and small game compatible with the cognitively demanding task of childrearing [@hagen2019], and men engage in higher risk, higher return big game hunting [@kelly2013].
+This suggests that sex differences in body composition could be due, at least in part, to a sexual division of labor.
 
-Females might also choose males who could provide more resources, which in ancestral hunter-gatherer populations might have been physically stronger men.
-It could also have been the case that mate choice was heavily influenced by parents, in which case parental preferences would have been important (see Hill et al.
-2017 for a brief review of alternative hypotheses for sexual strength dimorphism).
+:::{#fig-provisioning}
+![](Figures/plot_productivity.png)
 
-More generally, multiple mechanisms of sexual selection could have played a role in the evolution of human sexual dimorphism.
-Nevertheless, there is evidence that mating success due to formidability is mediated by male-rated dominance and not by female-rated attractiveness (Kordsmeyer et al., 2018; Hill et al., 2013).
+Net energy production in chimpanzees and human foragers over the lifespan. Solid lines are females. Dashed line is Hadza males. Data from @davison2021 and @davison2022.
+:::
+
+More generally, multiple forms of sexual selection could have played a role in the evolution of human sexual dimorphism [see @hill2017 for a brief review of alternative hypotheses for sexual strength dimorphism].
 
@@
 
-In a meta-analysis, Lidborg et al (2022) found that while voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations.
-The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported for men (Gallup & Fink, 2018).
-Of the four studies which have tested the association between grip strength and sexual behavior, only two included both women and men (Gallup et al., 2007; Varella et al., 2014) and the other two only men (Shoup and Gallup, 2008; Sneade and Furnham, 2016).
-In all four, hand grip strength was positively related to mating success (including number of sex partners) for men, but there was either no relationship or a negative relationship for women.
-These four studies were in convenience samples.
+A meta-analysis found that whereas voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations [@lidborg2022].
+In four studies of convenience samples (two of which included women), hand grip strength was positively related to mating success (including number of sex partners) for men, but there was either no relationship or a negative relationship for women [@gallup2007; @varella2014; @shoup2008; @sneade2016]
 
+The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported for men [@gallup2018].
 There have been very few tests of the hypotheses that some female traits, such as breast size and waist-hip ratio (WHR), have evolved by sexual selection.
-In a meta-analysis, Lidborg and Boothroyd (2023) found weak evidence that more feminine digit ratios predicted higher fertility, insufficient evidence for voice pitch, and none regarding facial femininity, breast size or waist-hip ratio (WHR).
-At present there is mixed evidence that strength is related to reproductive success in women (Lidborg & Boothroyd, 2023).
-One study among the Himba showed that women with higher strength/muscularity had more living children and grandchildren (Atkinson et al., 2012), while a similar study among the Hadza showed no relationship between strength and reproductive outcomes (Smith et al., 2017).
-The dearth of research on female traits and reproductive outcomes leaves us unable to draw any firm conclusions about their selection pressures over time.
+A meta-analysis found weak evidence that more feminine digit ratios predicted higher fertility, insufficient evidence for voice pitch, and none regarding facial femininity, breast size or waist-hip ratio (WHR).
+At present there is mixed evidence that strength is related to reproductive success in women [@lidborg2022a].
+One study among the Himba showed that women with higher strength/muscularity had more living children and grandchildren [@atkinson2012], while a similar study among the Hadza showed no relationship between strength and reproductive outcomes [@smith2017a].
+The dearth of research on female traits and reproductive outcomes leaves us unable to draw any firm conclusions about selection pressures over time.
 
-Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function (operationalized as C-reactive protein and white blood cell count) [@lassek2009], consistent with a tradeoff in mating success, immunity and energy costs.
-However, these relationships were not tested in women, for whom both selection hypotheses predict no association.
+Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function, operationalized as C-reactive protein and white blood cell count [@lassek2009], consistent with a tradeoff in mating success, immunity, and energy costs.
+However, these relationships were not tested in women, for whom both sexual selection hypotheses predict no association.
 
@@
 
-The aim of this study is to replicate the costs and benefits of formidability reported by @@lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
-It also aims to investigate if these costs and benefits are also experienced by women.
-
-In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success, with a significant interaction with sex such that strength will be a stronger predictor of male than female mating success.
-We also predict that there will be a negative association between male formidabillity and immune function, but little or no association for females.
-Finally, we predict that formidability will be positively associated with dietary energy intake for both males and females.
+According to the sexual selection hypothesis, males evolved greater physical formidability than females because physical formidability increased male reproductive success more than females.
+The aim of this study is to replicate the costs and benefits of formidability reported by @lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
+It also aims to investigate if these costs and benefits are also experienced by women, and specifically if the effect of physical strength on mating success, a proxy for reproductive success, is greater for men than women, as the sexual selection hypothesis predicts.
 
@@
 The pilot study uses the 2011-2012 dataset, whereas the 2013-2014 dataset has been held out for confirmatory analysis pending in principle acceptance of this registered report.
-Grip strength, our key predictor variable, was only collected in these years.
-We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable (Hogrel, 2015).
+Grip strength, our proxy for physical formidability and our key predictor variable, was only collected in these years.
+We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable [@hogrel2015].
 For each model, participants will be included if they have complete data for each predictor and outcome variable.
-
-### Key predictor variables
-
-Our main predictor variable is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
-Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
-All regression models also included sex (male/female), age in years, and partnered status (not included in models where partnered status was the outcome variable).
-The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
-Likewise, we also interacted age and partnered status with sex.
-We interacted age with sex since menopause occurs within the age range of our participants.
-For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
-Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable (Gelman, 2008).
 
@@
 
+Our main outcome was mating success.
 In industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).
-Our main outcome was mating success, which we operationalized in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
+Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with male reproductive success under ancestral conditions [@perusse1993].
+We operationalized mating success in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
 female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
@@
 
-Unlike @lassek2009, we also operationalized mating success in a fourth outcome measure as partnered status, as a proxy for longterm mating strategy since number of partners emphasizes success through short term mating strategies.
-Partnered status included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated.
+Number of partners emphasizes success through short term mating strategies, but it is an ambiguous measure as it could also represent repeated rejection, or a tendency to exaggerate or understate the actual number.
+It also does not clearly index female reproductive success because multiple sexual partners are likely to have increased male fitness more than female fitness [@Lehtonen_2022].
+We therefore also operationalized mating success in a fourth outcome measure as partnered status as a proxy for a long-term mating strategy that, as discussed earlier, benefits both sexes.
+Partnered status included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated [@lassek2009, included marital status as a control variable but not as an outcome variable].
 
-Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/ µL).
-In the original paper, @lassek2009 also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
+Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/µL).
+@lassek2009 also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
 
-Finally, to assess costs of greater strength, we investigated the relationship between strength and dietary energy intake.
+Finally, to assess costs of greater strength, we investigated the relationship between strength and dietary energy and protein intake.
 Participants reported all food and beverages they consumed in the 24 hour period prior to their interview.
@@
 For the outcome measure in this model we used the average number of calories per day calculated for each participant across their two recall days.
+
+### Key predictor variables
+
+Our main predictor variable of mating success is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
+Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
+All regression models also included sex (male/female), age in years, and partnered status (not included in models where partnered status was the outcome variable).
+The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
+Likewise, we also interacted sex with age and partnered status.
+We interacted age with sex since menopause occurs within the age range of our participants.
+For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
+Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable [@gelman2008].
 
@@
 Finally, we treated our pilot study as an exploratory study in which we fit numerous models not reported here.
-The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we have described above and given our constraints.
+The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we described above, the results of our exploratory analyses, and our constraints.
 
@@
 Our modelling strategy takes inspiration from @lassek2009.
-However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples (Smith, 2018).
+However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples [@smith2018a].
 Instead, we will fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses.
@@
 We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
-This model also includes an anthropometric control, body mass index (BMI), calculated as kg/m^2, since body size could impact either strength or mating outcomes.
+This model also includes an anthropometric control, body mass index (BMI), calculated as $kg/m^2$, since body size could impact either strength or mating outcomes.
 We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.
@@
 
-The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength (Johnson & Wilson, 2019) and sexual behavior (Fenton et al., 2005).
-Likewise, there is also variation in sexual behavior by education (Chandra et al., 2013), likewise there is variation in physical activity by education (He & Baker, 2005).
+The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength [@johnson2019] and sexual behavior [@fenton2005].
+Likewise, there is also variation in sexual behavior by education [@chandra2013], and there is variation in physical activity by education [@he2005].
 
@@
 
-An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability (Soysal et al., 2021).
-Furthermore, health status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women (Lindau & Gavrilova, 2010).
-Another study of men aged 45-59 found that lower sexual activity was associated with increased mortality (Smith et al., 1997).
+An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability [@soysal2021].
+Furthermore, healthy status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women [@lindau2010].
+Another study of men aged 45-59 found that lower sexual activity was associated with increased mortality [@daveysmith1997].
 The health model, therefore, included a number of variables related to health.
-White blood cell count (1000 cells/ µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression, which is negatively related to strength (Smith et al., 2022), was measured using the Patient Health Questionnaire-9 (PHQ-9; Kroenke et al., 2001), a validated nine-item screening instrument.
+White blood cell count (1000 cells/µL) and hemoglobin (g/dL) were included to control for acute infection.
+Depression, which is negatively related to strength [@smith2022], was measured using the Patient Health Questionnaire-9 [PHQ-9, @kroenke2001], a validated nine-item screening instrument.
 Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
@@
 
-There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women (Pollet et al., 2011; van Anders et al., 2007), but mixed evidence that it is related to sexual desire (van Anders, 2013).
-For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see van Anders, 2013.
-Likewise, although some studies have found a positive correlation between circulating testosterone and handgrip strength (Chiu et al., 2020; Gurlek et al., 2018) other studies have found no association (Gettler et al., 2010; Ribeiro et al., 2016).
-There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance (Guilherme et al., 2022).
+There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women [@pollet2011; @vananders2007], but mixed evidence that it is related to sexual desire [@vananders2013].
+For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see @vananders2013.
+Likewise, although some have found a positive correlation between circulating testosterone and handgrip strength [@Chiu_2019], other studies have found no association [@Gettler_2010; @Ribeiro_2016].
+There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance [@Guilherme_2021].
 The hormone control model included serum total testosterone (ng/dL).
-Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex.
+Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex, one of our key predictor variables.
 We therefore computed a sex-specific testosterone value as follows: for each sex we centered testosterone at the sex-specific mean and divided by two times the sex-specific standard deviation.
+Thus, for women, low or high values were relative to other women, and for men, low or high values were relative to other men.
 This differed from other centered and standardized variables which we centered and standardized across all male and female values.
@@
 
-Physical activity is related to sexual behavior and function in adults (Morris et al., 2022).
+Physical activity is related to sexual behavior and function in adults [@morris2022].
 Therefore, the physical activity control model included four dichotomous variables of vigorous and moderate work and recreational activity coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
@@
 We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.
+We then developed an expanded control model that also controlled for WBCC and food insecurity in addition to the variables in the model above, since either could influence energy intake and strength.
 
-We then developed an expanded control model that also controlled for WBCC and food insecurity in addition to the variables in the model above, since either could influence energy intake and strength.
+### A tradeoff between short- and long-term mating
+
+Exploratory analysis revealed a negative association between the lifetime number of sexual partners and the probability of being in a committed relationship (partnered status), which we interpreted as a possible tradeoff between short- and long-term mating strategies.
+In models of partnered status (one of our mating success outcome variables), we therefore included the lifetime number of sexual partners to assess if this negative association persisted after including the sets of control variables described above.
 
@@
 
-All analyses were completed in `r R.version.string`, using the survey package (version 4.1-1; Lumley, 2021) in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-For models of lifetime and past year numbers of partners we modeled the outcomes using a poisson distribution, since these variables are count data.
-Models of age at first sexual intercourse and white blood cell count utilized a normal distribution.
-Analyses include adults ages 18-60 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
+All analyses were completed in `{r} R.version.string`, using the survey package [version 4.1-1, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
+We modeled lifetime and past year numbers of partners using a Poisson distribution, since these variables are count data.
+We modeled of age at first sexual intercourse and white blood cell count with a normal distribution.
+We modeled partnered status using logistic regression.
+Analyses included adults ages 18-60 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
 
@@
 
+```{r}
+# Lifetime
+female_median <- 
+  svyquantile(~sex_partners, designsG$d.design.adult.female, 0.5, na.rm = T)$sex_partners[1,1]
+male_median <- 
+  svyquantile(~sex_partners, designsG$d.design.adult.male, 0.5, na.rm = T)$sex_partners[1,1]
+high_partners <- sum(d_G$sex_partners >= 100, na.rm=T)
+
+# Last year
+female_median_year <- 
+  svyquantile(~sex_partners_year, designsG$d.design.adult.female, 0.5, na.rm = T)$sex_partners_year[1,1]
+male_median_year <- 
+  svyquantile(~sex_partners_year, designsG$d.design.adult.male, 0.5, na.rm = T)$sex_partners_year[1,1]
+high_partners_year <- sum(d_G$sex_partners_year >= 10, na.rm=T)
+```
+
+Our initial analyses found little support for a sex difference in the effect of strength on mating success, contrary to the sexual selection hypothesis.
+We traced this failure to the unusually high number of lifetime sex partners reported by some men and women.
+We discovered that whereas women reported a median number of `{r} female_median` lifetime sex partners, and men `{r} male_median` partners, there were `{r} high_partners` individuals who reported 100 or more partners, a few of whom reported 1000 or more (@fig-ecdf).
+We decided to restrict our sample to individuals with fewer than 100 lifetime sex partners for several reasons.
+First, support for the sexual selection hypthesis was weak if these individuals were included, but much stronger if they were excluded.
+We suspect that because our sample is large and nationally representative, it probably includes sex workers, as well as individuals who regularly use the services of sex workers, and whose high partner numbers would therefore not reflect mate attraction or intrasexual competition.
+Second, there was an unusually high number of participants reporting exactly 100 lifetime sex partners, suggesting inaccurate recollections.
+Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
+We will use the same cutoff for our confirmatory study.
+
+```{r}
+#| label: fig-ecdf
+#| fig-width: 10
+#| fig-cap: Cumulative distribution of the number of sex partners, by sex.
+Participants who reported 100 or more partners were removed from the analysis.
+The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this plot and not for any analyses).
+
+plot_sexpartners
+```
+
 ### Descriptive statistics
 
-Table \@ref(tab:sexualdimorphism) shows weighted means and standard deviations of all variables for men and women.
-Mean age was 38 years for this sample.
-Sexual dimorphism is given as a ratio of male mean:female mean.
-Larger male size dimorphism is reflected in an 8% higher value for height and a 16% higher value for weight; females, however, had a 1% higher BMI than males on average.
-There is a much larger degree of dimorphism in combined grip strength (57%).
-Hemoglobin is 14% higher in males than females while white blood cell count is 3% higher in women.
+The mean age was 38 years.
+Sexual dimorphism---the ratio of male mean:female mean---was evident in an 8% higher value for male height and a 16% higher value for weight; females, however, had a 1% higher BMI than males on average.
+There was a much larger degree of male-biased dimorphism in combined grip strength (57%).
+Hemoglobin was 14% higher in males than females while white blood cell count was 3% higher in women.
+See @stbl-stats for weighted means, standard deviations, and standardized mean differences of all variables for men and women.
 
-```{r sexual dimorphism}
-getncat <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-                partnered,
-                edu,
-                race,
-                perceived_abnormal_weight,
-                special_equipment,
-                vigorous_work,
-                moderate_work,
-                vigorous_rec,
-                moderate_rec),
-    statistic = list(
-      all_categorical() ~ "{N_nonmiss_unweighted}"
-    ) , missing = "no" )
+### Coefficients of strength, and strength X sex interaction
 
-cat <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-      partnered,
-      edu,
-      race,
-      perceived_abnormal_weight,
-      special_equipment,
-      vigorous_work,
-      moderate_work,
-      vigorous_rec,
-      moderate_rec
-    ),
-    label  = list(
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
-    ),
-    statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(n_male = getncat$table_body$stat_1, digits = 2) %>%
-      dplyr::mutate(n_female = getncat$table_body$stat_2, digits = 2) %>%
-      dplyr::relocate(n_male, .before = stat_1) %>%
-      dplyr::relocate(n_female, .before = stat_2)
-  ) %>% 
-  modify_header(
-    label = "**Variable**",
-    n_male = "**N**",
-    n_female = "**N**",
-    stat_1 = "**Mean (SD) or n (%)**",
-    stat_2 = "**Mean (SD) or n (%)**",
-  ) %>%
-  modify_spanning_header(list(
-    c("stat_1", "n_male") ~ "**Male**",
-    c("stat_2", "n_female") ~ "**Female**"
-  )) %>% 
-  bold_labels()
+Our pilot results revealed strong positive significant main effects of strength on almost all mating success outcomes after controlling for all five sets of potential confounds.
+The only exceptions were that the strength coefficient for past year number of partners was not significant after controlling for tesosterone, and the coefficient for age at first sex was not significant after controlling for socioeconomic factors.
+Even in these cases, however, the sign and magnitude of these coefficients were very similar to those of these coefficients in the models with other control variables.
+See the left column of @fig-coefs.
 
-#####
+There were also clear sex differences in the effects of strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+The only exception was that the interaction term for lifetime partners was not significant in the anthropomentric control model.
+However, its sign and magnitude was similar to those in models with other control variables.
+There was also a persistent negative association of numbers of lifetime partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
+Finally there was no evidence for a sex difference in the effects of strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+See the right column of @fig-coefs.
 
-getncon <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age_first_sex,
-                sex_partners,
-                sex_partners_year,
-                strength,
-                age,
-                bmi,
-                height,
-                weight,
-                whitebloodcell,
-                hemoglobin,
-                testosterone, 
-                chronic_disease_score,
-                physical_disease_count,
-                depression,
-                avgcalories, 
-                avgprotein, 
-                foodinsecurity_adult, 
-                tot_MET),
-    statistic = list(
-      all_continuous() ~ "{N_nonmiss_unweighted}",
-      all_categorical() ~ "{N_nonmiss_unweighted}"
-    ),
-    digits = all_continuous() ~ 0,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous", 
-      foodinsecurity_adult ~ "continuous"
-    ),
-    missing = "no"
-  )
+```{r}
+#| label: fig-coefs
+#| fig-width: 10
+#| fig-height: 8
+#| fig-cap: Coefficients of strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasipoisson regressions, and age at first sex coefficients are from Gaussian regressions.
+Bars are 95% CIs.
+For the full models, see the SI.
 
-getratio <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age_first_sex,
-                sex_partners,
-                sex_partners_year,
-                strength,
-                age,
-                bmi,
-                height,
-                weight,
-                whitebloodcell,
-                hemoglobin,
-                testosterone, 
-                chronic_disease_score,
-                physical_disease_count,
-                depression,
-                avgcalories, 
-                avgprotein,
-                foodinsecurity_adult, 
-                tot_MET),
-    statistic = list(
-      all_continuous() ~ "{mean}",
-      all_categorical() ~ "{p}"
-    ),
-    digits = list(
-      all_continuous() ~ c(2)
-                  ), 
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous",
-      foodinsecurity_adult ~ "continuous"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  )
-
-con <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-      age_first_sex,
-      sex_partners,
-      sex_partners_year,
-      strength,
-      age,
-      bmi,
-      height,
-      weight,
-      whitebloodcell,
-      hemoglobin,
-      testosterone, 
-      chronic_disease_score,
-      physical_disease_count,
-      depression,
-      avgcalories, 
-      avgprotein,
-      foodinsecurity_adult, 
-      tot_MET
-    ),
-    label  = list(
-      age_first_sex = "Age at first sex (years)",
-      sex_partners = "Lifetime number of sexual partners",
-      sex_partners_year = "Past year number of sexual partners",
-      strength = "Combined Grip Strength (kg)",
-      age = "Age (Years)",
-      bmi = "Body mass index (kg/m^2)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      avgcalories = "Dietary energy intake (kcals)", 
-      avgprotein = "Dietary protein intake (grams)",
-      foodinsecurity_adult = "Food Insecurity Rating (1-4)", 
-      tot_MET = "Total MET"
-    ),
-    statistic = list(all_continuous() ~ "{mean} ({sd})",
-                     all_categorical() ~ "{n_unweighted} ({p}%)"),
-    digits = all_continuous() ~ 2,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous", 
-      foodinsecurity_adult ~ "continuous"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
-      dplyr::mutate(n_male = getncon$table_body$stat_1, digits = 2) %>%
-      dplyr::mutate(n_female = getncon$table_body$stat_2, digits = 2) %>%
-      dplyr::relocate(n_male, .before = stat_1) %>%
-      dplyr::relocate(n_female, .before = stat_2)
-  )  %>%
-  add_difference(test = everything() ~ "smd",
-                 estimate_fun = list(all_continuous() ~ purrr::partial(style_ratio, digits = 2))) %>%
-  modify_column_hide(columns = ci) %>%
-  modify_header(
-    label = "**Variable**",
-    ratio = "**Ratio**",
-    n_male = "**N**",
-    n_female = "**N**",
-    stat_1 = "**Mean (SD)**",
-    stat_2 = "**Mean (SD)**",
-    estimate = "*d*"
-  ) %>%
-  modify_spanning_header(list(
-    c("stat_1", "n_male") ~ "**Male**",
-    c("stat_2", "n_female") ~ "**Female**",
-    c("ratio", "estimate") ~ "**Sexual Dimorphism**"
-  )) %>%
-  modify_footnote(all_stat_cols() ~ "Weighted means and standard deviations shown for continuous variables, unweighted n (%) for categorical variables") %>%
-  modify_caption("**Sex differences for participants ages 18-60 using population weights**") %>%
-  bold_labels()
-
-tbl_stack(list(con, cat))
-```
-
-```{r sexualdimorphism}
-
-# getn <- designsG$d.design.adults %>%
-#   tbl_svysummary(
-#     by = sex,
-#     include = c(age,
-#                 height,
-#                 weight,
-#                 bmi,
-#                 strength,
-#                 whitebloodcell,
-#                 hemoglobin,
-#                 testosterone),
-#     statistic = list(
-#       all_continuous() ~ "{N_nonmiss_unweighted}"
-#     ),
-#     digits = all_continuous() ~ 0,
-#     missing = "no"
-#   )
-# 
-# getratio <- designsG$d.design.adults %>%
-#   tbl_svysummary(
-#     by = sex,
-#     include = c(age,
-#                 height,
-#                 weight,
-#                 bmi,
-#                 strength,
-#                 whitebloodcell,
-#                 hemoglobin,
-#                 testosterone),
-#     statistic = list(
-#       all_continuous() ~ "{mean}"
-#     ),
-#     missing = "no"
-#   ) %>%
-#   modify_table_body(
-#     ~ .x %>%
-#       dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-#   )
-# 
-# 
-# 
-# tab1 <- designsG$d.design.adults %>%
-#   tbl_svysummary(
-#     by = sex,
-#     include = c(
-#       age,
-#       height,
-#       weight,
-#       bmi,
-#       strength,
-#       whitebloodcell,
-#       hemoglobin,
-#       testosterone
-#     ),
-#     statistic = list(all_continuous() ~ "{mean} ({sd})"),
-#     digits = all_continuous() ~ 2,
-#     label  = list(
-#       strength = "Combined Grip Strength (kg)",
-#       height = "Height (cm)",
-#       weight = "Weight (kg)",
-#       age = "Age (years)",
-#       bmi = "Body mass index (kg/m^2^)",
-#       whitebloodcell = "White blood cell count (1000 cells/µL)",
-#       hemoglobin = "Hemoglobin (g/dL)",
-#       testosterone = "Testosterone (ng/dL)"
-#     ),
-#     missing = "no"
-#   ) %>%
-#   modify_table_body(
-#     ~ .x %>%
-#       dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
-#       dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
-#       dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
-#       dplyr::relocate(n_male, .before = stat_1) %>%
-#       dplyr::relocate(n_female, .before = stat_2)
-#   )  %>%
-#   add_difference(test = everything() ~ "smd",
-#                  estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
-#   modify_column_hide(columns = ci) %>%
-#   modify_header(
-#     label = "**Variable**",
-#     ratio = "**Ratio**",
-#     n_male = "**N**",
-#     n_female = "**N**",
-#     stat_1 = "**Mean (SD)**",
-#     stat_2 = "**Mean (SD)**",
-#     estimate = "*d*"
-#   ) %>%
-#   modify_spanning_header(list(
-#     c("stat_1", "n_male") ~ "**Male**",
-#     c("stat_2", "n_female") ~ "**Female**",
-#     c("ratio", "estimate") ~ "**Sexual Dimorphism**"
-#   )) %>%
-#   modify_footnote(update = everything() ~ NA) %>%
-#   modify_caption("**Sex differences for participants ages 18-59 using population weights**") %>%
-#   bold_labels()
-# 
-# tab1
-
-```
-
-### Anthropometric models of mating success
-
-In the anthropometric control model (Figure \@ref(fig:baseline)), strength was significantly associated with higher numbers of lifetime and past year sex partners and an earlier age at first sex, and the interaction with sex was not significant.
-
-```{r baseline, fig.width = 12, fig.height = 8, fig.cap = "Anthropometric models of mating success (Pilot): Coefficient plot from generalized linear models.
-Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
-
-Figure1 <- forestplot(
-  m_agefirst,
-  m_lifetime,
-  m_pastyear,
-  m_partnered,
-  intercept = F,
-  facet = F,
-  dodgewidth = .5,
-  modelnames = mnames1,
-  varnames = vnames, 
-  size = 1.2, 
-  linewidth = 1
-) + theme_minimal(17) +
-#  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
-#  labs(title = "Figure 1. Baseline Models: Regression coefficients from generalized linear models", color = "Model") +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-# 
-# 
-# annotate_figure(fig1,
-#                            top = text_grob("Figure 1. Baseline Models: Regression coefficients from generalized linear models\n",
-#                                            size = 20, face = "bold",
-#                                            hjust = 0,
-#                                            x = 0.01),
-#                            bottom = text_grob("Variables labelled (S) have been centered at the mean and standarized by 2 SD",
-#                                               size = 12))
-
-#visualize coefs from models derived from lassek and gaulin. 
-#add coef table below to match l and g
-
-ggsave("Figures/Figure1.pdf", Figure1, width = 14, height = 7)
-
-Figure1
-```
-
-### Models of mating success with socioeconomic, health, hormone, and physical activity controls
-
-We then used generalized linear regression to evaluate the impact of a host of other potential confounds on these results.
-For lifetime number of sex partners, strength remained a significant positive predictor when controlling for socioeconomic, physical activity, health, and hormone variables, with no interaction with sex.
-This was also the case for past year number of sex partners, except in the case of the hormone control model.
-Strength predicted an earlier age at first sex for all models except the socioeconomic.
-There was no significant interaction between sex and strength in any model.
-
-We also modeled partnered status.
-While strength was a significant predictor of currently being partnered in each model, a significant sex:strength interaction emerged in the physical activity and hormonal control models such that the strength effect was reduced for women.
-Coefficients are shown in Figure \@ref(fig:controls).
-
-```{r controls, fig.width = 12, fig.height = 42, fig.cap = "Additional control models (Pilot): Coefficient plots for predictors of mating success from generalized linear models.
-Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
-
-#include groups of potential confounds to explore lack of sex effect for all 4 mating outcomes
-
-fig1 <- forestplot(
-   msoc1, mheal1, mphys1, mhor1,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames, 
-  varnames = vnames, 
-  size = 0.7, 
-  linewidth = 0.7) + theme_minimal() +
-  theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11)) +
- # geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
-  labs(title = "Lifetime Number of Sexual Partners") 
- # coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
-
-fig2 <- forestplot(
- msoc2,mheal2, mphys2, mhor2,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
- size = 0.7, 
-  linewidth = 0.7
-) + theme_minimal() +
-#  geom_pointrange(size = 1.2, position = position_dodge(.8)) +
-  labs(title = "Past Year Number of Sexual Partners")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
-#  coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
-
-fig3 <- forestplot(
-  msoc3, mheal3, mphys3, mhor3,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
-  size = 0.7, 
-  linewidth = 0.7) +
-  theme_minimal() +
-#  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
-  labs(title = "Age at First Sex")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
- # coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
-
-fig4 <- forestplot(
-  msoc4, mheal4, mphys4, mhor4,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  modelnames = mnames,
-  varnames = vnames, 
-  size = 0.7, 
-  linewidth = 0.7) +
-  theme_minimal() + 
-  #geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
-  labs(title = "Currently Partnered")  +
-  theme(axis.text.y = element_blank(), plot.title = element_text(size = 11))
-#  coord_cartesian(clip = "off")
-# + xlim(-2.5, 3.55)
-
-Figure2 <- (fig1 + fig2 + fig3 + fig4) + plot_layout(guides = "collect", ncol = 4) & theme(legend.position = "bottom")
-
-# ggarrange(fig1, fig2, fig3, fig4,
-#                      labels = "AUTO",
-#                      font.label = list(size = 20),
-#                      ncol = 2, nrow = 2,
-#                      common.legend = TRUE, legend="bottom")
-
- # annotate_figure(fig1234,
- #                       top = text_grob("Figure 2. Regression coefficients from generalized linear models of mating success\n",
- #                                       size = 20, face = "bold",
- #                                       hjust = 0,
- #                                       x = 0.01),
- #                       bottom = text_grob("\nVariables labelled (S) have been centered at the mean and standarized by 2 SD Base level of race/ethnicity is Mexican American.",
- #                                          size = 12))
-
-ggsave("Figures/Figure2.pdf", Figure2, width = 14, height = 7)
-
-Figure2
-
+plot_coefs
 ```
@@
 Previous research found that higher musculature was associated with decreased investment in native immunity.
-We developed two models of immune investment in white blood cell count (Figure \@ref(fig:immune)).
+We developed two models of immune investment in white blood cell count (@fig-immune).
 In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
-In the expanded control model, this pattern remained even with additional controls (see Table ).
+In the expanded model with additional controls, this pattern remained.
 
-```{r immune, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
-Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
+```{r}
+#| label: fig-immune
+#| fig.width: 12
+#| fig.height: 8
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+
 Figure3 <- forestplot(
@@
   modelnames = mmnames) +
+  scale_color_binary() +
   theme_minimal(17) +
- # geom_pointrange(size = 2, position = position_dodge(width = .8)) +
- # labs(title = "White blood cell count") +
   guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
-  #      subtitle = "White Blood Cell Count",
-  #      caption = "Variables with (S) have been standarized by 2 SD.") +
-  # guides(colour = guide_legend(reverse = T))
 
@@
 
-Finally, we tested whether strength was related to increased energy (Figure \@ref(fig:dietary)) or protein intake (Figure \@ref(fig:protein)).
+Finally, we tested whether strength was related to increased energy (@fig-energy) or protein intake (@fig-protein).
 We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
-In the expanded control model we found the same pattern while controlling for additional variables.
+In the expanded models with additional controls we found the same pattern.
 
-```{r dietary, fig.width = 12, fig.height = 8, fig.cap = "Coefficient plot of predictors of dietary energy intake (kcals) from generalized linear models (Pilot).
-Variables labelled (S) have been centered at the mean and standardized by 2 SD."}
+```{r}
+#| label: fig-energy
+#| fig.width: 12
+#| fig.height: 8
+#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+
 Figure4 <- forestplot(
@@
   modelnames = mmnames) +
+  scale_color_binary() +
   theme_minimal(17) +
-  #geom_pointrange(size = 2, position = position_dodge(width = .8)) +
-  #labs(title = "Dietary energy intake") + 
   guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
-  #      subtitle = "White Blood Cell Count",
-  #      caption = "Variables with (S) have been standarized by 2 SD.") +
-  # guides(colour = guide_legend(reverse = T))
 
@@
 
-```{r protein}
+```{r}
+#| label: fig-protein
+#| fig.width: 12
+#| fig.height: 8
+#| fig.cap: Coefficient plot of predictors of dietary protein intake from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+
 Figure5 <- forestplot(
@@
   modelnames = mmnames) +
+  scale_color_binary() +
   theme_minimal(17) +
@@
 
-In this pilot study using grip strength as a proxy, we replicated findings from @lassek2009 that muscularity is associated with increased mating success, as well as with immune and energy costs, even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-The only exception was that strength was not a positive predictor of past year partners in the hormone control model; the coefficient, however, remained similar in magnitude to the significant strength coefficients in other models of past year partners.
-In addition, strength was a positive predictor of partnered status, an additional index of mating success.
+Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with all four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
+It was also associated with immune and energy costs.
+The only exceptions were that strength was not a significant predictor of past year partners in the hormone control model nor of age at first sex in the socioeconomic control model; the coefficients, however, were similar in sign and magnitude to the significant strength coefficients in the other models.
 
-We found that the negative effect of strength on age at first sex may be due to a confound with ethnicity.
-Specifically, Asian Americans have markedly lower grip strength compared to Mexican Americans and later age at first sex.
-For this reason, we do not predict an effect of strength on age at first sex controlling for ethnicity.
+The effect of strength on age at first sex was not signficant in our socioeconomic model probably because it is confounded with ethnicity.
+Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
+In addition, participants, whose grip strength was measured at the time of the study when they were adults, are reporting on an event that happened an average of 22 years in the past when most were adolescents and their grip strength had not yet reached its adult value.
+For these reasons, we do not predict an effect of strength on age at first sex after controlling for ethnicity.
 
-We extended these findings to include women in order to test the hypothesis that sexual dimorphism in strength is due to sexual selection for formidability.
-At the outset of our study we predicted that strength would be a strong predictor of mating success for men and a significantly weaker predictor for women.
-Contrary to predictions, in our pilot study we found no significant interaction between strength and sex for lifetime number of partners or past year number of partners.
-This calls into question the sexual selection hypothesis for sexual dimorphism in upper body strength, because our results confirm that strength has immune and energy costs that would be offset by the benefits of multiple mates for men but less plausibly for women.
-We did, however, find evidence that strength was a significantly stronger predictor of partnered status for men than for women.
-In humans biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships.
-We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability clearly applies to long-term partnerships (proxied by our partnered variable), where strength was a stronger predictor for men than women.
-Partner numbers are arguably better proxies for short term mating success.
+The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status, and also convincingly supported for numbers of lifetime partners.
+In humans, biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships.
+We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability applies to long-term partnerships (proxied by our partnered variable), whereas partner numbers are arguably better proxies for short term mating success.
+Thus, our pilot study supports the hypothesis that muscularity increases both short- and long-term mating success for men more than it does for women.
+Number of lifetime partners was also a negative predictor of partnered status, suggesting a tradeoff in pursuing these two reproductive strategies.
 
-The lack of a sex difference in the significant positive effect of strength on short term mating success is puzzling.
-Because men have greater strength than women their partner numbers are higher.
-Still, women with higher grip strength report more partners than women with lower grip strength even after controlling for a host of potential confounds.
+The lack of a sex difference in the significant positive effect of strength on short term mating success (proxied by number of partners in the last year) is puzzling.
+Because men have greater strength than women their self-reported partner numbers are higher, on average, although the median number of partners in the last year for both sexes is 1.
+Still, women with higher grip strength report more partners in the past year than women with lower grip strength even after controlling for a host of potential confounds.
 It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
-It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently their (stronger) mates would also necessarily have more mates.
-It may be that stronger women require less male investment and so instead benefit from greater partner numbers through e.g.
-genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for high quality long term mate, or through avoiding costly long term partnerships.
+It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
+It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
+genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
 It might also be the case that there are some sex-specific confounds that we failed to control for.
-Finally, greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
+Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
+Finally, our simulations revealed that we had relatively lower power to detect a significant interaction between strength and sex for this outcome (see Power section below).
 
-We now argue that these results emphasize the importance of biparental care in human reproduction (Gettler et al., 2020; Bribiescas et al., 2012).
-Across evolutionary time, fitness for both men and women was heavily dependent on obtaining successful long term partners, thereby reaping the efficiency benefits of a sexual division of labor (Quinlan, 2008; Quinlan & Quinlan, 2007).
-More specifically, humans require provisioning assistance for an extended period (Insert figure?) (Davison & Gurven, 2022; Davison & Gurven, 2021) and paternal care is an important predictor of child outcomes (Starkweather et al., 2021; Winking et al., 2011).
-Thus, our pilot study refines the sexual selection hypothesis for the relationship between strength and mating success to emphasize that mating success is best operationalized as long-term mating, rather than numbers of sexual partners.
-Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity and therefore the ability of stronger men to provision offspring through big-game hunting (Kaplan et al., 2000).
+Our pilot results support the importance of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
+More specifically, humans require provisioning for an extended period [@davison2021; @davison2022] (@fig-provisioning), and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
+Thus, our pilot study refines the sexual selection hypothesis for the relationship between strength and mating success to emphasize that mating success should be operationalized as long-term mating in addition to numbers of sexual partners.
+Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014].
 
@@
 
-Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is greater for males than females.
-Our aim was not to test all possible byproduct hypotheses, nor is it possible to do so.
-Therefore, we are unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS).
+Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females.
+Our aim is not to test all possible byproduct hypotheses, nor is it possible to do so.
+Therefore, we are unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success.
 Likewise, NHANES does not collect data on sociosexuality.
-Likewise, as described in our measures, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior for both men and women.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).
+As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women.
+Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
+Finally, all pilot results were obtained after considerable exploratory analysis, and must be replicated in a confirmatory study.
 
-# Confirmatory Analysis [not yet conducted]
+# Confirmatory Study (Not yet conducted)
 
-We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength, in Table X.
+In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime and last year number of sexual partners, but not age at first sex after controlling for ethnicity.
+We also predict a significant interaction with sex such that strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
+We also predict that there will be a negative association between male formidabillity and immune function, and that formidability will be positively associated with dietary energy intake for both males and females.
+Finally, we predict a significant tradeoff between engaging in a short-term vs.
+long-term mating strategy, specifically, the the coefficients of lifetime partner number in models of partnered status will be significantly negative.
+
+In @tbl-predictions we indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength for our four mating success outcomes.
 We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models.
-If the effect of strength on the outcome variable is not significant in any single model, especially if the size of the coefficient is markedly reduced, then we will consider that evidence against the sexual selection hypothesis.
+
+| Outcome            | Strength coefficient | Strength X Sex coefficient |
+|--------------------|:--------------------:|:--------------------------:|
+| Partnered status   |           +          |              -             |
+| Lifetime partners  |           +          |              -             |
+| Past year partners |           0          |              0             |
+| Age at first sex   |           0          |              0             |
+
+: Predictions.
+A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model.
+{#tbl-predictions}
+
+## Power
+
+We used simulations to estimate our power to detect strength effects that were 75% as large as the ones observed in our pilot study, and sex interaction effects that were equal in magnitude but opposite in sign to the strength effects (i.e., no effect of strength on mating success in women), with a sample size equal to that in our unobserved data.
+We are confident that we have sufficient power to detect true strength effects somewhat less than the ones observed in our pilot study (@tbl-power).
+We are less confident in our power to detect true sex interaction effects because, all else equal, power to detect interaction effects is generally lower than main effects, and in our simulations we used large interaction effects equal in magnitude to the main effects, which is probably unrealistic.
+
+| Outcome            | Strength coefficient | Strength X Sex coefficient |
+|--------------------|:--------------------:|:--------------------------:|
+| Partnered status   |         0.99         |            0.97            |
+| Lifetime partners  |         0.99         |            0.84            |
+| Past year partners |         0.95         |            0.52            |
+
+: Power to detect our predicted effects.
+Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients, sample size was set to that of our unobserved data, and $\alpha = 0.05$.
+{#tbl-power}
 
@@
 
-## Supplementary Information
+# Acknowledgements
 
-```{r vaginalsex}
+We thank two anonymous reviewers for their numerous helpful suggestions.
 
+# Supplementary Information
+
+## Pilot study
+
+::: {#stbl-stats}
+`{r} tbl_stack(list(con, cat))`
+
+Descriptive statistics of pilot study sample.
+:::
+
+### Models of mating success with socioeconomic, health, hormone, and physical activity controls
+
+::: {#sfig-lifetime}
+```{r}
+#| fig-height: 10
+forestplot2(m_lifetime, msoc1, mheal1, mphys1, mhor1, title = "Lifetime Number of Sexual Partners")
 ```
 
-```{r corrmatf}
-corr_f <- corrplot(cor_mat_cors_f, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
-                   title = "Correlation matrix of coninuous study variables, Females 18-60")
+Quasi-Poisson regression models of lifetime number of sex partners.
+:::
 
+::: {#sfig-year}
+```{r}
+#| fig-height: 10
+forestplot2(m_pastyear, msoc2, mheal2, mphys2, mhor2, title = "Past Year Number of Sexual Partners")
 ```
 
-```{r corrmatm}
+Quasi-Poisson regression models past year sex partners.
+:::
 
-corr_m <- corrplot(cor_mat_cors_m, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
-                   title = "Correlation matrix of coninuous study variables, Males 18-60")
-
+::: {#sfig-agefirstsex}
+```{r}
+#| fig-height: 10
+forestplot2(m_agefirst, msoc3, mheal3, mphys3, mhor3, title = "Age at First Sex")
 ```
 
-```{r corrmatcom}
+Gaussian regression models of age at first sex.
+:::
 
-corr_com <- corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75, 
-title = "Correlation matrix of coninuous study variables, Males and Females 18-60")
+::: {#sfig-partnered}
+```{r}
+#| fig-height: 10
+forestplot2(m_partnered, msoc4, mheal4, mphys4, mhor4, title = "Currently Partnered")
 ```
 
+Logistic regression models of partnered status.
+:::
+
+### The negative association of numbers of lifetime sexual partners on partnered status
+
+::: {#sfig-tradeoff}
+```{r}
+plot_lifetime_coefs
+```
+
+The adjusted odds ratios of numbers of lifetime partners on partnered status. In these models, lifetime partners was divided by 10, the interquartile range.
+:::
+
+### Correlation matrices of study variables
+
+::: {#sfig-cor}
+```{r}
+ggcorrplot(cor_mat_cors, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
+```
+
+Correlation matrix of study variables (both sexes). Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-fcor}
+```{r}
+ggcorrplot(cor_mat_cors_f, hc.order = T, hc.method = 'ward.D', title = 'Females')
+```
+
+Correlation matrix of study variables for females only. Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-mcor}
+```{r}
+ggcorrplot(cor_mat_cors_m, hc.order = T, hc.method = 'ward.D', title = 'Males')
+```
+
+Correlation matrix of study variables for males only. Rows and columns ordered by hierarchical clustering.
+:::
+
+### Regression tables
+
+**Lifetime number of partners**
+
+:::{#stbl-lifetime}
+```{r}
+modelsummary(
+  list(
+    "Anthropometric" = m_lifetime, 
+    "Socioeconomic" = msoc1, 
+    "Health" = mheal1, 
+    "Activity" = mphys1,
+    "Hormone" = mhor1
+    ), 
+  gof_map = "nobs"
+  )
+```
+
+Quasi-Poisson Regression Models of Lifetime Number of Sexual Partners
+:::
+
+**Past year number of partners**
+
+:::{#stbl-pastyear}
+```{r}
+modelsummary(
+  list(
+    "Anthropometric" = m_pastyear, 
+    "Socioeconomic" = msoc2, 
+    "Health" = mheal2, 
+    "Activity" = mphys2,
+    "Hormone" = mhor2
+    ), 
+  gof_map = "nobs"
+  )
+```
+
+Quasi-Poisson Regression Models of Past Year Number of Sexual Partners
+:::
+
+**Age at first sex**
+
+:::{#stbl-agefirstsex}
+```{r}
+modelsummary(
+  list(
+    "Anthropometric" = m_agefirst, 
+    "Socioeconomic" = msoc3, 
+    "Health" = mheal3, 
+    "Activity" = mphys3,
+    "Hormone" = mhor3
+    ), 
+  gof_map = "nobs"
+  )
+```
+
+Gaussian Regression Models of Age at First Sex
+:::
+
+**Partnered status**
+
+:::{#stbl-partnered}
+```{r}
+modelsummary(
+  list(
+    "Anthropometric" = m_partnered, 
+    "Socioeconomic" = msoc4, 
+    "Health" = mheal4, 
+    "Activity" = mphys4,
+    "Hormone" = mhor4
+    ), 
+  gof_map = "nobs"
+  )
+```
+
+Logistic Regression Models of Partnered Status
+:::
+
 # References

### 2024-03-30 paper.R1.qmd
Renamed to paper.R1.qmd

--- 2024-03-30 paper.v2.qmd
@@
 format:
+  docx:
+    number-sections: true
   html:
@@
 csl: apa.csl
-bibliography: regrep.bib
+bibliography: paper.bib
 crossref:
@@
 library(reshape2)
-library(gtsummary)
 library(ggcorrplot)
@@
 
-# Abstract
+Abstract
 
@@
 
-## Sexual dimorphism in modern humans:
+# Sexual dimorphism in modern humans
 
@@
 Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass [@kirchengast2010].
-At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass [@lassek2022], which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees [@plavcan2012; @smith1997].
+At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass [@lassek2022], which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees [@plavcan2012; @smith1997; @Puts_2023].
 Men also have larger and stronger bones [@wells2007].
-Other traits that are sexually dimorphic in humans include digit ratio, voice pitch, facial features, body and facial hair growth, and canine length.
+Other traits that are sexually dimorphic in humans include digit ratio, voice pitch, facial features, body and facial hair growth, and canine length [@Puts_2023].
 
@@
 The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past [@puts2010], and that instead male fitness was determined by physical contests with other males for access to mates.
-Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [@puts2016; @hill2017; @plavcan2012].
-Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression [@archer2009].
+Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [@puts2016; @Puts_2023; @hill2017; @plavcan2012].
+Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression [@archer2009; @Puts_2023].
 Particular emphasis is placed on sexual dimorphism in the upper-body due to the important role upper-body strength plays in both armed and unarmed fighting ability [@sell2012] and where sex differences in strength and muscularity are greatest as discussed above.
@@
 
-Although there is evidence in Western young adult populations that male mating success is mediated by male-rated dominance and not by female-rated attractiveness [@kordsmeyer2018; @hill2013], other researchers nevertheless argue that sexual dimorphism in humans is largely the result of female choice for high quality males [which was possibly heavily influenced by parents, in which case parental preferences would also have been important, @apostolou2007].
+Although there is evidence in Western young adult populations that male mating success is mediated by male-rated dominance and not by female-rated attractiveness [@kordsmeyer2018; @hill2013], others nevertheless argue that sexual dimorphism in humans is largely the result of female choice for high quality males [which was possibly heavily influenced by parents, in which case parental preferences would also have been important, @apostolou2007].
 The immunocompetence handicap hypothesis, for example, relies on evidence that testosterone is immunosuppressive, and therefore only males with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine secondary sexual traits [@folstad1992].
@@
 
-Females might be choosing males who could provide more resources, or men who could better protect them from attacks by other men or by predators, which in ancestral hunter-gatherer populations would have been physically stronger men [@apicella2014; @willems2017b].
+Females might therefore be choosing males who could provide more resources, or men who could better protect them from attacks by other men or by predators, which in ancestral hunter-gatherer populations would have been physically stronger men [@apicella2014; @willems2017b].
 There is substantial cross-cultural evidence of a female-biased preference for mates who can provide resources [@walter2020; @buss2019].
-Importantly, studies of contemporary hunter-gatherers and game-theoretic models find that a mutually beneficial sexual division of labor within long-term pair bonds is critical to provisioning offspring who remain dependent on caregivers for up to 20 years or more [@davison2021; @davison2022; @alger2020; @kaplan2000].
+Importantly, studies of contemporary hunter-gatherers and game-theoretic models find that a mutually beneficial sexual division of labor within long-term pair bonds, in which women engage in lower risk, lower return foraging of plants and small game compatible with the cognitively demanding task of childrearing [@hagen2019], and men engage in higher risk, higher return big game hunting [@kelly2013], is critical to provisioning offspring who remain dependent on caregivers for 20 years or more [@davison2021; @davison2022; @alger2020; @kaplan2000].
 See @fig-provisioning.
-Typically, forager women engage in lower risk, lower return foraging of plants and small game compatible with the cognitively demanding task of childrearing [@hagen2019], and men engage in higher risk, higher return big game hunting [@kelly2013].
 This suggests that sex differences in body composition could be due, at least in part, to a sexual division of labor.
@@
 Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with male reproductive success under ancestral conditions [@perusse1993].
-We operationalized mating success in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
+We operationalized mating success in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in the past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
 female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
@@
 
-Number of partners emphasizes success through short term mating strategies, but it is an ambiguous measure as it could also represent repeated rejection, or a tendency to exaggerate or understate the actual number.
-It also does not clearly index female reproductive success because multiple sexual partners are likely to have increased male fitness more than female fitness [@Lehtonen_2022].
+Lifetime and past year numbers of sexual partners index success through short term mating strategies.
+Lifetime numbers are probably a more precise proxy of the use of a short-term mating strategy as these are integrated over the entire life, whereas past year numbers likely fluctuate depending on partnered status and other transient factors.
+However, both are somewhat ambiguous measures as they could also represent repeated rejection, or a tendency to exaggerate or understate the actual numbers.
+Neither clearly indexes female reproductive success because multiple sexual partners are likely to have increased male fitness more than female fitness [i.e., the Bateman gradient, @Lehtonen_2022].
 We therefore also operationalized mating success in a fourth outcome measure as partnered status as a proxy for a long-term mating strategy that, as discussed earlier, benefits both sexes.
@@
 These data were then used to estimate energy intake in the form of kilocalories (kcal) calculated by matching reported foods to the USDA's Food and Nutrient Database for Dietary Studies.
-For the outcome measure in this model we used the average number of calories per day calculated for each participant across their two recall days.
+For the outcome measures in these models we used the average number of calories per day and grams of protein per day calculated for each participant across their two recall days.
 
-### Key predictor variables
+### Key explanatory variables
 
-Our main predictor variable of mating success is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
+Our main explanatory variable of mating success is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
 Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
-All regression models also included sex (male/female), age in years, and partnered status (not included in models where partnered status was the outcome variable).
+All regression models also included sex (male/female), our second main explanatory variable.
+All models also controlled for age in years and partnered status (not included in models where partnered status was the outcome variable).
+
 The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
-Likewise, we also interacted sex with age and partnered status.
+Grip strength is highly sexually dimorphic, however, with about 90% of men stronger than about 90% of women [@smith2022], and is thus highly collinear with sex, a problem for regression modeling and interpretation.
+We therefore computed a sex-specific grip strength value as follows: for each sex we centered grip strength at the sex-specific mean and divided by two times the sex-specific standard deviation.
+Thus, for women, low or high values are relative to other women, and for men, low or high values are relative to other men.
+The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high strength.
+Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes.
+
+We also interacted sex with age and partnered status.
 We interacted age with sex since menopause occurs within the age range of our participants.
@@
 
-Our modelling strategy takes inspiration from @lassek2009.
+Our modeling strategy takes inspiration from @lassek2009.
 However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples [@smith2018a].
@@
 
-There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women [@pollet2011; @vananders2007], but mixed evidence that it is related to sexual desire [@vananders2013].
+There is evidence that testosterone is positively associated with higher numbers of partners for men and less robustly so for women [@pollet2011; @vananders2007], but mixed evidence that it is related to sexual desire [@vananders2013].
 For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see @vananders2013.
@@
 The hormone control model included serum total testosterone (ng/dL).
-Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex, one of our key predictor variables.
-We therefore computed a sex-specific testosterone value as follows: for each sex we centered testosterone at the sex-specific mean and divided by two times the sex-specific standard deviation.
-Thus, for women, low or high values were relative to other women, and for men, low or high values were relative to other men.
+Since male mean testosterone is an order of magnitude higher than the female mean, testosterone, like grip strength, was highly confounded with sex, one of our key predictor variables.
+We therefore computed a sex-specific testosterone value in the same way we did for strength.
 This differed from other centered and standardized variables which we centered and standardized across all male and female values.
@@
 
-### Native Immune Function
+#### Native Immune Function
 
@@
 Second, the total metabolic equivalent (total MET) was calculated based on participants' responses to questions about the minutes they spent walking or bicycling, engaging in vigorous and moderate work, and vigorous and moderate recreation per day, using MET scores provided by NHANES.
-Finally we included dietary energy intake (average calories per day as discussed above).
+We included dietary energy intake (average calories per day as discussed above).
+Finally, we substituted separate height and weight variables for BMI.
 
-### Dietary Energy Intake
+#### Dietary Energy and Protein Intake
 
@@
 We then developed an expanded control model that also controlled for WBCC and food insecurity in addition to the variables in the model above, since either could influence energy intake and strength.
+Given the protein costs of muscle maintenance we likewise developed two models of dietary protein intake.
+We based the “replication model” on that of the @lassek2009 dietary energy model, and then applied the same expanded controls discussed above in the “expanded controls model,” including the substitution of separate height and weight variables for BMI.
 
@@
 All analyses were completed in `{r} R.version.string`, using the survey package [version 4.1-1, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-We modeled lifetime and past year numbers of partners using a Poisson distribution, since these variables are count data.
-We modeled of age at first sexual intercourse and white blood cell count with a normal distribution.
-We modeled partnered status using logistic regression.
+We modeled lifetime and past year numbers of partners using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
+We modeled age at first sexual intercourse and white blood cell count with a Gaussian glm.
+We modeled partnered status using a binomial glm.
 Analyses included adults ages 18-60 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
@@
 We decided to restrict our sample to individuals with fewer than 100 lifetime sex partners for several reasons.
-First, support for the sexual selection hypthesis was weak if these individuals were included, but much stronger if they were excluded.
+First, support for the sexual selection hypothesis was weak if these individuals were included, but much stronger if they were excluded.
 We suspect that because our sample is large and nationally representative, it probably includes sex workers, as well as individuals who regularly use the services of sex workers, and whose high partner numbers would therefore not reflect mate attraction or intrasexual competition.
@@
 
-Our pilot results revealed strong positive significant main effects of strength on almost all mating success outcomes after controlling for all five sets of potential confounds.
-The only exceptions were that the strength coefficient for past year number of partners was not significant after controlling for tesosterone, and the coefficient for age at first sex was not significant after controlling for socioeconomic factors.
-Even in these cases, however, the sign and magnitude of these coefficients were very similar to those of these coefficients in the models with other control variables.
+Our pilot results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners, and age at first sex after controlling for all five sets of potential confounds.
+The only exception was that the strength coefficient for age at first sex was not significant after controlling for socioeconomic factors, although the sign and magnitude was very similar to those of these coefficients in the models with other control variables.
+There was no significant effect of sex-specific strength on the number of partners in the past year.
 See the left column of @fig-coefs.
 
-There were also clear sex differences in the effects of strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
-The only exception was that the interaction term for lifetime partners was not significant in the anthropomentric control model.
-However, its sign and magnitude was similar to those in models with other control variables.
+There were also clear sex differences in the effects of sex-specific strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
 There was also a persistent negative association of numbers of lifetime partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
-Finally there was no evidence for a sex difference in the effects of strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
 See the right column of @fig-coefs.
@@
 #| fig-height: 8
-#| fig-cap: Coefficients of strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasipoisson regressions, and age at first sex coefficients are from Gaussian regressions.
+#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
 Bars are 95% CIs.
@@
 In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
-In the expanded model with additional controls, this pattern remained.
+In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was now not significant, albeit with a coefficient that was very similar to the replication model.
 
@@
 #| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot). Variables labeled (S) have been centered at the mean and standardized by 2 SD.
 
@@
 We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
-In the expanded models with additional controls we found the same pattern.
+In the expanded models with additional controls, strength was no longer a significant predictor of energy intake but remained a significant predictor of protein intake.
 
@@
 
-Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with all four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-It was also associated with immune and energy costs.
-The only exceptions were that strength was not a significant predictor of past year partners in the hormone control model nor of age at first sex in the socioeconomic control model; the coefficients, however, were similar in sign and magnitude to the significant strength coefficients in the other models.
+Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with three of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
+The only exceptions were that strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant strength coefficients in the other models.
+Strength was also associated with immune, and energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
+The expanded control models included height, which might be a proxy for aspects of strength, such as lower body strength, that are not accounted for by grip strength.
 
-The effect of strength on age at first sex was not signficant in our socioeconomic model probably because it is confounded with ethnicity.
+The effect of strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
 Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
@@
 
-The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status, and also convincingly supported for numbers of lifetime partners.
+The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status and numbers of lifetime partners.
 In humans, biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships.
-We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability applies to long-term partnerships (proxied by our partnered variable), whereas partner numbers are arguably better proxies for short term mating success.
+We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability applies to long-term partnerships (proxied by our partnered variable), as well as short-term matings (proxied by partner numbers).
 Thus, our pilot study supports the hypothesis that muscularity increases both short- and long-term mating success for men more than it does for women.
-Number of lifetime partners was also a negative predictor of partnered status, suggesting a tradeoff in pursuing these two reproductive strategies.
+Number of lifetime partners was also a negative predictor of partnered status (@sfig-tradeoff), suggesting a tradeoff in pursuing these two reproductive strategies.
 
-The lack of a sex difference in the significant positive effect of strength on short term mating success (proxied by number of partners in the last year) is puzzling.
+The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of partners in the last year) is puzzling.
 Because men have greater strength than women their self-reported partner numbers are higher, on average, although the median number of partners in the last year for both sexes is 1.
@@
 
-Our pilot results support the importance of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+Our pilot results support the importance of strength to male short-term mating success as well as to the long-term partnerships that are the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
 Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
 More specifically, humans require provisioning for an extended period [@davison2021; @davison2022] (@fig-provisioning), and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
-Thus, our pilot study refines the sexual selection hypothesis for the relationship between strength and mating success to emphasize that mating success should be operationalized as long-term mating in addition to numbers of sexual partners.
+Thus, our pilot study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.
 Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014].
@@
 
-In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime and last year number of sexual partners, but not age at first sex after controlling for ethnicity.
+In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime number of sexual partners, but not last year number of partners, nor age at first sex after controlling for ethnicity.
 We also predict a significant interaction with sex such that strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
-We also predict that there will be a negative association between male formidabillity and immune function, and that formidability will be positively associated with dietary energy intake for both males and females.
-Finally, we predict a significant tradeoff between engaging in a short-term vs.
-long-term mating strategy, specifically, the the coefficients of lifetime partner number in models of partnered status will be significantly negative.
+We also predict that there will be a negative association between formidability and immune function.
+Finally, we predict that formidability will be positively associated with dietary energy intake for both males and females.
 
-In @tbl-predictions we indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength for our four mating success outcomes.
+We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength, in @tbl-predictions.
 We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models.
@@
 |--------------------|:--------------------:|:--------------------------:|
-| Partnered status   |         0.99         |            0.97            |
-| Lifetime partners  |         0.99         |            0.84            |
-| Past year partners |         0.95         |            0.52            |
+| Partnered status   |         0.99         |            0.90 (0.67)            |
+| Lifetime partners  |         0.99         |            0.90 (0.66)            |
+| Past year partners |         0.80         |            0.30 (0.21)            |
 
 : Power to detect our predicted effects.
-Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients, sample size was set to that of our unobserved data, and $\alpha = 0.05$.
+Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients (value in parentheses is power to detect 3/4 of that effect), sample size was set to that of our unobserved data, and $\alpha = 0.05$.
 {#tbl-power}
@@
 
-We thank two anonymous reviewers for their numerous helpful suggestions.
+We thank the Editor and two anonymous reviewers for their numerous helpful comments and suggestions.
+
+# Data availability
+
+All data and code are available here: <https://github.com/grasshoppermouse/strengthmating>
 
@@
 
-Quasi-Poisson regression models of lifetime number of sex partners.
+Models of lifetime number of sex partners.
 :::
@@
 #| fig-height: 10
-forestplot2(m_pastyear, msoc2, mheal2, mphys2, mhor2, title = "Past Year Number of Sexual Partners")
+forestplot2(m_pastyear, msoc2,mheal2, mphys2, mhor2, title = "Past Year Number of Sexual Partners")
 ```
 
-Quasi-Poisson regression models past year sex partners.
+Models past year sex partners.
 :::
@@
 
-Gaussian regression models of age at first sex.
+Models of age at first sex
 :::
@@
 
-Logistic regression models of partnered status.
+Models of partnered status
 :::

### 2024-03-31 analysis.v2.R
Code cleanup

@@ -1,18 +1,64 @@
+library(tidyverse)
 library(survey)
-library(ggplot2)
 library(nhanesGH)
 library(hagenutils)
-library(ggpubr)
-library(tidyverse)
 library(broom)
+library(gtsummary)
+library(jtools)
 
-# Removing those with >=100 lifetime sex partners -----------------------------------
+# Cumulative distributions of sex partners ------------------------
 
+plot_sexpartners <-
+  ggplot(d_G, aes(sex_partners + 1, colour = sex)) +
+  stat_ecdf() +
+  geom_vline(xintercept = 100, linetype = 'dotted') +
+  annotate("text", label = 'Cutoff', x = 110, y = 0.80, hjust = 0) +
+  scale_x_log10(breaks = c(1, 5, 10, 25, 50, 100, 1000)) +
+  scale_color_binary() +
+  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
+  labs(x = "\nNumber of lifetime sex partners + 1", y = 'ECDF\n') +
+  theme_minimal(15)
+plot_sexpartners
+ggsave("Figures/plot_sexpartners.pdf", plot_sexpartners, width = 9, height = 6)
+
+plot_sexpartners_year <-
+  ggplot(d_G, aes(sex_partners_year + 1, colour = sex)) +
+  stat_ecdf() +
+  scale_x_log10(breaks = c(1, 5, 10, 25, 50)) +
+  scale_color_binary() +
+  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
+  labs(x = "\nNumber of last year sex partners + 1", y = 'ECDF\n') +
+  theme_minimal(15)
+plot_sexpartners_year
+ggsave("Figures/plot_sexpartners_year.pdf", plot_sexpartners, width = 9, height = 6)
+
+
+# Update design objects -----------------------------------
+
+design_all <- designsG$d.design.adults # Entire sample
+
+# Removing those with >=100 lifetime sex partners
 designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
-# designsG$d.design.adults <- update(designsG$d.design.adults, partnered = maritalstatus == 1)
 
-# Anthropometric models ---------------------------------------------------------
+# Scale sex_partners by the interquartile range
+designsG$d.design.adults <- update(designsG$d.design.adults, sex_partners_scaled = sex_partners/10)
 
+# Difference between current age and age at first sex
[…]

### 2024-03-31 paper.R1.qmd
Code cleanup

@@
 format:
-  docx:
-    number-sections: true
   html:
+    toc: true
     embed-resources: true
+  # docx:
+  #   number-sections: true
 execute:
@@
 
-```{r setup, include=FALSE}
-library(ggplot2)
-library(dplyr)
-library(psych)
-library(xtable)
-library(knitr)
-library(kableExtra)
-library(weights)
-library(jtools)
-library(reshape2)
+```{r}
 library(ggcorrplot)
-library(patchwork)
 library(modelsummary)
-source('analysis.v2.R')
+source('analysis.R1.R')
 ```
@@
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95% CIs.
+Bars are 95\% CIs.
 For the full models, see the SI.
@@
 We are confident that we have sufficient power to detect true strength effects somewhat less than the ones observed in our pilot study (@tbl-power).
-We are less confident in our power to detect true sex interaction effects because, all else equal, power to detect interaction effects is generally lower than main effects, and in our simulations we used large interaction effects equal in magnitude to the main effects, which is probably unrealistic.
+We are less confident in our power to detect true sex interaction effects because, all else equal, power to detect interaction effects is generally lower than main effects.
 

### 2024-03-31 simulate.R
Updated coefs

@@ -7,7 +7,8 @@ library(furrr)
 
 plan("multisession")
 
-d_adult <- d_G[d_G$age >= 18 & d_G$age < 60,]
+d_adult <- d_G[d_G$age >= 18 & d_G$age <= 60,]
+designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
 
 rqpois <- function(n, mu, theta) rnbinom(n=n, mu=mu, size=mu/(theta-1))
 
@@ -65,120 +66,87 @@ getstats_partnered <- function(params){
 
 # Coefs from this model:
 
-m_year <- svyglm(
-  sex_partners_year ~
-    strength_centered * sex + age_centered*sex + partnered*strength_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(m_year)
+summary(m_pastyear, df.resid = Inf)
 
 # Sample size in H series
-N_year = length(na.omit(d_H$sex_partners_year))
+N_year = length(na.omit(d_H$sex_partners_year[d_H$sex_partners<=100]))
 
 # With parameters equal to those in the G series data
 year_params <- list(
   N = N_year, # Sample size in H series
-  b0 = 0.43,
-  b_sex = -0.23,
-  b_age = -0.33,
-  b_partnered = -0.27,
-  b_strength = 0.43, # 0.43 is value in G series
-  b_sex_strength = -0.43, # Set to the negative of the male value
-  b_sex_age = -0.09,
-  b_strength_partnered = -0.50,
-  theta_year = 2.5
+  b0 = 0.40,
+  b_sex = -0.29,
+  b_age = -0.39,
+  b_partnered = -0.21,
+  b_strength = 0.30,
+  b_sex_strength = -0.225, # Set to the negative 3/4 of the male value
+  b_sex_age = 0.01,
+  b_strength_partnered = -0.23,
+  theta_year = 2
 )
 
-pvalues_year <- future_map_dfr(1:1000, ~getstats_year(year_params))
+pvalues_year <- future_map_dfr(1:1000, ~getstats_year(year_params), .options = furrr_options(seed = T))
 sum(pvalues_year$'strength_centered' < 0.05)/1000
 sum(pvalues_year$'sex:strength_centered' < 0.05)/1000
 
 # With strength parameter set to 3/4 of the G series
 year_params <- list(
   N = N_year, # Sample size in H series
-  b0 = 0.43,
-  b_sex = -0.23,
[…]

### 2024-03-31 paper.qmd
renamed paper.R1.qmd to paper.qmd

--- 2024-03-31 paper.R1.qmd
@@
 date: today
+lightbox: auto
 format:
@@
   #   number-sections: true
+  #   fig-format: pdf
 execute:
@@
 
-Figure3 <- forestplot(
+plot_immune <- forestplot(
   mwbc,
@@
 
-ggsave("Figures/Figure3.pdf", Figure3, width = 14, height = 7)
+ggsave("Figures/plot_immune.pdf", plot_immune, width = 14, height = 7)
 
-Figure3
-
+plot_immune
 ```
@@
 
-Figure4 <- forestplot(
+plot_energy <- forestplot(
   m_energy,
@@
 
-ggsave("Figures/Figure4.pdf", Figure4, width = 14, height = 7)
+ggsave("Figures/plot_energy.pdf", plot_energy, width = 14, height = 7)
 
-Figure4 
+plot_energy 
 ```
@@
 
-Figure5 <- forestplot(
+plot_protein <- forestplot(
   m_protein,
@@
 
-ggsave("Figures/Figure5.pdf", Figure4, width = 14, height = 7)
+ggsave("Figures/plot_protein.pdf", plot_protein, width = 14, height = 7)
 
-Figure5 
+plot_protein 
 ```
@@
 
-### Models of mating success with socioeconomic, health, hormone, and physical activity controls
+### Models of mating success with anthropometric, socioeconomic, health, hormone, and physical activity controls
 
-::: {#sfig-lifetime}
+::: {#sfig-allcoefs}
 ```{r}
-#| fig-height: 10
-forestplot2(m_lifetime, msoc1, mheal1, mphys1, mhor1, title = "Lifetime Number of Sexual Partners")
+plot_allcoefs
 ```
 
-Models of lifetime number of sex partners.
-:::
-
-::: {#sfig-year}
-```{r}
-#| fig-height: 10
-forestplot2(m_pastyear, msoc2,mheal2, mphys2, mhor2, title = "Past Year Number of Sexual Partners")
-```
-
-Models past year sex partners.
-:::
-
-::: {#sfig-agefirstsex}
-```{r}
-#| fig-height: 10
-forestplot2(m_agefirst, msoc3, mheal3, mphys3, mhor3, title = "Age at First Sex")
-```
-
-Models of age at first sex
-:::
-
-::: {#sfig-partnered}
-```{r}
-#| fig-height: 10
-forestplot2(m_partnered, msoc4, mheal4, mphys4, mhor4, title = "Currently Partnered")
-```
-
-Models of partnered status
+Coefficients from all models.
 :::
@@
 ```{r}
+#| fig-width: 11
 plot_lifetime_coefs
@@
 ```{r}
-ggcorrplot(cor_mat_cors, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
+cor_bothsexes <- ggcorrplot(cor_mat_cors, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
+ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes)
+cor_bothsexes
 ```
@@
 ```{r}
-ggcorrplot(cor_mat_cors_f, hc.order = T, hc.method = 'ward.D', title = 'Females')
+cor_females <- ggcorrplot(cor_mat_cors_f, hc.order = T, hc.method = 'ward.D', title = 'Females')
+ggsave("Figures/cor_females.pdf", cor_females)
+cor_females
 ```
@@
 ```{r}
-ggcorrplot(cor_mat_cors_m, hc.order = T, hc.method = 'ward.D', title = 'Males')
+cor_males <- ggcorrplot(cor_mat_cors_m, hc.order = T, hc.method = 'ward.D', title = 'Males')
+ggsave("Figures/cor_males.pdf", cor_males)
+cor_males
 ```

### 2024-03-31 analysis.R1.R
Mostly code for plot_allcoefs

@@ -30,8 +30,7 @@ plot_sexpartners_year <-
   labs(x = "\nNumber of last year sex partners + 1", y = 'ECDF\n') +
   theme_minimal(15)
 plot_sexpartners_year
-ggsave("Figures/plot_sexpartners_year.pdf", plot_sexpartners, width = 9, height = 6)
-
+# ggsave("Figures/plot_sexpartners_year.pdf", plot_sexpartners_year, width = 9, height = 6)
 
 # Update design objects -----------------------------------
 
@@ -517,7 +516,7 @@ extract_stats <- function(obj, terms = c('strength_centered', 'sexfemale:strengt
     dplyr::select(term, estimate, conf.low, conf.high)
 }
 
-# Plot strength and sex:strength coefficients -----------------------------
+# Plot model coefficients -----------------------------
 
 models <- tribble(
   ~Theme,               ~Outcome,             ~Model,
@@ -547,6 +546,30 @@ models <- tribble(
   "Socioeconomic",      "Partnered",          msoc4
 )
 
+d_allstats <-
+  models |>
+  mutate(
+    Stats = map(Model, \(m) {m$df.residual <- Inf; tidy(m, conf.int = T)})
+  ) |>
+  unnest(Stats) |>
+  dplyr::select(-Model) |>  # Remove due to RStudio bug
+  dplyr::filter(term != "(Intercept)") |>
+  mutate(
+    Outcome = factor(Outcome, levels = c('Partnered', 'Lifetime partners', 'Past year partners', 'Age of first sex')),
+    term = factor(term, levels=rev(names(vnames)), labels=rev(vnames))
+  )
+
+plot_allcoefs <-
+  ggplot(d_allstats, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Theme)) +
+  geom_pointrange(size = 0.4, linewidth = 0.7, position = position_dodge(width = 0.8)) +
+  geom_vline(xintercept=0, linetype='longdash') +
+  labs(title="Figure S1: Model coefficients", x='\nEstimate (95% CI)', y='') +
+  facet_wrap(~ Outcome, ncol = 4, scales = "free_x") +
+  theme_minimal() +
+  theme(legend.position = 'bottom', legend.title = element_blank())
+plot_allcoefs
+ggsave("Figures/plot_allcoefs.pdf", plot_allcoefs, width = 14, height = 8)
+
 d_strength_stats <-
   models |>
   mutate(
[…]

### 2024-03-31 analysis.R
Removed many old files

@@ -1,657 +1,0 @@
-library(survey)
-library(ggplot2)
-library(nhanesGH)
-library(hagenutils) # has a version of tidy modified to handle df.resid arg of summary.svyglm
-library(ggpubr)
-
-
-# Anthropometric models ---------------------------------------------------------
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-#visreg(m_lifetime,  xvar = "strength_centered", by = "sex", scale = "response")
-
-#plot(allEffects(m_lifetime))
-
-m_pastyear <- svyglm(
-  sex_partners_year ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered * strength_centered + #keeping partnered x strength interaction only for this model
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-m_agefirst <-  svyglm(
-  age_first_sex ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered  +
-    bmi_centered * sex,
-  family = gaussian(),
-  design = designsG$d.design.adults
-)
-
-m_partnered <-
-  svyglm(
-    partnered ~
-      age_centered * sex +
-      strength_centered * sex +
-      bmi_centered * sex,
-    family = quasibinomial(),
-    design = designsG$d.design.adults
-  )
-summary(m_partnered, df.resid = Inf)
-
-
-
-mnames1 <- c(
-  "Age at first intercourse",
-  "Lifetime Number of Sexual Partners",
-  "Past Year Number of Sexual Partners",
-  "Currently Partnered"
-)
-#
-# vnames1 <- c(
-#   "strength_centered" = "Strength (S)",
-#   "sexfemale" = "Sex (Female)",
-#   "sexfemale:strength_centered" = "Sex (Female) x Strength",
-#   "age_centered" = "Age (S)",
-#   "partneredTRUE" = "Partnered",
-#   "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
-#   "age_centered:sexfemale" = "Age x Sex (Female)",
-#   "height_centered" = "Height (S)",
-#   "weight_centered" = "Weight (S)",
[…]

### 2024-03-31 modelsregressiontables.R

@@ -1,2648 +1,0 @@
-library(survey)
-library(ggplot2)
-library(nhanesGH)
-library(hagenutils)
-library(ggpubr)
-library(tidyverse)
-library(gtsummary)
-
-# mating success models and tbl_regression objects using d.design.adults (NOT restricted to self-identified heterosexuals)
-
-
-# original models (adults with complete data 18-59) -------------------------------------------------
-
-# lifetime partners - baseline/anthropometric
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-# lifetime partners - socioeconomic
-msoc1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered + #* strength_centered + #added strength interaction 03.23 to match past year since lifetime partners is a function of past year; took out 4.6
-      edu +
-      race,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-
-
-# lifetime partners - health model
-mheal1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      perceived_abnormal_weight +
-      whitebloodcell_centered +
-      hemoglobin_centered +
-      special_equipment +
-      chronic_disease_score +
-      physical_disease_count +
-      depression,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-
-# lifetime partners - hormone
-mhor1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      testosterone_sex_centered * sex,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-
-
-# lifetime partners - phys activity
-mphys1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      vigorous_rec +
-      moderate_rec +
-      vigorous_work +
-      moderate_work,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-
-# past year partners - baseline/ anthropometric
-m_pastyear <- svyglm(
-  sex_partners_year ~
-    age_centered * sex +
[…]

### 2024-03-31 old table scripts/svytable.R

@@ -1,80 +1,0 @@
-#figure out how to add ns for male and female
-
-tabbsvy <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age,
-                height,
-                weight,
-                bmi,
-                strength,
-                whitebloodcell,
-                hemoglobin,
-                testosterone),
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  add_n() %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**") %>%
-  add_difference(everything() ~ "smd") %>%
-  modify_column_hide(columns = ci)
-
-tabbsvysd <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age,
-                height,
-                weight,
-                bmi,
-                strength,
-                whitebloodcell,
-                hemoglobin,
-                testosterone),
-    statistic = list(
-      all_continuous() ~ "{sd}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  )
-
-tabb3svy <- tabbsvy %>% modify_table_body(
-  ~ .x %>%
-    dplyr::mutate(sd1 = tabbsvysd$table_body$stat_1)
[…]

### 2024-03-31 old table scripts/tab1long.R

@@ -1,77 +1,0 @@
-
-
-designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-      age,
-      height,
-      weight,
-      bmi,
-      strength,
-      whitebloodcell,
-      hemoglobin,
-      testosterone,
-      depression,
-      chronic_disease_score,
-      physical_disease_count,
-      special_equipment,
-      perceived_abnormal_weight,
-      testosterone,
-      vigorous_work,
-      vigorous_rec,
-      moderate_work,
-      moderate_rec
-    ),
-    statistic = list(all_continuous() ~ "{mean} ({sd})"),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)",
-      depressiion = "Depression Score",
-      perceived_abnormal_weight = "Perceived abnormal weight",
-      special_equipment = "Special equipment needed to walk",
-      chronic_disease_score = "Chronic Disease Score (0-6)",
-      physical_disease_count = "Disease Impairment Score (0-5)",
-      depression = "Depression Score (0-27)",
-      vigorous_work = "Work involves vigorous activity",
-      moderate_work = "Work involves moderate activity",
-      vigorous_rec = "Recreation involves vigorous activity",
-      moderate_rec = "Recreation involves moderate activity"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
-      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
-      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
-      dplyr::relocate(n_male, .before = stat_1) %>%
-      dplyr::relocate(n_female, .before = stat_2)
-  )  %>%
-  add_difference(test = everything() ~ "smd",
-                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
[…]

### 2024-03-31 old table scripts/tab1simple.R

@@ -1,99 +1,0 @@
-getn <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age,
-                height,
-                weight,
-                bmi,
-                strength,
-                whitebloodcell,
-                hemoglobin,
-                testosterone),
-    statistic = list(
-      all_continuous() ~ "{N_nonmiss_unweighted}"
-    ),
-    digits = all_continuous() ~ 0,
-    missing = "no"
-  )
-
-getratio <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age,
-                height,
-                weight,
-                bmi,
-                strength,
-                whitebloodcell,
-                hemoglobin,
-                testosterone),
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  )
-
-
-
-tab1 <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-      age,
-      height,
-      weight,
-      bmi,
-      strength,
-      whitebloodcell,
-      hemoglobin,
-      testosterone
-    ),
-    statistic = list(all_continuous() ~ "{mean} ({sd})"),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
-      dplyr::mutate(n_male = getn$table_body$stat_1, digits = 2) %>%
-      dplyr::mutate(n_female = getn$table_body$stat_2, digits = 2) %>%
-      dplyr::relocate(n_male, .before = stat_1) %>%
[…]

### 2024-03-31 old table scripts/tabl1survey.R

@@ -1,136 +1,0 @@
-# has ratio and cohen's d but not sds
-
-tabbsvy <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age,
-                height,
-                weight,
-                bmi,
-                strength,
-                whitebloodcell,
-                hemoglobin,
-                testosterone),
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  add_n() %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**") %>%
-  add_difference(everything() ~ "smd") %>%
-  modify_column_hide(columns = ci)
-
-
-
-#get sds
-
-  tabbsvysd <- designsG$d.design.adults %>%
-    tbl_svysummary(
-      by = sex,
-      include = c(age,
-                  height,
-                  weight,
-                  bmi,
-                  strength,
-                  whitebloodcell,
-                  hemoglobin,
-                  testosterone),
-      statistic = list(
-        all_continuous() ~ "{sd}"
-      ),
-      digits = all_continuous() ~ 2,
-      label  = list(
-        strength = "Combined Grip Strength (kg)",
-        height = "Height (cm)",
-        weight = "Weight (kg)",
-        age = "Age (years)",
-        bmi = "Body mass index (kg/m^2)",
-        whitebloodcell = "White blood cell count (1000 cells/µL)",
-        hemoglobin = "Hemoglobin (g/dL)",
-        testosterone = "Testosterone (ng/dL)"
-      ),
-      missing = "no"
-    )
-
-  tabb3svy <- tabbsvy %>% modify_table_body(
-    ~ .x %>%
[…]

### 2024-03-31 old table scripts/table1.R

@@ -1,552 +1,0 @@
-
-# Table 1. sex differences using sample data -----------------------------
-
-
-# has ratio and cohen's d but not sds
-tabb <- d_G[d_G$age >= 18 &
-              d_G$age <= 60, ]  %>% select(
-                "sex",
-                "age",
-                "height",
-                "weight",
-                "bmi",
-                "strength",
-                "whitebloodcell",
-                "hemoglobin",
-                "testosterone",
-              ) %>%
-  tbl_summary(
-    by = sex,
-    statistic = list(
-      all_continuous() ~ "{mean}"
-    ),
-    digits = all_continuous() ~ 2,
-    label  = list(
-      strength = "Combined Grip Strength (kg)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      age = "Age (years)",
-      bmi = "Body mass index (kg/m^2)",
-      whitebloodcell = "White blood cell count (1000 cells/µL)",
-      hemoglobin = "Hemoglobin (g/dL)",
-      testosterone = "Testosterone (ng/dL)"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
-      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
-  ) %>%
-  modify_header(ratio = "**Ratio**") %>%
-  add_difference(everything() ~ "cohens_d")
-
-
-#has sds and ratio but not cohen's d
-
-sd_gts <- function(data, variable, ...) {
-  sd(data[[variable]], na.rm = TRUE)
-}
-
-
-tabb2 <- d_G[d_G$age >= 18 &
-               d_G$age <= 60, ]  %>% select(
-                 "sex",
-                 "age",
-                 "height",
-                 "weight",
-                 "bmi",
-                 "strength",
-                 "whitebloodcell",
-                 "hemoglobin",
-                 "testosterone",
-               ) %>%
-  tbl_strata(
-  strata = sex,
-  ~ .x %>%
-    tbl_summary(
- #     by = sex,
-      statistic = list(all_continuous() ~ "{mean}"),
-      digits = all_continuous() ~ 2,
-      label  = list(
-        strength = "Combined Grip Strength (kg)",
-        height = "Height (cm)",
-        weight = "Weight (kg)",
[…]

### 2024-03-31 old/analysis.R

@@ -1,604 +1,0 @@
-library(survey)
-library(tidyverse)
-library(nhanesGH)
-library(visreg)
-library(hagenutils)
-library(patchwork)
-library(foreign)
-library(effects)
-library(gtsummary)
-library(car)
-library(emmeans)
-#library(sjstats)
-
-#remove 10 males who report taking testosterone supplements
-designsG$d.design.adults.noT <- subset(designsG$d.design.adults, TESTOSTERONE == 0)
-
-m <- svyglm(whitebloodcell ~ bmi_centered + strength_centered + log(testosterone)*sex + age_centered*sex, family = gaussian(), design = designsG$d.design.adults.noT)
-summary(m)
-plot(allEffects(m))
-
-m1f <- svyglm(whitebloodcell ~ bmi + strength + testosterone + age, family = gaussian(), design = designsG$d.design.adult.female)
-summary(m1f)
-plot(allEffects(m1f))
-
-
-
-m <- svyglm(whitebloodcell ~ bmi_centered + strength_centered + testosterone*sex + age_centered*sex, family = quasipoisson(), design = designsG$d.design.adults)
-summary(m)
-plot(allEffects(m))
-
-m <- svyglm(whitebloodcell ~ bmi_centered*sex + strength_centered + log(testosterone)*sex + age_centered*sex + race, family = quasipoisson(), design = designsG$d.design.adults)
-summary(m)
-plot(allEffects(m))
-
-m2 <- svyglm(whitebloodcell ~  race*strength + age + bmi, family = gaussian(), design = designsG$d.design.adult.male)
-summary(m2)
-plot(allEffects(m2))
-plot(predictorEffects(m2, ~ strength))
-
-m3 <- svyglm(whitebloodcell ~  strength_centered*sex + race, family = quasipoisson(), design = designsG$d.design.adults.noT)
-summary(m3)
-plot(allEffects(m3))
-
-m2 <- svyglm(whitebloodcell ~ bmi + strength_centered + age, family = quasipoisson(), design = designsG$d.design.adults)
-summary(m2)
-
-m2 <- svyglm(whitebloodcell ~ bmi + strength_centered*sex + age, family = gaussian(), design = designsG$d.design.adults)
-summary(m2)
-
-# partnered status models -------------------------------------------------
-
-
-m <- svyglm(partnered ~ age_centered*sex + strength_centered*sex + edu + bmi, family = binomial, designsG$d.design.adults)
-summary(m)
-plot(allEffects(m))
[…]

### 2024-03-31 old/dailyenergy.R

@@ -1,225 +1,0 @@
-library(fitdistrplus)
-
-# Daily Energy Intake (Table 3) -------------------------------------------
-
-# bone mineral density not collected in G, collected in H in altered sample
-
-# compute occupational expenditure per Lassek and Gaulin
-# we have MET from participant reported occ exertion, a better var than this, also NHANES recommends using
-# diff weights for dietary data
-#
-# occ_exertion <- svyby(~avgcalories, ~census_code_best, design = designsG$d.design.adults, svymean)
-# occ_exertion_dict <- occ_exertion$avgcalories
-# names(occ_exertion_dict) <- occ_exertion$census_code_curr
-# designsG$d.design.adults <- update(designsG$d.design.adults, occ_exertion = occ_exertion_dict[census_code_curr])
-#
-#
-# x <- svyby(~avgcalories, ~race, design = designsG$d.design.adults, svymean, na.rm = T)
-
-# not including in rep because we do not have ffm
-
-
-x <- c(na.omit(d_G$avgcalories))
-x <- c(na.omit(d_G$whitebloodcell))
-
-descdist(x, discrete = T)
-
-
-adults = d_G$age>=18 & d_G$age<=60
-
-d.design.dietary1 <-
-  svydesign(
-    id = ~SDMVPSU ,
-    strata = ~SDMVSTRA ,
-    nest = TRUE ,
-    weights = ~WTMEC2YR, #~WTINT2YR ,
-    data = d_G
-  )
-
-d.design.dietary.adults1 <-
-  subset(
-    d.design.dietary1,
-    adults
-  )
-
-# m <- svyglm(
-#   log(avgcalories) ~
-#     age +
-#     sex * tot_MET +
-#     strength +
-#     weight +
-#     height,
-#   family = gaussian(),
-#   design = d.design.dietary.adults
-# )
-# summary(m)
-# plot(allEffects(m))
-#
-
-
-m <- svyglm(
-  log(avgcalories) ~
-    age +
-    total_work_MET  +
-    total_rec_MET +
-    wob_MET +
-    strength +
-    height +
-    weight,
-  family = gaussian(),
-  design = d.design.dietary.adults1
-)
-summary(m, df.resid = Inf)
-plot(allEffects(m))
-
-m_energy1 <- svyglm( #use this
-  avgcalories ~
-    age +
-    total_work_MET  +
-    total_rec_MET +
-    wob_MET +
-    strength +
-    height +
[…]

### 2024-03-31 old/differencesall.R

@@ -1,176 +1,0 @@
-getn <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age_first_sex,
-                sex_partners,
-                sex_partners_year,
-                strength,
-                age,
-                bmi,
-                height,
-                weight,
-                partnered,
-                edu,
-                race,
-                whitebloodcell,
-                hemoglobin,
-                testosterone,
-                perceived_abnormal_weight,
-                special_equipment,
-                chronic_disease_score,
-                physical_disease_count,
-                depression,
-                vigorous_work,
-                moderate_work,
-                vigorous_rec,
-                moderate_rec,
-                avgcalories),
-    statistic = list(
-      all_continuous() ~ "{N_nonmiss_unweighted}",
-      all_categorical() ~ "{N_nonmiss_unweighted}"
-    ),
-    digits = all_continuous() ~ 0,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
-    missing = "no"
-  )
-
-getratio <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age_first_sex,
-                sex_partners,
-                sex_partners_year,
-                strength,
-                age,
-                bmi,
-                height,
-                weight,
-                partnered,
-                edu,
-                race,
-                whitebloodcell,
-                hemoglobin,
-                testosterone,
-                perceived_abnormal_weight,
-                special_equipment,
-                chronic_disease_score,
-                physical_disease_count,
-                depression,
-                vigorous_work,
-                moderate_work,
-                vigorous_rec,
-                moderate_rec,
-                avgcalories),
-    statistic = list(
-      all_continuous() ~ "{mean}",
-      all_categorical() ~ "{p}"
-    ),
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous",
-      avgcalories ~ "continuous"
-    ),
-    missing = "no"
-  ) %>%
-  modify_table_body(
-    ~ .x %>%
[…]

### 2024-03-31 old/meetings_exploratory.R

@@ -1,487 +1,0 @@
-
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(m_lifetime, df.resid = Inf)
-
-
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    weight_centered * sex +
-    height_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-summary(m_lifetime, df.resid = Inf)
-
-
-
-#04.11.2023
-library(jtools)
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    strength_centered * sex +
-    age_centered * sex +
-    partnered +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-summary(m_lifetime, df.resid = Inf)
-sum <- summary(m_lifetime, df.resid = Inf)$coefficients
-
-summ(m_lifetime, df.resid = Inf)
-
-
-#04.07.2023
-library(hagenutils)
-fem <- svysmooth(whitebloodcell~age, designsG$d.design.adult.female)
-
-mal <- svysmooth(whitebloodcell~age, designsG$d.design.adult.male)
-
-d <- svysmooth2df(female = fem, male = mal)
-
-ggplot(d, aes(age, whitebloodcell, colour = Smooth)) + geom_line()
-
-
-library(mgcv)
-
-m <- gam(whitebloodcell ~ s(strength, by = sex) + s(age, by = sex), data = d_G)
-
-
-#03.31.23
-
-cordat <- d_G[, c("age", "strength", "bmi")]
-cordat <- cordat[cordat$age>=18 & cordat$age<60,]
-cormat <- round(cor(cordat, use = "complete.obs"),2)
-
-
-cordat <- d_Gadults[, c("age", "strength", "bmi", "testosterone")]
-cordat <- cordat[cordat$sex == "male",]
-cormat <- round(cor(cordat, use = "complete.obs"),2)
-
-
-ggplot(d_G) +
-     aes(x = "", y = testosterone) + facet_wrap(~sex) +
-      geom_boxplot(fill = "#0c4c8a") +
[…]

### 2024-03-31 old/old.R

@@ -1,132 +1,0 @@
-mm1_exact <- svyglm(
-  sex_partners ~
-    age_centered + #* sex + #include interaction since we include sex?
-    partnered +
-    sex * strength_centered + #include interaction since we include sex?
-    bmi_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-) %>%
-  tbl_regression(label = list(
-    age_centered = "Age",
-    sex = "Sex",
-    strength_centered = "Grip strength",
-    partnered = "Partnered",
-    bmi_centered = "BMI"
-  )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-
-# original tables ---------------------------------------------------------
-
-mm1_exact <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    sex * strength_centered +
-    partnered +
-    bmi_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-) %>%
-  tbl_regression() %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm1_exact <- svyglm(
-  sex_partners ~
-    age_centered  * sex + #include interaction since we include sex?
-    sex * strength_centered + #include interaction since we include sex?
-    partnered +
-    bmi_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-) %>%
-  # tbl_regression(
-  #                pvalue_fun = ~style_pvalue(.x, digits = 2),
-  # ) %>%
-  #   bold_p(t = 0.05) %>%
-  #   bold_labels() %>%
-  #   modify_header(label ~ "**Variable**") %>%
-  #   add_significance_stars(
-  #     hide_se = TRUE,
-  #     pattern = "{estimate}{stars}<br>({std.error})"
-  #   ) %>%
[…]

### 2024-03-31 old/pre_reg.R

@@ -1,309 +1,0 @@
-library(survey)
-library(tidyverse)
-library(nhanesGH)
-library(visreg)
-library(hagenutils)
-library(patchwork)
-library(foreign)
-library(effects)
-library(gtsummary)
-
-# compute occupational expenditure per Lassek and Gaulin
-# we have MET from participant reported occ exertion, a better var than this, also NHANES recommends using
-# diff weights for dietary data
-#
-# occ_exertion <- svyby(~avgcalories, ~census_code_best, design = designsG$d.design.adults, svymean)
-# occ_exertion_dict <- occ_exertion$avgcalories
-# names(occ_exertion_dict) <- occ_exertion$census_code_curr
-# designsG$d.design.adults <- update(designsG$d.design.adults, occ_exertion = occ_exertion_dict[census_code_curr])
-#
-#
-# x <- svyby(~avgcalories, ~race, design = designsG$d.design.adults, svymean, na.rm = T)
-
-
-# use both number of partners and num vaginal partners, missing 743 who report
-# having vaginal sex but are missing number of vaginal sex partners
-# much fewer (~56) missing number of total sex partners
-# around 686 ppl skipped whole module?
-
-
-# conceptual rep of Lassek and Gaulin, using grip strength and including males and females
-
-# Table 2 (Exact models: Total partners, Partners past year, age at first intercourse)
-
-mm1_exact <- svyglm(
-  sex_partners ~
-    age_centered * strength_centered + # * sex + #include interaction in SI
-    partnered + #* sex +
-    sex * strength_centered + #always include
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(mm1_exact)
-
-
-mm2_exact <- svyglm(
-  sex_partners_year ~
-    # age_centered * strength_centered +
-    sex * strength_centered +
-    partnered*strength_centered +
-    #  partnered*sex +
-    #  bmi_centered*sex  #add to SI
-    bmi_centered*partnered, #leave in
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(mm2_exact)
-
-mm3_exact <- svyglm(
-  age_first_sex ~
-    age_centered +
-    partnered +
[…]

### 2024-03-31 old/spsp_poster.R

@@ -1,531 +1,0 @@
-# spsp poster forestplot
-library(survey)
-library(ggplot2)
-library(nhanesGH)
-library(hagenutils)
-library(ggpubr)
-
-# Lifetime partner models -------------------------------------------------
-
-
-# lifetime partners (anthropometric model)
-manth1 <-
-  svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-   # height_centered * sex +
-   # weight_centered * sex,
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(manth1)
-
-# lifetime partners (socioeconomic model)
-msoc1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      edu +
-      race,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-summary(msoc1, df.resid=Inf)
-
-# lifetime partners (health model)
-mheal1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      perceived_abnormal_weight +
-      whitebloodcell_centered +
-      hemoglobin_centered +
-      special_equipment +
-      chronic_disease_score +
-      physical_disease_count +
-      depression,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-summary(mheal1)
-
-#lifetime partners (hormone model) #log(test) confounded with sex resulting in huge error bars, sub model below
-# mhor1 <-
-#   svyglm(
-#     sex_partners ~
-#       age_centered * sex +
-#       strength_centered * sex +
-#       partnered +
-#      log(testosterone) * sex,
-#     family = quasipoisson(),
-#     design = designsG$d.design.adults
-#   )
-# summary(mhor1)
-
-mhor1a <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      testosterone_sex_centered * sex,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-summary(mhor1a, df = Inf)
-
-#phys activity
-mphys1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
[…]

### 2024-03-31 old/tables.R

@@ -1,87 +1,0 @@
-mm1_exact <- svyglm(
-  sex_partners ~
-    age_centered + #* sex + #include interaction since we include sex?
-    partnered +
-    sex * strength_centered + #include interaction since we include sex?
-    bmi_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-mm2_exact <- svyglm(
-  sex_partners_year ~
-    sex * strength_centered +
-    partnered*strength_centered +
-    bmi_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-mm3_exact <- svyglm(
-  age_first_sex ~
-    age_centered +
-    partnered +
-    edu +
-    total_work_MET +
-    sex*strength_centered,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-
-
-mm1 <- tbl_regression(mm1_exact,
-                      label = list(
-                        age_centered = "Age",
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Partnered",
-                        bmi_centered = "BMI"
-                      )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-
-mm2 <- tbl_regression(mm2_exact,
-                      label = list(
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Partnered",
-                        bmi_centered = "BMI"
-                      )) %>%
-  add_significance_stars(
-    hide_se = TRUE,
-    pattern = "{estimate}{stars}<br>({std.error})"
-  ) %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm3 <- tbl_regression(mm3_exact,
-                      label = list(
-                        age_centered = "Age",
-                        sex = "Sex",
-                        strength_centered = "Grip strength",
-                        partnered = "Partnered",
-                        edu = "Education",
-                        total_work_MET = "Occupational exertion"
-                      )) %>%
-  add_significance_stars(
[…]

### 2024-03-31 participantcharacteristics.R

@@ -1,245 +1,0 @@
-
-getn <- designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(age_first_sex,
-                sex_partners,
-                sex_partners_year,
-                strength,
-                age,
-                bmi,
-                height,
-                weight,
-                partnered,
-                edu,
-                race,
-                whitebloodcell,
-                hemoglobin,
-                testosterone,
-                perceived_abnormal_weight,
-                special_equipment,
-                chronic_disease_score,
-                physical_disease_count,
-                depression,
-                vigorous_work,
-                moderate_work,
-                vigorous_rec,
-                moderate_rec,
-                avgcalories),
-    statistic = list(
-      all_continuous() ~ "{N_nonmiss_unweighted}",
-      all_categorical() ~ "{N_nonmiss_unweighted}"
-    ),
-    digits = all_continuous() ~ 0,
-    type = list(
-      chronic_disease_score ~ "continuous",
-      physical_disease_count ~ "continuous"
-    ),
-    missing = "no"
-  )
-
-designsG$d.design.adults %>%
-  tbl_svysummary(
-    by = sex,
-    include = c(
-      age_first_sex,
-      sex_partners,
-      sex_partners_year,
-      strength,
-      age,
-      bmi,
-      height,
-      weight,
-      partnered,
-      edu,
-      race,
-      whitebloodcell,
-      hemoglobin,
-      testosterone,
-      perceived_abnormal_weight,
-      special_equipment,
-      chronic_disease_score,
-      physical_disease_count,
-      depression,
-      vigorous_work,
-      moderate_work,
-      vigorous_rec,
-      moderate_rec,
-      avgcalories
-    ),
-    label  = list(
-      age_first_sex = "Age at first sex (years)",
-      sex_partners = "Lifetime number of sexual partners",
-      sex_partners_year = "Past year number of sexual partners",
-      strength = "Combined Grip Strength (kg)",
-      age = "Age (Years)",
-      bmi = "Body mass index (kg/m^2)",
-      height = "Height (cm)",
-      weight = "Weight (kg)",
-      partnered = "Partnered",
-      edu = "Education",
-      race = "Race and Ethnicity",
[…]

### 2024-03-31 revisions-exploratory.R

@@ -1,1280 +1,0 @@
-# new exploratory analyses based on reveiwers commments
-
-library(survey)
-library(tidyverse)
-library(nhanesGH)
-library(effects)
-
-# unweighted comparison ---------------------------------------------------
-# we cannot control for all variables accounted for through survey weights (eg response rate and income), and those we can control for (race/ethnicity)
-# are not accounted for in every model and would require completely restructuring our model design. Responded to reviewers that we will retain survey weights
-# but will report an unweighted analysis in the SI
-
-scale2 <- function(x) scale(x)[,1]/2
-
-m_lifetime_uw <- glm(
-  sex_partners ~
-    scale2(age) * sex +
-    scale2(strength) * sex +
-    partnered +
-    race +
-    scale2(bmi) * sex,
-  family = quasipoisson(),
-  data = d_G[d_G$age>=18 & d_G$age<=60,]
-)
-
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    race +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(m_lifetime, df.resid = Inf)
-summary(m_lifetime_uw)
-
-# lifetime partners socieconomic model
-
-msoc1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      edu +
-      race,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-summary(msoc1, df.resid = Inf)
-
-
-msoc1_uw <-
-  glm(
-    sex_partners ~
-      scale2(age) * sex +
-      scale2(strength) * sex +
-      partnered +
-      edu +
-      race,
-    family = quasipoisson(),
-    data = d_G[d_G$age>=18 & d_G$age<=60,]
-  )
-summary(msoc1_uw)
-
-# correlation matrices ----------------------------------------------------
-# reported in SI
-# need sex specific testosterone?
-
-cor_mat_f <- svycor(
-  ~ sex_partners +
-    sex_partners_year +
-    age_first_sex +
-    age +
-    strength +
-    bmi +
-    edu +
-    hemoglobin +
-    whitebloodcell +
-    depression +
-    chronic_disease_score +
[…]

### 2024-03-31 tables_df_pvalues.R

@@ -1,304 +1,0 @@
-
-
-
-# anthropometric model ----------------------------------------------------
-
-
-mm1 <- tbl_regression(
-  m_lifetime,
-  label = vnames
-) %>%
-  add_significance_stars(hide_se = TRUE,
-                         pattern = "{estimate}{stars}<br>({std.error})") %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-
-mm2 <- tbl_regression(
-  m_pastyear,
-  label = vnames
-) %>%
-  add_significance_stars(hide_se = TRUE,
-                         pattern = "{estimate}{stars}<br>({std.error})") %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm3 <- tbl_regression(
-  m_agefirst,
-  label = vnames
-) %>%
-  add_significance_stars(hide_se = TRUE,
-                         pattern = "{estimate}{stars}<br>({std.error})") %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-mm4 <- tbl_regression(
-    m_partnered,
-    label = vnames
-  ) %>%
-  add_significance_stars(hide_se = TRUE,
-                         pattern = "{estimate}{stars}<br>({std.error})") %>%
-  modify_header(label ~ "**Variable**", estimate ~ "**Estimate (SE)**") %>%
-  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
-
-tbl_merge(
-  tbls = list(mm1, mm2, mm3, mm4),
-  tab_spanner = c(
-    "Lifetime number of sexual partners",
-    "Past year number of sexual partners",
-    "Age at first sexual intercourse",
-    "Currently partnered"
-  )
-) %>%
-  as_gt() %>%
-  gt::tab_header(
-    title = "Table 2. Baseline Models of Mating Success",
-    subtitle = "Regression coefficients from generalized
-                 linear models for three indices of mating success for men and women ages
-                 18-59. Models derived from Lassek and Gaulin 2009"
-  )
-
-summary(m_lifetime, df.resid = Inf)
[…]

### 2024-03-31 tmp.R

@@ -1,10 +1,0 @@
-forestplot(
-  mwbc,
-  mwbc_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .8,
-  varnames = vnames)$plot +
-  theme_minimal(25) +
-  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
-  labs(title = "White blood cell count")

### 2024-03-31 unweighted.R

@@ -1,56 +1,0 @@
-scale2 <- function(x) scale(x)[,1]/2
-
-m_lifetime_uw <- glm(
-  sex_partners ~
-    scale2(age) * sex +
-    scale2(strength) * sex +
-    partnered +
-    race +
-    scale2(bmi) * sex,
-  family = quasipoisson(),
-  data = d_G[d_G$age>=18 & d_G$age<=60,]
-)
-
-
-m_lifetime <- svyglm(
-  sex_partners ~
-    age_centered * sex +
-    strength_centered * sex +
-    partnered +
-    race +
-    bmi_centered * sex,
-  family = quasipoisson(),
-  design = designsG$d.design.adults
-)
-summary(m_lifetime, df.resid = Inf)
-summary(m_lifetime_uw)
-
-# lifetime partners socieconomic model
-
-msoc1 <-
-  svyglm(
-    sex_partners ~
-      age_centered * sex +
-      strength_centered * sex +
-      partnered +
-      edu +
-      race,
-    family = quasipoisson(),
-    design = designsG$d.design.adults
-  )
-summary(msoc1, df.resid = Inf)
-
-
-msoc1_uw <-
-  glm(
-    sex_partners ~
-      scale2(age) * sex +
-      scale2(strength) * sex +
-      partnered +
-      edu +
-      race,
-    family = quasipoisson(),
-    data = d_G[d_G$age>=18 & d_G$age<=60,]
-  )
-summary(msoc1_uw)
-

### 2024-04-11 paper.qmd
renamed analysis.R1.R to analysis.R
Added computed power table to ms. code cleanup.

@@
 library(modelsummary)
-source('analysis.R1.R')
+library(tinytable)
+source('analysis.R')
+source('simulate.R')
 ```
@@
 We also predict a significant interaction with sex such that strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
-We also predict that there will be a negative association between formidability and immune function.
-Finally, we predict that formidability will be positively associated with dietary energy intake for both males and females.
+We also predict that there will be a significant negative association between formidability and immune function.
+Finally, we predict that formidability will be positively associated with dietary protein intake for both males and females.
 
@@
 | Age at first sex   |           0          |              0             |
+| Immunity           |           -          |              0             |
+| Energy             |           0          |                            |
+| Protein            |           +          |                            |
 
@@
 A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model.
+A blank indicates not relevant.
 {#tbl-predictions}
@@
 
-We used simulations to estimate our power to detect strength effects that were 75% as large as the ones observed in our pilot study, and sex interaction effects that were equal in magnitude but opposite in sign to the strength effects (i.e., no effect of strength on mating success in women), with a sample size equal to that in our unobserved data.
-We are confident that we have sufficient power to detect true strength effects somewhat less than the ones observed in our pilot study (@tbl-power).
-We are less confident in our power to detect true sex interaction effects because, all else equal, power to detect interaction effects is generally lower than main effects.
+Effects in replication studies are often smaller than in the original studies [@klein2018].
+We therefore used simulations to estimate our power to detect strength and sex X strength coefficients that are 25-100% as large as the ones observed in our pilot study, with a sample size equal to that in our unobserved data.
+We have high power to detect true effects at least 75% as large as the ones observed in our pilot study, and reasonable power to detect those that are at least 50% as large (@tbl-power)
 
-| Outcome            | Strength coefficient | Strength X Sex coefficient |
-|--------------------|:--------------------:|:--------------------------:|
-| Partnered status   |         0.99         |            0.90 (0.67)            |
-| Lifetime partners  |         0.99         |            0.90 (0.66)            |
-| Past year partners |         0.80         |            0.30 (0.21)            |
+:::{#tbl-power}
+```{r}
+v <- ifelse(as.matrix(df_pwr[-1])>0.8, 'red', 'black')
+tt(df_pwr) |>
+  group_tt(j = list("Partnered" = 2:3, "Lifetime partners" = 4:5, "Last year partners" = 6:7)) |>
+  format_tt(j = 1:7, digits = 2, num_zero = T) |>
+  style_tt(j = 1:7, align = c("ccccccc")) |> 
+  style_tt(i=1:4, j=2:7, color = v)
+```
 
-: Power to detect our predicted effects.
-Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients (value in parentheses is power to detect 3/4 of that effect), sample size was set to that of our unobserved data, and $\alpha = 0.05$.
-{#tbl-power}
+Power to detect strength and sex X strength coefficients that range from 25% to 100% of the size seen in the pilot study.
+**Scale**: the proportion of the effect seen in the pilot study.
+All other values are power (1-B).
+Values rounded to two significant digits.
+Red: power > 0.80.
+:::
 

### 2024-04-11 simulate.R

@@ -5,7 +5,7 @@ library(survey)
 library(boot)
 library(furrr)
 
-source("analysis.R")
+# source("analysis.R")
 
 plan("multisession")
 
@@ -16,15 +16,15 @@ rqpois <- function(n, mu, theta) rnbinom(n=n, mu=mu, size=mu/(theta-1))
 
 simdata <- function(
     N,
-    `(Intercept)`=0,
-    sexfemale=0,
-    age_centered=0,
-    partneredTRUE=0,
-    strength_centered=0,
-    sex_partners_scaled=0,
-    `sexfemale:strength_centered`=0,
-    `age_centered:sexfemale`=0,
-    `strength_centered:partneredTRUE`=0,
+    `(Intercept)` = 0,
+    sexfemale = 0,
+    age_centered = 0,
+    partneredTRUE = 0,
+    strength_centered = 0,
+    sex_partners_scaled = 0,
+    `sexfemale:strength_centered` = 0,
+    `age_centered:sexfemale` = 0,
+    `strength_centered:partneredTRUE` = 0,
     theta = 0,
     ...
 ){
@@ -34,20 +34,29 @@ simdata <- function(
     sex = rbinom(N, 1, 0.5),
     age = rnorm(N, 39, 12.5),
     agecentered = c(scale(age))/2,
-    partnered2 = rbinom(N, 1, prob = 0.56),
     strength = ifelse(sex == 1, rnorm(sum(sex), 58.5, 10.76), rnorm(N-sum(sex), 91.99, 17.37)),
-    strengthcentered = ifelse(sex ==1, c(scale(strength[sex==1]))/2,c(scale(strength[sex==0]))/2),
+    strengthcentered = ifelse(sex == 1, c(scale(strength[sex==1]))/2,c(scale(strength[sex==0]))/2),
+    strengthcentered2 = c(scale(strength))/2,
 
     # Models of outcome variables
+    partnered = rbinom(N, 1, prob = inv.logit(
+      `(Intercept)` +
+        sexfemale*sex +
+        age_centered*agecentered +
+        strength_centered*strengthcentered +
+        `sexfemale:strength_centered`*sex*strengthcentered +
+        `age_centered:sexfemale`*sex*agecentered
+    )
+    ),
     sexpartnersyear = rqpois(N, exp(
       `(Intercept)` +
         sexfemale*sex +
         age_centered*agecentered +
-        partneredTRUE*partnered2 +
+        partneredTRUE*partnered +
[…]

### 2024-04-11 paper.qmd
Switched from color to bold

@@
 ```{r}
-v <- ifelse(as.matrix(df_pwr[-1])>0.8, 'red', 'black')
+v <- ifelse(as.matrix(df_pwr[-1])>0.8, T, F)
 tt(df_pwr) |>
@@
 Values rounded to two significant digits.
-Red: power > 0.80.
+Bold: power > 0.80.
 :::

### 2024-05-28 analysis.R
Renamed analysis.R to pilot_analysis.R, and moved some code to functions.R

@@ -1,917 +1,0 @@
-library(tidyverse)
-library(survey)
-library(nhanesGH)
-library(hagenutils)
-library(broom)
-library(gtsummary)
-library(jtools)
-
-# Cumulative distributions of sex partners ------------------------
-
-plot_sexpartners <-
-  ggplot(d_G, aes(sex_partners + 1, colour = sex)) +
-  stat_ecdf() +
-  geom_vline(xintercept = 100, linetype = 'dotted') +
-  annotate("text", label = 'Cutoff', x = 110, y = 0.80, hjust = 0) +
-  scale_x_log10(breaks = c(1, 5, 10, 25, 50, 100, 1000)) +
-  scale_color_binary() +
-  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
-  labs(x = "\nNumber of lifetime sex partners + 1", y = 'ECDF\n') +
-  theme_minimal(15)
-plot_sexpartners
-ggsave("Figures/plot_sexpartners.pdf", plot_sexpartners, width = 9, height = 6)
-
-plot_sexpartners_year <-
-  ggplot(d_G, aes(sex_partners_year + 1, colour = sex)) +
-  stat_ecdf() +
-  scale_x_log10(breaks = c(1, 5, 10, 25, 50)) +
-  scale_color_binary() +
-  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
-  labs(x = "\nNumber of last year sex partners + 1", y = 'ECDF\n') +
-  theme_minimal(15)
-plot_sexpartners_year
-# ggsave("Figures/plot_sexpartners_year.pdf", plot_sexpartners_year, width = 9, height = 6)
-
-# Update design objects -----------------------------------
-
-design_all <- designsG$d.design.adults # Entire sample
-
-# Removing those with >=100 lifetime sex partners
-designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
-
-# Scale sex_partners by the interquartile range
-designsG$d.design.adults <- update(designsG$d.design.adults, sex_partners_scaled = sex_partners/10)
-
-# Difference between current age and age at first sex
-designsG$d.design.adults <- update(designsG$d.design.adults, age_diff = age - age_first_sex)
-
-# Create numeric version of sex for correlation matrix
-designsG$d.design.adults <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
-
-# set strength_centered2 to the values standardized across both sexes
[…]

### 2024-05-28 functions.R

@@ -1,0 +1,544 @@
+library(tidyverse)
+library(survey)
+library(nhanesGH)
+library(broom)
+library(modelsummary)
+
+name_dict <- c(
+  manth = 'Anthropomentric',
+  mheal = 'Health',
+  mhor = 'Hormone',
+  mphys = 'Activity',
+  msoc = 'Socioeconomic'
+)
+
+outcome_dict <- c(
+  "1" = "Lifetime partners",
+  "2" = "Past year partners",
+  "3" = "Age of first sex",
+  "4" = "Partnered"
+)
+
+# Coef plot labels --------------------------------------------------------------
+
+mnames1 <- c(
+  "Age at first intercourse",
+  "Lifetime Number of Sexual Partners",
+  "Past Year Number of Sexual Partners",
+  "Currently Partnered"
+)
+
+mnames <- c(
+  "Anthropometric",
+  "Socieoeconomic",
+  "Health",
+  "Physical Activity",
+  "Hormone"
+)
+
+mmnames <- c(
+  "Replication",
+  "Expanded Controls"
+)
+
+vnames <- c(
+  "strength_centered" = "Strength (S by sex)",
+  "strength_centered2" = "Strength (S)",
+  "strength" = "Strength",
+  "sexfemale" = "Sex (Female)",
+  "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "sexfemale:strength_centered2" = "Sex (Female) x Strength",
+  "age_centered" = "Age (S)",
+  "age" = "Age",
+  "partneredTRUE" = "Partnered",
+  "sex_partners" = "Lifetime sex partners",
+  "sex_partners_scaled" = "Lifetime sex partners (S)",
+  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
+  "age_centered:sexfemale" = "Age (S) x Sex (Female)",
+  "height_centered" = "Height (S)",
+  "weight_centered" = "Weight (S)",
+  "bmi_centered" = "BMI (S)",
+  "bmi" = "BMI",
+  "sexfemale:bmi_centered" = "Sex (Female) x BMI (S)",
+  "sexfemale:height_centered" = "Sex (Female) x Height (S)",
+  "sexfemale:weight_centered" = "Sex (Female) x Weight (S)",
+  'perceived_abnormal_weightTRUE' = "Perceived Abnormal Weight",
+  "whitebloodcell_centered" = "White Blood Cell Count (S)",
[…]

### 2024-05-28 pilot_analysis.R

@@ -1,0 +1,401 @@
+
+library(hagenutils)
+library(gtsummary)
+library(jtools)
+library(marginaleffects)
+library(tinytable)
+
+# Cumulative distributions of sex partners ------------------------
+
+plot_sexpartners <-
+  ggplot(d_G[!is.na(d_G$sex_partners),], aes(sex_partners + 1, colour = sex)) +
+  stat_ecdf() +
+  geom_vline(xintercept = 100, linetype = 'dotted') +
+  annotate("text", label = 'Cutoff', x = 110, y = 0.80, hjust = 0) +
+  scale_x_log10(breaks = c(1, 5, 10, 25, 50, 100, 1000)) +
+  scale_color_binary() +
+  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
+  labs(x = "\nNumber of lifetime sex partners + 1", y = 'ECDF\n') +
+  theme_minimal(15)
+plot_sexpartners
+ggsave("Figures/plot_sexpartners.pdf", plot_sexpartners, width = 9, height = 6)
+
+plot_sexpartners_year <-
+  ggplot(d_G[!is.na(d_G$sex_partners_year),], aes(sex_partners_year + 1, colour = sex)) +
+  stat_ecdf() +
+  scale_x_log10(breaks = c(1, 5, 10, 25, 50)) +
+  scale_color_binary() +
+  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
+  labs(x = "\nNumber of last year sex partners + 1", y = 'ECDF\n') +
+  theme_minimal(15)
+plot_sexpartners_year
+# ggsave("Figures/plot_sexpartners_year.pdf", plot_sexpartners_year, width = 9, height = 6)
+
+# Update pilot design objects -----------------------------------
+
+design_all <- designsG$d.design.adults # Entire sample
+
+# Removing those with >=100 lifetime sex partners
+designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
+
+# Scale sex_partners by the interquartile range
+designsG$d.design.adults <- update(designsG$d.design.adults, sex_partners_scaled = sex_partners/10)
+
+# Difference between current age and age at first sex
+designsG$d.design.adults <- update(designsG$d.design.adults, age_diff = age - age_first_sex)
+
+# Create numeric version of sex for correlation matrix
+designsG$d.design.adults <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
+
[…]

### 2024-05-28 simulate.R
Minor tweaks

@@ -1,7 +1,7 @@
 #+ message=F, warning=F
-library(tidyverse)
-library(nhanesGH)
-library(survey)
+# library(tidyverse)
+# library(nhanesGH)
+# library(survey)
 library(boot)
 library(furrr)
 
@@ -113,9 +113,9 @@ pwr <- function(N, model, getstatsfunction, outcome, scale_effect = 1){
 # Sample size
 N <- sum(d_H$sex_partners <= 100, na.rm = T)
 
-pwr_partnered <- map(seq(0.25, 1, 0.25), \(x) pwr(N, m_partnered, getstats_partnered, 'Partnered', scale_effect = x)) |> list_rbind()
-pwr_lifetime <- map(seq(0.25, 1, 0.25), \(x) pwr(N, m_lifetime, getstats_life, 'Lifetime partners', scale_effect = x)) |> list_rbind()
-pwr_pastyear <- map(seq(0.25, 1, 0.25), \(x) pwr(N, m_pastyear, getstats_year, 'Last year partners', scale_effect = x)) |> list_rbind()
+pwr_partnered <- map(seq(0.25, 1, 0.25), \(x) pwr(N, models$Model$manth4, getstats_partnered, 'Partnered', scale_effect = x)) |> list_rbind()
+pwr_lifetime <- map(seq(0.25, 1, 0.25), \(x) pwr(N, models$Model$manth1, getstats_life, 'Lifetime partners', scale_effect = x)) |> list_rbind()
+pwr_pastyear <- map(seq(0.25, 1, 0.25), \(x) pwr(N, models$Model$manth2, getstats_year, 'Last year partners', scale_effect = x)) |> list_rbind()
 
 df_pwr <- bind_cols(
   Scale = seq(0.25, 1, 0.25),

### 2024-05-28 paper.qmd
Removed one plot, added another per reviewer request. Tweaked some code.

@@
 library(ggcorrplot)
-library(modelsummary)
 library(tinytable)
-source('analysis.R')
+library(patchwork)
+source('functions.R')
+source('pilot_analysis.R')
 source('simulate.R')
+source('sexspecificmodels.R')
 ```
@@
 Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
-We will also test if there is a positive relationship between grip strength and dietary energy intake, and a negative relationship between grip strength and native immune function.
+We will also test if there is a positive relationship between grip strength and dietary energy and protein intake, and a negative relationship between grip strength and native immune function.
 This study replicates an earlier study of US males [@lassek2009] with numerous additional control variables, and it additionally tests these relationships in females.
@@
 Importantly, studies of contemporary hunter-gatherers and game-theoretic models find that a mutually beneficial sexual division of labor within long-term pair bonds, in which women engage in lower risk, lower return foraging of plants and small game compatible with the cognitively demanding task of childrearing [@hagen2019], and men engage in higher risk, higher return big game hunting [@kelly2013], is critical to provisioning offspring who remain dependent on caregivers for 20 years or more [@davison2021; @davison2022; @alger2020; @kaplan2000].
-See @fig-provisioning.
 This suggests that sex differences in body composition could be due, at least in part, to a sexual division of labor.
-
-:::{#fig-provisioning}
-![](Figures/plot_productivity.png)
-
-Net energy production in chimpanzees and human foragers over the lifespan. Solid lines are females. Dashed line is Hadza males. Data from @davison2021 and @davison2022.
-:::
-
 More generally, multiple forms of sexual selection could have played a role in the evolution of human sexual dimorphism [see @hill2017 for a brief review of alternative hypotheses for sexual strength dimorphism].
@@
 
-According to the sexual selection hypothesis, males evolved greater physical formidability than females because physical formidability increased male reproductive success more than females.
+According to the sexual selection hypothesis, males evolved greater physical formidability than females because physical formidability increased the mating success of males more than females.
 The aim of this study is to replicate the costs and benefits of formidability reported by @lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
@@
 The pilot study uses the 2011-2012 dataset, whereas the 2013-2014 dataset has been held out for confirmatory analysis pending in principle acceptance of this registered report.
-Grip strength, our proxy for physical formidability and our key predictor variable, was only collected in these years.
+Grip strength, our proxy for physical formidability and our key predictor variable, was only collected in these years [@smith2022 investigated the association of grip strength with depression in the 2013-14 data, but did not investigate or observe the relationship between grip strength and any of the mating success outcomes described below].
 We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable [@hogrel2015].
@@
 Bars are 95\% CIs.
-For the full models, see the SI.
+For the full regression tables, see the SI.
 
 plot_coefs
+```
+
+For a plot of all coefficients, see @fig-allcoefs.
+
+```{r}
+#| label: fig-allcoefs
+#| fig-cap: Coefficients from all models of our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95\% CIs.
+For the full regression tables, see the SI.
+
+plot_allcoefs
 ```
@@
 #| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot). Variables labeled (S) have been centered at the mean and standardized by 2 SD.
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
+Variables labeled (S) have been centered at the mean and standardized by 2 SD.
+Bars are 95% CIs.
 
 plot_immune <- forestplot(
-  mwbc,
-  mwbc_alt,
+  immune_models$mwbc,
+  immune_models$mwbc_alt,
   intercept = F,
@@
 
-Finally, we tested whether strength was related to increased energy (@fig-energy) or protein intake (@fig-protein).
+Finally, we tested whether strength was related to increased energy or protein intake (@fig-energy).
 We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
@@
 #| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) and dietary protein (g) from generalized linear models (Pilot).
+Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+Bars are 95% CIs.
 
 plot_energy <- forestplot(
-  m_energy,
-  m_energy_alt,
+  intake_models$m_energy,
+  intake_models$m_energy_alt,
   intercept = F,
@@
   scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-
-ggsave("Figures/plot_energy.pdf", plot_energy, width = 14, height = 7)
-
-plot_energy 
-```
-
-```{r}
-#| label: fig-protein
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of dietary protein intake from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+  ggtitle('Energy') +
+  theme_bw(17) +
+  guides(colour = guide_none(), shape = guide_none())
 
 plot_protein <- forestplot(
-  m_protein,
-  m_protein_alt,
+  intake_models$m_protein,
+  intake_models$m_protein_alt,
   intercept = F,
@@
   scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
+  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
+  ggtitle('Protein') +
+  theme_bw(17) +
+  theme(axis.text.y = element_blank())
 
-ggsave("Figures/plot_protein.pdf", plot_protein, width = 14, height = 7)
-
-plot_protein 
+p <- plot_energy + plot_protein + plot_layout(axes = 'collect') 
+ggsave("Figures/plot_energy_protein.pdf", p, width = 12, height = 6)
+p
 ```
@@
 
-Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with three of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-The only exceptions were that strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant strength coefficients in the other models.
-Strength was also associated with immune, and energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
+Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with two of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
+The exceptions were that strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant strength coefficients in the other models.
+Strength was also associated with immune, energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
 The expanded control models included height, which might be a proxy for aspects of strength, such as lower body strength, that are not accounted for by grip strength.
-
-The effect of strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
-Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
-In addition, participants, whose grip strength was measured at the time of the study when they were adults, are reporting on an event that happened an average of 22 years in the past when most were adolescents and their grip strength had not yet reached its adult value.
-For these reasons, we do not predict an effect of strength on age at first sex after controlling for ethnicity.
 
@@
 
-The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of partners in the last year) is puzzling.
+The effect of strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
+Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
+In addition, participants, whose grip strength was measured at the time of the study when they were adults, are reporting on an event that happened an average of 22 years in the past when most were adolescents and their grip strength had not yet reached its adult value.
+For these reasons, we do not predict an effect of strength on age at first sex after controlling for ethnicity.
+
+Strength was also not significantly associated with the number of partners in the last year, despite controlling for partnered status, which was a significant negative predictor.
+Approximately 60% of the sample was partnered (@stbl-stats), however, and stronger individuals, especially men, were more likely to be partnered.
+The pool of unpartnered individuals who might pursue multiple mating opportunities was therefore smaller and skewed toward lower strength individuals, and variation in this outcome over the relatively short time period (1 year) was low.
+Nevertheless, the non-significant strength coefficients were all positive, raising the possibility that a higher powered study might successfully detect a positive effect of strength on number of partners in the last year.
+
+<!-- The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of partners in the last year) is puzzling.
 Because men have greater strength than women their self-reported partner numbers are higher, on average, although the median number of partners in the last year for both sexes is 1.
@@
 Finally, our simulations revealed that we had relatively lower power to detect a significant interaction between strength and sex for this outcome (see Power section below).
+-->
 
@@
 Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
-More specifically, humans require provisioning for an extended period [@davison2021; @davison2022] (@fig-provisioning), and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
+More specifically, humans require provisioning for an extended period [@kaplan2000; @davison2021; @davison2022], and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
 Thus, our pilot study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.
-Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014].
+Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for protection from predators or other men in long-term relationships is possible too [@willems2017b; @kelly2005].
 
@@
 Effects in replication studies are often smaller than in the original studies [@klein2018].
-We therefore used simulations to estimate our power to detect strength and sex X strength coefficients that are 25-100% as large as the ones observed in our pilot study, with a sample size equal to that in our unobserved data.
-We have high power to detect true effects at least 75% as large as the ones observed in our pilot study, and reasonable power to detect those that are at least 50% as large (@tbl-power)
-
-:::{#tbl-power}
-```{r}
-v <- ifelse(as.matrix(df_pwr[-1])>0.8, T, F)
-tt(df_pwr) |>
-  group_tt(j = list("Partnered" = 2:3, "Lifetime partners" = 4:5, "Last year partners" = 6:7)) |>
-  format_tt(j = 1:7, digits = 2, num_zero = T) |>
-  style_tt(j = 1:7, align = c("ccccccc")) |> 
-  style_tt(i=1:4, j=2:7, color = v)
-```
-
-Power to detect strength and sex X strength coefficients that range from 25% to 100% of the size seen in the pilot study.
-**Scale**: the proportion of the effect seen in the pilot study.
-All other values are power (1-B).
-Values rounded to two significant digits.
-Bold: power > 0.80.
-:::
+We therefore used simulations to estimate our power to detect strength and sex X strength coefficients that are 25-100% as large as the ones observed in our pilot study for partnered and lifetime partners, with a sample size equal to that in our unobserved data.
+We have high power to detect effects at least 75% as large as the ones observed in our pilot study, and reasonable power to detect those that are at least 50% as large (@stbl-power).
 
@@
 Our methods and models will remain identical to those reported in the pilot study, using unobserved data from the 2013-2014 NHANES collection cycle.
+After testing and reporting our predictions using the unobserved data, we will fit our models on the combined data (2011-2014), and conduct other post hoc analyses.
 
@@
 
+## Power
+
+:::{#stbl-power}
+```{r}
+v <- ifelse(as.matrix(df_pwr[-1])>0.8, T, F)
+tt(df_pwr) |>
+  group_tt(j = list("Partnered" = 2:3, "Lifetime partners" = 4:5, "Last year partners" = 6:7)) |>
+  format_tt(j = 1:7, digits = 2, num_zero = T) |>
+  style_tt(j = 1:7, align = c("ccccccc")) |> 
+  style_tt(i=1:4, j=2:7, bold = v)
+```
+
+Power to detect strength and sex X strength coefficients that range from 25% to 100% of the size seen in the pilot study, based on models with anthropometric controls.
+**Scale**: the proportion of the effect seen in the pilot study.
+All other values are power (1-B).
+Values rounded to two significant digits.
+Bold: power > 0.80.
+:::
+
 ## Pilot study
+
+Descriptive statistics.
 
@@
 
-### Models of mating success with anthropometric, socioeconomic, health, hormone, and physical activity controls
+<!-- ### Models of mating success with anthropometric, socioeconomic, health, hormone, and physical activity controls -->
 
-::: {#sfig-allcoefs}
-```{r}
-plot_allcoefs
-```
-
-Coefficients from all models.
-:::
-
-### The negative association of numbers of lifetime sexual partners on partnered status
+### The negative association of numbers of lifetime sexual partners with partnered status
 
@@
 
+### Marginal effects
+
+::: {#stab-marginal}
+```{r}
+marginal |>
+  tt() |>
+  format_tt(fn = hagenutils::signif2) |>
+  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
+```
+
+The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
+:::
+
 ### Correlation matrices of study variables
@@
 ```{r}
-cor_bothsexes <- ggcorrplot(cor_mat_cors, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
-ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes)
+
+cor_bothsexes <- ggcorrplot(cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
+ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes, width = 12, height = 12)
 cor_bothsexes
@@
 ```{r}
-cor_females <- ggcorrplot(cor_mat_cors_f, hc.order = T, hc.method = 'ward.D', title = 'Females')
-ggsave("Figures/cor_females.pdf", cor_females)
+
+cor_females <- ggcorrplot(cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Females')
+ggsave("Figures/cor_females.pdf", cor_females, width = 12, height = 12)
 cor_females
@@
 ```{r}
-cor_males <- ggcorrplot(cor_mat_cors_m, hc.order = T, hc.method = 'ward.D', title = 'Males')
-ggsave("Figures/cor_males.pdf", cor_males)
+
+cor_males <- ggcorrplot(cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Males')
+ggsave("Figures/cor_males.pdf", cor_males, width = 12, height = 12)
 cor_males
@@
 
-**Lifetime number of partners**
-
-:::{#stbl-lifetime}
 ```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_lifetime, 
-    "Socioeconomic" = msoc1, 
-    "Health" = mheal1, 
-    "Activity" = mphys1,
-    "Hormone" = mhor1
-    ), 
-  gof_map = "nobs"
-  )
+#| results: asis
+for (i in 1:nrow(modelsummaries)){
+  cat(paste('####', modelsummaries$title[i]))
+  print(modelsummaries$summary[[i]])
+}
 ```
 
-Quasi-Poisson Regression Models of Lifetime Number of Sexual Partners
+### Sex-specific models
+
+At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
+
+:::{#sfig-sexspecific}
+```{r}
+#| fig-width: 7
+#| fig-height: 9
+plot_sexspecific_strength_coefs
+```
+
+Strength coefficients in each regression model fit separately by sex.
+Control variables are the same as for models fit with a sex X strength interaction term.
+Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
 :::
 
-**Past year number of partners**
+Regression tables for sex-specific models of the effect of strength on mating success.
 
-:::{#stbl-pastyear}
 ```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_pastyear, 
-    "Socioeconomic" = msoc2, 
-    "Health" = mheal2, 
-    "Activity" = mphys2,
-    "Hormone" = mhor2
-    ), 
-  gof_map = "nobs"
-  )
+#| results: asis
+for (i in 1:nrow(dfmodelsummaries)){
+  cat(paste('####', dfmodelsummaries$title[i]))
+  print(dfmodelsummaries$summary[[i]])
+}
 ```
 
-Quasi-Poisson Regression Models of Past Year Number of Sexual Partners
-:::
-
-**Age at first sex**
-
-:::{#stbl-agefirstsex}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_agefirst, 
-    "Socioeconomic" = msoc3, 
-    "Health" = mheal3, 
-    "Activity" = mphys3,
-    "Hormone" = mhor3
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Gaussian Regression Models of Age at First Sex
-:::
-
-**Partnered status**
-
-:::{#stbl-partnered}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_partnered, 
-    "Socioeconomic" = msoc4, 
-    "Health" = mheal4, 
-    "Activity" = mphys4,
-    "Hormone" = mhor4
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Logistic Regression Models of Partnered Status
-:::
-
 # References

### 2024-05-28 sexspecificmodels.R
Added sex-specific regression models, per reviewer request.

@@ -1,0 +1,568 @@
+# library(tidyverse)
+# library(nhanesGH)
+# library(survey)
+# library(broom)
+
+# These models were fit at the request of an anonymous reviewer
+
+# Removing those with >=100 lifetime sex partners
+designsG$d.design.adult.female <- subset(designsG$d.design.adult.female, sex_partners < 100)
+designsG$d.design.adult.male <- subset(designsG$d.design.adult.male, sex_partners < 100)
+
+# Scale sex_partners by the interquartile range of all adults
+designsG$d.design.adult.female <- update(designsG$d.design.adult.female, sex_partners_scaled = sex_partners/10)
+designsG$d.design.adult.male <- update(designsG$d.design.adult.male, sex_partners_scaled = sex_partners/10)
+
+sexspecificmodels <- list(
+  # Lifetime partners models -------------------------------------------------
+
+  # Anthropometric
+  manth_1_f = svyglm(
+    sex_partners ~
+      age +
+      strength +
+      partnered +
+      bmi,
+    family = quasipoisson(),
+    design = designsG$d.design.adult.female
+  ),
+
+  manth_1_m = svyglm(
+    sex_partners ~
+      age +
+      strength +
+      partnered +
+      bmi,
+    family = quasipoisson(),
+    design = designsG$d.design.adult.male
+  ),
+
+  # Socioeconomic model
+  msoc_1_f = svyglm(
+      sex_partners ~
+        age +
+        strength +
+        partnered +
+        edu +
+        race,
+      family = quasipoisson(),
+      design = designsG$d.design.adult.female
+    ),
+
+  msoc_1_m = svyglm(
+      sex_partners ~
+        age +
+        strength +
+        partnered +
+        edu +
+        race,
+      family = quasipoisson(),
+      design = designsG$d.design.adult.male
+    ),
+
+  # Health model
+  mheal_1_f = svyglm(
+      sex_partners ~
+        age +
+        strength +
+        partnered +
+        perceived_abnormal_weight +
+        whitebloodcell +
+        hemoglobin +
+        special_equipment +
+        chronic_disease_score +
+        physical_disease_count +
+        depression,
+      family = quasipoisson(),
+      design = designsG$d.design.adult.female
+    ),
+
+  mheal_1_m = svyglm(
+      sex_partners ~
+        age +
+        strength +
+        partnered +
+        perceived_abnormal_weight +
+        whitebloodcell +
+        hemoglobin +
+        special_equipment +
+        chronic_disease_score +
[…]

### 2024-06-03 paper.qmd
Preparing stage 2 registered report. Moved library calls and design object update code to functions.R. Compute years_sexually_mature. Move age term to offset in poisson models.

@@
 ```{r}
-library(ggcorrplot)
-library(tinytable)
-library(patchwork)
 source('functions.R')
@@
 source('sexspecificmodels.R')
+source('stage2_analysis.R')
 ```
@@
 
+### Effects plots
+
+:::{#sfig-effects}
+```{r}
+effects_plots(models)
+```
+
+Effects of sex-specific standardized strength on mating outcomes, by sex.
+Strength is centered and scaled by 2 SD, separately for each sex.
+Plots are based on the anthropometric control models, with control variables held at their mean values.
+:::
+
 ### Correlation matrices of study variables

### 2024-06-03 functions.R

@@ -3,9 +3,16 @@ library(survey)
 library(nhanesGH)
 library(broom)
 library(modelsummary)
+library(hagenutils)
+library(gtsummary)
+library(jtools)
+library(marginaleffects)
+library(tinytable)
+library(patchwork)
+library(ggcorrplot)
 
 name_dict <- c(
-  manth = 'Anthropomentric',
+  manth = 'Anthropometric',
   mheal = 'Health',
   mhor = 'Hormone',
   mphys = 'Activity',
@@ -47,6 +54,7 @@ vnames <- c(
   "strength" = "Strength",
   "sexfemale" = "Sex (Female)",
   "sexfemale:strength_centered" = "Sex (Female) x Strength",
+  "strength_centered:sexfemale" = "Sex (Female) x Strength",
   "sexfemale:strength_centered2" = "Sex (Female) x Strength",
   "age_centered" = "Age (S)",
   "age" = "Age",
@@ -91,6 +99,37 @@ vnames <- c(
   "avgcalories_centered" = "Average calories per day (S)"
 )
 
+# Update design objects -----------------------------------
+
+
+update_designs <- function(designs, cutoff = 100){
+
+  # Removing those with >=100 lifetime sex partners
+  designs$d.design.adults <- subset(designs$d.design.adults, sex_partners < cutoff)
+
+  # Scale sex_partners by the interquartile range
+  designs$d.design.adults <- update(designs$d.design.adults, sex_partners_scaled = sex_partners/10)
+
+  # Difference between current age and age at first sex
+  designs$d.design.adults <- update(designs$d.design.adults, age_diff = age - age_first_sex)
+
+  # Create numeric version of sex for correlation matrix
+  designs$d.design.adults <- update(designs$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
+
+  # set strength_centered2 to the values standardized across both sexes
+  designs$d.design.dietary.adults <- update(designs$d.design.dietary.adults, strength_centered2 = strength_centered)
+  designs$d.design.adults <- update(designs$d.design.adults, strength_centered2 = strength_centered)
+
+  # Set strength_centered to the sex-specific values
+  designs$d.design.adults <- update(designs$d.design.adults, strength_centered = strength_sex_centered)
+
+  designs$d.design.adults <- update(designs$d.design.adults, years_sexually_mature = age - 12)
+
+  return(designs)
+}
+
+
+
 # Mating regression models -------------------------------------------------------
 
 fitmodels <- function(design){
[…]

### 2024-06-03 pilot_analysis.R

@@ -1,9 +1,4 @@
 
-library(hagenutils)
-library(gtsummary)
-library(jtools)
-library(marginaleffects)
-library(tinytable)
 
 # Cumulative distributions of sex partners ------------------------
 
@@ -33,27 +28,7 @@ plot_sexpartners_year
 
 # Update pilot design objects -----------------------------------
 
-design_all <- designsG$d.design.adults # Entire sample
-
-# Removing those with >=100 lifetime sex partners
-designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
-
-# Scale sex_partners by the interquartile range
-designsG$d.design.adults <- update(designsG$d.design.adults, sex_partners_scaled = sex_partners/10)
-
-# Difference between current age and age at first sex
-designsG$d.design.adults <- update(designsG$d.design.adults, age_diff = age - age_first_sex)
-
-# Create numeric version of sex for correlation matrix
-designsG$d.design.adults <- update(designsG$d.design.adults, sex2 = ifelse(sex == "male", 1, 0))
-
-# set strength_centered2 to the values standardized across both sexes
-designsG$d.design.dietary.adults <- update(designsG$d.design.dietary.adults, strength_centered2 = strength_centered)
-designsG$d.design.adults <- update(designsG$d.design.adults, strength_centered2 = strength_centered)
-
-# Set strength_centered to the sex-specific values
-designsG$d.design.adults <- update(designsG$d.design.adults, strength_centered = strength_sex_centered)
-
+designsG <- update_designs(designsG)
 
 # Fit models --------------------------------------------------------------
 

### 2024-06-03 sexspecificmodels.R

@@ -13,13 +13,17 @@ designsG$d.design.adult.male <- subset(designsG$d.design.adult.male, sex_partner
 designsG$d.design.adult.female <- update(designsG$d.design.adult.female, sex_partners_scaled = sex_partners/10)
 designsG$d.design.adult.male <- update(designsG$d.design.adult.male, sex_partners_scaled = sex_partners/10)
 
+# Sexually mature
+designsG$d.design.adult.female <- update(designsG$d.design.adult.female, years_sexually_mature = age - 12)
+designsG$d.design.adult.male <- update(designsG$d.design.adult.male, years_sexually_mature = age - 12)
+
 sexspecificmodels <- list(
   # Lifetime partners models -------------------------------------------------
 
   # Anthropometric
   manth_1_f = svyglm(
     sex_partners ~
-      age +
+      offset(log(years_sexually_mature)) +
       strength +
       partnered +
       bmi,
@@ -29,7 +33,7 @@ sexspecificmodels <- list(
 
   manth_1_m = svyglm(
     sex_partners ~
-      age +
+      offset(log(years_sexually_mature)) +
       strength +
       partnered +
       bmi,
@@ -40,7 +44,7 @@ sexspecificmodels <- list(
   # Socioeconomic model
   msoc_1_f = svyglm(
       sex_partners ~
-        age +
+        offset(log(years_sexually_mature)) +
         strength +
         partnered +
         edu +
@@ -51,7 +55,7 @@ sexspecificmodels <- list(
 
   msoc_1_m = svyglm(
       sex_partners ~
-        age +
+        offset(log(years_sexually_mature)) +
         strength +
         partnered +
         edu +
@@ -63,7 +67,7 @@ sexspecificmodels <- list(
   # Health model
   mheal_1_f = svyglm(
       sex_partners ~
-        age +
+        offset(log(years_sexually_mature)) +
         strength +
         partnered +
         perceived_abnormal_weight +
@@ -79,7 +83,7 @@ sexspecificmodels <- list(
 
   mheal_1_m = svyglm(
       sex_partners ~
-        age +
+        offset(log(years_sexually_mature)) +
         strength +
         partnered +
         perceived_abnormal_weight +
@@ -95,7 +99,7 @@ sexspecificmodels <- list(
 
   mhor_1_f = svyglm(
       sex_partners ~
-        age +
+        offset(log(years_sexually_mature)) +
         strength +
         partnered +
         testosterone,
[…]

### 2024-06-03 simulate.R

@@ -1,11 +1,5 @@
-#+ message=F, warning=F
-# library(tidyverse)
-# library(nhanesGH)
-# library(survey)
 library(boot)
 library(furrr)
-
-# source("analysis.R")
 
 plan("multisession")
 
@@ -23,17 +17,22 @@ simdata <- function(
     strength_centered = 0,
     sex_partners_scaled = 0,
     `sexfemale:strength_centered` = 0,
+    `strength_centered:sexfemale` = 0,
     `age_centered:sexfemale` = 0,
     `strength_centered:partneredTRUE` = 0,
     theta = 0,
     ...
 ){
+
+  SEXSTRENGTH <- ifelse(length(`sexfemale:strength_centered`) > 0, `sexfemale:strength_centered`, `strength_centered:sexfemale`)
+
   tibble(
 
     # Explanatory variables
     sex = rbinom(N, 1, 0.5),
-    age = rnorm(N, 39, 12.5),
+    age = sample(d_adult$age, N, replace = T),
     agecentered = c(scale(age))/2,
+    years_sexually_mature = age - 12,
     strength = ifelse(sex == 1, rnorm(sum(sex), 58.5, 10.76), rnorm(N-sum(sex), 91.99, 17.37)),
     strengthcentered = ifelse(sex == 1, c(scale(strength[sex==1]))/2,c(scale(strength[sex==0]))/2),
     strengthcentered2 = c(scale(strength))/2,
@@ -44,7 +43,7 @@ simdata <- function(
         sexfemale*sex +
         age_centered*agecentered +
         strength_centered*strengthcentered +
-        `sexfemale:strength_centered`*sex*strengthcentered +
+        SEXSTRENGTH*sex*strengthcentered +
         `age_centered:sexfemale`*sex*agecentered
     )
     ),
@@ -54,7 +53,7 @@ simdata <- function(
         age_centered*agecentered +
         partneredTRUE*partnered +
         strength_centered*strengthcentered +
-        `sexfemale:strength_centered`*sex*strengthcentered +
+        SEXSTRENGTH*sex*strengthcentered +
         `age_centered:sexfemale`*sex*agecentered +
         `strength_centered:partneredTRUE`*partnered*strengthcentered),
       theta
@@ -62,9 +61,9 @@ simdata <- function(
     sexpartners = rqpois(N, exp(
       `(Intercept)` +
         sexfemale*sex +
-        age_centered*agecentered +
+        log(years_sexually_mature) +
[…]

### 2024-06-03 stage2_analysis.R

@@ -1,0 +1,123 @@
+
+designsH2 <- update_designs(designsH, cutoff = 100)
+
+models_stage2 <- fitmodels(designsH2$d.design.adults)
+modelsummaries_stage2 <- getmodelsummaries(models_stage2)
+immune_models_stage2 <- fit_immune_models(designsH2$d.design.adults) # Need to update to dietary design object
+intake_models_stage2 <- fit_intake_models(designsH2$d.design.dietary.adults)
+d_allstats_stage2 <- allstats(models_stage2)
+d_strength_stats_stage2 <- strength_stats(models_stage2)
+d_lifetime_stats_stage2 <- lifetime_stats(models_stage2)
+marginal_stage2 <- marginals(models_stage2)
+
+plot_allcoefs_stage2 <-
+  ggplot(d_allstats_stage2, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Controls)) +
+  geom_pointrange(size = 0.4, linewidth = 0.7, position = position_dodge(width = 0.8)) +
+  geom_vline(xintercept=0, linetype='longdash') +
+  labs(x='\nEstimate (95% CI)', y='') +
+  facet_wrap(~ Outcome, ncol = 4, scales = "free_x") +
+  theme_minimal() +
+  theme(legend.position = 'bottom', legend.title = element_blank())
+plot_allcoefs_stage2
+ggsave("Figures/plot_allcoefs_stage2.pdf", plot_allcoefs_stage2, width = 14, height = 8)
+
+plot_coefs_stage2 <-
+  ggplot(d_strength_stats_stage2, aes(estimate, Outcome, xmin = conf.low, xmax = conf.high, colour = Significant)) +
+  geom_pointrange() +
+  geom_vline(xintercept = 0, linetype = 'dotted') +
+  scale_color_binary() +
+  guides(colour = guide_legend(reverse = T)) +
+  labs(x = 'Estimate (95% CI)', y = '') +
+  facet_grid(Controls ~ term) +
+  theme_bw(15) +
+  theme(strip.text.y = element_text(angle = 0))
+plot_coefs_stage2
+ggsave("Figures/plot_coefs_stage2.pdf", plot_coefs_stage2, width = 10, height = 10)
+
+plot_lifetime_coefs_stage2 <-
+  ggplot(d_lifetime_stats_stage2, aes(estimate, Controls, xmin = conf.low, xmax = conf.high)) +
+  geom_pointrange() +
[…]

### 2024-06-03 paper.qmd
Preparing stage 2 registered report

@@
 ```{r}
+library(ggcorrplot)
+library(tinytable)
+library(patchwork)
 source('functions.R')
@@
 source('sexspecificmodels.R')
-source('stage2_analysis.R')
 ```

### 2024-06-03 functions.R

@@ -5,7 +5,7 @@ library(broom)
 library(modelsummary)
 
 name_dict <- c(
-  manth = 'Anthropomentric',
+  manth = 'Anthropometric',
   mheal = 'Health',
   mhor = 'Hormone',
   mphys = 'Activity',
@@ -541,4 +541,24 @@ marginals <- function(models){
     avg_comparisons(models$Model$manth3, variables = list(strength_centered = 1), by = 'sex', wts = "(weights)") |> mutate(Model = 'Age first sex')
   ) |>
     dplyr::select(sex, partnered, estimate)
+}
+
+
+# Effects plots -----------------------------------------------------------
+
+effects_plots <- function(models, controls = "Anthropometric"){
+  out <-
+    models |>
+    dplyr::filter(Controls == controls) |>
+    rowwise() |>
+    mutate(
+      Plot = list(
+        plot_predictions(Model, condition = c("strength_centered", "sex")) +
+          scale_color_binary() +
+          xlab("Strength (S by sex)") +
+          ylab(Outcome) +
+          theme_minimal()
+      )
+    )
+  wrap_plots(out$Plot, ncol = 2) + plot_layout(axes = 'collect', guides = 'collect')
 }

### 2024-06-06 paper.qmd
Mostly preparing plots for the stage 2 confirmatory study

--- 2024-06-03 paper.qmd
@@
 source('stage2_analysis.R')
+source('compare.R')
 ```
@@
 Bars are 95\% CIs.
+Left: Pilot study results.
+Right: Confirmatory study results.
 For the full regression tables, see the SI.
 
-plot_coefs
+plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
 ```
@@
 
-# Confirmatory Study (Not yet conducted)
+# Stage 2 Confirmatory Study
 
@@
 After testing and reporting our predictions using the unobserved data, we will fit our models on the combined data (2011-2014), and conduct other post hoc analyses.
+
+## Deviation from Stage 1 method
+
+Models and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study.
+However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include it as an exposure variable (an offset) in the model, which we now do.
+We also treat this exposure as years since sexual maturity (age - 12 years).
+The change in the strength and strength:sex coefficients from this change in model specification is minimal (see @sfig-comparison).
+
+## Results
+
+```{r}
+#| label: fig-stage2coef
+#| fig-width: 10
+#| fig-height: 8
+#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95\% CIs.
+For the full regression tables, see the SI.
+
+plot_coefs_stage2
+```
+
+:::{#fig-stage2effects}
+```{r}
+effects_plots(models_stage2)
+```
+
+Effects of sex-specific standardized strength on mating outcomes, by sex.
+Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
+Plots are based on the anthropometric control models, with control variables held at their mean values.
+Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
+:::
 
@@
 
-### Effects plots
+### Effects plots (Pilot)
 
@@
 Effects of sex-specific standardized strength on mating outcomes, by sex.
-Strength is centered and scaled by 2 SD, separately for each sex.
+Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
 Plots are based on the anthropometric control models, with control variables held at their mean values.
+Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
 :::
@@
 
+```{r}
+out_sexspecific_pilot <- sex_specific_models(designsG)
+plot_sexspecific_strength_coefs <- out_sexspecific_pilot$plot
+dfmodelsummaries <- out_sexspecific_pilot$summaries
+```
+
 At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
@@
 
+## Stage 2 confirmatory study
+
+:::{#sfig-comparison}
+```{r}
+plot_compare
+```
+
+A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (orignal), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+Both models are quasi-Poisson models.
+:::
+
 # References

### 2024-06-06 functions.R

@@ -20,7 +20,7 @@ name_dict <- c(
 )
 
 outcome_dict <- c(
-  "1" = "Lifetime partners",
+  "1" = "Lifetime partners (partners per year)",
   "2" = "Past year partners",
   "3" = "Age of first sex",
   "4" = "Partnered"
@@ -536,7 +536,7 @@ allstats <- function(models){
     dplyr::select(-Model) |>  # Remove due to RStudio bug
     dplyr::filter(term != "(Intercept)") |>
     mutate(
-      Outcome = factor(Outcome, levels = c('Partnered', 'Lifetime partners', 'Past year partners', 'Age of first sex')),
+      Outcome = factor(Outcome, levels = c('Partnered', ' (partners per year) partners', 'Past year partners', 'Age of first sex')),
       term = factor(term, levels=rev(names(vnames)), labels=rev(vnames))
     )
 }
@@ -550,7 +550,7 @@ strength_stats <- function(models){
     mutate(
       term = ifelse(term == 'strength_sex_centered', 'Strength (S)', 'Strength (S) X Sex (female)'),
       term = factor(term, levels = c('Strength (S)', 'Strength (S) X Sex (female)')),
-      Outcome = factor(Outcome, levels = rev(c('Partnered', 'Lifetime partners', 'Past year partners', 'Age of first sex'))),
+      Outcome = factor(Outcome, levels = rev(c('Partnered', 'Lifetime partners (partners per year)', 'Past year partners', 'Age of first sex'))),
       Significant = sign(conf.low) == sign(conf.high)
     )
 }
@@ -595,3 +595,24 @@ effects_plots <- function(models, controls = "Anthropometric"){
     )
   wrap_plots(out$Plot, ncol = 2) + plot_layout(axes = 'collect', guides = 'collect')
 }
+
+
+effects_plots2 <- function(df1, df2, controls = "Anthropometric"){
+  out <-
+    bind_rows(list(Pilot = df1, Confirmatory = df2), .id = 'Stage') |>
+    dplyr::filter(Controls == controls) |>
+    rowwise() |>
+    mutate(
+      Plot = list(
+        plot_predictions(Model, condition = c("strength_sex_centered", "sex")) +
+          scale_color_binary() +
+          xlab("Strength (S by sex)") +
+          ylab(Outcome) +
+          theme_minimal()
+      )
+    ) |>
+    ungroup() |>
[…]

### 2024-06-06 pilot_analysis.R

@@ -57,11 +57,14 @@ plot_coefs <-
   geom_pointrange() +
   geom_vline(xintercept = 0, linetype = 'dotted') +
   scale_color_binary() +
-  guides(colour = guide_legend(reverse = T)) +
-  labs(x = 'Estimate (95% CI)', y = '') +
+  guides(colour = 'none') + # guide_legend(reverse = T)
+  labs(title='Stage 1: Pilot study', x = 'Estimate (95% CI)', y = '') +
   facet_grid(Controls ~ term) +
   theme_bw(15) +
-  theme(strip.text.y = element_text(angle = 0))
+  theme(
+    strip.text.y = element_blank()
+  )
+  # theme(strip.text.y = element_text(angle = 0))
 plot_coefs
 ggsave("Figures/plot_coefs.pdf", plot_coefs, width = 10, height = 10)
 

### 2024-06-06 sexspecificmodels.R

@@ -5,44 +5,46 @@
 
 # These models were fit at the request of an anonymous reviewer
 
-# Removing those with >=100 lifetime sex partners
-designsG$d.design.adult.female <- subset(designsG$d.design.adult.female, sex_partners < 100)
-designsG$d.design.adult.male <- subset(designsG$d.design.adult.male, sex_partners < 100)
+sex_specific_models <- function(design){
 
-# Scale sex_partners by the interquartile range of all adults
-designsG$d.design.adult.female <- update(designsG$d.design.adult.female, sex_partners_scaled = sex_partners/10)
-designsG$d.design.adult.male <- update(designsG$d.design.adult.male, sex_partners_scaled = sex_partners/10)
+  # Removing those with >=100 lifetime sex partners
+  design$d.design.adult.female <- subset(design$d.design.adult.female, sex_partners < 100)
+  design$d.design.adult.male <- subset(design$d.design.adult.male, sex_partners < 100)
 
-# Sexually mature
-designsG$d.design.adult.female <- update(designsG$d.design.adult.female, years_sexually_mature = age - 12)
-designsG$d.design.adult.male <- update(designsG$d.design.adult.male, years_sexually_mature = age - 12)
+  # Scale sex_partners by the interquartile range of all adults
+  design$d.design.adult.female <- update(design$d.design.adult.female, sex_partners_scaled = sex_partners/10)
+  design$d.design.adult.male <- update(design$d.design.adult.male, sex_partners_scaled = sex_partners/10)
 
-sexspecificmodels <- list(
-  # Lifetime partners models -------------------------------------------------
+  # Sexually mature
+  design$d.design.adult.female <- update(design$d.design.adult.female, years_sexually_mature = age - 12)
+  design$d.design.adult.male <- update(design$d.design.adult.male, years_sexually_mature = age - 12)
 
-  # Anthropometric
-  manth_1_f = svyglm(
-    sex_partners ~
-      offset(log(years_sexually_mature)) +
-      strength +
-      partnered +
-      bmi,
-    family = quasipoisson(),
-    design = designsG$d.design.adult.female
-  ),
+  sexspecificmodels <- list(
+    # Lifetime partners models -------------------------------------------------
 
-  manth_1_m = svyglm(
-    sex_partners ~
-      offset(log(years_sexually_mature)) +
-      strength +
-      partnered +
-      bmi,
-    family = quasipoisson(),
-    design = designsG$d.design.adult.male
-  ),
[…]

### 2024-06-06 simulate.R

@@ -1,10 +1,11 @@
 library(boot)
 library(furrr)
 
+set.seed(765456)
+
 plan("multisession")
 
 d_adult <- d_G[d_G$age >= 18 & d_G$age <= 60,]
-designsG$d.design.adults <- subset(designsG$d.design.adults, sex_partners < 100)
 
 rqpois <- function(n, mu, theta) rnbinom(n=n, mu=mu, size=mu/(theta-1))
 

### 2024-06-06 stage2_analysis.R

@@ -27,12 +27,34 @@ plot_coefs_stage2 <-
   geom_vline(xintercept = 0, linetype = 'dotted') +
   scale_color_binary() +
   guides(colour = guide_legend(reverse = T)) +
-  labs(x = 'Estimate (95% CI)', y = '') +
+  labs(title='Stage 2: Confirmatory study', x = 'Estimate (95% CI)', y = '') +
   facet_grid(Controls ~ term) +
   theme_bw(15) +
   theme(strip.text.y = element_text(angle = 0))
 plot_coefs_stage2
 ggsave("Figures/plot_coefs_stage2.pdf", plot_coefs_stage2, width = 10, height = 10)
+
+d_strength_stats_combined <-
+bind_rows(
+  list(
+    Pilot = d_strength_stats,
+    Confirmatory = d_strength_stats_stage2
+  ),
+  .id = 'Stage'
+)
+
+plot_coefs_combined <-
+  ggplot(d_strength_stats_combined, aes(estimate, Outcome, xmin = conf.low, xmax = conf.high, shape = Significant, colour = Stage)) +
+  geom_pointrange(position = position_dodge(width=0.3)) +
+  geom_vline(xintercept = 0, linetype = 'dotted') +
+  scale_color_binary() +
+  guides(colour = guide_legend(reverse = T)) +
+  labs(x = 'Estimate (95% CI)', y = '') +
+  facet_grid(Controls ~ term) +
+  theme_bw(15) +
+  theme(strip.text.y = element_text(angle = 0))
+plot_coefs_combined
+
 
 plot_lifetime_coefs_stage2 <-
   ggplot(d_lifetime_stats_stage2, aes(estimate, Controls, xmin = conf.low, xmax = conf.high)) +
@@ -44,7 +66,7 @@ plot_lifetime_coefs_stage2 <-
   theme(strip.text.y = element_text(angle = 0))
 plot_lifetime_coefs_stage2
 
-effects_plots(models_stage2)
+# effects_plots(models_stage2)
 
 # Exploratory
 

### 2024-06-07 paper.qmd
Code for new combined effects plots. Improved coefficient plot. Used dietary weights for one immune model.

@@
 #| fig-height: 8
-#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
@@
 Right: Confirmatory study results.
-For the full regression tables, see the SI.
+For the full regression tables, see the SI."
 
@@
 ```{r}
-effects_plots(models_stage2)
+plot_effects_all
 ```

### 2024-06-07 functions.R

@@ -410,7 +410,7 @@ fitmodels <- function(design){
 # Immune models ------------------------------------------------------------
 
 # design.adults
-fit_immune_models <- function(design){
+fit_immune_models <- function(design1, design2){
 
   list(
 
@@ -420,7 +420,7 @@ fit_immune_models <- function(design){
         strength_centered * sex +
         bmi_centered,
       family= quasipoisson(),
-      design=design
+      design=design1
     ),
 
     mwbc_alt = svyglm(
@@ -435,7 +435,7 @@ fit_immune_models <- function(design){
         tot_MET_centered +
         depression,
       family= quasipoisson(),
-      design=design
+      design=design2
     )
   )
 }
@@ -548,8 +548,8 @@ strength_stats <- function(models){
     ) |>
     unnest(Stats) |>
     mutate(
-      term = ifelse(term == 'strength_sex_centered', 'Strength (S)', 'Strength (S) X Sex (female)'),
-      term = factor(term, levels = c('Strength (S)', 'Strength (S) X Sex (female)')),
+      term = ifelse(term == 'strength_sex_centered', 'Strength (S by sex)', 'Strength (S by sex) X Sex (female)'),
+      term = factor(term, levels = c('Strength (S by sex)', 'Strength (S by sex) X Sex (female)')),
       Outcome = factor(Outcome, levels = rev(c('Partnered', 'Lifetime partners (partners per year)', 'Past year partners', 'Age of first sex'))),
       Significant = sign(conf.low) == sign(conf.high)
     )
@@ -614,3 +614,52 @@ effects_plots2 <- function(df1, df2, controls = "Anthropometric"){
     arrange(Outcome, desc(Stage))
   wrap_plots(out$Plot, ncol = 2, byrow = T) + plot_layout(axes = 'collect', guides = 'collect')
 }
+
+plot_predictions2 <- function(model, stage){
+  d <- plot_predictions(model, condition = c("strength_sex_centered", "sex"), draw = F)
+  d$Stage <- stage
+  return(d)
+}
+effects_plots3 <- function(df1, df2, controls = "Anthropometric"){
+  out <-
+    bind_rows(list(Pilot = df1, Confirmatory = df2), .id = 'Stage') |>
+    dplyr::filter(Controls == controls) |>
+    rowwise() |>
+    mutate(
[…]

### 2024-06-07 pilot_analysis.R

@@ -34,7 +34,7 @@ designsG <- update_designs(designsG)
 
 models <- fitmodels(designsG$d.design.adults)
 modelsummaries <- getmodelsummaries(models)
-immune_models <- fit_immune_models(designsG$d.design.adults)
+immune_models <- fit_immune_models(designsG$d.design.adults, designsG$d.design.dietary.adults)
 intake_models <- fit_intake_models(designsG$d.design.dietary.adults)
 d_allstats <- allstats(models)
 d_strength_stats <- strength_stats(models)
@@ -53,13 +53,13 @@ plot_allcoefs
 ggsave("Figures/plot_allcoefs.pdf", plot_allcoefs, width = 14, height = 8)
 
 plot_coefs <-
-  ggplot(d_strength_stats, aes(estimate, Outcome, xmin = conf.low, xmax = conf.high, colour = Significant)) +
+  ggplot(d_strength_stats, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Significant)) +
   geom_pointrange() +
   geom_vline(xintercept = 0, linetype = 'dotted') +
   scale_color_binary() +
   guides(colour = 'none') + # guide_legend(reverse = T)
   labs(title='Stage 1: Pilot study', x = 'Estimate (95% CI)', y = '') +
-  facet_grid(Controls ~ term) +
+  facet_grid(Outcome ~ term) +
   theme_bw(15) +
   theme(
     strip.text.y = element_blank()

### 2024-06-07 stage2_analysis.R

@@ -3,7 +3,7 @@ designsH2 <- update_designs(designsH, cutoff = 100)
 
 models_stage2 <- fitmodels(designsH2$d.design.adults)
 modelsummaries_stage2 <- getmodelsummaries(models_stage2)
-immune_models_stage2 <- fit_immune_models(designsH2$d.design.adults) # Need to update to dietary design object
+immune_models_stage2 <- fit_immune_models(designsH2$d.design.adults, designsH2$d.design.dietary.adults) # Need to update to dietary design object
 intake_models_stage2 <- fit_intake_models(designsH2$d.design.dietary.adults)
 d_allstats_stage2 <- allstats(models_stage2)
 d_strength_stats_stage2 <- strength_stats(models_stage2)
@@ -22,13 +22,13 @@ plot_allcoefs_stage2
 ggsave("Figures/plot_allcoefs_stage2.pdf", plot_allcoefs_stage2, width = 14, height = 8)
 
 plot_coefs_stage2 <-
-  ggplot(d_strength_stats_stage2, aes(estimate, Outcome, xmin = conf.low, xmax = conf.high, colour = Significant)) +
+  ggplot(d_strength_stats_stage2, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Significant)) +
   geom_pointrange() +
   geom_vline(xintercept = 0, linetype = 'dotted') +
   scale_color_binary() +
   guides(colour = guide_legend(reverse = T)) +
   labs(title='Stage 2: Confirmatory study', x = 'Estimate (95% CI)', y = '') +
-  facet_grid(Controls ~ term) +
+  facet_grid(Outcome ~ term) +
   theme_bw(15) +
   theme(strip.text.y = element_text(angle = 0))
 plot_coefs_stage2

### 2024-06-07 paper.qmd
Added confirmatory discussion section

--- 2024-06-06 paper.qmd
@@
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95\% CIs.
-Left: Pilot study results.
-Right: Confirmatory study results.
+Bars are 95% CIs.
+Left Pilot study results.
+Right Confirmatory study results.
 For the full regression tables, see the SI.
@@
 
-Our methods and models will remain identical to those reported in the pilot study, using unobserved data from the 2013-2014 NHANES collection cycle.
-After testing and reporting our predictions using the unobserved data, we will fit our models on the combined data (2011-2014), and conduct other post hoc analyses.
+Our methods and models remained identical to those reported in the pilot study, other than the deviation reported below, using unobserved data from the 2013-2014 NHANES collection cycle.
+After testing and reporting our predictions using the unobserved data, we fit our models on the combined data (2011-2014), and conduct other post hoc analyses.
 
@@
 The change in the strength and strength:sex coefficients from this change in model specification is minimal (see @sfig-comparison).
+
+We also discovered that because we included dietary energy intake in the expanded controls immune model we needed to utilize the weights calculated for the dietary data rather than the weights calculated for other data for this model.
+This is corrected in the confirmatory results for that model.
 
@@
 ```
+Our confirmatory results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners after controlling for all five sets of potential confounds.
+There was no significant effect of sex-specific strength on the number of partners in the past year, and there were inconsistent effects of sex-specific strength on age at first sex across the potential confound models.
+
+Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+The interaction between sex and sex-specific strength on lifetime number of sexual partners was not replicated in the confirmatory data.
+Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
 
@@
 Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
+
+## Discussion
+The purpose of this study was to replicate as closely as possible, the results from @lassek2009, in new data that included women and used combined grip strength as a proxy for upper body strength.
+We found consistent support for a positive relationship between combined grip strength and two measures of mating success: being currently partnered, and lifetime number of sexual partners for both men and women, controlling for numerous possible confounds.
+We found no consistent association between strength and age at first sex or number of past year sexual partners.
+
+We also tested for interactions between sex and strength on these mating success outcomes.
+Across both pilot and confirmatory data, we found a consistent interaction between sex and strength on partnered status, such that grip strength was more important for predicting males' partnered status than females.
+Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not present in the confirmatory data.
+
 :::

### 2024-06-07 paper.qmd

@@
 #| fig-height: 8
-#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95% CIs.
-Left Pilot study results.
-Right Confirmatory study results.
-For the full regression tables, see the SI.
+Bars are 95\% CIs.
+Left: Pilot study results.
+Right: Confirmatory study results.
+For the full regression tables, see the SI."
 

### 2024-06-07 paper.qmd
Combining pilot and confirmatory plots, part I

--- 2024-06-07 paper.qmd
@@
 
-plot_immune <- forestplot(
-  immune_models$mwbc,
-  immune_models$mwbc_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
+plot_immune_both <- 
+  plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
 
-ggsave("Figures/plot_immune.pdf", plot_immune, width = 14, height = 7)
+ggsave("Figures/plot_immune_both.pdf", plot_immune, width = 14, height = 7)
 
-plot_immune
+plot_immune_both
 ```
@@
   scale_color_binary() +
-  ggtitle('Energy') +
+  ggtitle('Stage 1: Pilot Energy') +
   theme_bw(17) +
@@
   guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
-  ggtitle('Protein') +
+  ggtitle('Stage 1: Pilot Protein') +
   theme_bw(17) +

### 2024-06-07 pilot_analysis.R

@@ -79,6 +79,18 @@ plot_lifetime_coefs <-
 plot_lifetime_coefs
 ggsave("Figures/plot_lifetime_coefs.pdf", plot_lifetime_coefs)
 
+plot_immune <- forestplot(
+  immune_models$mwbc,
+  immune_models$mwbc_alt,
+  intercept = F,
+  facet = F,
+  dodgewidth = .4,
+  varnames = vnames,
+  modelnames = mmnames) +
+  scale_color_binary() +
+  ggtitle("Stage 1: Pilot") +
+  theme_minimal(17)
+
 # SI ----------------------------------------------------------------------
 
 # Descriptive statistics tables

### 2024-06-07 stage2_analysis.R

@@ -107,6 +107,7 @@ plot_immune_stage2 <- forestplot(
   varnames = vnames,
   modelnames = mmnames) +
   scale_color_binary() +
+  ggtitle("Stage 2: Confirmatory") +
   theme_minimal(17) +
   guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
 plot_immune_stage2
@@ -123,7 +124,7 @@ plot_energy_stage2 <- forestplot(
   varnames = vnames,
   modelnames = mmnames) +
   scale_color_binary() +
-  ggtitle('Energy') +
+  ggtitle('Stage 2: Confirmatory Energy') +
   theme_bw(17) +
   guides(colour = guide_none(), shape = guide_none())
 
@@ -137,7 +138,7 @@ plot_protein_stage2 <- forestplot(
   modelnames = mmnames) +
   scale_color_binary() +
   guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
-  ggtitle('Protein') +
+  ggtitle('Stage 2: Confirmatory Protein') +
   theme_bw(17) +
   theme(axis.text.y = element_blank())
 

### 2024-08-01 simulate.R
Removed use of furrr. Cleaned up code

@@ -1,13 +1,26 @@
 library(boot)
-library(furrr)
 
 set.seed(765456)
-
-plan("multisession")
 
 d_adult <- d_G[d_G$age >= 18 & d_G$age <= 60,]
 
 rqpois <- function(n, mu, theta) rnbinom(n=n, mu=mu, size=mu/(theta-1))
+
+param_dict <- c(
+  "(Intercept)" = "(Intercept)",
+  "sexfemale" = "sexfemale",
+  "age_centered" = "age_centered",
+  "partneredTRUE" = "partneredTRUE",
+  "strength_sex_centered" = "strength_sex_centered",
+  "sex_partners_scaled" = "sex_partners_scaled",
+
+  "sexfemale:strength_sex_centered" = "SEX_STRENGTH",
+  "strength_sex_centered:sexfemale" = "SEX_STRENGTH",
+  "age_centered:sexfemale" = "SEX_AGE",
+  "sexfemale:age_centered" = "SEX_AGE",
+  "strength_sex_centered:partneredTRUE" = "STRENGTH_PARTNERED",
+  "partneredTRUE:strength_sex_centered" = "STRENGTH_PARTNERED"
+)
 
 simdata <- function(
     N,
@@ -17,15 +30,12 @@ simdata <- function(
     partneredTRUE = 0,
     strength_sex_centered = 0,
     sex_partners_scaled = 0,
-    `sexfemale:strength_sex_centered` = 0,
-    `strength_sex_centered:sexfemale` = 0,
-    `age_centered:sexfemale` = 0,
-    `strength_sex_centered:partneredTRUE` = 0,
+    SEX_STRENGTH = 0,
+    SEX_AGE = 0,
+    STRENGTH_PARTNERED = 0,
     theta = 0,
     ...
 ){
-
-  SEXSTRENGTH <- ifelse(length(`sexfemale:strength_sex_centered`) > 0, `sexfemale:strength_sex_centered`, `strength_sex_centered:sexfemale`)
 
   tibble(
 
@@ -44,8 +54,8 @@ simdata <- function(
         sexfemale*sex +
         age_centered*agecentered +
         strength_sex_centered*strengthcentered +
-        SEXSTRENGTH*sex*strengthcentered +
-        `age_centered:sexfemale`*sex*agecentered
+        SEX_STRENGTH*sex*strengthcentered +
+        SEX_AGE*sex*agecentered
     )
     ),
[…]

### 2024-08-01 pilot_analysis.R
Moved table and some other code to functions.R

@@ -3,13 +3,13 @@
 # Cumulative distributions of sex partners ------------------------
 
 plot_sexpartners <-
-  ggplot(d_G[!is.na(d_G$sex_partners),], aes(sex_partners + 1, colour = sex)) +
+  ggplot(d_GH[!is.na(d_G$sex_partners),], aes(sex_partners + 1, colour = paste(sex, series))) +
   stat_ecdf() +
   geom_vline(xintercept = 100, linetype = 'dotted') +
   annotate("text", label = 'Cutoff', x = 110, y = 0.80, hjust = 0) +
   scale_x_log10(breaks = c(1, 5, 10, 25, 50, 100, 1000)) +
-  scale_color_binary() +
-  guides(colour = guide_legend(title = '', reverse = T, override.aes = list(linewidth=2))) +
+  # scale_color_binary() +
+  guides(colour = guide_legend(title = '', reverse = F, override.aes = list(linewidth=2))) +
   labs(x = "\nNumber of lifetime sex partners + 1", y = 'ECDF\n') +
   theme_minimal(15)
 plot_sexpartners
@@ -32,17 +32,18 @@ designsG <- update_designs(designsG)
 
 # Fit models --------------------------------------------------------------
 
-models <- fitmodels(designsG$d.design.adults)
+models <- fitmodels(designsG$d.design.adults, designsG$d.design.dietary.adults)
 modelsummaries <- getmodelsummaries(models)
-immune_models <- fit_immune_models(designsG$d.design.adults, designsG$d.design.dietary.adults)
-intake_models <- fit_intake_models(designsG$d.design.dietary.adults)
-d_allstats <- allstats(models)
-d_strength_stats <- strength_stats(models)
-d_lifetime_stats <- lifetime_stats(models)
 marginal <- marginals(models)
 
+d_allstats <- allstats(models)
+
+benefit_stats <-
+  d_allstats |>
+  dplyr::filter(benefit_cost == 'Benefit')
+
 plot_allcoefs <-
-  ggplot(d_allstats, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Controls)) +
+  ggplot(benefit_stats, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Controls)) +
   geom_pointrange(size = 0.4, linewidth = 0.7, position = position_dodge(width = 0.8)) +
   geom_vline(xintercept=0, linetype='longdash') +
[…]

### 2024-08-01 stage2_analysis.R
Code to combine results plots from pilot and confirmatory studies

@@ -1,17 +1,25 @@
 
 designsH2 <- update_designs(designsH, cutoff = 100)
 
-models_stage2 <- fitmodels(designsH2$d.design.adults)
+models_stage2 <- fitmodels(designsH2$d.design.adults, designsH2$d.design.dietary.adults)
 modelsummaries_stage2 <- getmodelsummaries(models_stage2)
-immune_models_stage2 <- fit_immune_models(designsH2$d.design.adults, designsH2$d.design.dietary.adults) # Need to update to dietary design object
-intake_models_stage2 <- fit_intake_models(designsH2$d.design.dietary.adults)
-d_allstats_stage2 <- allstats(models_stage2)
-d_strength_stats_stage2 <- strength_stats(models_stage2)
-d_lifetime_stats_stage2 <- lifetime_stats(models_stage2)
 marginal_stage2 <- marginals(models_stage2)
 
+d_allstats_stage2 <- allstats(models_stage2)
+
+d_allstats_combined <-
+  bind_rows(
+    "Stage 1: Pilot" = d_allstats,
+    "Stage 2: Confirmatory" = d_allstats_stage2,
+    .id = 'Stage'
+  )
+
+benefit_stats_stage2 <-
+  d_allstats_stage2 |>
+  dplyr::filter(benefit_cost == 'Benefit')
+
 plot_allcoefs_stage2 <-
-  ggplot(d_allstats_stage2, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Controls)) +
+  ggplot(benefit_stats_stage2, aes(estimate, term, xmin = conf.low, xmax = conf.high, colour = Controls)) +
   geom_pointrange(size = 0.4, linewidth = 0.7, position = position_dodge(width = 0.8)) +
   geom_vline(xintercept=0, linetype='longdash') +
   labs(x='\nEstimate (95% CI)', y='') +
@@ -21,8 +29,12 @@ plot_allcoefs_stage2 <-
 plot_allcoefs_stage2
 ggsave("Figures/plot_allcoefs_stage2.pdf", plot_allcoefs_stage2, width = 14, height = 8)
 
+benefit_stats_strength_stage2 <-
+  benefit_stats_stage2 |>
+  dplyr::filter(term2 %in% sex_strength_terms)
+
 plot_coefs_stage2 <-
-  ggplot(d_strength_stats_stage2, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Significant)) +
[…]

### 2024-08-01 paper.qmd
Saved time consuming simulation results to a file. Edited figure captions to reflect combination of pilot and confirmatory results.

@@
 source('pilot_analysis.R')
-source('simulate.R')
 source('sexspecificmodels.R')
@@
 source('compare.R')
+
+# source('simulate.R')
+load("df_pwr.rda")
 ```
@@
 Participants who reported 100 or more partners were removed from the analysis.
-The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this plot and not for any analyses).
+The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this Figure and not for any analyses).
 
@@
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95\% CIs.
-Left: Pilot study results.
-Right: Confirmatory study results.
+Bars are 95% CIs.
 For the full regression tables, see the SI."
 
-plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
+# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
+plot_coefs_combined
 ```
@@
 #| label: fig-allcoefs
-#| fig-cap: Coefficients from all models of our four mating success outcomes and our five sets of potential confounding variables.
+#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Mexican Amercians are the base level for the ethnicity control variable.
 Bars are 95\% CIs.
@@
 
-plot_allcoefs
+plot_allcoefs_stage2
 ```
@@
 
-plot_immune_both <- 
-  plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
+# plot_immune_both <- 
+#   plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
 
-ggsave("Figures/plot_immune_both.pdf", plot_immune, width = 14, height = 7)
+ggsave("Figures/plot_immune_strength_combined.pdf", plot_immune_strength_combined, width = 14, height = 7)
 
-plot_immune_both
+plot_immune_strength_combined
 ```
@@
 #| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) and dietary protein (g) from generalized linear models (Pilot).
+#| fig.cap: Coefficient plots of predictors of dietary energy (kcals) and protein (g) intake from generalized linear models for Stage 1 Pilot (left) and Stage 2 Confirmatory (right).
 Variables labelled (S) have been centered at the mean and standardized by 2 SD.
@@
 
-plot_energy <- forestplot(
-  intake_models$m_energy,
-  intake_models$m_energy_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  ggtitle('Stage 1: Pilot Energy') +
-  theme_bw(17) +
-  guides(colour = guide_none(), shape = guide_none())
-
-plot_protein <- forestplot(
-  intake_models$m_protein,
-  intake_models$m_protein_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
-  ggtitle('Stage 1: Pilot Protein') +
-  theme_bw(17) +
-  theme(axis.text.y = element_blank())
-
-p <- plot_energy + plot_protein + plot_layout(axes = 'collect') 
-ggsave("Figures/plot_energy_protein.pdf", p, width = 12, height = 6)
-p
+# p <- plot_energy_coefs / plot_protein_coefs + plot_layout(guides = 'collect', axis_titles = 'collect')
+ggsave("Figures/plot_intake_strength_combined.pdf", plot_intake_strength_combined, width = 12, height = 6)
+plot_intake_strength_combined
 ```
@@
 | Outcome            | Strength coefficient  | Strength X Sex coefficient |
-|--------------------|:--------------------:|:--------------------------:|
-| Partnered status   |           +          |              -             |
-| Lifetime partners  |           +          |              -             |
-| Past year partners |           0          |              0             |
-| Age at first sex   |           0          |              0             |
-| Immunity           |           -          |              0             |
-| Energy             |           0          |                            |
-| Protein            |           +          |                            |
+|--------------------|:---------------------:|:--------------------------:|
+| Partnered status   |           +   ✓      |              -   ✓         |
+| Lifetime partners  |           +   ✓      |              -   ❌         |
+| Past year partners |           0   ✓      |              0   ✓         |
+| Age at first sex   |           0   ✓      |              0   ✓         |
+| Immunity           |           -   ❌      |              0   ✓         |
+| Energy             |           0   ❌      |                            |
+| Protein            |           +   ✓      |                            |
 
@@
 A blank indicates not relevant.
+✓: Supported in confirmatory study.
+❌: Not supported in confirmatory study.
 {#tbl-predictions}
@@
 Models and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study.
-However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include it as an exposure variable (an offset) in the model, which we now do.
+However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, which we now do.
 We also treat this exposure as years since sexual maturity (age - 12 years).
-The change in the strength and strength:sex coefficients from this change in model specification is minimal (see @sfig-comparison).
+The changes in the strength and strength:sex coefficients from this change in model specification are minimal (see @sfig-comparison).
 
@@
 ```{r}
-#| label: fig-stage2coef
+#| label: fig-effects
 #| fig-width: 10
 #| fig-height: 8
-#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95\% CIs.
-For the full regression tables, see the SI.
-
-plot_coefs_stage2
-```
-
-:::{#fig-stage2effects}
-```{r}
-plot_effects_all
-```
-
-Effects of sex-specific standardized strength on mating outcomes, by sex.
+#| fig-cap: Effects of sex-specific standardized strength on mating outcomes, by sex.
 Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
@@
 Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
-:::
+
+plot_effects_all
+```
 
@@
 
-Descriptive statistics.
+### Descriptive statistics
 
 ::: {#stbl-stats}
-`{r} tbl_stack(list(con, cat))`
+```{r}
+pilot_summary_stats <- summary_tables(designsG$d.design.adults)
+tbl_stack(list(pilot_summary_stats$con, pilot_summary_stats$cat))
+```
 
-Descriptive statistics of pilot study sample.
+Pilot: Descriptive statistics and sex differences for participants ages 18-60 using population weights
 :::
@@
 #| fig-width: 11
-plot_lifetime_coefs
+plot_lifetime_coefs_combined
 ```
@@
 
-### Effects plots (Pilot)
-
-:::{#sfig-effects}
-```{r}
-effects_plots(models)
-```
-
-Effects of sex-specific standardized strength on mating outcomes, by sex.
-Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
-Plots are based on the anthropometric control models, with control variables held at their mean values.
-Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
-:::
-
 ### Correlation matrices of study variables
@@
 
-cor_bothsexes <- ggcorrplot(cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
+pilot_cor_mats <- cor_mats(designsG)
+
+cor_bothsexes <- ggcorrplot(pilot_cor_mats$cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Both sexes')
 ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes, width = 12, height = 12)
@@
 
-cor_females <- ggcorrplot(cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Females')
+cor_females <- ggcorrplot(pilot_cor_mats$cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Females')
 ggsave("Figures/cor_females.pdf", cor_females, width = 12, height = 12)
@@
 
-cor_males <- ggcorrplot(cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Males')
+cor_males <- ggcorrplot(pilot_cor_mats$cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Males')
 ggsave("Figures/cor_males.pdf", cor_males, width = 12, height = 12)
@@
 for (i in 1:nrow(modelsummaries)){
-  cat(paste('####', modelsummaries$title[i]))
+  cat(paste('####', str_replace(modelsummaries$title[i], '\n', ' '))) # Remove line feed char
   print(modelsummaries$summary[[i]])
@@
 for (i in 1:nrow(dfmodelsummaries)){
-  cat(paste('####', dfmodelsummaries$title[i]))
+  cat(paste('####', str_replace(dfmodelsummaries$title[i], '\n', ' ')))
   print(dfmodelsummaries$summary[[i]])
@@
 
+### Descriptive statistics
+
+::: {#stbl-confirmstats}
+```{r}
+confirm_summary_stats <- summary_tables(designsH2$d.design.adults)
+tbl_stack(list(confirm_summary_stats$con, confirm_summary_stats$cat))
+```
+
+Confirmatory: Descriptive statistics and sex differences for participants ages 18-60 using population weights
+:::
+
 :::{#sfig-comparison}
@@
 
-A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (orignal), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
 Both models are quasi-Poisson models.

### 2024-08-01 paper.qmd

@@
 The changes in the strength and strength:sex coefficients from this change in model specification are minimal (see @sfig-comparison).
+
+We also discovered that because we included dietary energy intake in the expanded controls immune model we needed to utilize the weights calculated for the dietary data rather than the weights calculated for other data for this model.
+This is corrected in the confirmatory results for that model.
 
@@
 ```
+Our confirmatory results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners after controlling for all five sets of potential confounds.
+There was no significant effect of sex-specific strength on the number of partners in the past year, and there were inconsistent effects of sex-specific strength on age at first sex across the potential confound models.
+
+Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+The interaction between sex and sex-specific strength on lifetime number of sexual partners was not replicated in the confirmatory data.
+Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+
+Effects of sex-specific standardized strength on mating outcomes, by sex.
+Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
+Plots are based on the anthropometric control models, with control variables held at their mean values.
+Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
+
+## Discussion
+The purpose of this study was to replicate as closely as possible, the results from @lassek2009, in new data that included women and used combined grip strength as a proxy for upper body strength.
+We found consistent support for a positive relationship between combined grip strength and two measures of mating success: being currently partnered, and lifetime number of sexual partners for both men and women, controlling for numerous possible confounds.
+We found no consistent association between strength and age at first sex or number of past year sexual partners.
+
+We also tested for interactions between sex and strength on these mating success outcomes.
+Across both pilot and confirmatory data, we found a consistent interaction between sex and strength on partnered status, such that grip strength was more important for predicting males' partnered status than females.
+Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not present in the confirmatory data.
 

### 2024-08-09 paper.qmd
edits to abstract and moved result figures to confirmatory section for full submission

@@
 
-Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
+<!-- Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
 It is unknown if females experience a similar tradeoff.
@@
 This study replicates an earlier study of US males [@lassek2009] with numerous additional control variables, and it additionally tests these relationships in females.
+-->
+
+Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
+It is unknown if females experience a similar tradeoff.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse, for both males and females.
+We found consistent evidence for a sex:strength interaction on partnered status, such that strength was significantly more important for male partnered status than female.
+Our results produced mixed evidence for a sex:strength interaction on lifetime number of sexual partners.
+We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein intake but no relationship with energy intake, and a negative relationship between grip strength and native immune function.
+This study replicates an earlier study of US men (Lassek & Gaulin, 2009) with numerous additional control variables, and the expansion of these models to include women.
 
@@
 Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
-We will use the same cutoff for our confirmatory study.
+We used the same cutoff for our confirmatory study.
+
+CHECK: edit key to refer to pilot and confirmatory data instead of G and H.
 
@@
 
-```{r}
-#| label: fig-coefs
-#| fig-width: 10
-#| fig-height: 8
-#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95% CIs.
-For the full regression tables, see the SI."
-
-# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
-plot_coefs_combined
-```
-
 For a plot of all coefficients, see @fig-allcoefs.
-
-```{r}
-#| label: fig-allcoefs
-#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Mexican Amercians are the base level for the ethnicity control variable.
-Bars are 95\% CIs.
-For the full regression tables, see the SI.
-
-plot_allcoefs_stage2
-```
 
@@
 
-```{r}
-#| label: fig-immune
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
-Variables labeled (S) have been centered at the mean and standardized by 2 SD.
-Bars are 95% CIs.
+Previous research found that higher musculature was associated with decreased investment in native immunity.
+We developed two models of immune investment in white blood cell count (@fig-immune).
+In the Lassek and Gaulin model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
+In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was no longer significant, albeit with a coefficient that was very similar to the replication model.
 
-# plot_immune_both <- 
-#   plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
-
-ggsave("Figures/plot_immune_strength_combined.pdf", plot_immune_strength_combined, width = 14, height = 7)
-
-plot_immune_strength_combined
-```
+CHECK: remind me why we are not using strength centered by sex here.
 
@@
 
-```{r}
-#| label: fig-energy
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plots of predictors of dietary energy (kcals) and protein (g) intake from generalized linear models for Stage 1 Pilot (left) and Stage 2 Confirmatory (right).
-Variables labelled (S) have been centered at the mean and standardized by 2 SD.
-Bars are 95% CIs.
+## Pilot Results Discussion
 
-# p <- plot_energy_coefs / plot_protein_coefs + plot_layout(guides = 'collect', axis_titles = 'collect')
-ggsave("Figures/plot_intake_strength_combined.pdf", plot_intake_strength_combined, width = 12, height = 6)
-plot_intake_strength_combined
-```
-
-## Pilot Results Discussion
+For ease of comparison, pilot results discussed here are displayed in conjunction with confirmatory results in the Confirmatory Results section below.
 
@@
 
-## Limitations
+CHECK: moved limitations section to confirmatory discussion. 
+<!-- ## Limitations -->
 
-Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females.
+<!-- Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females.
 Our aim is not to test all possible byproduct hypotheses, nor is it possible to do so.
@@
 Finally, all pilot results were obtained after considerable exploratory analysis, and must be replicated in a confirmatory study.
+-->
 
-# Stage 2 Confirmatory Study
+# Stage 2 Confirmatory Study Predictions
 
@@
 This is corrected in the confirmatory results for that model.
+CHECK: retroactively correct for pilot results?
 
 ## Results
+```{r}
+#| label: fig-coefs
+#| fig-width: 10
+#| fig-height: 8
+#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95% CIs.
+For the full regression tables, see the SI."
+
+# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
+plot_coefs_combined
+```
+
+```{r}
+#| label: fig-allcoefs
+#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Mexican Amercians are the base level for the ethnicity control variable.
+Bars are 95\% CIs.
+For the full regression tables, see the SI.
+
+plot_allcoefs_stage2
+```
 
@@
 
+```{r}
+#| label: fig-immune
+#| fig.width: 12
+#| fig.height: 8
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
+Variables labeled (S) have been centered at the mean and standardized by 2 SD.
+Bars are 95% CIs.
+
+# plot_immune_both <- 
+#   plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
+
+ggsave("Figures/plot_immune_strength_combined.pdf", plot_immune_strength_combined, width = 14, height = 7)
+
+plot_immune_strength_combined
+```
+
+```{r}
+#| label: fig-energy
+#| fig.width: 12
+#| fig.height: 8
+#| fig.cap: Coefficient plots of predictors of dietary energy (kcals) and protein (g) intake from generalized linear models for Stage 1 Pilot (left) and Stage 2 Confirmatory (right).
+Variables labelled (S) have been centered at the mean and standardized by 2 SD.
+Bars are 95% CIs.
+
+# p <- plot_energy_coefs / plot_protein_coefs + plot_layout(guides = 'collect', axis_titles = 'collect')
+ggsave("Figures/plot_intake_strength_combined.pdf", plot_intake_strength_combined, width = 12, height = 6)
+plot_intake_strength_combined
+```
+
 ## Discussion
@@
 Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not present in the confirmatory data.
+
+## Limitations
+
+Our registered report had one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females.
+Our aim was not to test all possible byproduct hypotheses, nor was it possible to do so.
+We were unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success.
+Likewise, NHANES does not collect data on sociosexuality.
+As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women.
+Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
+Finally, all pilot results were obtained after considerable exploratory analysis, and therefore our confirmatory results were critical to testing our predictions.
 

### 2024-08-09 functions.R
Don't remove rows with missing sex_partner values

@@ -118,9 +118,6 @@ sex_strength_terms <- c(
 
 update_designs <- function(designs, cutoff = 100){
 
-  # Removing those with very high numbers of lifetime sex partners
-  designs$d.design.adults <- subset(designs$d.design.adults, sex_partners < cutoff)
-
   # Scale sex_partners by the interquartile range
   designs$d.design.adults <- update(designs$d.design.adults, sex_partners_scaled = sex_partners/10)
 
@@ -130,13 +127,16 @@ update_designs <- function(designs, cutoff = 100){
   # Compute years since sexually maturity, defined as age 12
   designs$d.design.adults <- update(designs$d.design.adults, years_sexually_mature = age - 12)
 
+  # Removing those with very high numbers of lifetime sex partners
+  designs$d.design.adults <- subset(designs$d.design.adults, (sex_partners < cutoff) | is.na(sex_partners))
+
   return(designs)
 }
 
 
 # Regression models -------------------------------------------------------
 
-fitmodels <- function(design, design_dietary){
+fitmodels <- function(design, design_dietary, design_full){
 
   models0 <- list(
 
@@ -413,7 +413,7 @@ fitmodels <- function(design, design_dietary){
         strength_centered * sex +
         bmi_centered,
       family = quasipoisson(),
-      design = design
+      design = design_full
     ),
 
     mexp5 = svyglm(

### 2024-08-09 paper.qmd
Updated designs for cor_mats

--- 2024-08-01 paper.qmd
@@
 
-pilot_cor_mats <- cor_mats(designsG)
+pilot_cor_mats <- cor_mats(designsG2)
 
@@
 ```{r}
-out_sexspecific_pilot <- sex_specific_models(designsG)
+out_sexspecific_pilot <- sex_specific_models(designsG2)
 plot_sexspecific_strength_coefs <- out_sexspecific_pilot$plot

### 2024-08-13 combinedGH.R
Fit models on combined G and H series data

@@ -1,0 +1,93 @@
+# Fit models on combined G & H series data (2011-2014)
+
+designsGH2 <- update_designs(designsGH, cutoff = 100)
+
+modelsGH <- fitmodels(designsGH2$d.design.adults, designsGH2$d.design.dietary.adults, designsGH$d.design.adults)
+modelsummariesGH <- getmodelsummaries(modelsGH)
+marginalGH <- marginals(modelsGH)
+
+d_allstatsGH <- allstats(modelsGH)
+
+benefit_statsGH <-
+  d_allstatsGH |>
+  dplyr::filter(benefit_cost == "Benefit")
+
+benefit_stats_strengthGH <-
+  benefit_statsGH |>
+  dplyr::filter(term2 %in% sex_strength_terms) |>
+  mutate(
+    term = factor(term, levels = (unique(term)))
+  )
+
+plot_strengthGH <-
+  ggplot(benefit_stats_strengthGH, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Significant)) +
+  geom_pointrange() +
+  geom_vline(xintercept = 0, linetype = "dotted") +
+  scale_color_binary() +
+  guides(colour = guide_legend(reverse = T)) +
+  labs(title = "GH data (2011-2014)", x = "Estimate (95% CI)", y = "") +
+  facet_grid(Outcome ~ term) +
+  theme_bw(15) +
+  theme(strip.text.y = element_text(angle = 0))
+plot_strengthGH
+
+# Immunity ------------------------------------------------------
+
+immune_stats_GH <-
+  d_allstatsGH |>
+  dplyr::filter(Outcome == "Immunity")
+
+immune_stats_strength_GH <-
+  immune_stats_GH |>
+  dplyr::filter(str_detect(term, "Strength")) |>
+  mutate(
+    term = factor(term, levels = (unique(term)))
+  )
+
+plot_immune_strength_GH <-
+  ggplot(immune_stats_strength_GH, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Significant)) +
+  geom_pointrange(position = position_dodge(width = -0.5), linewidth = 2, size = 1.5) +
+  geom_vline(xintercept = 0, linetype = "dotted") +
+  scale_color_binary() +
+  guides(shape = guide_legend(override.aes = list(linetype = 0)), colour = guide_legend(reverse = T)) +
[…]

### 2024-08-13 functions.R
Changed replication to Lassek and Gaulin; added units to Energy and Protein

@@ -18,8 +18,8 @@ control_dict <- c(
   mhor = 'Hormone',
   mphys = 'Activity',
   msoc = 'Socioeconomic',
-  mrep = 'Replication',
-  mexp = 'Expanded'
+  mrep = 'Lassek and Gaulin',
+  mexp = 'Expanded controls'
 )
 
 outcome_dict <- c(
@@ -28,8 +28,8 @@ outcome_dict <- c(
   "3" = "Age of first sex",
   "4" = "Partnered",
   "5" = "Immunity",
-  "6" = "Energy",
-  "7" = "Protein"
+  "6" = "Energy (kcal)",
+  "7" = "Protein (g)"
 )
 
 # Coef plot labels --------------------------------------------------------------
@@ -493,7 +493,7 @@ fitmodels <- function(design, design_dietary, design_full){
     Outcome = outcome_dict[str_extract(modelnames, "\\d")],
     Model = models0,
     Stats = map(Model, \(x) {x$df.residual <- Inf; tidy(x, conf.int = T)}),
-    benefit_cost = ifelse(Outcome %in% c('Immunity', 'Energy', 'Protein'), 'Cost', 'Benefit')
+    benefit_cost = ifelse(Outcome %in% c('Immunity', 'Energy (kcal)', 'Protein (g)'), 'Cost', 'Benefit')
   )
 }
 

### 2024-08-13 sexspecificmodels.R
Fixed design update code for sexspecific models

@@ -5,11 +5,11 @@
 
 # These models were fit at the request of an anonymous reviewer
 
-sex_specific_models <- function(design){
+sex_specific_models <- function(design, cutoff = 100){
 
   # Removing those with >=100 lifetime sex partners
-  design$d.design.adult.female <- subset(design$d.design.adult.female, sex_partners < 100)
-  design$d.design.adult.male <- subset(design$d.design.adult.male, sex_partners < 100)
+  design$d.design.adult.female <- subset(design$d.design.adult.female, (sex_partners < cutoff) | is.na(sex_partners))
+  design$d.design.adult.male <- subset(design$d.design.adult.male, (sex_partners < cutoff) | is.na(sex_partners))
 
   # Scale sex_partners by the interquartile range of all adults
   design$d.design.adult.female <- update(design$d.design.adult.female, sex_partners_scaled = sex_partners/10)
@@ -560,15 +560,16 @@ sex_specific_models <- function(design){
     dplyr::filter(term == 'strength')
 
   plot_sexspecific_strength_coefs <-
-    ggplot(strengthstats, aes(estimate, Outcome, xmin = conf.low, xmax = conf.high, colour = Sex, shape = Sex, alpha = Significant)) +
+    ggplot(strengthstats, aes(estimate, Controls, xmin = conf.low, xmax = conf.high, colour = Sex, shape = Sex, alpha = Significant)) +
     geom_pointrange(position = position_dodge(width = 0.5), linewidth = 1) +
     geom_vline(xintercept = 0, linetype = 'dotted') +
     scale_alpha_ordinal(range = c(0.35, 1)) +
     hagenutils::scale_color_binary() +
     guides(colour = guide_legend(reverse = T), shape = 'none', alpha = 'none') +
     labs(x = '\nStrength coefficient (95% CI)', y = '') +
-    facet_wrap(~Controls, ncol = 1) +
-    theme_bw(15)
+    facet_wrap(~Outcome, ncol = 1, strip.position = 'right') +
+    theme_bw(15) +
+    theme(strip.text.y = element_text(angle = 0))
 
   return(list(summaries = dfmodelsummaries, stats = strengthstats, plot = plot_sexspecific_strength_coefs))
 }

### 2024-08-13 paper.qmd
Added Stage 2 SI results; minor tweaks to plots

--- 2024-08-09 paper.qmd
@@
 It is unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = `{r} nrow(designsH2$d.design.adults)`), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
 We will also test if there is a positive relationship between grip strength and dietary energy and protein intake, and a negative relationship between grip strength and native immune function.
@@
 
-All analyses were completed in `{r} R.version.string`, using the survey package [version 4.1-1, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
+All analyses were completed in `{r} R.version.string`, using the survey package [`{r} packageVersion("survey")`, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
 We modeled lifetime and past year numbers of partners using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
@@
 
-### Coefficients of strength, and strength X sex interaction
+### Coefficients of strength, and $strength \times sex$ interaction
 
@@
 #| label: fig-coefs
-#| fig-width: 10
-#| fig-height: 8
-#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+#| fig-width: 12
+#| fig-height: 10
+#| fig-cap: "Stage 1 pilot vs.
+stage 2 confirmatory coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
@@
 #| label: fig-allcoefs
+#| fig-width: 12
+#| fig-height: 8
 #| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
@@
 #| fig.width: 12
-#| fig.height: 8
+#| fig.height: 6
 #| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
@@
 
-## Pilot study
+## Stage 1: Pilot study
 
@@
 ```{r}
-#| fig-width: 7
+#| fig-width: 9
 #| fig-height: 9
@@
 
-## Stage 2 confirmatory study
+## Stage 2: Confirmatory study
 
@@
 ```{r}
-plot_compare
+#| fig-width: 10
+plot_compare_age_models
 ```
 
-A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+Pilot: A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
 Both models are quasi-Poisson models.
@@
 
+### Marginal effects
+
+::: {#stab-marginal-confirm}
+```{r}
+marginal_stage2 |>
+  tt() |>
+  format_tt(fn = hagenutils::signif2) |>
+  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
+```
+
+Confirmatory: The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
+:::
+
+### Correlation matrices of study variables
+
+::: {#sfig-cor-confirm}
+```{r}
+
+confirm_cor_mats <- cor_mats(designsH2)
+
+confirm_cor_bothsexes <- ggcorrplot(confirm_cor_mats$cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Both sexes')
+ggsave("Figures/confirm_cor_bothsexes.pdf", confirm_cor_bothsexes, width = 12, height = 12)
+confirm_cor_bothsexes
+```
+
+Confirmatory: Correlation matrix of study variables (both sexes). Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-fcor-confirm}
+```{r}
+
+confirm_cor_females <- ggcorrplot(confirm_cor_mats$cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Females')
+ggsave("Figures/confirm_cor_females.pdf", confirm_cor_females, width = 12, height = 12)
+confirm_cor_females
+```
+
+Confirmatory: Correlation matrix of study variables for females only. Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-mcor-confirm}
+```{r}
+
+confirm_cor_males <- ggcorrplot(confirm_cor_mats$cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Males')
+ggsave("Figures/confirm_cor_males.pdf", confirm_cor_males, width = 12, height = 12)
+confirm_cor_males
+```
+
+Confirmatory: Correlation matrix of study variables for males only. Rows and columns ordered by hierarchical clustering.
+:::
+
+### Regression tables
+
+```{r}
+#| results: asis
+for (i in 1:nrow(modelsummaries_stage2)){
+  cat(paste('####', str_replace(modelsummaries_stage2$title[i], '\n', ' '))) # Remove line feed char
+  print(modelsummaries_stage2$summary[[i]])
+}
+```
+
+### Sex-specific models
+
+```{r}
+out_sexspecific_confirm <- sex_specific_models(designsH2)
+plot_sexspecific_strength_coefs_confirm <- out_sexspecific_confirm$plot
+dfmodelsummaries_confirm <- out_sexspecific_confirm$summaries
+```
+
+Confirmatory: At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
+
+:::{#sfig-sexspecific-confirm}
+```{r}
+#| fig-width: 9
+#| fig-height: 9
+plot_sexspecific_strength_coefs_confirm
+```
+
+Confirmatory: Strength coefficients in each regression model fit separately by sex.
+Control variables are the same as for models fit with a sex X strength interaction term.
+Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
+:::
+
+Regression tables for sex-specific models of the effect of strength on mating success.
+
+```{r}
+#| results: asis
+for (i in 1:nrow(dfmodelsummaries_confirm)){
+  cat(paste('####', str_replace(dfmodelsummaries_confirm$title[i], '\n', ' ')))
+  print(dfmodelsummaries_confirm$summary[[i]])
+}
+```
+
 # References

### 2024-08-14 combinedGH.R
Updated plot titles

@@ -25,7 +25,7 @@ plot_strengthGH <-
   geom_vline(xintercept = 0, linetype = "dotted") +
   scale_color_binary() +
   guides(colour = guide_legend(reverse = T)) +
-  labs(title = "GH data (2011-2014)", x = "Estimate (95% CI)", y = "") +
+  labs(title = "G and H series data (2011-2014)", x = "Estimate (95% CI)", y = "") +
   facet_grid(Outcome ~ term) +
   theme_bw(15) +
   theme(strip.text.y = element_text(angle = 0))
@@ -84,10 +84,21 @@ plot_effects_lifetime_GH <-
   plot_predictions(modelsGH$Model$manth1, condition = c("strength_sex_centered", "sex")) +
   scale_color_binary() +
   scale_fill_binary() +
-  theme_minimal(20)
+  labs(
+    title = 'G and H series data (2011-2014)',
+    x = 'Strength (centered by sex)',
+    y = 'Lifetime sex\npartners'
+  ) +
+  theme_minimal(20) +
+  theme(axis.title.y = element_text(angle = 0))
 
 plot_effects_partners_GH <- 
   plot_predictions(modelsGH$Model$manth4, condition = c("strength_sex_centered", "sex")) +
   scale_color_binary() +
   scale_fill_binary() +
-  theme_minimal(20)
+    labs(
+      x = 'Strength (centered by sex)',
+      y = 'Partnered'
+    ) +
+  theme_minimal(20) +
+    theme(axis.title.y = element_text(angle = 0))

### 2024-08-14 paper.qmd
Edits to confirmatory results and discussion.

--- 2024-08-09 paper.qmd
@@
 Thus, for women, low or high values are relative to other women, and for men, low or high values are relative to other men.
-The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high strength.
-Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes.
+We denote the use of this predictor in figures below as 'Strength (S by sex).' The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high strength.
+Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes, denoted as Strength (S) in results below.
 
@@
 Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
+We applied this sample cutoff for all models of mating success (CHECK).
+For our tradeoff models of immunity, energy intake, and protein intake we did not restrict our sample based on reported number of lifetime partners.
 We used the same cutoff for our confirmatory study.
-
-CHECK: edit key to refer to pilot and confirmatory data instead of G and H.
 
@@
 
-<!-- The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of partners in the last year) is puzzling.
-Because men have greater strength than women their self-reported partner numbers are higher, on average, although the median number of partners in the last year for both sexes is 1.
-Still, women with higher grip strength report more partners in the past year than women with lower grip strength even after controlling for a host of potential confounds.
-It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
-It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
-It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
-genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
-It might also be the case that there are some sex-specific confounds that we failed to control for.
-Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
-Finally, our simulations revealed that we had relatively lower power to detect a significant interaction between strength and sex for this outcome (see Power section below).
--->
-
-Our pilot results support the importance of strength to male short-term mating success as well as to the long-term partnerships that are the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+<!-- Our pilot results support the importance of strength to male short-term mating success as well as to the long-term partnerships that are the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
 Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
@@
 Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for protection from predators or other men in long-term relationships is possible too [@willems2017b; @kelly2005].
+-->
 
-CHECK: moved limitations section to confirmatory discussion. 
 <!-- ## Limitations -->
@@
 We also treat this exposure as years since sexual maturity (age - 12 years).
-The changes in the strength and strength:sex coefficients from this change in model specification are minimal (see @sfig-comparison).
+The changes in the strength and $strength\times sex$ coefficients from this change in model specification are minimal (see @sfig-comparison).
 
-We also discovered that because we included dietary energy intake in the expanded controls immune model we needed to utilize the weights calculated for the dietary data rather than the weights calculated for other data for this model.
-This is corrected in the confirmatory results for that model.
-CHECK: retroactively correct for pilot results?
+As described in the methods section, a different set of weights is required in models that utilize dietary data.
+In one model of pilot data (the expanded controls immune model) we failed to do so, an oversight we correct here in both the pilot and confirmatory models.
 
 ## Results
+
 ```{r}
@@
 ```
+Our confirmatory results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners after controlling for all five sets of potential confounds.
+We found mixed evidence for a main effect of strength in the age of first sex models, as strength was not a significant predictor after controlling for socioeconomic, anthropometric, or physical activity controls.
+Strength was not a predictor of past year number of sexual partners in any control model.
+
+Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+The interaction between sex and sex-specific strength on lifetime number of sexual partners was not replicated in the confirmatory data.
+Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+See @fig-coefs for coefficients of sex-specific strength and strength X sex interaction for each proxy of mating success in both pilot and confirmatory data.
+See @fig-allcoefs for coefficients of potential confounds in each model in the confirmatory data.
+For effects plots of the effect of strength on each mating outcome in pilot and confirmatory studies (on the response scale), see @fig-effects.
 
@@
 #| fig-height: 8
-#| fig-cap: Effects of sex-specific standardized strength on mating outcomes, by sex.
+#| fig-cap: Effects of sex-specific strength on mating outcomes, by sex.
 Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
@@
 Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
+Outcomes are on the response scale.
 
@@
 ```
-Our confirmatory results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners after controlling for all five sets of potential confounds.
-There was no significant effect of sex-specific strength on the number of partners in the past year, and there were inconsistent effects of sex-specific strength on age at first sex across the potential confound models.
 
-Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
-The interaction between sex and sex-specific strength on lifetime number of sexual partners was not replicated in the confirmatory data.
-Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
-
-Effects of sex-specific standardized strength on mating outcomes, by sex.
-Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
-Plots are based on the anthropometric control models, with control variables held at their mean values.
-Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
+In our confirmatory data we found no significant effects of strength or a sex X strength interaction on WBCC in either our Lassek and Gaulin model, or the expanded controls model (@fig-immune).
 
@@
 
+Finally, we tested whether strength was related to increased energy or protein intake (@fig-energy).
+We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, strength remained a significant predictor of energy and protein intake in our confirmatory analysis.
+
 ```{r}
@@
 ## Discussion
-The purpose of this study was to replicate as closely as possible, the results from @lassek2009, in new data that included women and used combined grip strength as a proxy for upper body strength.
-We found consistent support for a positive relationship between combined grip strength and two measures of mating success: being currently partnered, and lifetime number of sexual partners for both men and women, controlling for numerous possible confounds.
+The purpose of this study was to replicate as closely as possible, the results from @lassek2009, in new data that used combined grip strength as a proxy for upper body strength.
+Unlike @lassek2009, we included women to test if the effects of upper body strength on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
+We also included numerous additional control variables, and used a registered report format with pre-specified models developed using pilot data in Stage 1 and then fit on held-out data in Stage 2.
+
+In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between combined grip strength and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
 We found no consistent association between strength and age at first sex or number of past year sexual partners.
@@
 Across both pilot and confirmatory data, we found a consistent interaction between sex and strength on partnered status, such that grip strength was more important for predicting males' partnered status than females.
-Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not present in the confirmatory data.
+Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not significant in the confirmatory data.
+
+The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of lifetime sexual partners) is puzzling.
+Because men have greater strength than women their self-reported partner numbers are higher, on average.
+Still, women with higher grip strength report more lifetime sexual partners than women with lower grip strength even after controlling for a host of potential confounds.
+
+It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
+It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
+It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
+genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
+It might also be the case that there are some sex-specific confounds that we failed to control for.
+Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
+Since this sex difference was significant in the Pilot Study (2011-12), but not in the Confirmatory Study (2013-14), it is possible that cultural changes influencing sexual behavior occurred between the data collection cycles.
 
@@
 Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
-Finally, all pilot results were obtained after considerable exploratory analysis, and therefore our confirmatory results were critical to testing our predictions.
+Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions.
+
+## Conclusion
+
+Our results most strongly support the importance of strength to male long-term mating success that is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
+More specifically, humans require provisioning for an extended period [@kaplan2000; @davison2021; @davison2022], and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
+Thus, our study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.
+Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for protection from predators or other men in long-term relationships is possible too [@willems2017b; @kelly2005].
 

### 2024-08-14 paper.qmd
Merged

--- 2024-08-13 paper.qmd
@@
 source('compare.R')
+source('combinedGH.R')
 
@@
 
+## Combined G and H series data (2011-2014)
+
+:::{#sfig-matingGH}
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+plot_strengthGH
+```
+
+Strength coefficients from models of mating fit on combined data.
+:::
+
+:::{#sfig-effectsGH}
+
+```{r}
+#| fig-width: 12
+#| fig-height: 12
+plot_effects_lifetime_GH / plot_effects_partners_GH + plot_layout(axes = 'collect', guides = 'collect')
+```
+
+Strength coefficients effects plots of mating success models fit on combined data.
+:::
+
+:::{#sfig-immunityGH}
+
+```{r}
+#| fig-width: 11
+#| fig-height: 6
+plot_immune_strength_GH
+```
+
+Strength coefficients from immunity models fit on combined data.
+:::
+
+:::{#sfig-intakeGH}
+
+```{r}
+#| fig-width: 9
+#| fig-height: 6
+plot_intake_strength_GH
+```
+
+Strength coefficients from energy and protein intake models fit on combined data.
+:::
+
 # References

### 2024-08-14 paper.qmd

--- 2024-08-14 paper.qmd
@@
 Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-It is unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse, for both males and females.
-We found consistent evidence for a sex:strength interaction on partnered status, such that strength was significantly more important for male partnered status than female.
-Our results produced mixed evidence for a sex:strength interaction on lifetime number of sexual partners.
-We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein intake but no relationship with energy intake, and a negative relationship between grip strength and native immune function.
-This study replicates an earlier study of US men (Lassek & Gaulin, 2009) with numerous additional control variables, and the expansion of these models to include women.
+Lassek & Gaulin (2009) demonstrated a positive association between male fat free mass and limb muscle volume and mating success, but did not investigate women.
+It is therefore unknown if females experience a similar tradeoff.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
+We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and native immune function.
+Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.
 
@@
 See the right column of @fig-coefs.
+```{r}
+#| label: fig-coefs
+#| fig-width: 12
+#| fig-height: 10
+#| fig-cap: "Stage 1 pilot vs.
+stage 2 confirmatory coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95% CIs.
+For the full regression tables, see the SI."
+
+# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
+plot_coefs_combined
+```
 
 For a plot of all coefficients, see @fig-allcoefs.
+
+```{r}
+#| label: fig-allcoefs
+#| fig-width: 12
+#| fig-height: 8
+#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Mexican Amercians are the base level for the ethnicity control variable.
+Bars are 95\% CIs.
+For the full regression tables, see the SI.
+
+plot_allcoefs_stage2
+```
 
@@
 
+```{r}
+#| label: fig-immune
+#| fig.width: 12
+#| fig.height: 6
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
+Variables labeled (S) have been centered at the mean and standardized by 2 SD.
+Bars are 95% CIs.
+```
+
 Previous research found that higher musculature was associated with decreased investment in native immunity.
@@
 In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was no longer significant, albeit with a coefficient that was very similar to the replication model.
-
-CHECK: remind me why we are not using strength centered by sex here.
 
@@
 
+### Marginal effects
+
+::: {#stab-marginal-confirm}
+```{r}
+marginal_stage2 |>
+  tt() |>
+  format_tt(fn = hagenutils::signif2) |>
+  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
+```
+
+Confirmatory: The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
+:::
+
+### Correlation matrices of study variables
+
+::: {#sfig-cor-confirm}
+```{r}
+
+confirm_cor_mats <- cor_mats(designsH2)
+
+confirm_cor_bothsexes <- ggcorrplot(confirm_cor_mats$cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Both sexes')
+ggsave("Figures/confirm_cor_bothsexes.pdf", confirm_cor_bothsexes, width = 12, height = 12)
+confirm_cor_bothsexes
+```
+
+Confirmatory: Correlation matrix of study variables (both sexes). Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-fcor-confirm}
+```{r}
+
+confirm_cor_females <- ggcorrplot(confirm_cor_mats$cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Females')
+ggsave("Figures/confirm_cor_females.pdf", confirm_cor_females, width = 12, height = 12)
+confirm_cor_females
+```
+
+Confirmatory: Correlation matrix of study variables for females only. Rows and columns ordered by hierarchical clustering.
+:::
+
+::: {#sfig-mcor-confirm}
+```{r}
+
+confirm_cor_males <- ggcorrplot(confirm_cor_mats$cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Males')
+ggsave("Figures/confirm_cor_males.pdf", confirm_cor_males, width = 12, height = 12)
+confirm_cor_males
+```
+
+Confirmatory: Correlation matrix of study variables for males only. Rows and columns ordered by hierarchical clustering.
+:::
+
+### Regression tables
+
+```{r}
+#| results: asis
+for (i in 1:nrow(modelsummaries_stage2)){
+  cat(paste('####', str_replace(modelsummaries_stage2$title[i], '\n', ' '))) # Remove line feed char
+  print(modelsummaries_stage2$summary[[i]])
+}
+```
+
+### Sex-specific models
+
+```{r}
+out_sexspecific_confirm <- sex_specific_models(designsH2)
+plot_sexspecific_strength_coefs_confirm <- out_sexspecific_confirm$plot
+dfmodelsummaries_confirm <- out_sexspecific_confirm$summaries
+```
+
+Confirmatory: At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
+
+:::{#sfig-sexspecific-confirm}
+```{r}
+#| fig-width: 9
+#| fig-height: 9
+plot_sexspecific_strength_coefs_confirm
+```
+
+Confirmatory: Strength coefficients in each regression model fit separately by sex.
+Control variables are the same as for models fit with a sex X strength interaction term.
+Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
+:::
+
+Regression tables for sex-specific models of the effect of strength on mating success.
+
+```{r}
+#| results: asis
+for (i in 1:nrow(dfmodelsummaries_confirm)){
+  cat(paste('####', str_replace(dfmodelsummaries_confirm$title[i], '\n', ' ')))
+  print(dfmodelsummaries_confirm$summary[[i]])
+}
+```
+
+## Combined G and H series data (2011-2014)
+
+:::{#sfig-matingGH}
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+plot_strengthGH
+```
+
+Strength coefficients from models of mating fit on combined data.
+:::
+
+:::{#sfig-effectsGH}
+
+```{r}
+#| fig-width: 12
+#| fig-height: 12
+plot_effects_lifetime_GH / plot_effects_partners_GH + plot_layout(axes = 'collect', guides = 'collect')
+```
+
+Strength coefficients effects plots of mating success models fit on combined data.
+:::
+
+:::{#sfig-immunityGH}
+
+```{r}
+#| fig-width: 11
+#| fig-height: 6
+plot_immune_strength_GH
+```
+
+Strength coefficients from immunity models fit on combined data.
+:::
+
+:::{#sfig-intakeGH}
+
+```{r}
+#| fig-width: 9
+#| fig-height: 6
+plot_intake_strength_GH
+```
+
+Strength coefficients from energy and protein intake models fit on combined data.
+:::
+
 # References

### 2024-08-15 paper.qmd
Fixed merge problems

@@
 See the right column of @fig-coefs.
-```{r}
-#| label: fig-coefs
-#| fig-width: 12
-#| fig-height: 10
-#| fig-cap: "Stage 1 pilot vs.
-stage 2 confirmatory coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95% CIs.
-For the full regression tables, see the SI."
-
-# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
-plot_coefs_combined
-```
 
 For a plot of all coefficients, see @fig-allcoefs.
-
-```{r}
-#| label: fig-allcoefs
-#| fig-width: 12
-#| fig-height: 8
-#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Mexican Amercians are the base level for the ethnicity control variable.
-Bars are 95\% CIs.
-For the full regression tables, see the SI.
-
-plot_allcoefs_stage2
-```
 
@@
 In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was now not significant, albeit with a coefficient that was very similar to the replication model.
-
-```{r}
-#| label: fig-immune
-#| fig.width: 12
-#| fig.height: 6
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
-Variables labeled (S) have been centered at the mean and standardized by 2 SD.
-Bars are 95% CIs.
-```
 
@@
 #| label: fig-coefs
-#| fig-width: 10
+#| fig-width: 12
 #| fig-height: 8
@@
 #| label: fig-allcoefs
+#| fig-width: 12
+#| fig-height: 10
 #| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.

### 2024-08-15 paper.qmd
updated version

@@
 It also aims to investigate if these costs and benefits are also experienced by women, and specifically if the effect of physical strength on mating success, a proxy for reproductive success, is greater for men than women, as the sexual selection hypothesis predicts.
+See @fig-sexualselection.
+
+:::{#fig-sexualselection}
+![](sexualselection.svg)
+
+Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
+According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by its positive effect on male mating success but not female mating success.
+:::
 

### 2024-08-23 paper.qmd
Latest edits

--- 2024-08-15 paper.qmd
@@
 ---
-title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
+title: "Strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
 author: 
@@
 The aim of this study is to replicate the costs and benefits of formidability reported by @lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
-It also aims to investigate if these costs and benefits are also experienced by women, and specifically if the effect of physical strength on mating success, a proxy for reproductive success, is greater for men than women, as the sexual selection hypothesis predicts.
+It also aims to investigate if these costs and benefits are also experienced by women, and specifically if there is a positive effect of physical strength on male mating success, a proxy for reproductive success, that is significantly greater than its effect on women's mating success, as the sexual selection hypothesis predicts.
+See @fig-sexualselection.
+
+::: {#fig-sexualselection}
+![](sexualselection.svg)
+
+Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
+According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success.
+:::
 
@@
 
-Our methods and models remained identical to those reported in the pilot study, other than the deviation reported below, using unobserved data from the 2013-2014 NHANES collection cycle.
+Our methods and models remained identical to those reported in the pilot study, other than the deviations reported below, using unobserved data from the 2013-2014 NHANES collection cycle.
 After testing and reporting our predictions using the unobserved data, we fit our models on the combined data (2011-2014), and conduct other post hoc analyses.
 
-## Deviation from Stage 1 method
+## Deviations from Stage 1 methods
 
@@
 However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, which we now do.
-We also treat this exposure as years since sexual maturity (age - 12 years).
-The changes in the strength and $strength\times sex$ coefficients from this change in model specification are minimal (see @sfig-comparison).
+We also treat this exposure as years since sexual maturity (defined this as age - 12 years because 95% of participants first had sex at age 12 or older).
+The changes  in the strength and $strength\times sex$ coefficients and their standard errors from this change in model specification are minimal (see @sfig-comparison).
 
 As described in the methods section, a different set of weights is required in models that utilize dietary data.
-In one model of pilot data (the expanded controls immune model) we failed to do so, an oversight we correct here in both the pilot and confirmatory models.
+In one of the models of pilot data using dietary variables (the expanded controls immune model) we used the MEC weights instead of the dietary weights, an oversight we correct here in both the pilot and confirmatory models.
+
+We downloaded the 2011-2014 NHANES data for this project from the CDC website on Feb 7, 2020.
+On March 20, 2020, Dual-Energy X-ray Absorptiometry (DXA) data were published for these series, which we unfortunatey failed to notice.
+These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
+Total body lean mass is similar to the bioelectrical impedence analysis fat-free mass variable used by @lassek2009.
+For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020.
+For details of the NHANES DXA measurements, see @DXXG2023.
+In addition to our grip strength models, we also fit our models exactly as specified in stage 1 except substituting four DXA lean masses for grip strength (all masses excluded bone mineral content).
+Specifically, we computed arm lean mass as the sum of left and right arm lean masses, and leg lean mass as the sum of left and right leg lean masses.
+We then fit our models using arm, leg, trunk, and total lean masses, each standardized by sex as described earlier for grip strength.
 
@@
 ```
-Our confirmatory results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners after controlling for all five sets of potential confounds.
-We found mixed evidence for a main effect of strength in the age of first sex models, as strength was not a significant predictor after controlling for socioeconomic, anthropometric, or physical activity controls.
-Strength was not a predictor of past year number of sexual partners in any control model.
+
+```{r}
+#| label: fig-leancoefs
+#| fig-width: 12
+#| fig-height: 8
+#| fig-cap: "Coefficients of sex-specific arm lean mass, and lean mass X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95% CIs."
+
+lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered")
+```
+
+Our confirmatory results revealed strong positive significant main effects of sex-specific grip strength and arm lean mass on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+We found mixed evidence for a main effect of strength  and lean mass in the age of first sex models, as these were not significant predictors after controlling for socioeconomic, anthropometric, or physical activity controls.
+Strength and arm lean mass were not predictors of past year number of sexual partners in any control model.
+Leg, trunk, and total lean mass were not consistently related to any mating success outcome.
+See @sfig-legs - @sfig-total.
 
@@
 #| fig.height: 6
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
-Variables labeled (S) have been centered at the mean and standardized by 2 SD.
-Bars are 95% CIs.
-
-# plot_immune_both <- 
-#   plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')
+#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models. Variables labeled (S) have been centered at the mean and standardized by 2 SD. Bars are 95% CIs.
 
@@
 
-# p <- plot_energy_coefs / plot_protein_coefs + plot_layout(guides = 'collect', axis_titles = 'collect')
 ggsave("Figures/plot_intake_strength_combined.pdf", plot_intake_strength_combined, width = 12, height = 6)
@@
 ## Discussion
-The purpose of this study was to replicate as closely as possible, the results from @lassek2009, in new data that used combined grip strength as a proxy for upper body strength.
-Unlike @lassek2009, we included women to test if the effects of upper body strength on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
+
+The purpose of this study was to replicate, as closely as possible, the @lassek2009 study with new data that used combined grip strength as a proxy for upper body strength, as well as the lean masses of arms, legs, trunk, and total body.
+Unlike @lassek2009, we included women to test if the effects of upper body strength and muscle masses on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
 We also included numerous additional control variables, and used a registered report format with pre-specified models developed using pilot data in Stage 1 and then fit on held-out data in Stage 2.
 
-In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between combined grip strength and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
-We found no consistent association between strength and age at first sex or number of past year sexual partners.
+In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between combined grip strength and arm lean mass, and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
+We found no consistent association between strength or lean masses, and age at first sex or number of past year sexual partners.
 
@@
 The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of lifetime sexual partners) is puzzling.
-Because men have greater strength than women their self-reported partner numbers are higher, on average.
-Still, women with higher grip strength report more lifetime sexual partners than women with lower grip strength even after controlling for a host of potential confounds.
-
+Men have greater strength than women and their average partner numbers are higher.
+Still, women with higher grip strength unexpectedly reported more lifetime sexual partners on average than women with lower grip strength even after controlling for a host of potential confounds.
 It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
-It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
+It could be that there is assortative mating on strength: if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
 It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
 genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
-It might also be the case that there are some sex-specific confounds that we failed to control for.
+It might be the case that there are some sex-specific confounds that we failed to control for, or that our models are otherwise misspecified.
+There also might be reverse causation, e.g., women who are interested in greater partner variety keep in better shape.
 Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
 Since this sex difference was significant in the Pilot Study (2011-12), but not in the Confirmatory Study (2013-14), it is possible that cultural changes influencing sexual behavior occurred between the data collection cycles.
+Or these results might simply be noise.
 
@@
 
-We thank the Editor and two anonymous reviewers for their numerous helpful comments and suggestions.
+We thank the Editor and two anonymous reviewers for their numerous helpful comments and suggestions. We thank Michael Barlev for pointing us to the DXA data.
 
@@
 
-:::{#stbl-power}
+::: {#stbl-power}
 ```{r}
@@
 :::
-
-<!-- ### Models of mating success with anthropometric, socioeconomic, health, hormone, and physical activity controls -->
 
@@
 
-:::{#sfig-sexspecific}
+::: {#sfig-sexspecific}
 ```{r}
@@
 
-:::{#sfig-comparison}
+::: {#sfig-comparison}
 ```{r}
@@
 
-:::{#sfig-sexspecific-confirm}
+::: {#sfig-sexspecific-confirm}
 ```{r}
@@
 
-:::{#sfig-matingGH}
+::: {#sfig-matingGH}
 
@@
 
-:::{#sfig-effectsGH}
+::: {#sfig-effectsGH}
 
@@
 
-:::{#sfig-immunityGH}
+::: {#sfig-immunityGH}
 
@@
 
-:::{#sfig-intakeGH}
+::: {#sfig-intakeGH}
 
@@
 
+## Lean mass coefficient plots
+
+### Arm lean mass
+
+::: {#sfig-arms}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for arm lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+### Leg lean mass
+
+::: {#sfig-legs}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Leg lean mass", "leg_lean_centered", "leg_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for leg lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+### Trunk lean mass
+
+::: {#sfig-trunk}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Trunk lean mass", "trunk_lean_centered", "trunk_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for trunk lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+### Total lean mass
+
+::: {#sfig-total}
+```{r}
+#| fig-width: 11
+lean_mass_plot("Total lean mass", "total_lean_centered", "total_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for total lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
 # References

### 2024-08-26 paper.qmd
Minor edits to ms abstract, title, and body

--- 2024-08-15 paper.qmd
@@
 ---
-title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
+title: "Grip strength, mating success, and immune and dietary costs in a population sample of US women and men: a registered report"
 author: 
@@
 source('compare.R')
-source('combinedGH.R')
 
@@
 
-Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-Lassek & Gaulin (2009) demonstrated a positive association between male fat free mass and limb muscle volume and mating success, but did not investigate women.
-It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
-We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and native immune function.
-Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.
+Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of innate immunity, likely due to an evolutionary history of sexual selection.
+It is unknown if females experience a similar tradeoff.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of US adults (N = 4384), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and partnered status, but not past year number of sexual partners or age at first intercourse, for both males and females.
+We found consistent evidence for a $strength\times sex$ interaction on partnered status, such that strength was significantly more important for male partnered status than female.
+Our results produced mixed evidence for a $strength\times sex$ interaction on lifetime number of sexual partners.
+We also tested for tradeoffs of strength with immune function and dietary intake and found a positive relationship between strength and protein intake but no relationship with energy intake, and a negative relationship between grip strength and innate immune function.
+This study replicates an earlier study of US men (Lassek & Gaulin, 2009) with numerous additional control variables, and the expansion of these models to include women.
 
@@
 
-Following @lassek2009 we also developed models of native immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day).
+Following @lassek2009 we also developed models of innate immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day).
 For each outcome we developed a model as similar as possible to those reported in @lassek2009, except that we always include sex and its interaction with other predictors where we have theoretical reasons to expect sex differences in the effect of the predictor on the outcome.
@@
 
-#### Native Immune Function
+#### Innate Immune Function
 
 Following the significant predictors reported in @lassek2009, Table 4, we first specified a replication model with age, sex, strength, and BMI.
-We also included an age:sex interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
-Finally, we included a strength:sex interaction to test the hypothesis that the immune costs of strength differ by sex.
+We also included an $age\times sex$ interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
+Finally, we included a $strength\times sex$ interaction to test the hypothesis that the immune costs of strength differ by sex.
 
@@
 
-All analyses were completed in `{r} R.version.string`, using the survey package [`{r} packageVersion("survey")`, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
+All analyses were completed in `{r} R.version.string`, using the survey package [version 4.1-1, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
 We modeled lifetime and past year numbers of partners using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
@@
 Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
-We applied this sample cutoff for all models of mating success (CHECK).
-For our tradeoff models of immunity, energy intake, and protein intake we did not restrict our sample based on reported number of lifetime partners.
+We applied this sample cutoff for all models of mating success.
+For our tradeoff models of innate immunity, energy intake, and protein intake we did not restrict our sample based on reported number of lifetime partners.
 We used the same cutoff for our confirmatory study.
@@
 
-### Coefficients of strength, and $strength \times sex$ interaction
+### Coefficients of strength, and strength X sex interaction
 
@@
 
-Previous research found that higher musculature was associated with decreased investment in native immunity.
+Previous research found that higher musculature was associated with decreased investment in innate immunity.
 We developed two models of immune investment in white blood cell count (@fig-immune).
@@
 #| label: fig-coefs
-#| fig-width: 12
+#| fig-width: 10
 #| fig-height: 8
@@
 #| label: fig-allcoefs
-#| fig-width: 12
-#| fig-height: 10
 #| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables.
@@
 #| fig.width: 12
-#| fig.height: 6
+#| fig.height: 8
 #| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot).
@@
 
-## Stage 1: Pilot study
+## Pilot study
 
@@
 ```{r}
-#| fig-width: 9
+#| fig-width: 7
 #| fig-height: 9
@@
 
-## Stage 2: Confirmatory study
+## Stage 2 confirmatory study
 
@@
 ```{r}
-#| fig-width: 10
-plot_compare_age_models
+plot_compare
 ```
 
-Pilot: A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+A comparison of strength and $strength\times sex$ coefficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
 Both models are quasi-Poisson models.
@@
 
-### Marginal effects
-
-::: {#stab-marginal-confirm}
-```{r}
-marginal_stage2 |>
-  tt() |>
-  format_tt(fn = hagenutils::signif2) |>
-  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
-```
-
-Confirmatory: The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
-:::
-
-### Correlation matrices of study variables
-
-::: {#sfig-cor-confirm}
-```{r}
-
-confirm_cor_mats <- cor_mats(designsH2)
-
-confirm_cor_bothsexes <- ggcorrplot(confirm_cor_mats$cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Both sexes')
-ggsave("Figures/confirm_cor_bothsexes.pdf", confirm_cor_bothsexes, width = 12, height = 12)
-confirm_cor_bothsexes
-```
-
-Confirmatory: Correlation matrix of study variables (both sexes). Rows and columns ordered by hierarchical clustering.
-:::
-
-::: {#sfig-fcor-confirm}
-```{r}
-
-confirm_cor_females <- ggcorrplot(confirm_cor_mats$cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Females')
-ggsave("Figures/confirm_cor_females.pdf", confirm_cor_females, width = 12, height = 12)
-confirm_cor_females
-```
-
-Confirmatory: Correlation matrix of study variables for females only. Rows and columns ordered by hierarchical clustering.
-:::
-
-::: {#sfig-mcor-confirm}
-```{r}
-
-confirm_cor_males <- ggcorrplot(confirm_cor_mats$cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Confirmatory: Males')
-ggsave("Figures/confirm_cor_males.pdf", confirm_cor_males, width = 12, height = 12)
-confirm_cor_males
-```
-
-Confirmatory: Correlation matrix of study variables for males only. Rows and columns ordered by hierarchical clustering.
-:::
-
-### Regression tables
-
-```{r}
-#| results: asis
-for (i in 1:nrow(modelsummaries_stage2)){
-  cat(paste('####', str_replace(modelsummaries_stage2$title[i], '\n', ' '))) # Remove line feed char
-  print(modelsummaries_stage2$summary[[i]])
-}
-```
-
-### Sex-specific models
-
-```{r}
-out_sexspecific_confirm <- sex_specific_models(designsH2)
-plot_sexspecific_strength_coefs_confirm <- out_sexspecific_confirm$plot
-dfmodelsummaries_confirm <- out_sexspecific_confirm$summaries
-```
-
-Confirmatory: At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
-
-:::{#sfig-sexspecific-confirm}
-```{r}
-#| fig-width: 9
-#| fig-height: 9
-plot_sexspecific_strength_coefs_confirm
-```
-
-Confirmatory: Strength coefficients in each regression model fit separately by sex.
-Control variables are the same as for models fit with a sex X strength interaction term.
-Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
-:::
-
-Regression tables for sex-specific models of the effect of strength on mating success.
-
-```{r}
-#| results: asis
-for (i in 1:nrow(dfmodelsummaries_confirm)){
-  cat(paste('####', str_replace(dfmodelsummaries_confirm$title[i], '\n', ' ')))
-  print(dfmodelsummaries_confirm$summary[[i]])
-}
-```
-
-## Combined G and H series data (2011-2014)
-
-:::{#sfig-matingGH}
-
-```{r}
-#| fig-width: 12
-#| fig-height: 9
-plot_strengthGH
-```
-
-Strength coefficients from models of mating fit on combined data.
-:::
-
-:::{#sfig-effectsGH}
-
-```{r}
-#| fig-width: 12
-#| fig-height: 12
-plot_effects_lifetime_GH / plot_effects_partners_GH + plot_layout(axes = 'collect', guides = 'collect')
-```
-
-Strength coefficients effects plots of mating success models fit on combined data.
-:::
-
-:::{#sfig-immunityGH}
-
-```{r}
-#| fig-width: 11
-#| fig-height: 6
-plot_immune_strength_GH
-```
-
-Strength coefficients from immunity models fit on combined data.
-:::
-
-:::{#sfig-intakeGH}
-
-```{r}
-#| fig-width: 9
-#| fig-height: 6
-plot_intake_strength_GH
-```
-
-Strength coefficients from energy and protein intake models fit on combined data.
-:::
-
 # References

### 2024-08-26 paper.qmd

--- 2024-08-23 paper.qmd
@@
 ---
-title: "Strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
+title: "Strength, mating success, and immune and dietary costs in a population sample of US women and men: a registered report"
+
 author: 
@@
 
-Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-Lassek & Gaulin (2009) demonstrated a positive association between male fat free mass and limb muscle volume and mating success, but did not investigate women.
+Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of innate immunity, likely due to an evolutionary history of sexual selection.
+Lassek & Gaulin (2009) demonstrated a positive association between male fat free mass and limb muscle volume and mating success but did not investigate women.
 It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
 We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
-We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and native immune function.
+We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant relationship between grip strength and innate immune function.
 Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.

### 2024-08-29 paper.qmd
added lean masses

--- 2024-08-23 paper.qmd
@@
 bibliography: paper.bib
+tbl-cap-location: bottom
 crossref:
@@
 It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
-We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and native immune function.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength and arm lean mass, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength and arm lean mass on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
+We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength and arm lean mass were significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+We also tested for tradeoffs of strength and lean masses with immune and dietary intake and found a positive relationship between strength an lean masses and protein and energy intake, and a negative association between lean masses and native immune function, but no significant  relationship between grip strength and native immune function.
 Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.
@@
 Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
-According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success.
+According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., an interaction (the hypothesis does not necessarily predict no relationship between female strength and mating success, however).
 :::
@@
 Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes, denoted as Strength (S) in results below.
+(See below for a deviation from registration where, in addition to grip strength, we also fit models using lean body masses.)
 
@@
 
-| Outcome            | Strength coefficient  | Strength X Sex coefficient |
-|--------------------|:---------------------:|:--------------------------:|
-| Partnered status   |           +   ✓      |              -   ✓         |
-| Lifetime partners  |           +   ✓      |              -   ❌         |
-| Past year partners |           0   ✓      |              0   ✓         |
-| Age at first sex   |           0   ✓      |              0   ✓         |
-| Immunity           |           -   ❌      |              0   ✓         |
-| Energy             |           0   ❌      |                            |
-| Protein            |           +   ✓      |                            |
-
-: Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model.
-A blank indicates not relevant.
+```{r}
+#| label: tbl-predictions
+#| tbl-cap: "Predictions.
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength and arm lean mass; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
 ✓: Supported in confirmatory study.
 ❌: Not supported in confirmatory study.
-{#tbl-predictions}
+*Supported for lean masses only."
+
+tribble(
+              ~Outcome, ~Prediction, ~Result, ~Prediction, ~Result,
+    "Partnered status",     "+",       "✓",       "-",       "✓",
+   "Lifetime partners",     "+",       "✓",       "-",       "❌",
+  "Past year partners",     "0",       "✓",       "0",       "✓",
+    "Age at first sex",     "0",       "✓",       "0",       "✓",
+            "Immunity",     "-",       "✓*",      "0",       "✓",
+              "Energy",     "0",       "✓",        "",        "",
+             "Protein",     "+",       "✓",        "",        ""
+  ) |>      
+  tt() |> 
+  style_tt(j=c(1,3), align = 'l') |> 
+  group_tt(j = list("Strength coefficient" = 2:3, "Strength X sex coefficient" = 4:5))
+```
 
@@
 
-In our confirmatory data we found no significant effects of strength or a sex X strength interaction on WBCC in either our Lassek and Gaulin model, or the expanded controls model (@fig-immune).
+In our confirmatory data we found no significant effects of grip strength or a sex X grip strength interaction on WBCC in either our Lassek and Gaulin model, or the expanded controls model (@fig-immune).
+However, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC.
+See @sfig-arms - @sfig-total.
 
@@
 
-Finally, we tested whether strength was related to increased energy or protein intake (@fig-energy).
-We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, strength remained a significant predictor of energy and protein intake in our confirmatory analysis.
+Finally, we tested whether strength was related to increased energy or protein intake.
+We found that strength and all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, strength remained a significant predictor of energy and protein intake in our confirmatory analysis and for all lean masses.
+See @fig-energy and @sfig-arms - @sfig-total.
 
@@
 The purpose of this study was to replicate, as closely as possible, the @lassek2009 study with new data that used combined grip strength as a proxy for upper body strength, as well as the lean masses of arms, legs, trunk, and total body.
-Unlike @lassek2009, we included women to test if the effects of upper body strength and muscle masses on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
+Unlike @lassek2009, we included women to test if the effects of upper body strength and lean masses on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
 We also included numerous additional control variables, and used a registered report format with pre-specified models developed using pilot data in Stage 1 and then fit on held-out data in Stage 2.
@@
 Men have greater strength than women and their average partner numbers are higher.
-Still, women with higher grip strength unexpectedly reported more lifetime sexual partners on average than women with lower grip strength even after controlling for a host of potential confounds.
+Still, women with higher grip strength and arm lean mass unexpectedly reported more lifetime sexual partners on average than women with lower grip strength or arm lean mass, even after controlling for a host of potential confounds.
+A meta-analysis of 77 species, including humans and 7 other mammal species, found a positive Bateman gradient for females, i.e., a positive correlation between mating success and reproductive success, although authors and commentators both caution that correlation is not causation [@fromonteil2023; @kokko2023].
+It is also not clear why women's strength would be positively associated with partner numbers.
+
 It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
@@
 Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
-Since this sex difference was significant in the Pilot Study (2011-12), but not in the Confirmatory Study (2013-14), it is possible that cultural changes influencing sexual behavior occurred between the data collection cycles.
+Since a sex difference was significant in the Pilot Study (2011-12), but not in the Confirmatory Study (2013-14), it is possible that cultural changes influencing sexual behavior occurred between the data collection cycles.
 Or these results might simply be noise.
+
+Finally, we found consistent evidence for a tradeoff between grip strength and all lean masses, and immunity and energy and protein intake.
+Grip strength and all lean masses were positively associated with energy and protein intake in our confirmatory models, and our proxy for immunity (WBCC) was significantly negatively associated with all lean masses (but was not significantly associated with grip strength).
 
@@
 
-Our results most strongly support the importance of strength to male long-term mating success that is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+Our results most strongly support the importance of strength to male long-term mating success, which is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+This benefit would compensate for the energy, protein, and immune costs associated with strength and lean masses in our confirmatory results.
 Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
-More specifically, humans require provisioning for an extended period [@kaplan2000; @davison2021; @davison2022], and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
+More specifically, human children require provisioning for an extended period [@kaplan2000; @davison2021; @davison2022], and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
 Thus, our study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.

### 2024-08-29 combinedGH.R

@@ -102,3 +102,7 @@ plot_effects_partners_GH <-
     ) +
   theme_minimal(20) +
     theme(axis.title.y = element_text(angle = 0))
+
+# Correlation of grip strength with lean masses --------------
+
+cor_out <- svycor(~ArmLeanexclBMC+LegLeanexclBMC+TrunkLeanexclBMC+TotalLeanexclBMC+strength, designsGH2$d.design.adults, na.rm=T)

### 2024-08-29 functions.R

@@ -24,9 +24,13 @@ control_dict <- c(
 
 outcome_dict <- c(
   "1" = "Lifetime partners\n(partners per year)",
+  "1b" = "Lifetime partners\n(partners per year)",
   "2" = "Past year partners",
+  "2b" = "Past year partners",
   "3" = "Age of first sex",
+  "3b" = "Age of first sex",
   "4" = "Partnered",
+  "4b" = "Partnered",
   "5" = "Immunity",
   "6" = "Energy (kcal)",
   "7" = "Protein (g)"
@@ -102,16 +106,65 @@ vnames <- c(
   "raceOtherRace" = "Other Race",
   "foodinsecurity_adult" = "Food Insecurity",
   "tot_MET_centered" = "Total MET (S)",
-  "avgcalories_centered" = "Average calories per day (S)"
+  "avgcalories_centered" = "Average calories per day (S)",
+
+  "arm_lean_centered" = "Arm lean mass (S)",
+  "leg_lean_centered" = "Legs lean mass (S)",
+  "trunk_lean_centered" = "Trunk lean mass (S)",
+  "upper_lean_centered" = "Upper lean mass (S)",
+  "total_lean_centered" = "Total lean mass (S)",
+
+  "arm_lean_sex_centered" = "Arm lean mass (S by sex)",
+  "leg_lean_sex_centered" = "Legs lean mass (S by sex)",
+  "trunk_lean_sex_centered" = "Trunk lean mass (S by sex)",
+  "total_lean_sex_centered" = "Total lean mass (S by sex)",
+
+  "arm_lean_sex_centered:sexfemale" = "Sex (female) x Arm lean mass (S by sex)",
+  "leg_lean_sex_centered:sexfemale" = "Sex (female) x Legs lean mass (S by sex)",
+  "trunk_lean_sex_centered:sexfemale" = "Sex (female) x Trunk lean mass (S by sex)",
+  "upper_lean_sex_centered:sexfemale" = "Sex (female) x Upper lean mass (S by sex)",
+  "total_lean_sex_centered:sexfemale" = "Sex (female) x Total lean mass (S by sex)",
+
+  "sexfemale:arm_lean_sex_centered" = "Sex (female) x Arm lean mass (S by sex)",
+  "sexfemale:leg_lean_sex_centered" = "Sex (female) x Legs lean mass (S by sex)",
[…]

### 2024-08-29 stage2_analysis.R

@@ -62,7 +62,7 @@ plot_coefs_combined <-
   geom_vline(xintercept = 0, linetype = 'dotted') +
   scale_color_binary() +
   guides(colour = guide_legend(reverse = T)) +
-  labs(x = 'Estimate (95% CI)', y = '') +
+  labs(title = 'Grip strength', x = 'Estimate (95% CI)', y = '') +
   facet_grid(Outcome ~ term) +
   theme_bw(15) +
   theme(strip.text.y = element_text(angle = 0))

### 2024-08-29 paper.qmd

@@
 
-Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
+Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of innate immunity, likely due to an evolutionary history of sexual selection.
 Lassek & Gaulin (2009) demonstrated a positive association between male fat free mass and limb muscle volume and mating success, but did not investigate women.
 It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength and arm lean mass, on mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength and arm lean mass on mating success proxied by lifetime number of sexual partners and current partnership status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength and arm lean mass were significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
-We also tested for tradeoffs of strength and lean masses with immune and dietary intake and found a positive relationship between strength an lean masses and protein and energy intake, and a negative association between lean masses and native immune function, but no significant  relationship between grip strength and native immune function.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
+We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and innate immune function.
 Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.

### 2024-09-06 paper.qmd
edits made to conclusion (added projectile stuff) and moved prediction table to results

@@
 We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $sex \times strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
-We also tested for tradeoffs of strength with immune and dietary intake and found a positive relationship between strength and protein and energy intake, but no significant  relationship between grip strength and innate immune function.
+We found consistent evidence for a $sex \times grip strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+We also tested for tradeoffs of upper body strength with immune and dietary intake and found a positive relationship between grip strength and protein and energy intake, but no significant  relationship between grip strength and innate immune function.
 Our results suggest that sexually dimorphic upper body strength might have evolved, in part, by increasing male long-term mating success.
@@
 At present there is mixed evidence that strength is related to reproductive success in women [@lidborg2022a].
-One study among the Himba showed that women with higher strength/muscularity had more living children and grandchildren [@atkinson2012], while a similar study among the Hadza showed no relationship between strength and reproductive outcomes [@smith2017a].
+One study among the Himba showed that women with higher handgrip strength/muscularity had more living children and grandchildren [@atkinson2012], while a similar study among the Hadza showed no relationship between strength and reproductive outcomes [@smith2017a].
 The dearth of research on female traits and reproductive outcomes leaves us unable to draw any firm conclusions about selection pressures over time.
@@
 
-Finally, to assess costs of greater strength, we investigated the relationship between strength and dietary energy and protein intake.
+Finally, to assess costs of greater upper body strength, we investigated the relationship between grip strength and dietary energy and protein intake.
 Participants reported all food and beverages they consumed in the 24 hour period prior to their interview.
@@
 
-The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
+The interaction between sex and grip strength was crucial for our models since we want to test if the effect of grip strength on mating and immune outcomes is different for women than for men.
 Grip strength is highly sexually dimorphic, however, with about 90% of men stronger than about 90% of women [@smith2022], and is thus highly collinear with sex, a problem for regression modeling and interpretation.
@@
 White blood cell count (1000 cells/µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression, which is negatively related to strength [@smith2022], was measured using the Patient Health Questionnaire-9 [PHQ-9, @kroenke2001], a validated nine-item screening instrument.
+Depression, which is negatively related to grip strength [@smith2022], was measured using the Patient Health Questionnaire-9 [PHQ-9, @kroenke2001], a validated nine-item screening instrument.
 Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
@@
 For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see @vananders2013.
-Likewise, although some have found a positive correlation between circulating testosterone and handgrip strength [@Chiu_2019], other studies have found no association [@Gettler_2010; @Ribeiro_2016].
+Likewise, although some have found a positive correlation between circulating testosterone and grip strength [@Chiu_2019], other studies have found no association [@Gettler_2010; @Ribeiro_2016].
 There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance [@Guilherme_2021].
@@
 Since male mean testosterone is an order of magnitude higher than the female mean, testosterone, like grip strength, was highly confounded with sex, one of our key predictor variables.
-We therefore computed a sex-specific testosterone value in the same way we did for strength.
+We therefore computed a sex-specific testosterone value in the same way we did for grip strength.
 This differed from other centered and standardized variables which we centered and standardized across all male and female values.
@@
 #| tbl-cap: "Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength and arm lean mass; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
 ✓: Supported in confirmatory study.
-❌: Not supported in confirmatory study.
-*Supported for lean masses only."
+❌: Not supported in confirmatory study."
 
 tribble(
-              ~Outcome, ~Prediction, ~Result, ~Prediction, ~Result,
-    "Partnered status",     "+",       "✓",       "-",       "✓",
-   "Lifetime partners",     "+",       "✓",       "-",       "❌",
-  "Past year partners",     "0",       "✓",       "0",       "✓",
-    "Age at first sex",     "0",       "✓",       "0",       "✓",
-            "Immunity",     "-",       "✓*",      "0",       "✓",
-              "Energy",     "0",       "✓",        "",        "",
-             "Protein",     "+",       "✓",        "",        ""
+              ~Outcome, ~`Grip Strength Coefficient`, ~`Grip Strength X Sex Coefficient`,
+    "Partnered status",     "+",             "-",       
+   "Lifetime partners",     "+",             "-",       
+  "Past year partners",     "0",             "0",       
+    "Age at first sex",     "0",             "0",       
+            "Immunity",     "-",            "0",       
+              "Energy",     "0",              "",       
+             "Protein",     "+",              "",       
   ) |>    
   tt() |> 
-  style_tt(j=c(1,3), align = 'l') |> 
-  group_tt(j = list("Strength coefficient" = 2:3, "Strength X sex coefficient" = 4:5))
-```
-
+  style_tt(j=c(1), align = 'l') 
+```
 ## Power
@@
 
-We downloaded the 2011-2014 NHANES data for this project from the CDC website on Feb 7, 2020.
-On March 20, 2020, Dual-Energy X-ray Absorptiometry (DXA) data were published for these series, which we unfortunatey failed to notice.
+We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study.
 These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
-Total body lean mass is similar to the bioelectrical impedence analysis fat-free mass variable used by @lassek2009.
+Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009.
 For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020.
@@
 plot_intake_strength_combined
+```
+
+```{r}
+#| label: tbl-predictionsresults
+#| tbl-cap: "Predictions.
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
+✓: Supported in confirmatory study.
+❌: Not supported in confirmatory study."
+
+tribble(
+              ~Outcome, ~Prediction, ~Result, ~Prediction, ~Result,
+    "Partnered status",     "+",       "✓",       "-",       "✓",
+   "Lifetime partners",     "+",       "✓",       "-",       "❌",
+  "Past year partners",     "0",       "✓",       "0",       "✓",
+    "Age at first sex",     "0",       "✓",       "0",       "✓",
+            "Immunity",     "-",       "❌",      "0",       "✓",
+              "Energy",     "0",       "❌",        "",        "",
+             "Protein",     "+",       "✓",        "",        ""
+  ) |>      
+  tt() |> 
+  style_tt(j=c(1,3), align = 'l') |> 
+  group_tt(j = list("Strength coefficient" = 2:3, "Strength X sex coefficient" = 4:5))
 ```
@@
 Our results most strongly support the importance of strength to male long-term mating success, which is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
-This benefit would compensate for the energy, protein, and immune costs associated with strength and lean masses in our confirmatory results.
+This benefit would compensate for the energy, protein, and immune costs associated with strength in our confirmatory results and lean masses in our *post hoc* results.
 Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
@@
 Thus, our study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.
-Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for protection from predators or other men in long-term relationships is possible too [@willems2017b; @kelly2005].
+
+Moreover, in addition to formidability, which is important in male-male competition for short and long-term partners, upper body strength indexes productivity.
+Specifically, strength in forearm and shoulder is critical for effective use of projectile hunting technology such as the bow and arrow [@ertan2003] and javelin [@kim2014].
+Indeed, @apicella2014 found that upper body strength predicted both hunting reputation and reproductive success in Hadza men.
+Given the similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular.
+Therefore, our results could support an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for long-term partners who can confer protection from predators or other men is possible too [@willems2017b; @kelly2005].
 

### 2024-09-06 paper.qmd
Changed sexualselection plot. Minor edits.

--- 2024-08-29 paper.qmd
@@
 Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
-According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., an interaction (the hypothesis does not necessarily predict no relationship between female strength and mating success, however).
+According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex:strength interaction (the hypothesis does allow some positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
 :::
@@
 All analyses were completed in `{r} R.version.string`, using the survey package [`{r} packageVersion("survey")`, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-We modeled lifetime and past year numbers of partners using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
-We modeled age at first sexual intercourse and white blood cell count with a Gaussian glm.
+We modeled lifetime past year numbers of partners, and white blood cell count using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
+We modeled age at first sexual intercourse and energy and protein intakes with a Gaussian glm.
 We modeled partnered status using a binomial glm.

### 2024-09-13 paper.qmd
Edits made to text, clarifying grip strength and sex-specific grip strength. Moved post hoc analyses to separate section. Added Kim2014 to bib file.

@@
 We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $sex \times grip strength$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+We found consistent evidence for a $grip\ strength\times sex$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
 We also tested for tradeoffs of upper body strength with immune and dietary intake and found a positive relationship between grip strength and protein and energy intake, but no significant relationship between grip strength and innate immune function.
@@
 Our main explanatory variable of mating success is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
-Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
+Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer (hereafter 'grip strength').
 All regression models also included sex (male/female), our second main explanatory variable.
@@
 Thus, for women, low or high values are relative to other women, and for men, low or high values are relative to other men.
-We denote the use of this predictor in figures below as 'Strength (S by sex).' The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high strength.
-Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes, denoted as Strength (S) in results below.
-(See below for a deviation from registration where, in addition to grip strength, we also fit models using lean body masses.)
+We denote the use of this predictor in text as 'sex-specific grip strength' and figures below as 'Grip Strength (S by sex).' The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high upper body strength.
+Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes, denoted as Grip Strength (S) in results below.
+(See the post hoc analyses section, where in addition to grip strength, we also fit models using lean body masses.)
 
@@
 We interacted age with sex since menopause occurs within the age range of our participants.
-For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
+For models of past year number of sexual partners we included an interaction between partnered status and sex-specific grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
 Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable [@gelman2008].
@@
 Control variables included a much wider range of theoretically motivated potential confounds than used by @lassek2009, organized into themes.
-We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated with strength and sexual behavior.
+We specified five models for each mating success outcome measure in order to determine if the effect of sex-specific grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated with strength and sexual behavior.
 
@@
 
-We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
+We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, sex-specific grip strength, and partnered status.
 This model also includes an anthropometric control, body mass index (BMI), calculated as $kg/m^2$, since body size could impact either strength or mating outcomes.
-We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.
+We interacted age, sex-specific grip strength, and BMI with sex to assess differences in these predictors for men and women.
 
@@
 
-Following the significant predictors reported in @lassek2009, Table 4, we first specified a replication model with age, sex, strength, and BMI.
+Following the significant predictors reported in @lassek2009, Table 4, we first specified a replication model with age, sex, grip strength, and BMI.
 We also included an $age\times sex$ interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
-Finally, we included a $strength\times sex$ interaction to test the hypothesis that the immune costs of strength differ by sex.
+Finally, we included a $grip\ strength\times sex$ interaction to test the hypothesis that the immune costs of strength differ by sex.
 
@@
 
-Based on @lassek2009, Table 3, we developed a replication model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
+Based on @lassek2009, Table 3, we developed a replication model of dietary energy intake (average calories per day) and included age, sex, grip strength, BMI, and total MET.
 We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.
@@
 All analyses were completed in `{r} R.version.string`, using the survey package [`{r} packageVersion("survey")`, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-We modeled lifetime past year numbers of partners, and white blood cell count using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
+We modeled lifetime number of partners, past year numbers of partners, and white blood cell count using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
 We modeled age at first sexual intercourse and energy and protein intakes with a Gaussian glm.
@@
 
-Our initial analyses found little support for a sex difference in the effect of strength on mating success, contrary to the sexual selection hypothesis.
+Our initial analyses found little support for a sex difference in the effect of sex-specific grip strength on mating success, contrary to the sexual selection hypothesis.
 We traced this failure to the unusually high number of lifetime sex partners reported by some men and women.
@@
 Sexual dimorphism---the ratio of male mean:female mean---was evident in an 8% higher value for male height and a 16% higher value for weight; females, however, had a 1% higher BMI than males on average.
-There was a much larger degree of male-biased dimorphism in combined grip strength (57%).
+There was a much larger degree of male-biased dimorphism in grip strength (57%).
 Hemoglobin was 14% higher in males than females while white blood cell count was 3% higher in women.
@@
 
-### Coefficients of strength, and $strength \times sex$ interaction
-
-Our pilot results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners, and age at first sex after controlling for all five sets of potential confounds.
-The only exception was that the strength coefficient for age at first sex was not significant after controlling for socioeconomic factors, although the sign and magnitude was very similar to those of these coefficients in the models with other control variables.
-There was no significant effect of sex-specific strength on the number of partners in the past year.
+### Coefficients of sex-specific grip strength, and sex-specific grip strength x sex interaction
+
+Our pilot results revealed strong positive significant main effects of sex-specific grip strength on partnered status, numbers of lifetime partners, and age at first sex after controlling for all five sets of potential confounds.
+The only exception was that the sex-specific grip strength coefficient for age at first sex was not significant after controlling for socioeconomic factors, although the sign and magnitude was very similar to those of these coefficients in the models with other control variables.
+There was no significant effect of sex-specific grip strength on the number of partners in the past year.
 See the left column of @fig-coefs.
 
-There were also clear sex differences in the effects of sex-specific strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+There were also clear sex differences in the effects of sex-specific grip strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
 There was also a persistent negative association of numbers of lifetime partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
-Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
 See the right column of @fig-coefs.
@@
 We developed two models of immune investment in white blood cell count (@fig-immune).
-In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
-In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was now not significant, albeit with a coefficient that was very similar to the replication model.
-
-Previous research found that higher musculature was associated with decreased investment in innate immunity.
-We developed two models of immune investment in white blood cell count (@fig-immune).
-In the Lassek and Gaulin model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
-In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was no longer significant, albeit with a coefficient that was very similar to the replication model.
+In the Lassek and Gaulin model, grip strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and grip strength) controlling for age and BMI.
+In the expanded model with additional controls, including the substitution of height and weight for BMI, grip strength was no longer significant, albeit with a coefficient that was very similar to the replication model.
 
@@
 
-Finally, we tested whether strength was related  to increased energy or protein intake (@fig-energy).
-We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, strength was no longer a significant predictor of energy intake but remained a significant predictor of protein intake.
+Finally, we tested whether grip strength was related to increased energy or protein intake (@fig-energy).
+We found that grip strength was a significant, positive predictor of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, grip strength was no longer a significant predictor of energy intake but remained a significant predictor of protein intake.
 
@@
 
-Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with two of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-The exceptions were that strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant strength coefficients in the other models.
-Strength was also associated with immune, energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
+Using sex-specific grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with two of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
+The exceptions were that sex-specific grip strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant sex-specific grip strength coefficients in the other models.
+Grip strength was also associated with immune, energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
 The expanded control models included height, which might be a proxy for aspects of strength, such as lower body strength, that are not accounted for by grip strength.
@@
 
-The effect of strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
+The effect of sex-specific grip strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
 Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
 In addition, participants, whose grip strength was measured at the time of the study when they were adults, are reporting on an event that happened an average of 22 years in the past when most were adolescents and their grip strength had not yet reached its adult value.
-For these reasons, we do not predict an effect of strength on age at first sex after controlling for ethnicity.
-
-Strength was also not significantly associated with the number of partners in the last year, despite controlling for partnered status, which was a significant negative predictor.
+For these reasons, we do not predict an effect of sex-specific grip strength on age at first sex after controlling for ethnicity.
+
+Sex-specific grip strength was also not significantly associated with the number of partners in the last year, despite controlling for partnered status, which was a significant negative predictor.
 Approximately 60% of the sample was partnered (@stbl-stats), however, and stronger individuals, especially men, were more likely to be partnered.
-The pool of unpartnered individuals who might pursue multiple mating opportunities was therefore smaller and skewed toward lower strength individuals, and variation in this outcome over the relatively short time period (1 year) was low.
-Nevertheless, the non-significant strength coefficients were all positive, raising the possibility that a higher powered study might successfully detect a positive effect of strength on number of partners in the last year.
+The pool of unpartnered individuals who might pursue multiple mating opportunities was therefore smaller and skewed toward lower grip strength individuals, and variation in this outcome over the relatively short time period (1 year) was low.
+Nevertheless, the non-significant sex-specific grip strength coefficients were all positive, raising the possibility that a higher powered study might successfully detect a positive effect of sex-specific grip strength on number of partners in the last year.
 
@@
 In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime number of sexual partners, but not last year number of partners, nor age at first sex after controlling for ethnicity.
-We also predict a significant interaction with sex such that strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
+We also predict a significant interaction with sex such that sex-specific grip strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
 We also predict that there will be a significant negative association between formidability and immune function.
@@
 
-We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength, in @tbl-predictions.
+We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of grip strength and of the interaction between sex and grip strength, in @tbl-predictions.
 We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models.
@@
 #| tbl-cap: "Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
-✓: Supported in confirmatory study.
-❌: Not supported in confirmatory study."
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant."
 
@@
 Effects in replication studies are often smaller than in the original studies [@klein2018].
-We therefore used simulations to estimate our power to detect strength and sex X strength coefficients that are 25-100% as large as the ones observed in our pilot study for partnered and lifetime partners, with a sample size equal to that in our unobserved data.
+We therefore used simulations to estimate our power to detect grip strength and $grip\ strength\times sex$ coefficients that are 25-100% as large as the ones observed in our pilot study for partnered and lifetime partners, with a sample size equal to that in our unobserved data.
 We have high power to detect effects at least 75% as large as the ones observed in our pilot study, and reasonable power to detect those that are at least 50% as large (@stbl-power).
 
-## Method
+## Confirmatory Study Methods
 
@@
 However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, which we now do.
-We also treat this exposure as years since sexual maturity (defined this as age - 12 years because 95% of participants first had sex at age 12 or older).
-The changes  in the strength and $strength\times sex$ coefficients and their standard errors from this change in model specification are minimal (see @sfig-comparison).
+We also treat this exposure as years since sexual maturity (defined as age - 12 years because 95% of participants first had sex at age 12 or older).
+The changes  in the grip strength and $grip\ strength\times sex$ coefficients and their standard errors from this change in model specification are minimal (see @sfig-comparison).
 
@@
 
-We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study.
-These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
-Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009.
-For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020.
-For details of the NHANES DXA measurements, see @DXXG2023.
-In addition to our grip strength models, we also fit our models exactly as specified in stage 1 except substituting four DXA lean masses for grip strength (all masses excluded bone mineral content).
-Specifically, we computed arm lean mass as the sum of left and right arm lean masses, and leg lean mass as the sum of left and right leg lean masses.
-We then fit our models using arm, leg, trunk, and total lean masses, each standardized by sex as described earlier for grip strength.
-
-## Results
+## Confirmatory Study Results
 
@@
 #| fig-height: 8
-#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+#| fig-cap: "Coefficients of sex-specific grip strength, and sex-specific grip strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
 Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
@@
 
-```{r}
-#| label: fig-leancoefs
-#| fig-width: 12
-#| fig-height: 8
-#| fig-cap: "Coefficients of sex-specific arm lean mass, and lean mass X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95% CIs."
-
-lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered")
-```
-
-Our confirmatory results revealed strong positive significant main effects of sex-specific grip strength and arm lean mass on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
-We found mixed evidence for a main effect of strength  and lean mass in the age of first sex models, as these were not significant predictors after controlling for socioeconomic, anthropometric, or physical activity controls.
-Strength and arm lean mass were not predictors of past year number of sexual partners in any control model.
-Leg, trunk, and total lean mass were not consistently related to any mating success outcome.
-See @sfig-legs - @sfig-total.
-
-Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
-The interaction between sex and sex-specific strength on lifetime number of sexual partners was not replicated in the confirmatory data.
-Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
-See @fig-coefs for coefficients of sex-specific strength and strength X sex interaction for each proxy of mating success in both pilot and confirmatory data.
+### Relationship between sex-specific grip strength and four measures of mating success
+
+Our confirmatory results revealed a strong positive significant main effect of sex-specific grip strength on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+We found mixed evidence for a main effect of sex-specific grip strength in the age of first sex models, as it was not a significant predictor after controlling for socioeconomic, anthropometric, or physical activity controls.
+Strength was not a predictor of past year number of sexual partners in any control model.
+
+Unlike findings from the pilot study, a clear sex difference in the effects of sex-specific grip strength only emerged for one mating success outcome---partnered status---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+The interaction between sex and sex-specific grip strength on lifetime number of sexual partners was not replicated in the confirmatory data.
+Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+See @fig-coefs for coefficients of sex-specific grip strength and $grip\ strength\times sex$ interaction for each proxy of mating success in both pilot and confirmatory data.
 See @fig-allcoefs for coefficients of potential confounds in each model in the confirmatory data.
-For effects plots of the effect of strength on each mating outcome in pilot and confirmatory studies (on the response scale), see @fig-effects.
+For effects plots of the effect of sex-specific grip strength on each mating outcome in pilot and confirmatory studies (on the response scale), see @fig-effects.
 
@@
 #| fig-height: 8
-#| fig-cap: Effects of sex-specific strength on mating outcomes, by sex.
-Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
+#| fig-cap: Effects of sex-specific grip strength on mating outcomes, by sex.
+Grip strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1).
 Plots are based on the anthropometric control models, with control variables held at their mean values.
@@
 
-In our confirmatory data we found no significant effects of grip strength or a sex X grip strength interaction on WBCC in either our Lassek and Gaulin model, or the expanded controls model (@fig-immune).
-However, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC.
-See @sfig-arms - @sfig-total.
+### Immune costs of strength
+In our confirmatory data we found no significant effects of grip strength or a $grip\ strength\times sex$ interaction on WBCC in either our Lassek and Gaulin model, or the expanded controls model (@fig-immune).
 
@@
 
+### Dietary costs of strength 
 Finally, we tested whether strength was related to increased energy or protein intake.
-We found that strength and all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, strength remained a significant predictor of energy and protein intake in our confirmatory analysis and for all lean masses.
-See @fig-energy and @sfig-arms - @sfig-total.
+We found that grip strength was a significant, positive predictor of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, grip strength remained a significant predictor of energy and protein intake in our confirmatory analysis.
+See @fig-energy.
 
@@
 
+## Predictions and Results 
 ```{r}
@@
   tt() |> 
-  style_tt(j=c(1,3), align = 'l') |> 
-  group_tt(j = list("Strength coefficient" = 2:3, "Strength X sex coefficient" = 4:5))
+  style_tt(j=c(1,3,5), align = 'l') |> 
+  group_tt(j = list("Grip strength coefficient" = 2:3, "Grip strength X sex coefficient" = 4:5))
+```
+
+## Post Hoc Analyses
+We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study.
+These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
+Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009.
+For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020.
+For details of the NHANES DXA measurements, see @DXXG2023.
+In addition to our grip strength models, we also fit our models exactly as specified in stage 1 except substituting four DXA lean masses for grip strength (all masses excluded bone mineral content).
+Specifically, we computed arm lean mass as the sum of left and right arm lean masses, and leg lean mass as the sum of left and right leg lean masses.
+We then fit our models using arm, leg, trunk, and total lean masses, each standardized by sex as described earlier for grip strength.
+
+Our post hoc analyses revealed that coefficients for sex-specific arm lean mass were generally similar in direction to those of sex-specific grip strength discussed above, including a strong positive significant main effect on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+We found mixed evidence for a main effect of sex-specific lean mass in the age of first sex models, as these were not significant predictors after controlling for socioeconomic or physical activity controls.
+As with sex-specific grip strength, sex-specific arm lean mass was not a predictor of past year number of sexual partners in any control model.
+Sex-specific leg, trunk, and total lean mass were not consistently related to any mating success outcome.
+See @sfig-legs - @sfig-total.
+
+Unlike grip strength, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC.
+We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, all lean masses remained a significant predictor of energy and protein intake.
+See @fig-energy and @sfig-arms - @sfig-total.
+
+```{r}
+#| label: fig-leancoefs
+#| fig-width: 12
+#| fig-height: 8
+#| fig-cap: "Coefficients of sex-specific arm lean mass, and lean mass X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
+Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
+Bars are 95% CIs."
+
+lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered")
 ```
@@
 
-In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between combined grip strength and arm lean mass, and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
-We found no consistent association between strength or lean masses, and age at first sex or number of past year sexual partners.
+In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between both sex-specific combined grip strength and sex-specific arm lean mass, and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
+We found no consistent association between either sex-specific combined grip strength or any sex-specific lean masses and age at first sex or number of past year sexual partners.
 
 We also tested for interactions between sex and strength on these mating success outcomes.
-Across both pilot and confirmatory data, we found a consistent interaction between sex and strength on partnered status, such that grip strength was more important for predicting males' partnered status than females.
-Although in the pilot data we did see evidence of an interaction between sex and strength on the outcome lifetime number of sexual partners, such that strength was more important for predicting males' lifetime sexual partners, this effect was not significant in the confirmatory data.
-
-The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of lifetime sexual partners) is puzzling.
+Across both pilot and confirmatory data, we found a consistent interaction between sex and sex-specific grip strength on partnered status, such that sex-specific grip strength was more important for predicting males' partnered status than females.
+Although in the pilot data we did see evidence of an interaction between sex and sex-specific grip strength on the outcome lifetime number of sexual partners, such that sex-specific grip strength was more important for predicting males' lifetime sexual partners, this effect was not significant in the confirmatory data.
+
+The lack of a sex difference in the significant positive effect of sex-specific grip strength on short term mating success (proxied by number of lifetime sexual partners) is puzzling.
 Men have greater strength than women and their average partner numbers are higher.
-Still, women with higher grip strength and arm lean mass unexpectedly reported more lifetime sexual partners on average than women with lower grip strength or arm lean mass, even after controlling for a host of potential confounds.
+Still, women with higher sex-specific grip strength and sex-specific arm lean mass unexpectedly reported more lifetime sexual partners on average than women with lower grip strength or arm lean mass, even after controlling for a host of potential confounds.
 A meta-analysis of 77 species, including humans and 7 other mammal species, found a positive Bateman gradient for females, i.e., a positive correlation between mating success and reproductive success, although authors and commentators both caution that correlation is not causation [@fromonteil2023; @kokko2023].
@@
 
-Finally, we found consistent evidence for a tradeoff between grip strength and all lean masses, and immunity and energy and protein intake.
+Finally, we found consistent evidence for a tradeoff between both grip strength and all lean masses, and immune investment and energy and protein intake.
 Grip strength and all lean masses were positively associated with energy and protein intake in our confirmatory models, and our proxy for immunity (WBCC) was significantly negatively associated with all lean masses (but was not significantly associated with grip strength).
@@
 
-Our results most strongly support the importance of strength to male long-term mating success, which is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
+Our results most strongly support the importance of upper body strength to male long-term mating success, which is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
 This benefit would compensate for the energy, protein, and immune costs associated with strength in our confirmatory results and lean masses in our *post hoc* results.
@@
 Indeed, @apicella2014 found that upper body strength predicted both hunting reputation and reproductive success in Hadza men.
-Given the similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular.
+The similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular.
 Therefore, our results could support an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for long-term partners who can confer protection from predators or other men is possible too [@willems2017b; @kelly2005].
@@
 
-Power to detect strength and sex X strength coefficients that range from 25% to 100% of the size seen in the pilot study, based on models with anthropometric controls.
+Power to detect grip strength and sex X grip strength coefficients that range from 25% to 100% of the size seen in the pilot study, based on models with anthropometric controls.
 **Scale**: the proportion of the effect seen in the pilot study.
@@
 
-The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
+The average marginal effect of a 2 standard deviation increase in grip strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
 :::
@@
 
-At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
+At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the grip strength coefficients for each model for each sex.
 
@@
 
-Strength coefficients in each regression model fit separately by sex.
-Control variables are the same as for models fit with a sex X strength interaction term.
+Grip strength coefficients in each regression model fit separately by sex.
+Control variables are the same as for models fit with a sex X grip strength interaction term.
 Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
@@
 
-Regression tables for sex-specific models of the effect of strength on mating success.
+Regression tables for sex-specific models of the effect of grip strength on mating success.
 
@@
 
-Pilot: A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+Pilot: A comparison of grip strength and $grip\ strength\times sex$ coefficients with the original model specification for lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
 Both models are quasi-Poisson models.
@@
 
-Confirmatory: The average marginal effect of a 2 standard deviation increase in strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
+Confirmatory: The average marginal effect of a 2 standard deviation increase in grip strength on each mating outcome, by sex and partnered status, for the models with anthropometric controls.
 :::
@@
 
-Confirmatory: At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the strength coefficients for each model for each sex.
+Confirmatory: At the suggestion of an anonymous reviewer, we fit all our regression models separately by sex. We plotted the grip strength coefficients for each model for each sex.
 
@@
 Confirmatory: Strength coefficients in each regression model fit separately by sex.
-Control variables are the same as for models fit with a sex X strength interaction term.
-Note that these models test if each strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
-:::
-
-Regression tables for sex-specific models of the effect of strength on mating success.
+Control variables are the same as for models fit with a $grip\ strength\times sex$ interaction term.
+Note that these models test if each grip strength coefficient is significantly different from 0, not if the male and female coefficients are significantly different from each other.
+:::
+
+Regression tables for sex-specific models of the effect of grip strength on mating success.
 
@@
 
-Strength coefficients from models of mating fit on combined data.
+Grip trength coefficients from models of mating fit on combined data.
 :::
@@
 
-Strength coefficients effects plots of mating success models fit on combined data.
+Grip strength coefficients effects plots of mating success models fit on combined data.
 :::
@@
 
-Strength coefficients from immunity models fit on combined data.
+Grip strength coefficients from immunity models fit on combined data.
 :::
@@
 
-Strength coefficients from energy and protein intake models fit on combined data.
+Grip strength coefficients from energy and protein intake models fit on combined data.
 :::

### 2024-09-13 paper.qmd
Added note to caption regarding posthoc results

--- 2024-09-06 paper.qmd
@@
 ✓: Supported in confirmatory study.
-❌: Not supported in confirmatory study."
+❌: Not supported in confirmatory study.
+*The posthoc results found a significant negative association between lean mass and immunity, and a positive association with energy."
 
@@
     "Age at first sex",     "0",       "✓",       "0",       "✓",
-            "Immunity",     "-",       "❌",      "0",       "✓",
-              "Energy",     "0",       "❌",        "",        "",
+            "Immunity",     "-",       "❌*",      "0",       "✓",
+              "Energy",     "0",       "❌*",        "",        "",
              "Protein",     "+",       "✓",        "",        ""

### 2024-09-13 extract-bib.R
Added comment about arm lean mass interaction for partnered status

@@ -1,0 +1,30 @@
+library(tidyverse)
+
+clean_bib <- function(input_file, input_bib){
+  lines <- paste(readLines(input_file), collapse = "")
+  entries <- unique(str_match_all(lines, "@([a-zA-Z0-9_]+)[,;\\. \\?\\!\\]]")[[1]][, 2])
+  return(entries)
+  # bib <- paste(readLines(input_bib), collapse = "\n")
+  # bib <- unlist(strsplit(bib, "\n@"))
+  #
+  # output <- sapply(entries, grep, bib, value = T)
+  # output <- paste("@", output, sep = "")
+  #
+  # writeLines(unlist(output), output_bib)
+}
+
+# now call the function
+# clean_bib(...)
+
+
+keys <- clean_bib("paper.v2.qmd", "regrep.bib")
+
+library(bib2df)
+d <- bib2df("regrep.bib")
+
+d_used <-
+  d |>
+  dplyr::filter(BIBTEXKEY %in% keys) |>
+  dplyr::select(-ABSTRACT, -FILE)
+
+df2bib(d_used, "paper.bib")

### 2024-09-13 quantile.R

@@ -1,0 +1,74 @@
+# Quantile regression -----------------------------------------------------
+
+library(Qtools)
+library(quantreg)
+
+d <-
+  d_GH |>
+  dplyr::filter(age >=18, age <= 60) |>
+  mutate(
+    age_centered = c(scale(age))/2,
+    whitebloodcell_centered = c(scale(whitebloodcell))/2,
+    hemoglobin_centered = c(scale(hemoglobin))/2,
+    years_sexually_mature = age - 12
+  ) |>
+  group_by(sex) |>
+  mutate(
+    strength_sex_centered = c(scale(strength))/2,
+    testosterone_sex_centered = c(scale(testosterone))/2
+  ) |>
+  ungroup()
+
+formulas <- c(
+  anthro_lifetime  = "sex_partners ~ offset(log(years_sexually_mature)) + strength_sex_centered*sex + partnered + bmi*sex",
+  anthro_year      = "sex_partners_year ~ age_centered * sex + strength_sex_centered*sex + partnered + bmi*sex",
+  socio_lifetime   = "sex_partners ~ offset(log(years_sexually_mature)) + strength_sex_centered * sex + partnered + edu + race",
+  socio_year       = "sex_partners_year ~ age_centered * sex + strength_sex_centered * sex + partnered + edu + race",
+  health_lifetime  = "sex_partners ~ offset(log(years_sexually_mature)) + strength_sex_centered * sex + partnered + perceived_abnormal_weight + whitebloodcell_centered + hemoglobin_centered + special_equipment + chronic_disease_score + physical_disease_count + depression",
+  health_year      = "sex_partners_year ~ age_centered * sex + strength_sex_centered * sex + partnered + perceived_abnormal_weight + whitebloodcell_centered + hemoglobin_centered + special_equipment + chronic_disease_score + physical_disease_count + depression",
+  phys_lifetime    = "sex_partners ~ offset(log(years_sexually_mature)) + strength_sex_centered * sex + partnered + vigorous_rec + moderate_rec + vigorous_work + moderate_work",
+  phys_year        = "sex_partners_year ~ age_centered * sex + strength_sex_centered * sex + partnered + vigorous_rec + moderate_rec + vigorous_work + moderate_work",
[…]

### 2024-09-13 sexspecific.R

@@ -1,0 +1,22 @@
+library(tidyverse)
+
+N = 3000
+d <- tibble(
+  v1 = rnorm(N),
+  v2 = rnorm(N),
+  v3 = rnorm(N),
+  v4 = rnorm(N),
+  sex = rbinom(N, 1, 0.5), # 0: females, 1: males
+  males = sum(sex),
+  strength = ifelse(sex == 0, rnorm(N - males, -1), rnorm(males, 1)),
+  partners = strength * sex + v1 + v2 + v3 + v4 + rnorm(N, 0, 5)
+)
+
+m <- lm(partners ~ strength*sex + v1 + v2 + v3 + v4, d)
+summary(m)
+
+mf <- lm(partners ~ strength + v1 + v2 + v3 + v4, d, subset = sex == 0)
+summary(mf)
+
+mm <- lm(partners ~ strength + v1 + v2 + v3 + v4, d, subset = sex == 1)
+summary(mm)

### 2024-09-13 paper.qmd
Added comment about arm lean mass interaction for partnered status
Fixed fig ref
Several minor edits

@@
 It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the effect of upper body strength, proxied by grip strength, on mating success is significantly positive for males and significantly less so for females.
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the association of upper body strength, proxied by grip strength, with mating success is significantly positive for males and significantly less so for females.
 We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
@@
 
-# Stage 2 Confirmatory Study Predictions
+# Stage 2 Confirmatory Study
 
@@
 
-Models and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study.
-However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, which we now do.
+Model specification and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study.
+However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, an adjustment we apply to both the pilot and confirmatory models.
 We also treat this exposure as years since sexual maturity (defined as age - 12 years because 95% of participants first had sex at age 12 or older).
@@
 Our post hoc analyses revealed that coefficients for sex-specific arm lean mass were generally similar in direction to those of sex-specific grip strength discussed above, including a strong positive significant main effect on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+In addition, sex-specific arm lean mass had a significantly smaller association with partnered status for women than men, consistent with the sexual selection hypothesis.
 We found mixed evidence for a main effect of sex-specific lean mass in the age of first sex models, as these were not significant predictors after controlling for socioeconomic or physical activity controls.
@@
 Sex-specific leg, trunk, and total lean mass were not consistently related to any mating success outcome.
-See @sfig-legs - @sfig-total.
-
-Unlike grip strength, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC.
-We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, all lean masses remained a significant predictor of energy and protein intake.
-See @fig-energy and @sfig-arms - @sfig-total.
+See @fig-leancoefs and @sfig-legs - @sfig-total.
 
@@
 
+Unlike grip strength, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC.
+We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
+In the expanded models with additional controls, all lean masses remained a significant predictor of energy and protein intake.
+See @sfig-arms - @sfig-total.
+
 ## Discussion
@@
 The purpose of this study was to replicate, as closely as possible, the @lassek2009 study with new data that used combined grip strength as a proxy for upper body strength, as well as the lean masses of arms, legs, trunk, and total body.
-Unlike @lassek2009, we included women to test if the effects of upper body strength and lean masses on mating success were greater for men than women, as predicted by the sexual selection hypothesis.
+Unlike @lassek2009, we included women to test if the associations of upper body strength and lean masses with mating success were greater for men than women, as predicted by the sexual selection hypothesis.
 We also included numerous additional control variables, and used a registered report format with pre-specified models developed using pilot data in Stage 1 and then fit on held-out data in Stage 2.
@@
 We also tested for interactions between sex and strength on these mating success outcomes.
-Across both pilot and confirmatory data, we found a consistent interaction between sex and sex-specific grip strength on partnered status, such that sex-specific grip strength was more important for predicting males' partnered status than females.
-Although in the pilot data we did see evidence of an interaction between sex and sex-specific grip strength on the outcome lifetime number of sexual partners, such that sex-specific grip strength was more important for predicting males' lifetime sexual partners, this effect was not significant in the confirmatory data.
+Across both pilot and confirmatory analyses, we found a consistent interaction between sex and sex-specific grip strength on partnered status, such that sex-specific grip strength was a better predictor of males' partnered status than females.
+Although in the pilot analysis we did see evidence of an interaction between sex and sex-specific grip strength on the outcome lifetime number of sexual partners, such that sex-specific grip strength was a better predictor of males' lifetime sexual partners, this effect was not significant in the confirmatory analysis.
 
@@
 Men have greater strength than women and their average partner numbers are higher.
-Still, women with higher sex-specific grip strength and sex-specific arm lean mass unexpectedly reported more lifetime sexual partners on average than women with lower grip strength or arm lean mass, even after controlling for a host of potential confounds.
+Still, women with higher grip strength and arm lean mass unexpectedly reported more lifetime sexual partners on average than women with lower grip strength or arm lean mass, even after controlling for a host of potential confounds.
 A meta-analysis of 77 species, including humans and 7 other mammal species, found a positive Bateman gradient for females, i.e., a positive correlation between mating success and reproductive success, although authors and commentators both caution that correlation is not causation [@fromonteil2023; @kokko2023].
-It is also not clear why women's strength would be positively associated with partner numbers.
-
+
+Still, it is not clear why women's strength would be positively associated with partner numbers.
 It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
 It could be that there is assortative mating on strength: if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
-It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
-genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
+It might be that stronger women require less male investment, or can take more physical risks, and so instead benefit from greater partner numbers through, e.g.
+genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding or leaving costly long term partnerships.
 It might be the case that there are some sex-specific confounds that we failed to control for, or that our models are otherwise misspecified.
@@
 Since a sex difference was significant in the Pilot Study (2011-12), but not in the Confirmatory Study (2013-14), it is possible that cultural changes influencing sexual behavior occurred between the data collection cycles.
+The sexual sexual hypothesis also might be wrong.
 Or these results might simply be noise.
@@
 Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
+In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis.
+Ideally, however, analyses would account for the full range of variation in sexual behavior.
 Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions.
+However, even our confirmatory results were based on generalized linear models, yet there might be important non-linear relationships that would alter interpretations.
 
@@
 ```{r}
-pilot_summary_stats <- summary_tables(designsG$d.design.adults)
+pilot_summary_stats <- summary_tables(designsG2$d.design.adults)
 tbl_stack(list(pilot_summary_stats$con, pilot_summary_stats$cat))

### 2024-09-13 paper.qmd
Changed to correct confirmatory desc stats table

--- 2024-09-13 paper.qmd
@@
 
+### Descriptive statistics
+Sample characteristics were similar in the confirmatory dataset to those in the pilot dataset.
+See @stbl-confirmstats for weighted means, standard deviations, and standardized mean differences of all variables for men and women.
+
 ```{r}
@@
 ✓: Supported in confirmatory study.
-❌: Not supported in confirmatory study.
-*The posthoc results found a significant negative association between lean mass and immunity, and a positive association with energy."
+❌: Not supported in confirmatory study."
 
@@
     "Age at first sex",     "0",       "✓",       "0",       "✓",
-            "Immunity",     "-",       "❌*",      "0",       "✓",
-              "Energy",     "0",       "❌*",        "",        "",
+            "Immunity",     "-",       "❌",      "0",       "✓",
+              "Energy",     "0",       "❌",        "",        "",
              "Protein",     "+",       "✓",        "",        ""

### 2024-09-13 misc.qmd

--- 2024-03-31 paper.R1.qmd
@@
----
-title: "Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report"
-author: 
-  - name: Caroline B. Smith
-    affiliation:
-     - id: wsu
-       department: Department of Anthropology
-       name: Washington State University
-  - name: Edward H. Hagen
-    affiliation:
-     - ref: wsu
-date: today
-format:
-  html:
-    toc: true
-    embed-resources: true
-  # docx:
-  #   number-sections: true
-execute:
-  echo: false
-  warning: false
-csl: apa.csl
-bibliography: paper.bib
-crossref:
-  custom:
-    - kind: float
-      reference-prefix: Figure S
-      key: sfig
-      space-before-numbering: false
-    - kind: float
-      reference-prefix: Table S
-      key: stbl
-      space-before-numbering: false
----
+### Anthropometric models of mating success
 
+In the anthropometric control model (@sfig-anthro), strength was significantly associated with higher numbers of lifetime and past year sex partners and an earlier age at first sex, and the interaction with sex was not significant.
+
+::: {#sfig-anthro}
 ```{r}
-library(ggcorrplot)
-library(modelsummary)
-source('analysis.R1.R')
+#| fig-width: 12
+#| fig-height: 8
+
+plot_anthro <- forestplot(
+  m_agefirst,
+  m_lifetime,
+  m_pastyear,
+  m_partnered,
+  intercept = F,
+  facet = F,
+  dodgewidth = .5,
+  modelnames = mnames1,
+  varnames = vnames, 
+  size = 1.2, 
+  linewidth = 1
+) + theme_minimal(17) +
+  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
+
+ggsave("Figures/plot_anthro.pdf", plot_anthro, width = 14, height = 7)
+
+plot_anthro
 ```
 
-Abstract
-
-Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-It is unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
-We will also test if there is a positive relationship between grip strength and dietary energy intake, and a negative relationship between grip strength and native immune function.
-This study replicates an earlier study of US males [@lassek2009] with numerous additional control variables, and it additionally tests these relationships in females.
-
-# Sexual dimorphism in modern humans
-
-The mechanisms that have shaped human sexual dimorphism and whether they have primarily operated on male or female traits remain debated.
-Sexual dimorphism refers to sex differences in morphological and behavioral traits, excluding reproductive organs [@plavcan2001].
-Unlike most animals, mammalian sexual size dimorphism, when it exists, tends to be male biased [@andersson1994; @Tombak_2024].
-In humans, sexual size dimorphism is observable as early as the first trimester in utero where male embryos are larger than female embryos [@bukowski2007].
-Body composition dimorphism is evident through childhood such that, adjusted for height, males have higher lean body mass and females have higher fat mass [@kirchengast2010].
-At puberty, sexual dimorphism in size and body composition significantly increases such that in adults, adjusting for a 7-8% dimorphism for height, men have 12-25% higher body mass [@lassek2022], which is slightly greater than body mass dimorphism in gibbons, and slightly less than chimpanzees [@plavcan2012; @smith1997; @Puts_2023].
-Men also have larger and stronger bones [@wells2007].
-Other traits that are sexually dimorphic in humans include digit ratio, voice pitch, facial features, body and facial hair growth, and canine length [@Puts_2023].
-
-While humans are only moderately dimorphic in terms of overall body mass, this is not the case for fat and muscle allocation, which are highly dimorphic [@puts2010] due to women’s copious storage of fat and male investment in muscle mass.
-Men have between 30% and 42% more fat-free mass [@lassek2022], 61% more overall muscle mass, and 78% more muscle mass in the upper arms.
-This concentrated muscle dimorphism in the arms and back translates to greater upper body strength in men than women.
-Studies of sex differences in strength show that men are stronger on average than women on all tests of muscle strength, but especially for tests of upper-body muscle strength, where female upper-body strength is 50-60% of male upper-body strength while lower-body strength is 60-70% of male values, and trunk strength is 60% of male values [@nuzzo2023].
-Muscle mass alone does not explain the sex difference in strength, as strength assessments are often greater in males than females even when pair-matched on muscle thickness [@kataoka2023], and likewise men have greater strength-to-body mass ratios than women [@nuzzo2023].
-
-There is no evidence for a sex difference in the ability of the nervous system to drive the muscle (voluntary action) but instead differences in strength lie in muscle characteristics including mass, size, and fiber type.
-Men have both more absolute muscle mass and greater mass proportional to body size, and these muscles have higher volume and cross-sectional area size.
-These differences are greatest in the upper body.
-Finally, a greater proportion of male to female muscle is occupied by type II muscle fibers, which create greater force than type I [@nuzzo2023].
-
-## Sexual dimorphism hypotheses
-
-Two hypotheses for the evolution of human sexual size dimorphism and body composition dimorphism dominate the literature: 1) intrasexual selection through male contests and 2) intersexual selection through female mate choice.
-Both assert that men with more masculine traits had higher biological fitness over evolutionary history: intrasexual selection if more formidable men were able to physically outcompete other men for access to mates, and intersexual selection if females preferentially mated with men who displayed masculine traits.
-The male contest and the female choice hypotheses concur that sexual size and strength dimorphism are proximately (that is, developmentally) caused by sex differences in androgen hormones in the uterine environment and during differentiation at puberty.
-They differ, however, in the ultimate (functional) explanations for these differences, as discussed below.
-
-### Male contest competition
-
-The intrasexual selection hypothesis emphasizes evidence that female mate choice was limited in the ancestral past [@puts2010], and that instead male fitness was determined by physical contests with other males for access to mates.
-Under these circumstances more physically formidable males (i.e., those having the ability to inflict costs on competitors) had higher reproductive success [@puts2016; @Puts_2023; @hill2017; @plavcan2012].
-Therefore, human males have been sexually selected to be formidable resulting in the male bias in morphological traits like bone density, height, weight, muscle mass, and strength, as well as behavioral traits such as aggression [@archer2009; @Puts_2023].
-Particular emphasis is placed on sexual dimorphism in the upper-body due to the important role upper-body strength plays in both armed and unarmed fighting ability [@sell2012] and where sex differences in strength and muscularity are greatest as discussed above.
-
-In extant nonhuman primate species, sexual dimorphism in body weight and canine size are strongly associated with the degree of intrasexual competition, a relationship which, if used to infer the behavior of extinct australopithecines, suggests a high degree of intrasexual competition in ancestral species [@plavcan1997].
-Evidence of sexual size dimorphism in *Homo* is less certain, but in general shows reduced size dimorphism compared to earlier hominin species [@plavcan2012].
-
-### Female choice for masculine traits
-
-Although there is evidence in Western young adult populations that male mating success is mediated by male-rated dominance and not by female-rated attractiveness [@kordsmeyer2018; @hill2013], others nevertheless argue that sexual dimorphism in humans is largely the result of female choice for high quality males [which was possibly heavily influenced by parents, in which case parental preferences would also have been important, @apostolou2007].
-The immunocompetence handicap hypothesis, for example, relies on evidence that testosterone is immunosuppressive, and therefore only males with highly competent immune systems can afford to pay the costs of high levels of testosterone required to develop masculine secondary sexual traits [@folstad1992].
-Androgen-dependent traits could thus be costly signals of good genetic quality, with females increasing their fitness by preferentially mating with males displaying these traits, thereby conferring heritable immunity to their offspring and securing investment from a healthy partner [@folstad1992].
-Evidence that testosterone is immunosuppressive is mixed, however [@nowak2018], and although testosterone is associated with mating success, it does not seem to be through the mechanism of female choice for testosterone-derived face or body features since testosterone does not predict female rated masculinity or attractiveness [@peters2008].
-
-Similar arguments have therefore been proposed that masculine traits could be more general cues of quality and condition, or an ability to bear other costs [@kokko2003; @frederick2007].
-Cues of upper body strength indeed account for most of the variation in female-rated male body attractiveness [@sell2017].
-Yet masculine trait expression is not consistently related to retrospective or prospective health [@boothroyd2013].
-
-Females might therefore be choosing males who could provide more resources, or men who could better protect them from attacks by other men or by predators, which in ancestral hunter-gatherer populations would have been physically stronger men [@apicella2014; @willems2017b].
-There is substantial cross-cultural evidence of a female-biased preference for mates who can provide resources [@walter2020; @buss2019].
-Importantly, studies of contemporary hunter-gatherers and game-theoretic models find that a mutually beneficial sexual division of labor within long-term pair bonds, in which women engage in lower risk, lower return foraging of plants and small game compatible with the cognitively demanding task of childrearing [@hagen2019], and men engage in higher risk, higher return big game hunting [@kelly2013], is critical to provisioning offspring who remain dependent on caregivers for 20 years or more [@davison2021; @davison2022; @alger2020; @kaplan2000].
-See @fig-provisioning.
-This suggests that sex differences in body composition could be due, at least in part, to a sexual division of labor.
-
-:::{#fig-provisioning}
-![](Figures/plot_productivity.png)
-
-Net energy production in chimpanzees and human foragers over the lifespan. Solid lines are females. Dashed line is Hadza males. Data from @davison2021 and @davison2022.
+Anthropometric models of mating success (Pilot). Coefficient plot from generalized linear models. Variables labelled (S) have been centered at the mean and standardized by 2 SD.
 :::
-
-More generally, multiple forms of sexual selection could have played a role in the evolution of human sexual dimorphism [see @hill2017 for a brief review of alternative hypotheses for sexual strength dimorphism].
-
-## Strength and reproductive success
-
-A meta-analysis found that whereas voice pitch, height, and testosterone levels were associated with mating success in low fertility populations, only muscularity was associated with actual reproductive success in high fertility populations [@lidborg2022].
-In four studies of convenience samples (two of which included women), hand grip strength was positively related to mating success (including number of sex partners) for men, but there was either no relationship or a negative relationship for women [@gallup2007; @varella2014; @shoup2008; @sneade2016]
-
-The literature on the relationship between grip strength and mating and reproductive success is disproportionately reported for men [@gallup2018].
-There have been very few tests of the hypotheses that some female traits, such as breast size and waist-hip ratio (WHR), have evolved by sexual selection.
-A meta-analysis found weak evidence that more feminine digit ratios predicted higher fertility, insufficient evidence for voice pitch, and none regarding facial femininity, breast size or waist-hip ratio (WHR).
-At present there is mixed evidence that strength is related to reproductive success in women [@lidborg2022a].
-One study among the Himba showed that women with higher strength/muscularity had more living children and grandchildren [@atkinson2012], while a similar study among the Hadza showed no relationship between strength and reproductive outcomes [@smith2017a].
-The dearth of research on female traits and reproductive outcomes leaves us unable to draw any firm conclusions about selection pressures over time.
-
-Previous research has shown that physical formidability in males (operationalized as fat free mass and limb muscle volume) predicts numbers of total and past-year self-reported sex partners, but that it also involves costs such as increased daily energy intake and decreased immune function, operationalized as C-reactive protein and white blood cell count [@lassek2009], consistent with a tradeoff in mating success, immunity, and energy costs.
-However, these relationships were not tested in women, for whom both sexual selection hypotheses predict no association.
-
-## Experimental aims and hypotheses
-
-According to the sexual selection hypothesis, males evolved greater physical formidability than females because physical formidability increased male reproductive success more than females.
-The aim of this study is to replicate the costs and benefits of formidability reported by @lassek2009 using similar nationally representative data, controlling for a wider range of possible confounds, and precisely prespecifying the statistical models in a registered report format.
-It also aims to investigate if these costs and benefits are also experienced by women, and specifically if the effect of physical strength on mating success, a proxy for reproductive success, is greater for men than women, as the sexual selection hypothesis predicts.
-
-# Pilot study
-
-In order to refine our hypotheses and statistical models for our confirmatory study, which will use data we have not yet observed, we first conducted a pilot study whose results we report here.
-
-## Methods
-
-To assess the relationship between formidability and mating benefits, and immune and energy costs, we used data from the Centers for Disease Control (CDC) National Health and Nutrition Examination Survey (NHANES).
-NHANES utilizes a complex, multi-stage sampling strategy in order to collect data representative of the civilian, non-institutionalized U.S.
-population.
-NHANES combines interview, examination, and laboratory data to assess health status and identify health risks for adults and children in the United States.
-Data collection occurs in new cycles every two years.
-
-The pilot study uses the 2011-2012 dataset, whereas the 2013-2014 dataset has been held out for confirmatory analysis pending in principle acceptance of this registered report.
-Grip strength, our proxy for physical formidability and our key predictor variable, was only collected in these years.
-We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable [@hogrel2015].
-For each model, participants will be included if they have complete data for each predictor and outcome variable.
-
-### Outcome variables
-
-Our main outcome was mating success.
-In industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior.
-Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with male reproductive success under ancestral conditions [@perusse1993].
-We operationalized mating success in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in the past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
-female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
-3) Age at first sexual intercourse reported from: "How old were you when you had sex for the first time?" The sexual behavior questionnaire was self-administered on a computer in a private room at the examination center, using the Audio Computer Assisted Self Interview system, which allows participants to hear questions through headphones as well as read them on screen.
-Only respondents who could self-report were asked these questions.
-
-Lifetime and past year numbers of sexual partners index success through short term mating strategies.
-Lifetime numbers are probably a more precise proxy of the use of a short-term mating strategy as these are integrated over the entire life, whereas past year numbers likely fluctuate depending on partnered status and other transient factors.
-However, both are somewhat ambiguous measures as they could also represent repeated rejection, or a tendency to exaggerate or understate the actual numbers.
-Neither clearly indexes female reproductive success because multiple sexual partners are likely to have increased male fitness more than female fitness [i.e., the Bateman gradient, @Lehtonen_2022].
-We therefore also operationalized mating success in a fourth outcome measure as partnered status as a proxy for a long-term mating strategy that, as discussed earlier, benefits both sexes.
-Partnered status included participants who reported either being married or living with a partner whereas unpartnered included participants who reported being single, widowed, divorced, or separated [@lassek2009, included marital status as a control variable but not as an outcome variable].
-
-Our second outcome measure was immune investment, which we operationalized as white blood cell count (1000 cells/µL).
-@lassek2009 also used C-reactive protein (CRP) as an outcome variable, but CRP was not measured in the 2011-2014 data collection years and so is not included here.
-
-Finally, to assess costs of greater strength, we investigated the relationship between strength and dietary energy and protein intake.
-Participants reported all food and beverages they consumed in the 24 hour period prior to their interview.
-This interview was repeated for a different 24 hour period 3-10 days later to obtain dietary recalls for two separate days.
-These data were then used to estimate energy intake in the form of kilocalories (kcal) calculated by matching reported foods to the USDA's Food and Nutrient Database for Dietary Studies.
-For the outcome measures in these models we used the average number of calories per day and grams of protein per day calculated for each participant across their two recall days.
-
-### Key explanatory variables
-
-Our main explanatory variable of mating success is formidability, which @lassek2009 operationalized as fat-free mass and limb muscle volume.
-Since those variables were not available in our data, we operationalized formidability as combined grip strength (kg), the sum of the highest of three readings taken on each hand using a dynamometer.
-All regression models also included sex (male/female), our second main explanatory variable.
-All models also controlled for age in years and partnered status (not included in models where partnered status was the outcome variable).
-
-The interaction between sex and strength was crucial for our models since we want to test if the effect of strength on mating and immune outcomes is different for women than for men.
-Grip strength is highly sexually dimorphic, however, with about 90% of men stronger than about 90% of women [@smith2022], and is thus highly collinear with sex, a problem for regression modeling and interpretation.
-We therefore computed a sex-specific grip strength value as follows: for each sex we centered grip strength at the sex-specific mean and divided by two times the sex-specific standard deviation.
-Thus, for women, low or high values are relative to other women, and for men, low or high values are relative to other men.
-The sex-specific values were used in all models except the immune function and dietary intake models described below, which assess the costs of high strength.
-Here, an index of absolute rather than relative strength is called for, and we used grip strength standardized across both sexes.
-
-We also interacted sex with age and partnered status.
-We interacted age with sex since menopause occurs within the age range of our participants.
-For models of past year number of sexual partners we included an interaction between partnered status and grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
-Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable [@gelman2008].
-
-## Models
-
-We attempted to replicate the regression models in @lassek2009 as closely as possible.
-There were some unavoidable differences, however.
-First, our study aimed to investigate the role of sex; thus we include sex as a main effect as well as in interactions with key predictor variables.
-Second, some variables included in @lassek2009 were not available in our data.
-Third, we included additional control variables as noted below.
-Finally, we treated our pilot study as an exploratory study in which we fit numerous models not reported here.
-The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we described above, the results of our exploratory analyses, and our constraints.
-
-### Control models
-
-Our modeling strategy takes inspiration from @lassek2009.
-However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples [@smith2018a].
-Instead, we will fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses.
-Control variables included a much wider range of theoretically motivated potential confounds than used by @lassek2009, organized into themes.
-We specified five models for each mating success outcome measure in order to determine if the effect of grip strength on those outcomes was due to confounds with socioeconomic, health, hormone, or physical activity variables that have been associated with strength and sexual behavior.
-
-### Anthropometric control model
-
-We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, grip strength, and partnered status.
-This model also includes an anthropometric control, body mass index (BMI), calculated as $kg/m^2$, since body size could impact either strength or mating outcomes.
-We interacted age, strength, and BMI with sex to assess differences in these predictors for men and women.
-
-### Socioeconomic control model
-
-The socioeconomic control model included education and race as categorical variables, since there is evidence that race is related to variation in both strength [@johnson2019] and sexual behavior [@fenton2005].
-Likewise, there is also variation in sexual behavior by education [@chandra2013], and there is variation in physical activity by education [@he2005].
-
-### Health control model
-
-An umbrella review found the grip strength is a useful indicator of general health status, early all-cause mortality, cardiovascular mortality, and disability [@soysal2021].
-Furthermore, healthy status is associated with higher likelihood of sexual activity, frequency of sexual activity, and reported quality of sex in men and women [@lindau2010].
-Another study of men aged 45-59 found that lower sexual activity was associated with increased mortality [@daveysmith1997].
-The health model, therefore, included a number of variables related to health.
-White blood cell count (1000 cells/µL) and hemoglobin (g/dL) were included to control for acute infection.
-Depression, which is negatively related to strength [@smith2022], was measured using the Patient Health Questionnaire-9 [PHQ-9, @kroenke2001], a validated nine-item screening instrument.
-Each item represents a symptom of depression, and for each one participants were asked to consider how frequently they had been bothered by that symptom over the past two weeks, rated on a scale from 0 (not at all) to 3 (nearly everyday).
-These ratings were summed to produce a depression score ranging from 0-27.
-Chronic illness was included using Chronic Disease Score (0-6) a count of chronic diseases participants reported having been diagnosed with including diabetes, cancer, stroke, arthritis, heart disease and respiratory disease.
-A point was added for each disease a participant reported being diagnosed with, regardless of any impairment due to the disease.
-We then controlled for impairment resulting from chronic illness (Disease Impairment Score; 0-5) calculated from a different NHANES question, which asked participants to list up to five health conditions that specifically cause them to have difficulties with physical activities.
-We also controlled for physical disability and using the item "special equipment" which referred to participants' report that they needed special equipment to walk.
-Finally the 'perceived abnormal weight' variable was coded as true or false depending on whether a participant reported that they perceived their weight to be abnormal.
-
-### Hormone control model
-
-There is evidence that testosterone is positively associated with higher numbers of partners for men and less robustly so for women [@pollet2011; @vananders2007], but mixed evidence that it is related to sexual desire [@vananders2013].
-For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see @vananders2013.
-Likewise, although some have found a positive correlation between circulating testosterone and handgrip strength [@Chiu_2019], other studies have found no association [@Gettler_2010; @Ribeiro_2016].
-There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance [@Guilherme_2021].
-The hormone control model included serum total testosterone (ng/dL).
-Since male mean testosterone is an order of magnitude higher than the female mean, testosterone, like grip strength, was highly confounded with sex, one of our key predictor variables.
-We therefore computed a sex-specific testosterone value in the same way we did for strength.
-This differed from other centered and standardized variables which we centered and standardized across all male and female values.
-
-### Physical activity control model
-
-Physical activity is related to sexual behavior and function in adults [@morris2022].
-Therefore, the physical activity control model included four dichotomous variables of vigorous and moderate work and recreational activity coded as 1 if participants reported that their work and/or recreation caused large increases in heart rate or breathing for at least 10 minutes continuously (vigorous work and/or rec), or small increases in breathing or heart rate for 10 minutes continuously (moderate work and/or rec), and zero if they reported their work and/or recreation did not.
-
-### Costs of strength
-
-Following @lassek2009 we also developed models of native immune function (operationalized as white blood cell count) and dietary energy intake (operationalized as average calories per day).
-For each outcome we developed a model as similar as possible to those reported in @lassek2009, except that we always include sex and its interaction with other predictors where we have theoretical reasons to expect sex differences in the effect of the predictor on the outcome.
-We then also specified an alternative model for each outcome with additional control variables that could confound strength and energy or immune investment.
-These models were based in part on exploratory analyses of multiple models not reported here.
-
-#### Native Immune Function
-
-Following the significant predictors reported in @lassek2009, Table 4, we first specified a replication model with age, sex, strength, and BMI.
-We also included an age:sex interaction because we expected that immune investment over the lifespan might differ between males and females due to menopause in women.
-Finally, we included a strength:sex interaction to test the hypothesis that the immune costs of strength differ by sex.
-
-Based on exploratory analysis we developed an expanded control model of WBCC that included a few additional controls.
-First, we controlled for sex-specific testosterone and an interaction of sex and sex-specific testosterone since testosterone possibly has immunosuppressive effects, which may differ by sex.
-We also controlled for three variables related to energy availability that could impact immune investment.
-First, food security status for adults in the participant's household was assessed using 10 items from the U.S.
-Food Security Survey Module.
-The resulting value can range from 1-4, where 1 represents full food security and 4 represents very low food security.
-Second, the total metabolic equivalent (total MET) was calculated based on participants' responses to questions about the minutes they spent walking or bicycling, engaging in vigorous and moderate work, and vigorous and moderate recreation per day, using MET scores provided by NHANES.
-We included dietary energy intake (average calories per day as discussed above).
-Finally, we substituted separate height and weight variables for BMI.
-
-#### Dietary Energy and Protein Intake
-
-Based on @lassek2009, Table 3, we developed a replication model of dietary energy intake (average calories per day) and included age, sex, strength, BMI, and total MET.
-We did not include interactions with sex in this model because we do not anticipate sex differences in the effects of any predictor variable, nor did we see any in exploratory analyses.
-We then developed an expanded control model that also controlled for WBCC and food insecurity in addition to the variables in the model above, since either could influence energy intake and strength.
-Given the protein costs of muscle maintenance we likewise developed two models of dietary protein intake.
-We based the “replication model” on that of the @lassek2009 dietary energy model, and then applied the same expanded controls discussed above in the “expanded controls model,” including the substitution of separate height and weight variables for BMI.
-
-### A tradeoff between short- and long-term mating
-
-Exploratory analysis revealed a negative association between the lifetime number of sexual partners and the probability of being in a committed relationship (partnered status), which we interpreted as a possible tradeoff between short- and long-term mating strategies.
-In models of partnered status (one of our mating success outcome variables), we therefore included the lifetime number of sexual partners to assess if this negative association persisted after including the sets of control variables described above.
-
-## Analysis
-
-All analyses were completed in `{r} R.version.string`, using the survey package [version 4.1-1, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-We modeled lifetime and past year numbers of partners using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
-We modeled age at first sexual intercourse and white blood cell count with a Gaussian glm.
-We modeled partnered status using a binomial glm.
-Analyses included adults ages 18-60 because this is the period when the majority of reproduction takes place and because these are the years when grip strength is the most stable.
-
-## Pilot Study Results
-
-```{r}
-# Lifetime
-female_median <- 
-  svyquantile(~sex_partners, designsG$d.design.adult.female, 0.5, na.rm = T)$sex_partners[1,1]
-male_median <- 
-  svyquantile(~sex_partners, designsG$d.design.adult.male, 0.5, na.rm = T)$sex_partners[1,1]
-high_partners <- sum(d_G$sex_partners >= 100, na.rm=T)
-
-# Last year
-female_median_year <- 
-  svyquantile(~sex_partners_year, designsG$d.design.adult.female, 0.5, na.rm = T)$sex_partners_year[1,1]
-male_median_year <- 
-  svyquantile(~sex_partners_year, designsG$d.design.adult.male, 0.5, na.rm = T)$sex_partners_year[1,1]
-high_partners_year <- sum(d_G$sex_partners_year >= 10, na.rm=T)
-```
-
-Our initial analyses found little support for a sex difference in the effect of strength on mating success, contrary to the sexual selection hypothesis.
-We traced this failure to the unusually high number of lifetime sex partners reported by some men and women.
-We discovered that whereas women reported a median number of `{r} female_median` lifetime sex partners, and men `{r} male_median` partners, there were `{r} high_partners` individuals who reported 100 or more partners, a few of whom reported 1000 or more (@fig-ecdf).
-We decided to restrict our sample to individuals with fewer than 100 lifetime sex partners for several reasons.
-First, support for the sexual selection hypothesis was weak if these individuals were included, but much stronger if they were excluded.
-We suspect that because our sample is large and nationally representative, it probably includes sex workers, as well as individuals who regularly use the services of sex workers, and whose high partner numbers would therefore not reflect mate attraction or intrasexual competition.
-Second, there was an unusually high number of participants reporting exactly 100 lifetime sex partners, suggesting inaccurate recollections.
-Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
-We will use the same cutoff for our confirmatory study.
-
-```{r}
-#| label: fig-ecdf
-#| fig-width: 10
-#| fig-cap: Cumulative distribution of the number of sex partners, by sex.
-Participants who reported 100 or more partners were removed from the analysis.
-The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this plot and not for any analyses).
-
-plot_sexpartners
-```
-
-### Descriptive statistics
-
-The mean age was 38 years.
-Sexual dimorphism---the ratio of male mean:female mean---was evident in an 8% higher value for male height and a 16% higher value for weight; females, however, had a 1% higher BMI than males on average.
-There was a much larger degree of male-biased dimorphism in combined grip strength (57%).
-Hemoglobin was 14% higher in males than females while white blood cell count was 3% higher in women.
-See @stbl-stats for weighted means, standard deviations, and standardized mean differences of all variables for men and women.
-
-### Coefficients of strength, and strength X sex interaction
-
-Our pilot results revealed strong positive significant main effects of sex-specific strength on partnered status, numbers of lifetime partners, and age at first sex after controlling for all five sets of potential confounds.
-The only exception was that the strength coefficient for age at first sex was not significant after controlling for socioeconomic factors, although the sign and magnitude was very similar to those of these coefficients in the models with other control variables.
-There was no significant effect of sex-specific strength on the number of partners in the past year.
-See the left column of @fig-coefs.
-
-There were also clear sex differences in the effects of sex-specific strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
-There was also a persistent negative association of numbers of lifetime partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
-Finally there was no evidence for a sex difference in the effects of sex-specific strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
-See the right column of @fig-coefs.
-
-```{r}
-#| label: fig-coefs
-#| fig-width: 10
-#| fig-height: 8
-#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables.
-Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions.
-Bars are 95\% CIs.
-For the full models, see the SI.
-
-plot_coefs
-```
-
-### Immune costs of strength
-
-Previous research found that higher musculature was associated with decreased investment in native immunity.
-We developed two models of immune investment in white blood cell count (@fig-immune).
-In the replication model, strength was a significant negative predictor of WBCC for both men and women (there was no significant interaction of sex and strength) controlling for age and BMI.
-In the expanded model with additional controls, including the substitution of height and weight for BMI, strength was now not significant, albeit with a coefficient that was very similar to the replication model.
-
-```{r}
-#| label: fig-immune
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of immune investment (WBCC) from generalized linear models (Pilot). Variables labeled (S) have been centered at the mean and standardized by 2 SD.
-
-Figure3 <- forestplot(
-  mwbc,
-  mwbc_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-
-ggsave("Figures/Figure3.pdf", Figure3, width = 14, height = 7)
-
-Figure3
-
-```
-
-### Models of energy and protein intake
-
-Finally, we tested whether strength was related to increased energy (@fig-energy) or protein intake (@fig-protein).
-We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, strength was no longer a significant predictor of energy intake but remained a significant predictor of protein intake.
-
-```{r}
-#| label: fig-energy
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
-
-Figure4 <- forestplot(
-  m_energy,
-  m_energy_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-
-ggsave("Figures/Figure4.pdf", Figure4, width = 14, height = 7)
-
-Figure4 
-```
-
-```{r}
-#| label: fig-protein
-#| fig.width: 12
-#| fig.height: 8
-#| fig.cap: Coefficient plot of predictors of dietary protein intake from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD.
-
-Figure5 <- forestplot(
-  m_protein,
-  m_protein_alt,
-  intercept = F,
-  facet = F,
-  dodgewidth = .4,
-  varnames = vnames, 
-  modelnames = mmnames) +
-  scale_color_binary() +
-  theme_minimal(17) +
-  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))
-
-ggsave("Figures/Figure5.pdf", Figure4, width = 14, height = 7)
-
-Figure5 
-```
-
-## Pilot Results Discussion
-
-Using grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with three of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-The only exceptions were that strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant strength coefficients in the other models.
-Strength was also associated with immune, and energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
-The expanded control models included height, which might be a proxy for aspects of strength, such as lower body strength, that are not accounted for by grip strength.
-
-The effect of strength on age at first sex was not significant in our socioeconomic model probably because it is confounded with ethnicity.
-Specifically, Asian Americans had a markedly lower grip strength compared to the other ethnic groups, and a later age at first sex.
-In addition, participants, whose grip strength was measured at the time of the study when they were adults, are reporting on an event that happened an average of 22 years in the past when most were adolescents and their grip strength had not yet reached its adult value.
-For these reasons, we do not predict an effect of strength on age at first sex after controlling for ethnicity.
-
-The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status and numbers of lifetime partners.
-In humans, biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships.
-We therefore propose that the sexual selection hypothesis for the sexual dimorphism in physical formidability applies to long-term partnerships (proxied by our partnered variable), as well as short-term matings (proxied by partner numbers).
-Thus, our pilot study supports the hypothesis that muscularity increases both short- and long-term mating success for men more than it does for women.
-Number of lifetime partners was also a negative predictor of partnered status (@sfig-tradeoff), suggesting a tradeoff in pursuing these two reproductive strategies.
-
-The lack of a sex difference in the significant positive effect of sex-specific strength on short term mating success (proxied by number of partners in the last year) is puzzling.
-Because men have greater strength than women their self-reported partner numbers are higher, on average, although the median number of partners in the last year for both sexes is 1.
-Still, women with higher grip strength report more partners in the past year than women with lower grip strength even after controlling for a host of potential confounds.
-It could be that there was selection for more formidable men to prefer more partner variety, and stronger women have a similar preference as a byproduct of selection on men.
-It could be that there is assortative mating on strength; thus if stronger men are motivated to switch partners more frequently, their (stronger) mates would also likely have more mates.
-It might be that stronger women require less male investment and so instead benefit from greater partner numbers through, e.g.
-genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding costly long term partnerships.
-It might also be the case that there are some sex-specific confounds that we failed to control for.
-Greater partner numbers might indicate mating failures rather than mating successes for males or females, although why strength would be associated with mating failures is not clear.
-Finally, our simulations revealed that we had relatively lower power to detect a significant interaction between strength and sex for this outcome (see Power section below).
-
-Our pilot results support the importance of strength to male short-term mating success as well as to the long-term partnerships that are the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
-Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
-More specifically, humans require provisioning for an extended period [@davison2021; @davison2022] (@fig-provisioning), and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
-Thus, our pilot study underscores that the sexual selection hypothesis for the relationship between strength and mating success should include long-term mating in addition to numbers of sexual partners.
-Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014].
-
-## Limitations
-
-Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females.
-Our aim is not to test all possible byproduct hypotheses, nor is it possible to do so.
-Therefore, we are unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success.
-Likewise, NHANES does not collect data on sociosexuality.
-As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women.
-Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
-Finally, all pilot results were obtained after considerable exploratory analysis, and must be replicated in a confirmatory study.
-
-# Confirmatory Study (Not yet conducted)
-
-In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime number of sexual partners, but not last year number of partners, nor age at first sex after controlling for ethnicity.
-We also predict a significant interaction with sex such that strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
-We also predict that there will be a negative association between formidability and immune function.
-Finally, we predict that formidability will be positively associated with dietary energy intake for both males and females.
-
-We indicate our predictions for the results of the confirmatory tests, specifically, our predictions for significant coefficients of strength and of the interaction between sex and strength, in @tbl-predictions.
-We indicate for each outcome whether we predict a significant positive, negative, or no effect of these variables across our control models.
-
-| Outcome            | Strength coefficient | Strength X Sex coefficient |
-|--------------------|:--------------------:|:--------------------------:|
-| Partnered status   |           +          |              -             |
-| Lifetime partners  |           +          |              -             |
-| Past year partners |           0          |              0             |
-| Age at first sex   |           0          |              0             |
-
-: Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model.
-{#tbl-predictions}
-
-## Power
-
-We used simulations to estimate our power to detect strength effects that were 75% as large as the ones observed in our pilot study, and sex interaction effects that were equal in magnitude but opposite in sign to the strength effects (i.e., no effect of strength on mating success in women), with a sample size equal to that in our unobserved data.
-We are confident that we have sufficient power to detect true strength effects somewhat less than the ones observed in our pilot study (@tbl-power).
-We are less confident in our power to detect true sex interaction effects because, all else equal, power to detect interaction effects is generally lower than main effects.
-
-| Outcome            | Strength coefficient | Strength X Sex coefficient |
-|--------------------|:--------------------:|:--------------------------:|
-| Partnered status   |         0.99         |            0.90 (0.67)            |
-| Lifetime partners  |         0.99         |            0.90 (0.66)            |
-| Past year partners |         0.80         |            0.30 (0.21)            |
-
-: Power to detect our predicted effects.
-Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients (value in parentheses is power to detect 3/4 of that effect), sample size was set to that of our unobserved data, and $\alpha = 0.05$.
-{#tbl-power}
-
-## Method
-
-Our methods and models will remain identical to those reported in the pilot study, using unobserved data from the 2013-2014 NHANES collection cycle.
-
-# Acknowledgements
-
-We thank the Editor and two anonymous reviewers for their numerous helpful comments and suggestions.
-
-# Data availability
-
-All data and code are available here: <https://github.com/grasshoppermouse/strengthmating>
-
-# Supplementary Information
-
-## Pilot study
-
-::: {#stbl-stats}
-`{r} tbl_stack(list(con, cat))`
-
-Descriptive statistics of pilot study sample.
-:::
-
-### Models of mating success with socioeconomic, health, hormone, and physical activity controls
 
@@
 
-### The negative association of numbers of lifetime sexual partners on partnered status
+```{r}
+forestplot2 <- function(..., title=""){
+  models <- list(...)
+  models <- map(models, \(m) {m$df.residual = Inf; return(m)})
+  models <- c(models, list(
+    intercept = F,
+    facet = F,
+    dodgewidth = .8,
+    modelnames = mnames,
+    varnames = vnames,
+    size = 0.7,
+    linewidth = 0.7
+  ))
 
-::: {#sfig-tradeoff}
-```{r}
-plot_lifetime_coefs
+  p <- do.call(forestplot, models)
+
+  p +
+    guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
+    labs(title = title) +
+    theme_minimal() +
+    theme(axis.text.y = element_text(size = 9), plot.title = element_text(size = 11))
+}
 ```
 
-The adjusted odds ratios of numbers of lifetime partners on partnered status. In these models, lifetime partners was divided by 10, the interquartile range.
-:::
+| Outcome            | Strength coefficient | Strength X Sex coefficient |
+|--------------------|:--------------------:|:--------------------------:|
+| Partnered status   |         0.99         |            0.90 (0.67)            |
+| Lifetime partners  |         0.99         |            0.90 (0.66)            |
+| Past year partners |         0.80         |            0.30 (0.21)            |
 
-### Correlation matrices of study variables
-
-::: {#sfig-cor}
-```{r}
-ggcorrplot(cor_mat_cors, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
-```
-
-Correlation matrix of study variables (both sexes). Rows and columns ordered by hierarchical clustering.
-:::
-
-::: {#sfig-fcor}
-```{r}
-ggcorrplot(cor_mat_cors_f, hc.order = T, hc.method = 'ward.D', title = 'Females')
-```
-
-Correlation matrix of study variables for females only. Rows and columns ordered by hierarchical clustering.
-:::
-
-::: {#sfig-mcor}
-```{r}
-ggcorrplot(cor_mat_cors_m, hc.order = T, hc.method = 'ward.D', title = 'Males')
-```
-
-Correlation matrix of study variables for males only. Rows and columns ordered by hierarchical clustering.
-:::
-
-### Regression tables
-
-**Lifetime number of partners**
-
-:::{#stbl-lifetime}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_lifetime, 
-    "Socioeconomic" = msoc1, 
-    "Health" = mheal1, 
-    "Activity" = mphys1,
-    "Hormone" = mhor1
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Quasi-Poisson Regression Models of Lifetime Number of Sexual Partners
-:::
-
-**Past year number of partners**
-
-:::{#stbl-pastyear}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_pastyear, 
-    "Socioeconomic" = msoc2, 
-    "Health" = mheal2, 
-    "Activity" = mphys2,
-    "Hormone" = mhor2
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Quasi-Poisson Regression Models of Past Year Number of Sexual Partners
-:::
-
-**Age at first sex**
-
-:::{#stbl-agefirstsex}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_agefirst, 
-    "Socioeconomic" = msoc3, 
-    "Health" = mheal3, 
-    "Activity" = mphys3,
-    "Hormone" = mhor3
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Gaussian Regression Models of Age at First Sex
-:::
-
-**Partnered status**
-
-:::{#stbl-partnered}
-```{r}
-modelsummary(
-  list(
-    "Anthropometric" = m_partnered, 
-    "Socioeconomic" = msoc4, 
-    "Health" = mheal4, 
-    "Activity" = mphys4,
-    "Hormone" = mhor4
-    ), 
-  gof_map = "nobs"
-  )
-```
-
-Logistic Regression Models of Partnered Status
-:::
-
-# References
+: Power to detect our predicted effects.
+Strength coefficients (effect sizes) were set to 3/4 of those we found in our pilot study, interaction coefficients were set to the same magnitude (but opposite sign) of the strength coefficients (value in parentheses is power to detect 3/4 of that effect), sample size was set to that of our unobserved data, and $\alpha = 0.05$.
+{#tbl-power}

### 2024-09-14 paper.qmd
Minor edits. Updated index.html

--- 2024-09-13 paper.qmd
@@
 We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study.
-These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
+DXA data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body.
 Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009.
@@
 We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex.
-In the expanded models with additional controls, all lean masses remained a significant predictor of energy and protein intake.
+In the expanded models with additional controls, all lean masses remained significant predictors of energy and protein intake.
 See @sfig-arms - @sfig-total.
@@
 Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993].
-In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis.
+In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners (`{r} signif(100*sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)/nrow(designsH$d.design.adults$variables), 3)`%) on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis.
 Ideally, however, analyses would account for the full range of variation in sexual behavior.
 Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions.
-However, even our confirmatory results were based on generalized linear models, yet there might be important non-linear relationships that would alter interpretations.
+However, even our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.
 
@@
 Specifically, strength in forearm and shoulder is critical for effective use of projectile hunting technology such as the bow and arrow [@ertan2003] and javelin [@kim2014].
-Indeed, @apicella2014 found that upper body strength predicted both hunting reputation and reproductive success in Hadza men.
+Indeed, upper body strength predicted both hunting reputation and reproductive success in Hadza men [@apicella2014].
 The similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular.

### 2024-09-15 paper.qmd
Minor adjustment of heading levels and reorganization of the SI.

@@
 
-## Post Hoc Analyses
+# Post Hoc Analyse
+
 We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study.
@@
 
-## Discussion
+Finally, we fit our grip strength models on the combined data (2011-2014).
+Results qualitatively correspond to those seen in the pilot study, which is not surprising because our models were developed on the pilot data.
+See @sfig-matingGH - @sfig-intakeGH.
+
+# Discussion
 
@@
 
-## Conclusion
+# Conclusion
 
@@
 
+## Post hoc results
+
+### Lean mass coefficient plots
+
+#### Arm lean mass
+
+::: {#sfig-arms}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for arm lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+#### Leg lean mass
+
+::: {#sfig-legs}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Leg lean mass", "leg_lean_centered", "leg_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for leg lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+#### Trunk lean mass
+
+::: {#sfig-trunk}
+
+```{r}
+#| fig-width: 11
+lean_mass_plot("Trunk lean mass", "trunk_lean_centered", "trunk_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for trunk lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
+#### Total lean mass
+
+::: {#sfig-total}
+```{r}
+#| fig-width: 11
+lean_mass_plot("Total lean mass", "total_lean_centered", "total_lean_sex_centered", mating_only = F)
+```
+
+Coefficient plots for total lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
+:::
+
 ### Sex-specific models
@@
 
-## Combined G and H series data (2011-2014)
+### Combined G and H series data (2011-2014)
 
@@
 
-## Lean mass coefficient plots
-
-### Arm lean mass
-
-::: {#sfig-arms}
-
-```{r}
-#| fig-width: 11
-lean_mass_plot("Arm lean mass", "arm_lean_centered", "arm_lean_sex_centered", mating_only = F)
-```
-
-Coefficient plots for arm lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
-:::
-
-### Leg lean mass
-
-::: {#sfig-legs}
-
-```{r}
-#| fig-width: 11
-lean_mass_plot("Leg lean mass", "leg_lean_centered", "leg_lean_sex_centered", mating_only = F)
-```
-
-Coefficient plots for leg lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
-:::
-
-### Trunk lean mass
-
-::: {#sfig-trunk}
-
-```{r}
-#| fig-width: 11
-lean_mass_plot("Trunk lean mass", "trunk_lean_centered", "trunk_lean_sex_centered", mating_only = F)
-```
-
-Coefficient plots for trunk lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
-:::
-
-### Total lean mass
-
-::: {#sfig-total}
-```{r}
-#| fig-width: 11
-lean_mass_plot("Total lean mass", "total_lean_centered", "total_lean_sex_centered", mating_only = F)
-```
-
-Coefficient plots for total lean mass vs. mating success outcomes, energy and protein intake, and immunity (white blood cell count).
-:::
-
 # References

### 2024-09-20 paper.qmd
Minor change of title and other edits.

@@
 ---
-title: "Strength, mating success, and immune and dietary costs in a population sample of US women and men: a registered report"
+title: "Strength, mating success, and immune and nutritional costs in a population sample of US women and men: a registered report"
 
@@
 
-Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
+Theoretically predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
 According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex:strength interaction (the hypothesis does allow some positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
@@
 #| tbl-cap: "Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant."
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; ns indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant."
 
@@
    "Lifetime partners",     "+",             "-",       
-  "Past year partners",     "0",             "0",       
-    "Age at first sex",     "0",             "0",       
-            "Immunity",     "-",            "0",       
-              "Energy",     "0",              "",       
+  "Past year partners",     "ns",            "ns",       
+    "Age at first sex",     "ns",            "ns",       
+            "Immunity",     "-",             "ns",       
+              "Energy",     "ns",             "",       
              "Protein",     "+",              "",       
@@
 #| tbl-cap: "Predictions.
-A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
+A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; ns indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant.
 ✓: Supported in confirmatory study.
 ❌: Not supported in confirmatory study.
-*The posthoc results found a significant negative association between lean mass and immunity, and a positive association with energy."
+*The posthoc results found a significant negative association between lean mass and immunity, and a positive association grip strength and lean mass with energy."
 
 tribble(
-              ~Outcome, ~Prediction, ~Result, ~Prediction, ~Result,
+              ~Outcome, ~Prediction, ~Supported, ~Prediction, ~Supported,
     "Partnered status",     "+",       "✓",       "-",       "✓",
    "Lifetime partners",     "+",       "✓",       "-",       "❌",
-  "Past year partners",     "0",       "✓",       "0",       "✓",
-    "Age at first sex",     "0",       "✓",       "0",       "✓",
-            "Immunity",     "-",       "❌*",      "0",       "✓",
-              "Energy",     "0",       "❌*",        "",        "",
+  "Past year partners",     "ns",      "✓",       "ns",       "✓",
+    "Age at first sex",     "ns",      "✓",       "ns",       "✓",
+            "Immunity",     "-",       "❌*",     "ns",       "✓",
+              "Energy",     "ns",      "❌*",      "",        "",
              "Protein",     "+",       "✓",        "",        ""
@@
 
-# Post Hoc Analyse
+# Post Hoc Analyses
 
@@
 Finally, we fit our grip strength models on the combined data (2011-2014).
-Results qualitatively correspond to those seen in the pilot study, which is not surprising because our models were developed on the pilot data.
+Results qualitatively correspond to those in the pilot study, which is not surprising because our models were developed on the pilot data.
 See @sfig-matingGH - @sfig-intakeGH.
@@
 In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners (`{r} signif(100*sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)/nrow(designsH$d.design.adults$variables), 3)`%) on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis.
-Ideally, however, analyses would account for the full range of variation in sexual behavior.
+Ideally, however, analyses would account for the full range of variation in sexual behavior (we fit our confirmatory models on the complete data, and all results were very similar to those we report).
 Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions.
-However, even our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.
+However, our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.
 

### 2024-09-21 paper.qmd
Deleted old abstract

@@
 
-<!-- Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection.
-It is unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US (N = 4384), we will test whether grip strength, a proxy for upper body strength and physical formidability, is a positive predictor of self-reported measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, age at first intercourse, and current partnership status) in both males and females, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related potential confounds.
-We will also test if there is a positive relationship between grip strength and dietary energy and protein intake, and a negative relationship between grip strength and native immune function.
-This study replicates an earlier study of US males [@lassek2009] with numerous additional control variables, and it additionally tests these relationships in females.
--->
-
 Theory and evidence suggest that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of innate immunity, likely due to an evolutionary history of sexual selection.

### 2024-09-23 paper.qmd
Grammar fixes. Changed 'lifetime number of partners' to 'lifetime sexual partners' and 'past year number of partners' to 'past-year sexual partners' throughout.

--- 2024-09-20 paper.qmd
@@
 It is therefore unknown if females experience a similar tradeoff.
-Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of the US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the association of upper body strength, proxied by grip strength, with mating success is significantly positive for males and significantly less so for females.
-We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past year number of sexual partners or age at first intercourse.
-We found consistent evidence for a $grip\ strength\times sex$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime number of partners).
+Using data from the 2013-2014 phase of the National Health and Nutrition Examination Survey (NHANES), a large nationally representative sample of US adults (N = `{r} nrow(designsH2$d.design.adults)`), we tested the prediction from the sexual selection hypothesis that the association of upper body strength, proxied by grip strength, with mating success is significantly positive for males and significantly less so for females.
+We found a main effect of strength on mating success proxied by lifetime number of sexual partners and current partnered status, but not past-year number of sexual partners or age at first intercourse.
+We found consistent evidence for a $grip\ strength\times sex$ interaction on partnered status, such that strength was significantly more important for male partnered status than female (but no significant interaction for lifetime sexual partners).
 We also tested for tradeoffs of upper body strength with immune and dietary intake and found a positive relationship between grip strength and protein and energy intake, but no significant relationship between grip strength and innate immune function.
@@
 
-Theoretically predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women.
-According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex:strength interaction (the hypothesis does allow some positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
+Theoretically predicted relationships between upper body strength and various indices of mating success, such as numbers of sexual partners, partnered status, and age at first sex, for both men and women.
+According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex x strength interaction (the hypothesis does allow for some degree of positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
 :::
@@
 
-In order to refine our hypotheses and statistical models for our confirmatory study, which will use data we have not yet observed, we first conducted a pilot study whose results we report here.
+In order to refine our hypotheses and statistical models for our confirmatory study, we first conducted a pilot study whose results we report here.
 
@@
 
-The pilot study uses the 2011-2012 dataset, whereas the 2013-2014 dataset has been held out for confirmatory analysis pending in principle acceptance of this registered report.
+The pilot study used the 2011-2012 dataset, whereas the 2013-2014 dataset was held out for confirmatory analysis pending in principle acceptance of this registered report.
 Grip strength, our proxy for physical formidability and our key predictor variable, was only collected in these years [@smith2022 investigated the association of grip strength with depression in the 2013-14 data, but did not investigate or observe the relationship between grip strength and any of the mating success outcomes described below].
 We include data from US adults between the ages of 18 and 60, the years when grip strength is the most stable [@hogrel2015].
-For each model, participants will be included if they have complete data for each predictor and outcome variable.
+For each model, participants were included if they had complete data for each predictor and outcome variable.
 
@@
 Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with male reproductive success under ancestral conditions [@perusse1993].
-We operationalized mating success in three ways following @lassek2009: 1) Total sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Total sexual partners in the past year, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
+We operationalized mating success in three ways following @lassek2009: 1) Lifetime sexual partners was a count based on responses to the question: "In your lifetime, with how many men/women have you had any kind of sex?" 2) Past-year sexual partners, from the question: "In the past 12 months, with how many men/women have you had any kind of sex?" Because of the way these questions are framed, (e.g.
 female participants were asked about their male partners and vice versa) these variables represent heterosexual partners (however, it is not necessarily the case that all of the participants included in these models identified as heterosexual).
@@
 
-Lifetime and past year numbers of sexual partners index success through short term mating strategies.
-Lifetime numbers are probably a more precise proxy of the use of a short-term mating strategy as these are integrated over the entire life, whereas past year numbers likely fluctuate depending on partnered status and other transient factors.
+Lifetime and past-year numbers of sexual partners index success through short term mating strategies.
+Lifetime numbers are probably a more precise proxy of the use of a short-term mating strategy as these are integrated over the entire life, whereas past-year numbers likely fluctuate depending on partnered status and other transient factors.
 However, both are somewhat ambiguous measures as they could also represent repeated rejection, or a tendency to exaggerate or understate the actual numbers.
@@
 All models also controlled for age in years and partnered status (not included in models where partnered status was the outcome variable).
-
-The interaction between sex and grip strength was crucial for our models since we want to test if the effect of grip strength on mating and immune outcomes is different for women than for men.
+Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable [@gelman2008].
+Standardized variables are denoted in figures as 'Variable Name (S)'.
+
+The interaction between sex and grip strength was crucial for our models since we wanted to test if the effect of grip strength on mating and immune outcomes is different for women than for men.
 Grip strength is highly sexually dimorphic, however, with about 90% of men stronger than about 90% of women [@smith2022], and is thus highly collinear with sex, a problem for regression modeling and interpretation.
@@
 We interacted age with sex since menopause occurs within the age range of our participants.
-For models of past year number of sexual partners we included an interaction between partnered status and sex-specific grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
-Continuous-valued predictor variables were centered and standardized by 2 standard deviations which approximately matches the variation in binary variables like sex, thus making the regression coefficients more comparable [@gelman2008].
+For models of past-year sexual partners we included an interaction between partnered status and sex-specific grip strength since being currently partnered would plausibly impact mating behavior in the past year regardless of strength.
 
@@
 Finally, we treated our pilot study as an exploratory study in which we fit numerous models not reported here.
-The models we report in the pilot study and which we will test exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we described above, the results of our exploratory analyses, and our constraints.
+The models we report in the pilot study and which we then tested exactly in the confirmatory study represent those that we consider to best test the hypotheses in @lassek2009 based on the theories we described above, the results of our exploratory analyses, and our constraints.
 
@@
 However, @lassek2009 utilized stepwise regression to automatically eliminate candidate predictors from their models which has since been found to overfit data and therefore estimated coefficients fail to replicate in future samples [@smith2018a].
-Instead, we will fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses.
+Instead, we fit generalized linear regression models (GLMs) with prespecified treatment and control variables chosen based on theoretical considerations and our exploratory analyses.
 Control variables included a much wider range of theoretically motivated potential confounds than used by @lassek2009, organized into themes.
@@
 
-We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (total number of partners, past year number of partners, and age at first intercourse) predicted by age, sex, sex-specific grip strength, and partnered status.
+We first derived a simple model based on that reported in Table 2 in @lassek2009 for each outcome measure (lifetime sexual partners, past-year sexual partners, and age at first intercourse) predicted by age, sex, sex-specific grip strength, and partnered status.
 This model also includes an anthropometric control, body mass index (BMI), calculated as $kg/m^2$, since body size could impact either strength or mating outcomes.
@@
 All analyses were completed in `{r} R.version.string`, using the survey package [`{r} packageVersion("survey")`, @lumley2021] in order to incorporate the survey sampling weights and to preserve the representative structure of the sample.
-We modeled lifetime number of partners, past year numbers of partners, and white blood cell count using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
+We modeled lifetime sexual partners, past-year sexual partners, and white blood cell count using a quasi-Poisson generalized linear regression model (glm), since these variables are overdispersed count data.
 We modeled age at first sexual intercourse and energy and protein intakes with a Gaussian glm.
@@
 ```
+For ease of comparison, pilot results described here are displayed in conjunction with confirmatory results in the Confirmatory Results section below.
 
 Our initial analyses found little support for a sex difference in the effect of sex-specific grip strength on mating success, contrary to the sexual selection hypothesis.
-We traced this failure to the unusually high number of lifetime sex partners reported by some men and women.
-We discovered that whereas women reported a median number of `{r} female_median` lifetime sex partners, and men `{r} male_median` partners, there were `{r} high_partners` individuals who reported 100 or more partners, a few of whom reported 1000 or more (@fig-ecdf).
-We decided to restrict our sample to individuals with fewer than 100 lifetime sex partners for several reasons.
+We traced this failure to the unusually high number of lifetime sexual partners reported by some men and women.
+We discovered that whereas women reported a median number of `{r} female_median` lifetime sexual partners, and men `{r} male_median` partners, there were `{r} high_partners` individuals who reported 100 or more partners, a few of whom reported 1000 or more (@fig-ecdf).
+We decided to restrict our sample to individuals with fewer than 100 lifetime sexual partners for several reasons.
 First, support for the sexual selection hypothesis was weak if these individuals were included, but much stronger if they were excluded.
 We suspect that because our sample is large and nationally representative, it probably includes sex workers, as well as individuals who regularly use the services of sex workers, and whose high partner numbers would therefore not reflect mate attraction or intrasexual competition.
-Second, there was an unusually high number of participants reporting exactly 100 lifetime sex partners, suggesting inaccurate recollections.
+Second, there was an unusually high number of participants reporting exactly 100 lifetime sexual partners, suggesting inaccurate recollections.
 Finally, our statistical models do not easily accommodate the high degree of overdispersion caused by these few exceptionally high partner numbers, and our extensive simulations showed that high overdispersion would reduce our statistical power to detect sex differences.
@@
 #| fig-width: 10
-#| fig-cap: Cumulative distribution of the number of sex partners, by sex.
+#| fig-cap: Cumulative distribution of the number of sexual partners, by sex.
 Participants who reported 100 or more partners were removed from the analysis.
-The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this Figure and not for any analyses).
+The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sexual partners (only for this Figure and not for any analyses).
 
@@
 
-Our pilot results revealed strong positive significant main effects of sex-specific grip strength on partnered status, numbers of lifetime partners, and age at first sex after controlling for all five sets of potential confounds.
+Our pilot results revealed strong positive significant main effects of sex-specific grip strength on partnered status, lifetime sexual partners, and age at first sex after controlling for all five sets of potential confounds.
 The only exception was that the sex-specific grip strength coefficient for age at first sex was not significant after controlling for socioeconomic factors, although the sign and magnitude was very similar to those of these coefficients in the models with other control variables.
-There was no significant effect of sex-specific grip strength on the number of partners in the past year.
+There was no significant effect of sex-specific grip strength on the number of sexual partners in the past year.
 See the left column of @fig-coefs.
 
-There were also clear sex differences in the effects of sex-specific grip strength on two mating success outcomes---partnered status and lifetime number of partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
-There was also a persistent negative association of numbers of lifetime partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
-Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+There were also clear sex differences in the effects of sex-specific grip strength on two mating success outcomes---partnered status and lifetime sexual partners---with smaller effects for women compared to men, as predicted by the sexual selection hypothesis.
+There was also a persistent negative association of lifetime sexual partners on the probability of partnered status in all models of this outcome (@sfig-tradeoff).
+Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past-year sexual partners or age at first sex, contrary to the sexual selection hypothesis.
 See the right column of @fig-coefs.
@@
 
-For ease of comparison, pilot results discussed here are displayed in conjunction with confirmatory results in the Confirmatory Results section below.
-
 Using sex-specific grip strength as a proxy for muscularity, this pilot study replicated findings from @lassek2009 that muscularity is significantly positively associated with two of the four indices of mating success, including our additional index of mating success---partnered status---even after controlling for a wider range of potential confounds including socioeconomic, health, hormone, and physical activity variables.
-The exceptions were that sex-specific grip strength was not a significant predictor of past year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant sex-specific grip strength coefficients in the other models.
-Grip strength was also associated with immune, energy, and protein costs in the replication models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
+The exceptions were that sex-specific grip strength was not a significant predictor of past-year partners, nor of age at first sex in the socioeconomic control model; the coefficient of the latter, however, was similar in sign and magnitude to the significant sex-specific grip strength coefficients in the other models.
+Grip strength was also associated with immune, energy, and protein costs in the Lassek and Gaulin models, but not in the expanded control models of immune or energy costs; it remained a significant predictor of protein intake, however.
 The expanded control models included height, which might be a proxy for aspects of strength, such as lower body strength, that are not accounted for by grip strength.
 
-The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status and numbers of lifetime partners.
+The sexual selection hypothesis predicts that the effect of strength on mating success will be greater for men than for women, and this prediction was exceptionally well-supported for partnered status and lifetime sexual partners.
 In humans, biological fitness depends critically on extensive investment in offspring which typically takes place in long-term partnerships.
@@
 Thus, our pilot study supports the hypothesis that muscularity increases both short- and long-term mating success for men more than it does for women.
-Number of lifetime partners was also a negative predictor of partnered status (@sfig-tradeoff), suggesting a tradeoff in pursuing these two reproductive strategies.
+Lifetime sexual partners was also a negative predictor of partnered status (@sfig-tradeoff), suggesting a tradeoff in pursuing these two reproductive strategies.
 
@@
 
-Sex-specific grip strength was also not significantly associated with the number of partners in the last year, despite controlling for partnered status, which was a significant negative predictor.
+Sex-specific grip strength was also not significantly associated with the number of sexual partners in the past year, despite controlling for partnered status, which was a significant negative predictor.
 Approximately 60% of the sample was partnered (@stbl-stats), however, and stronger individuals, especially men, were more likely to be partnered.
 The pool of unpartnered individuals who might pursue multiple mating opportunities was therefore smaller and skewed toward lower grip strength individuals, and variation in this outcome over the relatively short time period (1 year) was low.
-Nevertheless, the non-significant sex-specific grip strength coefficients were all positive, raising the possibility that a higher powered study might successfully detect a positive effect of sex-specific grip strength on number of partners in the last year.
+Nevertheless, the non-significant sex-specific grip strength coefficients were all positive, raising the possibility that a higher powered study might successfully detect a positive effect of sex-specific grip strength on number of partners in the past year.
 
@@
 
-In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime number of sexual partners, but not last year number of partners, nor age at first sex after controlling for ethnicity.
-We also predict a significant interaction with sex such that sex-specific grip strength will be a stronger predictor of male than female partnered status and number of lifetime partners, but not number of last year partners nor age at first sex.
+In a large, nationally representative US sample, we predict that formidability will be significantly positively associated with mating success as indexed by partnered status and lifetime number of sexual partners, but not past-year number of partners, nor age at first sex after controlling for ethnicity.
+We also predict a significant interaction with sex such that sex-specific grip strength will be a stronger predictor of male than female partnered status and lifetime number of sexual partners, but not number of past-year sexual partners nor age at first sex.
 We also predict that there will be a significant negative association between formidability and immune function.
@@
    "Lifetime partners",     "+",             "-",       
-  "Past year partners",     "ns",            "ns",       
+  "Past-year partners",     "ns",            "ns",       
     "Age at first sex",     "ns",            "ns",       
@@
 Effects in replication studies are often smaller than in the original studies [@klein2018].
-We therefore used simulations to estimate our power to detect grip strength and $grip\ strength\times sex$ coefficients that are 25-100% as large as the ones observed in our pilot study for partnered and lifetime partners, with a sample size equal to that in our unobserved data.
+We therefore used simulations to estimate our power to detect grip strength and $grip\ strength\times sex$ coefficients that are 25-100% as large as the ones observed in our pilot study for partnered and lifetime sexual partners, with a sample size equal to that in our unobserved data.
 We have high power to detect effects at least 75% as large as the ones observed in our pilot study, and reasonable power to detect those that are at least 50% as large (@stbl-power).
@@
 ## Confirmatory Study Results
+
+### Descriptive statistics
+Sample characteristics were similar in the confirmatory dataset to those in the pilot dataset.
+See @stbl-confirmstats for weighted means, standard deviations, and standardized mean differences of all variables for men and women.
 
@@
 
-Our confirmatory results revealed a strong positive significant main effect of sex-specific grip strength on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+Our confirmatory results revealed a strong positive significant main effect of sex-specific grip strength on partnered status and lifetime number of partners after controlling for all five sets of potential confounds.
 We found mixed evidence for a main effect of sex-specific grip strength in the age of first sex models, as it was not a significant predictor after controlling for socioeconomic, anthropometric, or physical activity controls.
-Strength was not a predictor of past year number of sexual partners in any control model.
+Grip strength was not a predictor of past-year number of sexual partners in any control model.
 
@@
 The interaction between sex and sex-specific grip strength on lifetime number of sexual partners was not replicated in the confirmatory data.
-Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past year number of partners or age at first sex, contrary to the sexual selection hypothesis.
+Finally there was no evidence for a sex difference in the effects of sex-specific grip strength on past-year number of partners or age at first sex, contrary to the sexual selection hypothesis.
 See @fig-coefs for coefficients of sex-specific grip strength and $grip\ strength\times sex$ interaction for each proxy of mating success in both pilot and confirmatory data.
@@
    "Lifetime partners",     "+",       "✓",       "-",       "❌",
-  "Past year partners",     "ns",      "✓",       "ns",       "✓",
+  "Past-year partners",     "ns",      "✓",       "ns",       "✓",
     "Age at first sex",     "ns",      "✓",       "ns",       "✓",
@@
 
-Our post hoc analyses revealed that coefficients for sex-specific arm lean mass were generally similar in direction to those of sex-specific grip strength discussed above, including a strong positive significant main effect on partnered status and numbers of lifetime partners after controlling for all five sets of potential confounds.
+Our post hoc analyses revealed that coefficients for sex-specific arm lean mass were generally similar in direction to those of sex-specific grip strength discussed above, including a strong positive significant main effect on partnered status and lifetime number of partners after controlling for all five sets of potential confounds.
 In addition, sex-specific arm lean mass had a significantly smaller association with partnered status for women than men, consistent with the sexual selection hypothesis.
 We found mixed evidence for a main effect of sex-specific lean mass in the age of first sex models, as these were not significant predictors after controlling for socioeconomic or physical activity controls.
-As with sex-specific grip strength, sex-specific arm lean mass was not a predictor of past year number of sexual partners in any control model.
+As with sex-specific grip strength, sex-specific arm lean mass was not a predictor of past-year number of sexual partners in any control model.
 Sex-specific leg, trunk, and total lean mass were not consistently related to any mating success outcome.
@@
 In the Stage 2 Confirmatory Study, we found consistent support for a positive relationship between both sex-specific combined grip strength and sex-specific arm lean mass, and two measures of mating success: being currently partnered, and lifetime number of sexual partners for men, controlling for numerous possible confounds.
-We found no consistent association between either sex-specific combined grip strength or any sex-specific lean masses and age at first sex or number of past year sexual partners.
+We found no consistent association between either sex-specific combined grip strength or any sex-specific lean masses and age at first sex or number of past-year sexual partners.
 
@@
 It might be that stronger women require less male investment, or can take more physical risks, and so instead benefit from greater partner numbers through, e.g.
-genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding or leaving costly long term partnerships.
+genetic bet-hedging, forging relationships with multiple males, ability to conduct a more extensive search for a high quality long term mate, or through avoiding or leaving costly long-term partnerships.
 It might be the case that there are some sex-specific confounds that we failed to control for, or that our models are otherwise misspecified.
@@
 Our results most strongly support the importance of upper body strength to male long-term mating success, which is the basis of biparental care in human reproduction [@Gettler_2020; @Bribiescas_2012].
-This benefit would compensate for the energy, protein, and immune costs associated with strength in our confirmatory results and lean masses in our *post hoc* results.
-Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
+This benefit would compensate for the energy, protein, and immune costs associated with strength in our confirmatory results and lean masses in our post hoc results.
+Across evolutionary time, fitness for both men and women was heavily dependent on successfully obtaining long-term partners, thereby reaping the efficiency benefits of a sexual division of labor [@Quinlan_2007; @Quinlan_2008].
 More specifically, human children require provisioning for an extended period [@kaplan2000; @davison2021; @davison2022], and paternal care is an important predictor of child outcomes [@Winking_2011; @Starkweather_2021].
@@
 tt(df_pwr) |>
-  group_tt(j = list("Partnered" = 2:3, "Lifetime partners" = 4:5, "Last year partners" = 6:7)) |>
+  group_tt(j = list("Partnered" = 2:3, "Lifetime partners" = 4:5, "Past-year partners" = 6:7)) |>
   format_tt(j = 1:7, digits = 2, num_zero = T) |>
@@
 
-### The negative association of numbers of lifetime sexual partners with partnered status
+### The negative association of lifetime number of sexual partners with partnered status
 
@@
 
-The adjusted odds ratios of numbers of lifetime partners on partnered status. In these models, lifetime partners was divided by 10, the interquartile range.
+The adjusted odds ratios of lifetime number of sexual partners on partnered status. In these models, lifetime partners was divided by 10, the interquartile range.
 :::
@@
   format_tt(fn = hagenutils::signif2) |>
-  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
+  group_tt(i = list("ΔPartnered probability" = 1, "ΔLifetime partners" = 3, "ΔPast-year partners" = 7, "ΔAge first sex (years)" = 11), indent = 2)
 ```
@@
 
-Pilot: A comparison of grip strength and $grip\ strength\times sex$ coefficients with the original model specification for lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
+Pilot: A comparison of grip strength and $grip\ strength\times sex$ coefficients with the original model specification for lifetime sexual partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new).
 Both models are quasi-Poisson models.
